### 2021-06-07 docs/presentations/Xia_horizontal_genetic_effects/presentation.Rmd
add first presentation from meeting with Zoltan and setting up project

---
title: "ProxyMR Project: Update and Review of Tenesa et al. IGE manuscript"
author: "Jenny Sjaarda"
date: "04/06/2021 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    nature:
      highlightLines: true
      countIncrementalSlides: false
---

class: inverse, center, middle

<center>
<img src = "./images/title_and_abstract.png" align="middle" style="border: none; box-shadow: none; text-align: center;">
</center>

---

## Introduction

- **Indirect genetic effects (IGEs)**: the effect of a genotype of one individual on the phenotype of other individuals.

- **Vertical IGEs**: IGEs across generation.
  - For example, maternal genetic effects on their offspring phenotypes (such as birthweight).
  - Non-transmitted parental genotypes have also been described to influence offspring phenotypes (e.g. educational attainment).
  - Suggest vertical IGEs can be mediated through both *biological* and *social beahviors*.

- **Horizontal IGEs**: IGEs within generation. 
  - Reported for school friends in relation to educational attainment.
  - Unknown how much these observations are spread across traits, particularly in *partners* and *siblings*.

---

## Objectives and study overview

- Search for evidence of IGEs in UKB couples: 
  - 105 complex traits. 
  - In 80,889 couples. 
  - European ancestry. 
- Replication:
  - In 8,144 sibling pairs of European ancestry.
  - 3,752 couples of mixed ancestry (one European and the other non-European).
- Linear Mixed Model (LMM) approach to estimate **Partner indirect heritability** ( $h^2_{Partner}$ ): proportion of phenotypic variance observed for individuals that is explained by the additive genetic effect of their social partners.
- **Traits studied**: anthropometric traits, lifestyle choices, mental heath and late-onset diseases. 
- **Demonstrate**: $h^2_{Partner}$ captures IGEs beyond just assortative mating.
- ~50% of traits had a significant $h^2_{Partner}$ (this could be simply AM) and about ~25% of those showed evidence of IGEs.

---

## Couple definitions

- As described previously in:
```{r, out.width='50%', fig.align='center', fig.cap='', echo=F}
knitr::include_graphics('./images/Genetic_determinants_height-med_mate_choice.png')
```
- 138 couple pairs with a genomic coefficient of relationship above 0.025 were removed.
- 7 related couples (for example, two couples made up of two relatives from one family and two relatives from another family) were removed also from analysis *(we did not apply this filter)*.
- Resulting in **80,889** couples for discovery.

???

*Tenesa, A., Rawlik, K., Navarro, P. & Canela-Xandri, O. Genetic determination of height-mediated mate choice. Genome Biol. 16, 269 (2016).*

So far we have not created a discovery set nor did we removed related couples.

---

## Maximum-likelihood analysis

- Estimated proportion of phenotypic variation explained by individuals' own + partner's genotype, respectively.

- Used genomic restricted maximum-likelihood (GREML) analyis is DISSECT.

- The main **model** takes the following form:

$$
\begin{aligned}
\textbf{y} &= X\beta + \textbf{g}_o + \textbf{g}_p + \varepsilon
\end{aligned}
$$

- Where:
  - $\textbf{y}$ = vector of phenotypes
  - $\textbf{X}$ = design matrix of fixed effect B
  - $\textbf{g}_o$ = effect of the individual's own genotype
  - $\textbf{g}_p$ = effect of the partners genotypes 
  - $\varepsilon$ = residual effect
  - $\textbf{g}_o$, $\textbf{g}_p$ and $\varepsilon$ are random effects.

---

## Maximum-likelihood analysis

- **Fixed-effect covariates**:
  - sex + 
  - genotype batch + 
  - assessment centre + 
  - age + age-squared + 
  - PC 1-20

- **Model interpretation**: 
  - Proportion of phenotypic heritability explained by individuals' own genotype:

    $$h^2_{Own} =  \dfrac{\sigma^2_{o}}{\sigma^2_{o} + \sigma^2_{p} + \sigma^2_{\epsilon}}$$

  - Proportion of phenotypic heritability explained by individuals' partner's genotype:

    $$h^2_{Partner} =  \dfrac{\sigma^2_{p}}{\sigma^2_{o} + \sigma^2_{p} + \sigma^2_{\epsilon}}$$

---

## Maximum-likelihood analysis

- **Statistical test**: 
  - For each trait, tested against two null hypotheses. 
  - $\sigma^2_{o}= 0$ and $\sigma^2_{p} = 0$. 
  - Likelihood ratio test against the appropriate reduced model. 
  - Two component mixture, a delta spike at 0 and $\chi^2_{d.f.=1}$.
  - Each component with probability of 0.5 as null distribution using DISSECT.

- **Sex-specific analysis**:
  - Fitted sex-specific models for sex-specific phenotypes. 
  - Fitted sex-stratified models for non-sex-specific phenotypes with estimates of $h^2_{Partner}$ significant with *P* < 0.05.

- **Couple-shared phenotype**: $h^2_{Male}$ and $h^2_{Female}$ were defined as SNP-heritability contributed by male genotype and female genotype, respectively.

---

## Prediction and replication:

- Learned the marker effects in 80K, EU couples.

- Used  the direct PGS ( $\hat g_{o}$ ) and indirect PGS ( $\hat g_{p}$ ) of sibling and mixed-couples in DISSECT.

- *Non-binary traits*. Pearson correlation between phenotype and predictions, comparing Cor (${y}$, $\hat g_{o}$)  with Cor( ${y}$, $\hat g_{o}$ + $\hat g_{p}$), to see if including partner IGEs increases correlation.

- *Binary traits*. Prediction accuracy assessed with ROC curves using AUC statistic. Comparing model with and without $\hat g_{p}$.

- *Couple and sibling replication analysis*. Attempted to remove correlations due to own and partner's direct genetic effects on focal phenotype and partner's focal phenotype. Conservative approach as it removes IGEs mediated through the same trait.

---

#### Couple and sibling replication analysis.

```{r, out.width='80%', fig.align='center', fig.cap='', echo=F}
knitr::include_graphics('./images/couple_sib_analysis.png')
```

---

## Results: Evidence for partner indirect heritability

- Estimated $h^2_{Partner}$ for **91** traits expressed in women and men. 
- 28 (31%) found to have $h^2_{Partner}$ BF sig (*P* < 105; 105 total traits tested including sex-specific traits).
- 6/6 anthropometric traits.
- 11/18 dietary trait were categories most enriched. 
- Other notable significant traits: 
  - Educational attainment
  - Time spent watching TV
  - Smoking status
  - Disease and mental-health related traits. 
- Given traits are correlated, also applied FDR threshold, and 47 traits pass significance. 
- Used *P* < 0.05 as threshold for inclusion in downstream analyses, comprising 51 traits.

---

## Table 1

```{r, out.width='100%', fig.align='center', fig.cap='Table 1 (abbreviated)', echo=F}
knitr::include_graphics('./images/Table1_1.png')
```

---

#### Figure 1

```{r, out.width='60%', fig.align='center', fig.cap='', echo=F}
knitr::include_graphics('./images/Figure1.png')
```

Estimates of direct ( $h^2_{Own}$ ) and indirect ( $h^2_{Partner}$ ) genetic effects in couples for 91 traits expressed in females and males. Each dot represents a trait. The blue dots represent traits with a significant estimate of $h^2_{Partner}$ (*P* < 0.05) and the blue dotted line shows the associated linear trend line. The grey dots represent traits with a non-significant estimate of $h^2_{Partner}$ and the grey dotted line shows the associated linear trend line. Liability-scale heritability is shown for binary traits.

---

## Results: Evidence for partner indirect heritability

- The variance explained by the partner genotype mainly came from the environment of the individual (that is, from the environmental residual variance)

- Whereas the estimate of individuals’ own genetic variance was largely unchanged compared to the base model containing only the person’s genotype.

- Also found that, generally, including the partner’s genotype in a prediction model improved prediction accuracy

???

(Supplementary Fig.2)

---

## Results: Differences between sexes

- Restricting to only 51 nominal traits.
- Found that $h^2_{Partner (Female \to Male)}$ was significantly larger than $h^2_{Partner (Male \to Female)}$ for 6 traits. 
- Mainly dietary- and obesity-related traits such as cereal intake (P = 3.89 × 10−3), beef intake frequency (P = 9.45 × 10−3) or waist-to-hip ratio (P = 5.53 × 10−5). 
```{r, out.width='90%', fig.align='center', fig.cap='Figure 2', echo=F}
knitr::include_graphics('./images/Figure2.png')
```

---

## Results: Differences between sexes

- No evidence of partner contribution on sex-specific diseases.
- Found significantly different estimates between male and females on couple-shared traits (e.g. number of children).

```{r, out.width='100%', fig.align='center', fig.cap='', echo=F}
knitr::include_graphics('./images/Table1_2.png')
```

---

## Results: IGES under assortative mating

- Definitions: 
  - **Direct assortative mating**: assortment based on trait of interest.
  - **Indirect assortative mating**: assortment based on traits other than the trait of interest. 
- Sought to assess to what degree the observed $h^2_{Partner}$ could be explained by a model that does not invoke the existence of IGEs:
  1. Assessed the fit of the data to theoretical predictions under a pure model of assortative mating where $h^2_{Partner}$is generated exclusively though direct assortment. 
  2. Used permutations to estimate the expected $h^2_{Partner}$ in the European couples under the same strength of assortment. 
  3. Examine whether the IGEs learned in EU couples also explained variation in other ancestries and siblings.

- Performed meta anlaysis of all results.

---

## Results: Theoretical expectation of the partner heritability under direct assortative mating.

- Under direct assortative mating, the expectation is: $h^2_{Partner} = r^2_{CP}h^2_{Own}$.

- Inflated values of $h^2_{Partner}$ compared to theoretical imply partner genotypes explain **more** than under direct assortative mating.

- Among the 51, 29 deviated for theoretical (*P* < 0.05).

- Most  significant were educational attainment, time spent watching tv and smoking status.

- Simulations suggest that estimates are explained by models involving a combination of direct assortative mating **and** IGEs.

???

Don't defined r2cp (assuming it's correlation of traits in couples)

---

## Results: Partner heritability in randomized couples with similar strength of assortment.

- Again under direct assortative mating on the focal trait, randomizing couples but keeping similar $r^2_{CP}$ should recapitulate the estimates of $h^2_{Partner}$.

- Tested the within-trait differences between these permuted couples couples compared to the real couples.

- Found marginally significant difference for 14 traits (*P* < 0.05).

- Most  differentiatied traits were time spent watching tv, smoking status and educational attainment.

- On average, estimates were 40% less in the randomized couples that in the real couples.

???

This permutation is biased towards discovering cases where the focal traits are mediated through an uncorrelated trait. When the mediator of IGEs is correlated with the focal trait (for example, the same trait), the permutation is unable to break down the correlation structure made up by assortment and IGEs, thus making it hard to detect those traits and to separate them from direct assortment (Supplementary Fig. 3). As a consequence, the permutation results are probably conservative and underesti- mate the number of traits affected by IGEs.

---

## Results: Evidence of IGEs over the expectation under assortative mating.

```{r, out.width='120%', fig.align='center', fig.cap='Figure 3.', echo=F}
knitr::include_graphics('./images/Figure3.png')
```

---

## Results: Replication in couples and evidence of IGEs in siblings.

- For the 51 traits, estimated best linear unbiased prediction (BLUPs) of the marker effect for both the direct and indirect effects using DISSECT.
- BLUPs were used to predict direct and indirect PGSs of non-EUR couples and sibling pairs. 
- Found partner indirect PGS was marginally significant with 19 out 51 traits. 
- Also removed correlations due to own and partner’s direct genetic effects on the focal phenotype as well as partner’s focal phenotype (i.e. removing effects of assortment). *This is a conservative approach, since it would also remove IGEs if they are mediated through the same trait.*
- In siblings, found that couple-estimated indirect PGS was marginally significant in 8 out of 51 traits. These could be due to vertical IGEs in parents captured by proxy.

---

## Results: Evidence of IGEs under assortative mating across follow-up analyses.

- Combined all 4 results by meta-analysis using Fisher's method:

  1. Estimate of $h^2_{Partner}$ in real couples deviates from theoretical expectation. 
  2. Estimates of $h^2_{Partner}$ in real couples is different from permuted couples
  3. Evidence of IGEs on top of assortment in mixed couples. 
  4. Evidence of IGEs on top of assortment in siblings.

- Used BF correction (# of tests = 51)

- 13 traits significant evidence of IGEs.

---

## Theoretical and simulation study

- Derived expectation of $h^2_{Own}$ and $h^2_{Partner}$ and $r^2_{cp}$ under different scenerios.
- Considered a model under two general configurations:
  1. Direct assortative mating on the trait affected by IGEs (e.g. BMI under assortment and affected by dietary preferences). 
  2. Indirect assortative mating through assortment on the trait mediating IGEs (e.g. social economics status under assortment and affecting disease risk).
```{r, out.width='80%', fig.align='center', fig.cap='', echo=F}
knitr::include_graphics('./images/Supp_Figure6.png')
```

???

We considered a model under two general configurations, direct assortative mating on the trait affected by IGEs (e.g. BMI under assortment and affected by dietary preferences) or indirect assortative mating through assortment on the trait mediating IGEs (e.g. social economics status under assortment and affecting disease risk). Our objective is to obtain an analytic expression for $h^2_{Own}$, $h^2_{Partner}$, and $r_{cp}$ , under different assumptions, which we will then compare with empirical results. The models are summarised in Supplementary Figure 6 (below).

To this end, we define a model with three traits. The trait of interest, Py, and two intermediate traits: Pa andPb which would be the trait affected by IGEs and the trait mediating IGEs, respectively (see Fig. S1). These intermediate traits are a composed of a genetic and an environmental component: P = G + E. For example, the phenotypic value of partner 1 would be: Py = Pa + phi(Pb)

---

## Theoretical and simulation study

- Derived expectation of: 
  - $h^2_{Partner}$
  - $h^2_{Own}$
  - $r_{cp}$
- In different scenarios, covering: 
  - Direct vs. indirect assortment. 
  - Presence vs absence of IGEs. 
  - Combination of these.

```{r, out.width='100%', fig.align='center', fig.cap='Table 1 (abbreviated)', echo=F}
knitr::include_graphics('./images/Supp_Table20.png')
```

---

## Theoretical and simulation study

```{r, out.width='100%', fig.align='center', fig.cap='', echo=F}
knitr::include_graphics('./images/Supp_Figure8.png')
```

---

## Theoretical and simulation study

```{r, out.width='100%', fig.align='center', fig.cap='', echo=F}
knitr::include_graphics('./images/Supp_Table21.png')
```

---

## Theoretical and simulation study

```{r, out.width='100%', fig.align='center', fig.cap='', echo=F}
knitr::include_graphics('./images/Supp_Figure9.png')
```

---

## Discussion: Summary.

- Investigated 105 traits. 
- Found IGEs are common and present among partners and relatives.
- Presence of IGEs are not simply due assortative mating.
- Results suggest that trait mediating IGEs is unlikely to be the same as the facal trait. 
- Consistent with previous data: 
  - *Kong et al.* found evidence of IGEs for height mediated through direct effects on educational attainment. 
  - *Cheesaman et al.* found no evidence of IGEs for height mediated by direct genetic effects for height.

```{r, echo=FALSE,out.width="49%", out.height="20%",fig.cap="caption", fig.show='hold',fig.align='center'}
knitr::include_graphics(c('./images/Kong_Nature_Nurture.png','./images/Cheesman_Adopted_Education2.png'))
```

---

## Discussion: Future directions and suggestions.

1. Disentangle IGEs from assortative mating:
  - Assortment creates genetic-phenotypic assocation between partners, results in false-discovery of IGEs. 
  - Attempted to whether there is evidence of IGEs on top of assortment.
  - Methods favor identifying instances here the trait mediating IGEs is **different** from focal trait.
  - For this, they suggest longitudinal studies recruiting participants earlier in life.

2. Identify genetic variants that mediate the IGEs:
  - Are these the same variants that affect the phenotype.
  - Require larger studies as $h^2_{Partner}$ << $h^2_{Own}$.

3. Understand how IGEs mediate trait of interest: 
  - Likely through behavioural changes? 
  - Potential solution to investigate adopted relatives (no relatedness or assortment challenges).

---

## Summary of previous analyses.

- Initial idea was to compare MR using PRS as proxy for phenotype at mate choice (i.e. true assortative mating phenotype).

```{r, out.width='60%', fig.align='center', fig.cap='', echo=F}
knitr::include_graphics('./images/AM_vs_convergence_MR.png')
```

- PRS was too weak a proxy to be used as phenotype at mate choice. 
- Assortative mating (AM) MR:
  - In both directions (male to female, and vice versa).
  - Binned by age.
  - Binned by time-together.

---

## Summary of status of project.

```{r, out.width='100%', fig.align='center', fig.cap='', echo=F}
knitr::include_graphics('./images/shiny_education_MR1.png')
```

---

## Summary of status of project.

```{r, out.width='100%', fig.align='center', fig.cap='', echo=F}
knitr::include_graphics('./images/shiny_education_MR2.png')
```

---

## Summary of status of project.

```{r, out.width='100%', fig.align='center', fig.cap='', echo=F}
knitr::include_graphics('./images/shiny_education_MR3.png')
```

---

## Comparison with our analysis.

- Slightly different couple definitions and filtering (to more closely compare).

- Use MR approach rather than LMM to estimate heritability contribution for both individuals own and partner (i.e. $h^2_{Partner}$ << $h^2_{Own}$).

- Created replication / validation group.

- Do not consider time together or age of couples (age included as co-variate).

- Our analysis better teases apart the AM (or IGE) effect in males vs females (especially since we are using sex-specific IVs).

- Our analysis better teases out direct AM from indirect?

- Traits analyzed: 
  - To date we have not investigated diet. 
  - We only include traits with observed phenotypic correlation in UKB.

---

## Next steps.

1. Compile all various results and (restrict to BF-sig / FDR / nominal results for downstream analyses). 
2. **Which traits have significant differences between age bins?** Run a heterogeneity test between age binned results.
3. **Which traits have significant differences between 'age'time-together bins?** Run a heterogeneity test between time-together binned results.
4. Combine age and time-together bins since these two variables are strongly correlated, so we essentially would have a grid of MR effects with one axis as average age, and the other as time together.
--

5. Employ subset of UKB with multiple time-points to evaluate influence on partners overtime (for significant phenotypes?).
--

6. Current results are all using PHESANT transformed phenotypes, we had discussed running with the raw and ranked phenotypes to see if there is an effect of the INT transformation. 
--

7. **Spatial MR**: do IGEs also extend beyond couples (and siblings?) into neighborhood. *Note: I have already computed distance between each person in UKB*
--

8. Explore dietary traits?

---

### 2021-06-07 R/functions.R

@@ -1,0 +1,2182 @@
+
+
+download_neale_files  <-  function(phenotype_ids,
+                                   irnt = TRUE,
+                                   max_regex = 200,
+                                   chunk_size = 10,
+                                   max_downloads = 500,
+                                   output_path = '/data/sgg3/data/neale_files',
+                                   reference_file = list.files(path = output_path,
+                                                               pattern = 'UKBB.*Manifest.*[.]tsv')[2],
+                                   sex = c('both_sexes', 'female', 'male'),
+                                   category = 'unsorted') {
+  if (length(phenotype_ids) > max_regex) {
+    n_downloads  =  download_neale_files(
+      phenotype_ids[1:max_regex],
+      irnt = irnt,
+      max_regex = max_regex,
+      chunk_size = chunk_size,
+      max_downloads = max_downloads,
+      output_path = output_path,
+      reference_file = reference_file,
+      sex = sex,
+      category = category
+    )
+
+    download_neale_files(
+      phenotype_ids[(max_regex + 1):length(phenotype_ids)],
+      irnt = irnt,
+      max_regex = max_regex,
+      chunk_size = chunk_size,
+      max_downloads = max_downloads - n_downloads,
+      output_path = output_path,
+      reference_file = reference_file,
+      sex = sex,
+      category = category
+    )
+  }
+  if (length(sex) > 1)
+    sex = paste(sex, collapse = '|') %>%
+      paste0('(', ., ')')
+
+  phenotype_ids  =  paste0('^', phenotype_ids, ifelse(irnt, '(_irnt|)$', '(_raw|)$')) %>%
+    paste(collapse = '|')
+
+  existing_files  =  list.files(output_path,
+                                recursive = TRUE,
+                                pattern = '[.]gz') %>%
+    str_match('[^/]+$') %>%
+    c
+
+  to_download  =  read_tsv(reference_file) %>%
+    filter(str_detect(.$'Phenotype Code', phenotype_ids)) %>%
[…]

### 2021-06-07 _targets.R

@@ -1,0 +1,171 @@
+library(targets)
+library(tarchetypes)
+library(data.table)
+source("R/functions.R")
+options(tidyverse.quiet = TRUE)
+tar_option_set(packages = c("biglm", "tidyverse", "data.table"))
+
+UKBB_dir <- "/data/sgg2/jenny/data/UKBB_raw/"
+
+household_relationships_field <- "6141_1"
+#time_at_address_file <- "/data/sgg2/jenny/data/UKBB_processed/PHESANT/ukb31459/bin1/out_bin1..tsv"
+#time_at_address_raw_file <- "/data/sgg2/jenny/data/UKBB_raw/pheno/ukb31459.csv"
+time_at_address_field <- "699"
+time_at_address_raw_field <- "699-0.0"
+
+source("code/settings.R")
+
+list(
+
+  ##################################
+  ### MAKE HOUSEHOLD PAIRS FILE ####
+  ##################################
+  tar_target(
+    household_info_file,
+    paste0(UKBB_dir,"/pheno/ukb6881.csv"),
+    format = "file"
+  ),
+  tar_target(
+    household_info,
+    read.table(household_info_file, header=T),
+  ),
+  tar_target(
+    relatives_file,
+    paste0(UKBB_dir,"/geno/","ukb1638_rel_s488366.dat", header=T),
+    format = "file"
+  ),
+  tar_target(
+    relatives,
+    read.table(relatives_file),
+  ),
+  tar_target(
+    phesant_directory_file,
+    paste0(UKBB_processed,"/PHESANT/","PHESANT_file_directory.txt"),
+    format = "file"
+  ),
+  tar_target(
+    phesant_directory,
+    read.table(phesant_directory_file, header=T),
+  ),
+  tar_target(
+    hh_pairs,
+    pairs_only(household_info),
+  ),
+  tar_target(
+    hh_pairs_kin,
+    find_kinship(hh_pairs, relatives),
+  ),
+  tar_target(
+    household_relationships_file,
+    as.character(phesant_directory[which(phesant_directory[,2]==household_relationships_field),"File"]),
+    format = "file"
+  ),
+  tar_target(
+    household_relationships,
+    fread(household_relationships_file, header=T, select=c("userId","sex",household_relationships_field), data.table=F),
+  ),
+  tar_target(
+    hh_pairs_filter,
+    filter_pairs(hh_pairs_kin, household_relationships, household_relationships_field),
+  ),
+
[…]

### 2021-06-07 code/settings.R

@@ -1,0 +1,67 @@
+#####################################################################################
+## Author: Jenny Sjaarda
+## Project: proxyMR
+##
+## Time-stamp: <[settings.R] by JS 2021-06-01 10:42:03 CEST>
+##
+## Description:
+##
+##
+## History:
+##
+#####################################################################################
+
+### Set data and project directories
+
+project_dir <- "/data/sgg2/jenny/projects/proxyMR/"
+SGG_generic <- "/data/sgg2/jenny/SGG_generic/"
+
+## register clustermq and future plans
+#options(clustermq.scheduler = "slurm", clustermq.template = "slurm_clustermq.tmpl")
+#future::plan(batchtools_slurm, template = "slurm_batchtools.tmpl")
+
+#source(paste0(SGG_generic,"/scripts/settings.r"))
+
+traits <- c()
+MR_method_list <- c("mr_wald_ratio","mr_ivw","mr_ivw_fe","mr_egger_regression","mr_weighted_median")
+
+### define variables
+IV_threshold <- 5e-08
+prune_threshold <- 0.001
+GRS_thresholds <- c(0.1,0.01,0.001)
+num_household_bins <- 5
+
+num_IVs_threshold <- 5
+household_correlation_threshold <-0.1
+irnt=TRUE
+
+#phesant_directory <- read.table(paste0(UKBB_processed,"/PHESANT/","PHESANT_file_directory.txt"), header=T)
+#phesant_file_list <- unique(phesant_directory$File)
+
+# Data-Field 6141
+# Description:	How are people in household related to participant
+# limit to only individuals who responded YES to husband, wife, or partner
+
+relatedness_field <- "6141_1"
+#relations_file <- as.character(phesant_directory[which(phesant_directory[,2]==relatedness_field),"File"])
+
+# name of first phesant file
+#first_phesant_file <- as.character(phesant_directory[1,"File"])
+
+
+#time_at_address_file <- "/data/sgg2/jenny/data/UKBB_processed/PHESANT/ukb31459/bin1/out_bin1..tsv"
+#time_at_address_raw_file <- "/data/sgg2/jenny/data/UKBB_raw/pheno/ukb31459.csv"
+time_at_address_field <- "699"
+time_at_address_raw_field <- "699-0.0"
+
+# -10 represents "Less than a year"
+# -1 represents "Do not know"
+# -3 represents "Prefer not to answer"
+time_together_min <- 0
+time_together_max <- 50
[…]

### 2021-06-07 master.R

@@ -1,0 +1,19 @@
+#master script
+
+library(workflowr)
+wflow_start("proxyMR")
+
+renv::init()
+
+install.packages("remotes")
+remotes::install_github("ropensci/targets")
+remotes::install_github("ropensci/tarchetypes")
+
+library(targets)
+#manually create R/ folder
+#create 'R/function.R' file
+#add '_drake.R' file
+
+wflow_use_github("jennysjaarda")
+
+

### 2021-06-16 R/functions.R
run until end of pre-pipeline

@@ -757,16 +757,14 @@ write_sex_het <- function(sex_het_summary, traits_to_calc_het, traits_corr3,dir)
   }
 }
 
-sex_het_filter <- function(corr_traits, sex_het_summary, traits_to_calc_het, num_IVs_threshold){
+sex_het_filter <- function(corr_traits, sex_het_summary, num_IVs_threshold){
   list_length <- 4
   build_df <- numeric()
-  for(i in 1:dim(traits_to_calc_het)[1]){
-    data_out <- readd(sex_het_summary,subtargets=i)
-    row <- data_out[[1]]$sex_het_summary
-    build_df <- rbind(build_df, row)
+  for(i in 1:dim(corr_traits)[1]){
+    data_out <- sex_het_summary[[(i*4)]] #sex_het_summary is in spot 4
+    build_df <- rbind(build_df, data_out)
   }
   colnames(build_df) <- c(colnames(corr_traits),"num_IVs_pass_het")
-  #sex_het_summary is in spot 4
   result_df <- as.data.frame(build_df)
   result_df$num_IVs_pass_het <- as.numeric(as.character(result_df$num_IVs_pass_het))
   to_run <- result_df[which(result_df$num_IVs_pass_het >=num_IVs_threshold),]
@@ -778,6 +776,11 @@ sex_het_filter <- function(corr_traits, sex_het_summary, traits_to_calc_het, num
   with significant evidence of heterogeneity between sexes (p<0.05/[number of GW-sig IVs]),
   filtered phenotypic correlations have been saved to:
   'output/tables/4.household_correlations.sexhet_filter.csv'.\n"))
+}
+
+continuous_filter <- function(traits){
+  output <- traits %>% filter(variable_type=="ordinal" | variable_type=="continuous_irnt")
+  return(output)
 }
 
 check_valid_GRS_input <- function(traits,reference_file, Neale_output_path, Neale_summary_dir){

### 2021-06-16 _targets.R

@@ -27,14 +27,15 @@ tar_option_set(
 input_data <- tar_map(
   values = list(
     custom_names = c("household_info", "phesant_directory", "relatives", "fam", "sqc", "time_at_address",
-                     "time_at_address_raw", "UKBB_directory", "Neale_SGG_dir", "Neale_manifest", "code_process_Neale", "Neale_variants"),
+                     "time_at_address_raw", "UKBB_directory", "Neale_SGG_dir", "Neale_manifest", "code_process_Neale", "Neale_variants",
+                     "UKBB_sample"),
     files = c(paste0(UKBB_dir,"/pheno/ukb6881.csv"), paste0(UKBB_processed_dir,"/PHESANT/","PHESANT_file_directory.txt"),
               paste0(UKBB_dir,"/geno/","ukb1638_rel_s488366.dat"),  paste0(UKBB_dir,"/plink/_001_ukb_cal_chr9_v2.fam"),
               paste0(UKBB_dir,"/geno/ukb_sqc_v2.txt"),
               paste0(UKBB_processed_dir, "/PHESANT/ukb31459/bin1/out_bin1..tsv"), paste0(UKBB_dir, "/pheno/ukb31459.csv"),
               paste0(UKBB_processed_dir,"/UKBB_pheno_directory.csv"),
               Neale_SGG_dir_file_cp, paste0(Neale_output_dir,"/",Neale_manifest_file), "code/process_Neale.sh",
-              Neale_variant_file)
+              Neale_variant_file, paste0(UKBB_dir, "/imp/ukb1638_imp_chr1_v2_s487398.sample"))
   ),
   names = custom_names,
   unlist = FALSE,
@@ -110,7 +111,10 @@ list(
     data_Neale_manifest,
     read_tsv(path_Neale_manifest, col_types = cols())
   ),
-
+  tar_target(
+    data_UKBB_sample,
+    fread(path_UKBB_sample, skip=2, header=F,data.table=F)
+  ),
 
   ##################################
   ### MAKE HOUSEHOLD PAIRS FILE ####
@@ -304,21 +308,28 @@ list(
     write_IV_info(IV_data_summary, traits_to_calc_het$Neale_pheno_ID),
     format = "file",
     pattern = map(IV_data_summary, traits_to_calc_het)
+  ),
+  tar_target(
+    traits_corr4,
+    sex_het_filter(traits_corr3$to_run, IV_data_summary, num_IVs_threshold)
+  ),
+  # filter for continuous
[…]

### 2021-06-16 analysis/setup.Rmd
add website docs

---
title: "Project setup"
author: "jennysjaarda"
date: "2021-06-14"
output: workflowr::wflow_html
editor_options:
    toc: true
    number_sections: true
---

**Last updated:** `r Sys.Date()`

**Code version:** `r system("git log -1 --format='%H'", intern = TRUE)`

**To reproduce the results from this project, please follow these instructions.**

*In general, [`targets`](https://books.ropensci.org/targets/dynamic.html) was used to manage long-running code and [`workflowr`](https://jdblischak.github.io/workflowr/index.html) was used to manage the [website](https://jennysjaarda.github.io/proxymr/index.html).*

***

# Initiate project on remote server.

All processing scripts were run from the root sgg directory.
Project was initialized using `workflowr` rpackage, see [here](https://jdblischak.GitHub.io/workflowr/articles/wflow-01-getting-started.html).

On personal computer:

```{r eval=FALSE}
project_name <- "proxyMR"
library("workflowr")

wflow_start(project_name) # creates directory called `project_name`

options("workflowr.view" = FALSE) # if using cluster
options("workflowr.view" = TRUE) # defaule, if using Rstudio on personal computer

wflow_build() # create directories
options(workflowr.sysgit = "")

wflow_use_github("jennysjaarda") # select option 2: manually create new repository

wflow_git_push()
```

You have now successfully created a GitHub repository for your project that is accessible on GitHub and the servers.

Next setup a local copy.

# Create local copy on personal computer.

Within terminal of personal computer, clone the git repository.

```{r, engine = 'bash', eval = FALSE}
cd ~/Dropbox/UNIL/projects/
git clone https://GitHub.com/jennysjaarda/proxyMR.git proxyMR
```

Open project in atom (or preferred text editor) and modify the following files:

- Because Terminal cannot generate a preview and `workflowr` doesn't like the sysgit, to the `.Rprofile` file, add:
  - `options(workflowr.sysgit = "")`
  - `options("workflowr.view" = FALSE)`
- To ensure git hub isn't manaaging large files, modify the `.gitignore` file, by adding the following lines:
  - `analysis/*`
  - `data/*`
  - `!analysis/*.Rmd`
  - `!data/*.md`
  - `.git/`
  - `.drake/`
- Save and push these changes to GitHub.
- Pull to the server.

### 2021-06-16 master.R

@@ -4,24 +4,39 @@ library(workflowr)
 wflow_start("proxyMR")
 
 renv::init()
+tar_renv()
 
 install.packages("remotes")
 remotes::install_github("ropensci/targets")
 remotes::install_github("ropensci/tarchetypes")
 
 library(targets)
+
 #manually create R/ folder
 #create 'R/function.R' file
 #add '_drake.R' file
 
 wflow_use_github("jennysjaarda")
 
-
 ## On terminal
 
 #git clone https://github.com/jennysjaarda/proxyMR.git
 
 # only do this once
+
 Neale_SGG_dir_org <- paste0(Neale_summary_dir,"/Neale_SGG_directory.csv") ## not tracking this as input because it changes overtime
 Neale_SGG_dir_cp <- paste0("data/Neale_SGG_directory_", format(Sys.Date(), "%d_%m_%Y"), ".csv")
-Neale_SGG_dir_cp <- file.copy(Neale_SGG_dir_org, Neale_SGG_dir_cp)
+file.copy(Neale_SGG_dir_org, Neale_SGG_dir_cp)
+
+dir.create("output/figures")
+dir.create("output/tables")
+
+
+stats1 <- 2
+stats2 <- 2
+while ( (length(stats1) > 1 | length(stats2) > 1)){
+  stats1 <- suppressWarnings(system(paste("squeue -n", "process_Neale"), intern = TRUE))
+  stats2 <- suppressWarnings(system(paste("squeue -n", "clump_Neale_IVs"), intern = TRUE))
+  print("Still running...")
+  Sys.sleep(60)
+}

### 2021-06-16 analysis/update_meeting_14_06_2021.Rmd

---
title: "Update Meeting: 14-06-2021"
author: "jennysjaarda"
date: "2021-06-14"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
  toc: true
---

## Last week summary.

### Next steps:

1. Summarize all our results to get a feel for top traits (let's call this: trait/trait MR). A logical addition that we didn't discuss would be to meta-analyze across sexes and see what traits are significant in pooled analysis and then check which of these are significantly different between sexes.

2. Test linear association across bins (both age and time-together bins) using IVW estimated inversely weighted by SE.

3. Combine age and time-together bins into a grid of effect sizes to see if we see any pattern.

4. Move beyond our trait/trait MR into investigating how one trait impacts other traits to tease out direct vs indirect effects, two parts here:
    (a) as univariate model and 
    (b) as multivariate model (selecting traits for multivariate model that were nominal in univariate).

5. Replicate permutation procedure from Tenesa paper. Do you still want to explore this?

6. Explore dietary traits - request estimates from Ninon?

### Follow-up questions:

Re (#4), for the trait/trait MR, I essentially followed a 2-sample MR procedure (just as a refresher for us both), so for example BMI partner vs BMI trait:

- Find IVs for BMI in Neale and take these estimates as betaX.
- Test association of IVs with partner BMI in couples (this is betaY). 
- Combine effects together in standard MR framework (betaY ~ betaX).

For the univariate analysis between two traits, I think what you were suggesting is to follow exactly the same procedure as above. So for example, disease partner vs BMI trait:

- Same IVs as above and take these estimates as betaX.
- Test association of these IVs with partner "different phenotype" (some disease, let's say) (betaY). 
- Combine in MR framework.

For the multivariate model, so if the model of interest is BMI ~ BMI + education + diet + activity level:

- IVs are the union of the IVs for each individual trait (estimates taken from Neale: beta_BMI, beta_edu, beta_diet). 
- **Additional step**: prune these SNPs to ensure independence.
- Test association of these pruned SNPs with BMI in couples (betaY). 
- Combine effects in multivariate MR framework: betaY ~ beta_BMI + beta_edu + beta_diet, etc...

Does this look right (particularly for the multivariate model)? i.e. the final MR model only includes beta estimates, no raw genetic data.

I also understood that one of the traits to explore would be PCs. To me it seems logical that this would be a causal variable to explore (for eg, an MR being: BMI_beta ~ PCX_beta) rather than the other way around. If I have that right, then that means we would need IVs for PCs. At quick glance, Neale doesn't have this available. Do you know if anyone in the group would have done this, otherwise I should run this GWAS.

### Responses:

- Re. #3: also test heterogeneity (in case there is discrepancy, but no trend), using Cochran's Q stat.
- Re. combining time-together  and age bins: Do this only if the previous point gave something promising. Otherwise drop it. 
- Not urgent to replicate Tenesa paper.
- To explore dietary traits, got in touch again with Nicola Pirastu.
- Raw genetic data is only used to derive the betaY, but from then it is a 2-sample MR.
-  PCs are only exposures, not outcomes. Need to run GWAS.

### 2021-06-16 R/functions.R
run until prep pheno data

@@ -498,6 +498,11 @@ pull_traits_to_calc_het <- function(traits_corr3_run){
   return(output)
 }
 
+pull_traits_to_run <- function(traits_corr5){
+  output <- tibble(Neale_pheno_ID =traits_corr5[,"Neale_pheno_ID"])
+  return(output)
+}
+
 get_IV_list <- function(corr_traits, Neale_pheno_ID, reference_file, IV_threshold, Neale_output_path, Neale_summary_dir){
 
   corr_traits_both <- corr_traits[which(corr_traits[["Neale_file_sex"]]=="both"),]
@@ -876,12 +881,9 @@ valid_GRS_filter <- function(traits,valid_GRS_summary){
 
 }
 
-prep_data <- function(traits,i,phesant_directory,GRS_thresholds,reference_file,sqc,fam, UKBB_dir, Neale_summary_dir, Neale_output_path){
+create_trait_dirs <- function(Neale_pheno_ID){
 
-  category <- as.character(traits[i,"category"])
-  trait_ID <- as.character(traits[i,"Neale_pheno_ID"]) ## this is the Neale_id, used to be pheno_description
-  phes_ID <- as.character(traits[i,"SGG_PHESANT_ID"])
-  variable_type <- as.character(traits[i,"variable_type"])
+  trait_ID <- Neale_pheno_ID ## this is the Neale_id, used to be pheno_description
 
   print(trait_ID)
   dir.create(paste0("output/tables/traitMR/", trait_ID), showWarnings = FALSE)
@@ -892,9 +894,9 @@ prep_data <- function(traits,i,phesant_directory,GRS_thresholds,reference_file,s
   ### create_relevant directories
   dir.create(paste0(pheno_dir, "/household_MR/", trait_ID), showWarnings = FALSE)
   dir.create(paste0(pheno_dir, "/household_GWAS/", trait_ID), showWarnings = FALSE)
-  dir.create(paste0(pheno_dir, "/GRS/", trait_ID), showWarnings = FALSE)
-  dir.create(paste0(pheno_dir, "/GRS/", trait_ID, "/male"), showWarnings = FALSE)
-  dir.create(paste0(pheno_dir, "/GRS/", trait_ID, "/female"), showWarnings = FALSE)
+  #dir.create(paste0(pheno_dir, "/GRS/", trait_ID), showWarnings = FALSE)
+  #dir.create(paste0(pheno_dir, "/GRS/", trait_ID, "/male"), showWarnings = FALSE)
[…]

### 2021-06-16 _targets.R

@@ -313,11 +313,41 @@ list(
     traits_corr4,
     sex_het_filter(traits_corr3$to_run, IV_data_summary, num_IVs_threshold)
   ),
-  # filter for continuous
+  # filter for continuous traits only
   tar_target(
     traits_corr5,
     continuous_filter(traits_corr4$to_run)
+  ),
+  # final list of traits to run in pipeline, just a copy of above target
+  tar_target(
+    traits_final,
+    traits_corr5
+  ),
+  tar_target(
+    path_correlations_final_filter,
+    write.csv(traits_final, "output/tables/household_correlations.final_filter.csv", row.names=F),
+    format = "file"
+  ),
+  tar_target(
+    traits_to_run,
+    pull_traits_to_run(traits_final), iteration = "list"
+  ),
+  ## data prep
+  tar_target(
+    path_trait_dirs,
+    create_trait_dirs(traits_to_run$Neale_pheno_ID), pattern = map(traits_to_run), iteration = "list"
+  ),
+  tar_target(
+    pheno_data,
+    {
+      path_phesant
+      path_trait_dirs
+      prep_pheno_data(traits_final, traits_to_run$Neale_pheno_ID,
+                data_phesant_directory, data_Neale_manifest, data_sqc, data_fam, data_relatives,
+                UKBB_dir, Neale_summary_dir, Neale_output_dir)
+    }, pattern = map(traits_to_run), iteration = "list"
   )
+
 
 
 
@@ -327,7 +357,7 @@ list(
 #   traits_corr4 = sex_het_filter(traits_corr3$to_run, sex_het_summary, traits_to_calc_het, !!num_IVs_threshold),
 #   write_traits_corr3 = write.csv(traits_corr4$non_filtered, file_out("output/tables/3.household_correlations.numIVs_filter.csv"), row.names=F),
 #   write_traits_corr4 = write.csv(traits_corr4$to_run, file_out("output/tables/4.household_correlations.sexhet_filter.csv"), row.names=F),
-#   write_traits_corr5 = write.csv(traits_corr6, file_out("output/tables/6.household_correlations.nonbinary_filter.csv"), row.names=F)
+#   write_traits_corr5 = write.csv(traits_corr5, file_out("output/tables/6.household_correlations.nonbinary_filter.csv"), row.names=F)
 
 
 

### 2021-06-16 _targets.R
remove list iteration from some targets

@@ -330,12 +330,12 @@ list(
   ),
   tar_target(
     traits_to_run,
-    pull_traits_to_run(traits_final), iteration = "list"
+    pull_traits_to_run(traits_final)
   ),
   ## data prep
   tar_target(
     path_trait_dirs,
-    create_trait_dirs(traits_to_run$Neale_pheno_ID), pattern = map(traits_to_run), iteration = "list"
+    create_trait_dirs(traits_to_run$Neale_pheno_ID), pattern = map(traits_to_run)
   ),
   tar_target(
     pheno_data,

### 2021-06-18 R/functions.R
run to household_GWAS

@@ -1450,10 +1450,10 @@ plink_clump <- function(bfile, fn, prune_threshold){
   return(a_out)
 }
 
-load_IV_geno <- function(i,sample_file,pheno_dir, IV_data_file){ # UKBB_dir should be added to variable list
+load_IV_geno <- function(IV_data, sample_file, UKBB_imp_dir){ # UKBB_dir should be added to variable list
 
   ####### load IV list
-  IV_data <- fread(IV_data_file, header=T, data.table=F) ## IV list is the same for males and females, but the effects will be different
+  # IV_data <- fread(IV_data_file, header=T, data.table=F) ## IV list is the same for males and females, but the effects will be different
 
   ####### Extract SNP froms bgen files
   snp_map <- numeric()
@@ -1462,7 +1462,7 @@ load_IV_geno <- function(i,sample_file,pheno_dir, IV_data_file){ # UKBB_dir shou
   {
     snps <- IV_data[which(IV_data[,"chr"]==chr),"rsid"]
     if(length(snps)==0) next
-    bgen_file=paste(UKBB_dir, "/imp/", "_001_ukb_imp_chr", chr, "_v2.bgen", sep="")
+    bgen_file=paste(UKBB_imp_dir, "/_001_ukb_imp_chr", chr, "_v2.bgen", sep="")
     bgen_chr_extract <- extract_bgen(snps,bgen_file)
     IV_geno <- cbind(IV_geno, bgen_chr_extract$geno_data)
     snp_map <- rbind(snp_map,bgen_chr_extract$snp_map)
@@ -1561,111 +1561,116 @@ define_models <- function(traits){
 
 }
 
-household_GWAS <- function(i, trait_ID, exposure_sex, phenotype_file, phenotype_col, phenotype_description,IV_file,sample_file,pheno_cov,grouping_var,household_time,output){
+run_household_GWAS <- function(trait_info, summ_stats, pheno_data, path_UKBB_imp_data, data_UKBB_sample, joint_model_adjustments, grouping_var, household_time_munge){
 
+  trait_ID <- as.character(trait_info["trait_ID",1])
   cat(paste0("Calculating household GWAS for phenotype: ", trait_ID, "...\n"))
   cat(paste0("Loading genotype data...\n"))
   pheno_dir <- paste0("analysis/traitMR/")
[…]

### 2021-06-18 _targets.R

@@ -14,7 +14,7 @@ tar_option_set(
                                                         cpus = 1, partition = "sgg",
                                                         log_file="/data/sgg2/jenny/projects/proxyMR/proxymr_%a_clustermq.out"))
   ),
-  packages = c("tidyverse", "data.table", "cutr", "ukbtools"),
+  packages = c("tidyverse", "data.table", "cutr", "ukbtools", "rbgen"),
   error = "workspace"
 
 )
@@ -28,14 +28,15 @@ input_data <- tar_map(
   values = list(
     custom_names = c("household_info", "phesant_directory", "relatives", "fam", "sqc", "time_at_address",
                      "time_at_address_raw", "UKBB_directory", "Neale_SGG_dir", "Neale_manifest", "code_process_Neale", "Neale_variants",
-                     "UKBB_sample"),
+                     "UKBB_sample", "UKBB_imp_data"),
     files = c(paste0(UKBB_dir,"/pheno/ukb6881.csv"), paste0(UKBB_processed_dir,"/PHESANT/","PHESANT_file_directory.txt"),
               paste0(UKBB_dir,"/geno/","ukb1638_rel_s488366.dat"),  paste0(UKBB_dir,"/plink/_001_ukb_cal_chr9_v2.fam"),
               paste0(UKBB_dir,"/geno/ukb_sqc_v2.txt"),
               paste0(UKBB_processed_dir, "/PHESANT/ukb31459/bin1/out_bin1..tsv"), paste0(UKBB_dir, "/pheno/ukb31459.csv"),
               paste0(UKBB_processed_dir,"/UKBB_pheno_directory.csv"),
               Neale_SGG_dir_file_cp, paste0(Neale_output_dir,"/",Neale_manifest_file), "code/process_Neale.sh",
-              Neale_variant_file, paste0(UKBB_dir, "/imp/ukb1638_imp_chr1_v2_s487398.sample"))
+              Neale_variant_file, paste0(UKBB_dir, "/imp/ukb1638_imp_chr1_v2_s487398.sample"),
+              paste0(UKBB_dir, "/imp"))
   ),
   names = custom_names,
   unlist = FALSE,
@@ -366,23 +367,28 @@ list(
   ),
 
 
-  tar_target(IV_data_summary_run,
-             IV_data_summary[[IV_indices_to_run]],
-             pattern = map(IV_indices_to_run), iteration= "list"),
+  tar_target(
+    IV_data_summary_run,
+    IV_data_summary[[IV_indices_to_run]],
[…]

### 2021-06-18 _targets.R
don't track imputed data

@@ -18,7 +18,7 @@ tar_option_set(
   error = "workspace"
 
 )
-
+path_UKBB_imp_data <- paste0(UKBB_dir, "/imp") #too large to track
 
 ##################################
 ### INPUT DATA ####
@@ -28,15 +28,14 @@ input_data <- tar_map(
   values = list(
     custom_names = c("household_info", "phesant_directory", "relatives", "fam", "sqc", "time_at_address",
                      "time_at_address_raw", "UKBB_directory", "Neale_SGG_dir", "Neale_manifest", "code_process_Neale", "Neale_variants",
-                     "UKBB_sample", "UKBB_imp_data"),
+                     "UKBB_sample"),
     files = c(paste0(UKBB_dir,"/pheno/ukb6881.csv"), paste0(UKBB_processed_dir,"/PHESANT/","PHESANT_file_directory.txt"),
               paste0(UKBB_dir,"/geno/","ukb1638_rel_s488366.dat"),  paste0(UKBB_dir,"/plink/_001_ukb_cal_chr9_v2.fam"),
               paste0(UKBB_dir,"/geno/ukb_sqc_v2.txt"),
               paste0(UKBB_processed_dir, "/PHESANT/ukb31459/bin1/out_bin1..tsv"), paste0(UKBB_dir, "/pheno/ukb31459.csv"),
               paste0(UKBB_processed_dir,"/UKBB_pheno_directory.csv"),
               Neale_SGG_dir_file_cp, paste0(Neale_output_dir,"/",Neale_manifest_file), "code/process_Neale.sh",
-              Neale_variant_file, paste0(UKBB_dir, "/imp/ukb1638_imp_chr1_v2_s487398.sample"),
-              paste0(UKBB_dir, "/imp"))
+              Neale_variant_file, paste0(UKBB_dir, "/imp/ukb1638_imp_chr1_v2_s487398.sample"))
   ),
   names = custom_names,
   unlist = FALSE,

### 2021-06-18 analysis/update_meeting_14_06_2021.Rmd
wflow_publish(c("analysis/index.Rmd", "analysis/update_meeting_14_06_2021.Rmd"))

@@
 ---
+
+```{r setup, include = FALSE}
+
+options(scipen=999)
+source("code/settings.R")
+load("summary.Rdata")
+library(DT)
+library(knitr)
+
+```
+
+```{r, echo=FALSE,include = FALSE}
+# You need this code to conduct the magic dependences attaching...
+DT::datatable(matrix())
+```
 
@@
 -  PCs are only exposures, not outcomes. Need to run GWAS.
+
+## Results.
+
+### Traits analyzed.
+
+```{r traits, echo = FALSE}
+
+DT::datatable(traits, width = "100%")
+
+```
+
+### Summary Heterogeneity between sexes.
+
+```{r sex_het, echo = FALSE}
+
+reduced_mr_summary <- calc_mr_sex_het %>% dplyr::select(trait_ID, trait_description, 
+                                                        exposure_sex, N_snps,"IVW/Wald_summary","IVW/Wald_pval", 
+                                                        "MR_meta_est", "MR_meta_pval", "sex_het")
+
+DT::datatable(reduced_mr_summary, width = "100%")
+
+```
+
+### Testing discrepancy amongst bins.
+
+```{r mr_summary_age, echo = FALSE}
+
+mr_summary_age_reduced <- mr_summary_age %>% dplyr::select(trait_ID, exposure_sex, Q_stat, Q_pval)
+DT::datatable(mr_summary_age_reduced, width = "100%")
+
+mr_summary_tt_reduced <- mr_summary_tt %>% dplyr::select(trait_ID, exposure_sex, Q_stat, Q_pval)
+DT::datatable(mr_summary_tt_reduced, width = "100%")
+
+```

### 2021-06-18 analysis/update_meeting_14_06_2021.Rmd
wflow_publish(c("analysis/index.Rmd", "analysis/update_meeting_14_06_2021.Rmd"))

@@
 
-### Summary Heterogeneity between sexes.
+### Summary of results
+
+- IVW estimates report sex-specific results
+- MR_meta statistics refer to the meta of both-sexes combined. 
+- Sex heterogeneity column reports the p-value of the difference between the MR estimates from the two sexes.
 
@@
 
+- The first table represents the results of individuals binned by average age. 
+- The second table represent the results of individuals binned by time together.
+- Q p-values report the difference amongst bins.
+
 ```{r mr_summary_age, echo = FALSE}

### 2021-06-18 R/functions.R
divide AF by 2

@@ -1620,7 +1620,7 @@ run_household_GWAS <- function(trait_info, summ_stats, pheno_data, path_UKBB_imp
 
       mod_extract <- extract_model_stats(model_summary, c("geno_index"))
       n <- length(fitted(mod))
-      group_AF <- mean(final_data$geno_index)
+      group_AF <- mean(final_data$geno_index)/2
       k_GWAS <- numeric()
       k_GWAS <- rbind(k_GWAS,cbind(outcome, as.character(snp), grouping_var, "all", exposure_sex, outcome_sex, mod_extract,n, group_AF))
 
@@ -1633,7 +1633,7 @@ run_household_GWAS <- function(trait_info, summ_stats, pheno_data, path_UKBB_imp
         bin_pheno_run$outcome <- scale(bin_pheno_run$outcome)
         bin_pheno_run <- bin_pheno_run[complete.cases(bin_pheno_run),] #if all values are the same then NA's will be produced
         if(dim(bin_pheno_run)[1]!=0){
-          group_AF <- mean(bin_pheno_run$geno_index)
+          group_AF <- mean(bin_pheno_run$geno_index)/2
           bin_mod <- glm(outcome ~ ., data=bin_pheno_run, family="gaussian")
 
           bin_model_summary <- full_model_summary(bin_mod)

### 2021-06-18 _targets.R
allocate 2 cpus per task

@@ -11,7 +11,7 @@ options(clustermq.scheduler = "slurm", clustermq.template = "slurm_clustermq.tmp
 tar_option_set(
   resources = tar_resources(
     clustermq = tar_resources_clustermq(template = list(num_cores = 1, account = "sgg",
-                                                        cpus = 1, partition = "sgg",
+                                                        cpus = 2, partition = "sgg",
                                                         log_file="/data/sgg2/jenny/projects/proxyMR/proxymr_%a_clustermq.out"))
   ),
   packages = c("tidyverse", "data.table", "cutr", "ukbtools", "rbgen"),

### 2021-07-05 R/functions.R
write MR results

@@ -2063,26 +2063,30 @@ household_MR_all_outcomes <- function(exposure_info, summ_stats, outcomes_to_run
 
 }
 
-write_household_MR <- function(exposure_ID, outcomes_to_run, MR_result){
+write_household_MR <- function(exposure_info, outcomes_to_run, household_MR){
+
+  exposure_ID <- as.character(exposure_info["trait_ID",1])
 
   output_files <- numeric()
   pheno_dir <- paste0("analysis/traitMR")
 
+  cat(paste0("\Writing household MR results for all outcomes with phenotype `", exposure_ID, "` as exposure.\n\n"))
+
   for(i in 1:dim(outcomes_to_run)[1]){
 
     outcome_ID <- outcomes_to_run$Neale_pheno_ID[[i]]
-    cat(paste0("Loading GWAS results for outcome `", outcome_ID, "` and IVs from exposure `", exposure_ID, "`...\n"))
     MR_file_i <- paste0(pheno_dir, "/household_MR/", outcome_ID, "/univariate_MR/", outcome_ID, "_vs_", exposure_ID, "_MR.csv")
 
     output_files <- c(output_files, MR_file_i)
 
-    MR_result_i <- MR_result[[paste0(outcome_ID, "_vs_", exposure_ID, "_MR")]]
+    MR_result_i <- household_MR[[paste0(outcome_ID, "_vs_", exposure_ID, "_MR")]]
     write.csv(household_MR_result, MR_file_i, row.names = F)
+
+    cat(paste0("Finished writing MR results for outcome ", i, " of ", dim(outcomes_to_run)[1], ".\n\n" ))
 
   }
 
   return(output_files)
-
 
 }
 

### 2021-07-05 _targets.R

@@ -435,7 +435,7 @@ list(
     household_MR,
     {
       household_MR_all_outcomes(exposure_info, summ_stats, outcomes_to_run, gwas_files = path_household_GWAS,
-                       traits_corr2_update, grouping_var, MR_method_list) #could add MR_method_list = MR_method_list
+                       traits_corr2_update, grouping_var, MR_method_list)
     },
 
     pattern = map(exposure_info, summ_stats, IV_genetic_data, path_household_GWAS), iteration = "list"
@@ -445,7 +445,7 @@ list(
     path_household_MR,
     {
       path_MR_dirs
-      write_household_MR(exposure_info, summ_stats, outcomes_to_run, household_MR) #could add MR_method_list = MR_method_list
+      write_household_MR(exposure_info, outcomes_to_run, household_MR)
     },
 
     pattern = map(exposure_info, household_MR), format = "file"

### 2021-07-05 R/functions.R
Typo in output message for household MR

@@ -2070,7 +2070,7 @@ write_household_MR <- function(exposure_info, outcomes_to_run, household_MR){
   output_files <- numeric()
   pheno_dir <- paste0("analysis/traitMR")
 
-  cat(paste0("\Writing household MR results for all outcomes with phenotype `", exposure_ID, "` as exposure.\n\n"))
+  cat(paste0("\nWriting household MR results for all outcomes with phenotype `", exposure_ID, "` as exposure.\n\n"))
 
   for(i in 1:dim(outcomes_to_run)[1]){
 

### 2021-07-05 R/functions.R
fix `object 'household_MR_result' not found`

@@ -2080,7 +2080,7 @@ write_household_MR <- function(exposure_info, outcomes_to_run, household_MR){
     output_files <- c(output_files, MR_file_i)
 
     MR_result_i <- household_MR[[paste0(outcome_ID, "_vs_", exposure_ID, "_MR")]]
-    write.csv(household_MR_result, MR_file_i, row.names = F)
+    write.csv(MR_result_i, MR_file_i, row.names = F)
 
     cat(paste0("Finished writing MR results for outcome ", i, " of ", dim(outcomes_to_run)[1], ".\n\n" ))
 

### 2021-07-05 R/functions.R
run extensive/sensitivity MR analyses for each combination but not in each bin

@@ -1787,10 +1787,16 @@ household_MR_plot <- function(harmonise_dat, MR_method_list, exposure_sex, outco
 
 }
 
-household_MR_complete <- function(harmonise_dat, MR_method_list, exposure_ID, outcome_ID, exposure_sex, outcome_sex){
+household_MR_complete <- function(harmonise_dat, MR_method_list){
+
 
   exposure_ID <- harmonise_dat$exposure[1]
   outcome_ID <- harmonise_dat$outcome[1]
+  exposure_sex <- harmonise_dat$exposure_sex[1]
+  outcome_sex <- harmonise_dat$outcome_sex[1]
+  exposure_description <- harmonise_dat$exposure_description[1]
+  outcome_description <- harmonise_dat$outcome_description[1]
+
 
   het_test <- mr_heterogeneity(harmonise_dat, method_list=c("mr_egger_regression", "mr_ivw"))
   egger_test <- mr_pleiotropy_test(harmonise_dat)
@@ -1800,12 +1806,13 @@ household_MR_complete <- function(harmonise_dat, MR_method_list, exposure_ID, ou
   nsnps <- dim(harmonise_dat)[1]
   n_neale <- max(harmonise_dat$samplesize.exposure, na.rm=T)
 
-  temp_summary <- summarize_mr_result (paste0(exposure_ID,"_INDEX"), paste0(exposure_ID,"_HOUSEHOLD"), exposure_sex, outcome_sex, nsnps,MR_res, het_test, egger_test,n_neale, ngwas)
 
   ## Test for reverse causality
   harmonise_dat_sensitivity <- harmonise_dat[which(harmonise_dat[["pval.exposure"]] < harmonise_dat[["pval.outcome"]]),]
   MR_res <- include_MR_NA(mr(harmonise_dat_sensitivity, method=MR_method_list))
   #write.csv(MR_res, paste0(MR_dir, "/", phenotype_description, "_", exposure_sex,"-",outcome_sex, "_MR-sensitivity.csv"), row.names=F)
+
+  temp_summary <- summarize_mr_result (paste0(exposure_ID,"_INDEX"), paste0(outcome_ID,"_HOUSEHOLD"), exposure_sex, outcome_sex, nsnps, MR_res, het_test, egger_test,n_neale, ngwas)
 
   MR_ivw_row <- which(MR_res[,"method"]=="Inverse variance weighted")
   MR_wald_row <- which(MR_res[,"method"]=="Wald ratio")
[…]

### 2021-07-05 _targets.R

@@ -14,7 +14,7 @@ tar_option_set(
                                                         ntasks = 1, partition = "sgg",
                                                         log_file="/data/sgg2/jenny/projects/proxyMR/proxymr_%a_clustermq.out"))
   ),
-  packages = c("tidyverse", "data.table", "cutr", "ukbtools", "rbgen", "bigsnpr", "TwoSampleMR"),
+  packages = c("tidyverse", "data.table", "cutr", "ukbtools", "rbgen", "bigsnpr", "TwoSampleMR", "ggplot2"),
   error = "workspace",
   memory = "transient",
   garbage_collection = TRUE
@@ -457,18 +457,17 @@ list(
       harmonise_household_data_all_outcomes(exposure_info, summ_stats, outcomes_to_run, gwas_files = path_household_GWAS, traits_corr2_update)
     },
 
-    pattern = map(exposure_info, summ_stats, IV_genetic_data, path_household_GWAS), iteration = "list"
+    pattern = map(exposure_info, summ_stats, IV_genetic_data, path_household_GWAS), iteration = "list" #could remove IV_genetic data
 
   ),
 
   tar_target(
     household_MR_exhaustive,
     {
-      path_MR_dirs
-      household_MR_complete(harmonised_data)
+      household_MR_complete_all_outcomes(exposure_info, harmonised_data, outcomes_to_run, MR_method_list)
     },
 
-    pattern = map(exposure_info, summ_stats, IV_genetic_data, path_household_GWAS), format = "file"
+    pattern = map(exposure_info, harmonised_data), iteration = "list"
 
   )
 

### 2021-07-05 R/functions.R
add more complete cat message for progress of houshole_MR_complete_all_outcome

@@ -1861,6 +1861,8 @@ household_MR_complete_all_outcomes <- function(exposure_info, harmonised_data, o
 
     output_list[[paste0(outcome_ID, "_vs_", exposure_ID, "_MR_complete")]] <- MR_complete_i
 
+    cat(paste0("Finished computing full MR results for outcome ", i, " of ", dim(outcomes_to_run)[1], ".\n\n" ))
+
   }
 
   return(output_list)

### 2021-07-05 R/functions.R
fix axes in MR plot, wrap into one function

@@ -1769,24 +1769,6 @@ household_GWAS_all_outcomes <- function(exposure_info, summ_stats, outcomes_to_r
 }
 
 
-household_MR_plot <- function(harmonise_dat, MR_method_list, exposure_sex, outcome_sex, exposure_description, outcome_description){
-
-  ## Make MR plot
-
-  #output_figure_dir <- paste0(project_dir, "/output/figures/traitMR/",trait_ID)
-  #pdf(file=paste0(output_figure_dir,"/", trait_ID, "_", exposure_sex, "-",outcome_sex, "_household_MR", ".pdf"))
-
-  original_MR <- mr(harmonise_dat, method=MR_method_list)
-
-  mr_title <- bquote(atop(.(paste0("Estimate of the assortative mating effect of ")),
-                          italic(.(exposure_description)) ~ 'on' ~ italic(.(outcome_description)) ~ .(paste0(' for ', exposure_sex, "s on ", outcome_sex,"s"))))
-  mr_plot <- mr_scatter_plot_custom(original_MR, harmonise_dat, mr_title, exposure_sex, outcome_sex)
-  #dev.off()
-
-  return(mr_plot)
-
-}
-
 household_MR_complete <- function(harmonise_dat, MR_method_list){
 
 
@@ -1806,11 +1788,13 @@ household_MR_complete <- function(harmonise_dat, MR_method_list){
   nsnps <- dim(harmonise_dat)[1]
   n_neale <- max(harmonise_dat$samplesize.exposure, na.rm=T)
 
+  # Standard MR
+  original_MR <- mr(harmonise_dat, method=MR_method_list)
+  MR_res <- include_MR_NA(original_MR)
 
   ## Test for reverse causality
   harmonise_dat_sensitivity <- harmonise_dat[which(harmonise_dat[["pval.exposure"]] < harmonise_dat[["pval.outcome"]]),]
-  MR_res <- include_MR_NA(mr(harmonise_dat_sensitivity, method=MR_method_list))
-  #write.csv(MR_res, paste0(MR_dir, "/", phenotype_description, "_", exposure_sex,"-",outcome_sex, "_MR-sensitivity.csv"), row.names=F)
+  MR_res_sensitivity <- include_MR_NA(mr(harmonise_dat_sensitivity, method=MR_method_list))
 
[…]

### 2021-07-05 R/functions.R
summarize exhaustive analyses

@@ -1853,6 +1853,24 @@ household_MR_complete_all_outcomes <- function(exposure_info, harmonised_data, o
 
 }
 
+
+household_MR_complete_summary <- function(household_MR_complete_result){
+
+  result <- numeric()
+  for(i in 1:length(household_MR_complete_result)){
+    male_result <- household_MR_complete_result[[i]][["exp_male_MR_complete"]][["MR_summary"]]
+    female_result <- household_MR_complete_result[[i]][["exp_female_MR_complete"]][["MR_summary"]]
+    result_i <- rbind(male_result, female_result)
+    result <- rbind(result, result_i)
+
+  }
+
+  result <- as_tibble(result)
+  return(result)
+
+}
+
+
 harmonise_household_data <- function(exposure_info, summ_stats, traits_corr2_update, outcome_ID, household_GWAS_result, grouping_var = "age_even_bins") {
 
   output_list <- list()

### 2021-07-05 _targets.R

@@ -408,7 +408,6 @@ list(
     pattern = map(summ_stats)
   ),
 
-
   tar_target(
     path_household_GWAS,
     {
@@ -421,7 +420,8 @@ list(
 
   ),
 
-  ## MR file prep
+  ## MR
+
   tar_target(
     path_MR_dirs,
     {
@@ -431,6 +431,7 @@ list(
     pattern = map(outcomes_to_run), iteration = "list"
   ),
 
+  ## Binned results -> could change the name
   tar_target(
     household_MR,
     {
@@ -461,6 +462,7 @@ list(
 
   ),
 
+  ## RUN in full sample only, not binned
   tar_target(
     household_MR_exhaustive,
     {
@@ -469,25 +471,21 @@ list(
 
     pattern = map(exposure_info, harmonised_data), iteration = "list"
 
+  ),
+
+  ##summarize into one table, ignore leave-1-out analyses for now
+  tar_target(
+    household_MR_exhaustive_summary,
+    {
+      household_MR_complete_summary(household_MR_exhaustive)
+    },
+
+    pattern = map(household_MR_exhaustive)
+
   )
 
-  ## NEXT RUN SENSITIVITY ANALYSES WITH: plot, leave-one-out, egger, etc. not in bins
-  ## WRITE MR RESULTS
 
 
-  # household GWAS produces a DF with results for each group,
-  # so in MR function we can just filter to relevant groups.
-
-# household_MR(outcome_trait = "", household_GWAS, exposure_trait = "", IVs, grouping_var),
-
-# cross(cross(map(outcome_traits, household_GWAS), grouping_var), map(exposure_traits, IVs))
-# trait folder represents outcome trait
-# for each outcome find all trait < 0.05
-# combine
-
-  ## removed all `readd` commands from next set of functions
-
-#
 #   traits_corr4 = sex_het_filter(traits_corr3$to_run, sex_het_summary, traits_to_calc_het, !!num_IVs_threshold),
 #   write_traits_corr3 = write.csv(traits_corr4$non_filtered, file_out("output/tables/3.household_correlations.numIVs_filter.csv"), row.names=F),
 #   write_traits_corr4 = write.csv(traits_corr4$to_run, file_out("output/tables/4.household_correlations.sexhet_filter.csv"), row.names=F),

### 2021-07-05 analysis/update_meeting_05_07_2021.Rmd
update meeting July05/21

---
title: "Update Meeting: 05-07-2021"
author: "jennysjaarda"
date: "2021-07-02"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
  toc: true
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
library(targets)
library(dplyr)
library(DT)
library(knitr)

```

```{r, echo=FALSE,include = FALSE}
# You need this code to conduct the magic dependences attaching...
DT::datatable(matrix())
```

## Last week summary.

- Discussed PolyMR project with Jonathan. 
  - To start by writing introduction. 
  - Going through paper in-depth.

## Results.

- Univariate MR.

```{r sex_het, echo = FALSE}

tar_load(household_MR_exhaustive_summary)

household_MR_exhaustive_summary <- household_MR_exhaustive_summary %>% mutate(same_traits = case_when(exposure_ID == outcome_ID ~ TRUE, 
                                                                   TRUE ~ FALSE))

DT::datatable(household_MR_exhaustive_summary, width = "100%")

```

## Questions.

- Expand list of outcomes? 
  - Initially only performed MR with a single trait (i.e. trait_partner ~ trait_index). 
  - Now that we are expanding to two traits in the analysis we could be less stringent with our trait selection.
  - Traits were filtered as follows: 
    1. UKBB phenotypes in SGG database.
    2. Link with Neale summary stats and include only phenotypes which have been processed in both male, female and joint GWAS. 
    3. Evidence of correlation among couples. 
    4. &gt;5 IVs at GW significance. 
    5. No evidence of sex heterogeneity among valid IVs. 
  - Could add more phenotpyes as outcomes

## Next steps.

-

### 2021-07-14 analysis/update_meeting_14_07_2021.Rmd
update meeting

---
title: "Update Meeting: 14-07-2021"
author: "jennysjaarda"
date: "2021-06-14"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
  toc: true
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
load("summary.Rdata")
library(DT)
library(knitr)
library(targets)

```

```{r, echo=FALSE,include = FALSE}
# You need this code to conduct the magic dependences attaching...
DT::datatable(matrix())
```

## Last week summary.

- Update phenotypes, make sure blood chemistry values are included (see list of biomarkers included [here](https://biobank.ndph.ox.ac.uk/ukb/label.cgi?id=17518)).
- Non-urgent analysis: cross partner heritability. 
  - For this we need a full GWAS of index genotype on partner trait (up until this point only run on IVs of interest).
- GWAS of PCs 1-40, adjusted for age and sex, to get IVs (Neale has not run this).
- Run MR in same individual (i.e. standard MR).

## Update.

### Reprocessed PHESANT

- For all new phenotype files since starting the project and added them to the pipeline.

```{r new_phenos, echo = FALSE}

tar_load(data_UKBB_directory)
new_files <- c("ukb39955", "ukb42673", "ukb44073", "ukb44073")

data_UKBB_directory_new <- data_UKBB_directory %>% dplyr::filter(file %in% !!new_files)

DT::datatable(data_UKBB_directory_new, width = "100%")

```

### Identified a problem with the Neale summary stats.
- Only affecting the `low_confidence_variant` column. 
- This filter was applied before generating IVs for each summary file that had been downloaded.
- See a description of the issue on [the Neale github issues page](https://github.com/Nealelab/UK_Biobank_GWAS/issues/20).
- Currently in the process of updating the impacted files:
  - Updated Neale GWAS files can be found here: `/data/sgg2/jenny/data/Neale_UKBB_GWAS/neale_files/variant_flag_update`.  
  - Also will recreate the IVs for these files (~70 of our files), under the same name. 
-  As far as I can tell the issue is only in one direction, where some SNPs that were not really low confidence variants (by their definition) were listed as `low_confidence_variant = TRUE`, not the other way around. In other words, the mistake means that we would only be filtering out variants that we should keep.

### Adding PCs to the pipeline.

- Currently running GWAS of first 40 PCs, residualized for `age` and `sex`. Applied inverse normal rank transformation first.  
- Test between partners.

```{r pc_corr, echo = FALSE}

tar_load(PCs_corr)

DT::datatable(PCs_corr, width = "100%")

```

## Next steps.

Define the following:

**Goal**: To use Mendelian Randomization (MR) to estimate a causal effect of ${X}$ on ${Y}$ between household pairs (i.e. *couples*).

Definitions:

- Let ${X}$ and ${Y}$ denote two random variables representing complex traits. 
- Let ${X}_i$ and ${X}_p$ denote ${X}$ in an index case and in their partner. 
- The genotype data of the SNPs to be used as instrumental variables (IVs) is denoted by ${G}$. 
- Let ${G}_i$ and ${G}_p$ denote ${G}$ in an index case and in their partner. 
- To simplify notation we assume that $E(X) = E(Y) = E(G) =  0$ and $Var(X) = Var(Y) = Var(G) = 1$. 
- The effect sizes of the instruments on $X$ is denoted by ${\beta}$. 
- The causal effect is denoted by ${\alpha}$.
- Let us assume the following models.

$$
\begin{eqnarray*}
	(1) \; X_{i} &=& G_{i}\cdot {\beta_{x, i\rightarrow i} } + \epsilon \\
	(2) \; X_{p} &=& G_{i}\cdot {\beta_{x, i\rightarrow p} } + \epsilon \\
	(3) \; Y_{i} &=& G_{i}\cdot {\beta_{y, i\rightarrow i} } + \epsilon \\
	(4) \; Y_{p} &=& G_{i}\cdot {\beta_{y, i\rightarrow p} } + \epsilon \\
	(5) \; Y_{i} &=& \alpha_j\cdot X_{i} +\epsilon \\
	(6) \; X_{p} &=& \alpha_{x, am}\cdot X_{i{}} +\epsilon \\
	(7) \; Y_{p} &=& \alpha_{y, am}\cdot Y_{i{}} +\epsilon \\
	(8) \; Y_{p} &=& \alpha_{k}\cdot X_{i} +\epsilon

\end{eqnarray*}
$$

- Equation (1) and (3) are simply same-person GWAS result (obtained from Neale).
- Equation (2) and (4) are household GWAS results (calculated in-house).
- Equation (5), and ${\alpha_{j}}$, in the causal effect of a trait ${X}$ on ${Y}$ from a standard MR single-person MR.
- Equation (6) and (7), represent assortative mating effects for traits ${X}$ and ${Y}$, respectively.
- Equation (8) ${\alpha_{k}}$ is the result of phenotype ${X}$ in the index, having a causal effect on a phenotype ${Y}$ in the partner. 
- This effect is either due to the product of one of the two:
  - ${\alpha_{j}}$ with ${\alpha_{x, am}}$ 
  - ${\alpha_{j}}$ with ${\alpha_{y, am}}$

### 2021-07-22 analysis/update_meeting_14_06_2021.Rmd
wflow_publish("analysis/update_meeting_14_06_2021.Rmd")

@@
 ```{r, echo=FALSE,include = FALSE}
-# You need this code to conduct the magic dependences attaching...
+# You need this code to conduct the magic dependencies attaching...
 DT::datatable(matrix())
@@
 
-```{r traits, echo = FALSE}
+```{r traits, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
 
@@
 
-```{r sex_het, echo = FALSE}
+```{r sex_het, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
 
@@
 
-```{r mr_summary_age, echo = FALSE}
+```{r mr_summary_age, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
 

### 2021-07-22 analysis/index.Rmd
Updating website

---
title: "Home"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

Welcome to my research website.

Check-out my latex tests [here](latex_test.html)

### 2021-07-22 analysis/update_meeting_05_07_2021.Rmd
Updating website

@@
 
-```{r sex_het, echo = FALSE}
+```{r univar_mr, echo = FALSE}
 
-tar_load(household_MR_exhaustive_summary)
+household_MR_exhaustive_summary <- readRDS("household_MR_exhaustive_summary.rds")
+traits_corr2_update <- readRDS("traits_corr2_update.rds")
+traits_corr2_update <- traits_corr2_update[which(traits_corr2_update[["Neale_file_sex"]]=="both"),]
 
-household_MR_exhaustive_summary <- household_MR_exhaustive_summary %>% mutate(same_traits = case_when(exposure_ID == outcome_ID ~ TRUE, 
-                                                                   TRUE ~ FALSE))
+description_temp <- left_join(household_MR_exhaustive_summary, traits_corr2_update[, c("Neale_pheno_ID", "description")], by=c("exposure_ID"="Neale_pheno_ID")) %>%
+  rename(exposure_description = description)
+
+description <- left_join(description_temp, traits_corr2_update[, c("Neale_pheno_ID", "description")], by=c("outcome_ID"="Neale_pheno_ID")) %>%
+  rename(outcome_description = description)
+
+household_MR_exhaustive_summary <- description %>% mutate(same_traits = case_when(exposure_ID == outcome_ID ~ TRUE, 
+                                                                   TRUE ~ FALSE)) %>% dplyr::select(-exposure, -outcome) %>%
+  dplyr::select(-Egger_summary, -Egger_pval, -Egger_int_pval, -Het_IVW_Qpval, -N_snps_sensitivity, -`IVW/Wald_summary_sensitivity`, -`IVW/Wald_pval_sensitivity`, -outcome_sex)
 
@@
 
+- Number of BF-significant results: 
+```{r BF_sig, echo = FALSE}
+
+BF_sig <- household_MR_exhaustive_summary %>% rename(IVW_Wald_pval= `IVW/Wald_pval`) %>% 
+  mutate_at("IVW_Wald_pval", as.numeric) %>% filter(IVW_Wald_pval < 0.05/dim(.)[1])
+
+dim(BF_sig)
+
+```
+
+- Significant results not in the same trait
+
+```{r non_same_traits, echo = FALSE}
+
+diff_traits <- BF_sig %>% filter(same_traits==FALSE)
+DT::datatable(diff_traits, width = "100%")
+
+```
 ## Questions.

### 2021-07-22 analysis/update_meeting_14_06_2021.Rmd
Updating website

@@
 ## Last week summary.
+
+### Getting back up to speed.
+
+- Project had been dormant for > 1 year. 
+- Since then, one key paper has been published in Nature Communications by Tenesa et al: [Evidence of horizontal indirect genetic effects in humans](https://www.nature.com/articles/s41562-020-00991-9).
+- Discussed where we left off and how to build on Tenesa paper.
+- Slides can be found [here](presentations/Xia_horizontal_genetic_effects/presentation.html#1).
 

### 2021-07-22 analysis/update_meeting_14_07_2021.Rmd
update meeting
Updating website

@@
 source("code/settings.R")
-load("summary.Rdata")
+
 library(DT)
@@
 library(targets)
-
 ```
@@
 
-```{r new_phenos, echo = FALSE}
+```{r new_phenos, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
 
@@
 
-```{r pc_corr, echo = FALSE}
+```{r pc_corr, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
 
@@
 
+BREAK1
+
 $$
-\begin{eqnarray*}
-	(1) \; X_{i} &=& G_{i}\cdot {\beta_{x, i\rightarrow i} } + \epsilon \\
-	(2) \; X_{p} &=& G_{i}\cdot {\beta_{x, i\rightarrow p} } + \epsilon \\
-	(3) \; Y_{i} &=& G_{i}\cdot {\beta_{y, i\rightarrow i} } + \epsilon \\
-	(4) \; Y_{p} &=& G_{i}\cdot {\beta_{y, i\rightarrow p} } + \epsilon \\
-	(5) \; Y_{i} &=& \alpha_j\cdot X_{i} +\epsilon \\
-	(6) \; X_{p} &=& \alpha_{x, am}\cdot X_{i{}} +\epsilon \\
-	(7) \; Y_{p} &=& \alpha_{y, am}\cdot Y_{i{}} +\epsilon \\
-	(8) \; Y_{p} &=& \alpha_{k}\cdot X_{i} +\epsilon
+    \(1) \; X_{i} = G_{i}\cdot {\beta_{x, i\rightarrow i} } + \epsilon \\
+    \(2) \; X_{p} = G_{i}\cdot {\beta_{x, i\rightarrow p} } + \epsilon \\
+    (3) \; Y_{i} = G_{i}\cdot {\beta_{y, i\rightarrow i} } + \epsilon \\
+	  (4) \; Y_{p} = G_{i}\cdot {\beta_{y, i\rightarrow p} } + \epsilon \\
+	  (5) \; Y_{i} = \alpha_j\cdot X_{i} +\epsilon \\
+	  (6) \; X_{p} = \alpha_{x, am}\cdot X_{i{}} +\epsilon \\
+	  (7) \; Y_{p} = \alpha_{y, am}\cdot Y_{i{}} +\epsilon \\
+	  (8) \; Y_{p} = \alpha_{k}\cdot X_{i} +\epsilon
 
-\end{eqnarray*}
+$$
+
+BREAK2
+
+\begin{equation}
+	 \; X_{p} = G_{i}\cdot {\beta_{x, i\rightarrow p} } + \epsilon \\
+	 \; Y_{i} = G_{i}\cdot {\beta_{y, i\rightarrow i} } + \epsilon \\
+	 \; Y_{p} = G_{i}\cdot {\beta_{y, i\rightarrow p} } + \epsilon \\
+	 \; Y_{i} = \alpha_j\cdot X_{i} +\epsilon \\
+	 \; X_{p} = \alpha_{x, am}\cdot X_{i{}} +\epsilon \\
+	 \; Y_{p} = \alpha_{y, am}\cdot Y_{i{}} +\epsilon \\
+	 \; Y_{p} = \alpha_{k}\cdot X_{i} +\epsilon
+\end{equation}
+
+BREAK3
+
+$$
+	 \; X_{p} = G_{i}\cdot {\beta_{x, i\rightarrow p} } + \epsilon \\
+	 \; Y_{i} = G_{i}\cdot {\beta_{y, i\rightarrow i} } + \epsilon \\
+	 \; Y_{p} = G_{i}\cdot {\beta_{y, i\rightarrow p} } + \epsilon \\
+	 \; Y_{i} = \alpha_j\cdot X_{i} +\epsilon \\
+	 \; X_{p} = \alpha_{x, am}\cdot X_{i{}} +\epsilon \\
+	 \; Y_{p} = \alpha_{y, am}\cdot Y_{i{}} +\epsilon \\
+	 \; Y_{p} = \alpha_{k}\cdot X_{i} +\epsilon
 $$
@@
   - ${\alpha_{j}}$ with ${\alpha_{y, am}}$
+
+$$
+\sqrt{p}
+$$
+
+$$
+x = \frac{-b \pm \sqrt{b^2 - 4ac}}{2a}
+$$

### 2021-08-24 R/functions.R
meta analyze slopes amongst sexes.

@@ -2298,16 +2298,21 @@ calc_binned_household_MR_het <- function(exposure_info, outcomes_to_run, househo
           mutate_if(is.factor,as.character) %>%
           as_tibble()
 
-        cat(paste0("Finished calculating heterogeneity statistics for outcome ", i, " of ", dim(outcomes_to_run)[1], ".\n\n" ))
 
         exposure_i_result <- rbind( exposure_i_result, out_temp)
       }
 
     }
+    cat(paste0("Finished calculating heterogeneity statistics for outcome ", i, " of ", dim(outcomes_to_run)[1], ".\n\n" ))
 
   }
 
-  return(exposure_i_result)
+  meta_result <- exposure_i_result %>% group_by(outcome_ID, grouping_var) %>%
+    group_modify(~ summarize_sex_specific_results(as.numeric(.x$bin_slope_beta), as.numeric(.x$bin_slope_se)))
+
+  output <- left_join(exposure_i_result, meta_result)
+
+  return(output)
 
 
 }
@@ -2669,7 +2674,6 @@ run_proxyMR_comparison <- function(exposure_info, standard_MR_summary_BF_sig, ho
     }
 
 
-    ### now meta-analyze across sexes and calculate heterogeneity difference
     summarized_result <- as_tibble(summarized_result) %>%
       mutate_at(vars(-c("exposure_ID", "outcome_ID", "exposure_description", "outcome_description", "exposure_sex", "outcome_sex")),as.numeric)
 

### 2021-08-24 R/functions.R
fix typo in col names of meta result in `calc_binned_household_MR_het`

@@ -2310,9 +2310,10 @@ calc_binned_household_MR_het <- function(exposure_info, outcomes_to_run, househo
   meta_result <- exposure_i_result %>% group_by(outcome_ID, grouping_var) %>%
     group_modify(~ summarize_sex_specific_results(as.numeric(.x$bin_slope_beta), as.numeric(.x$bin_slope_se)))
 
+  colnames(meta_result)[-c(1:2)] <- paste0("bin_slope_", colnames(meta_result)[-c(1:2)])
   output <- left_join(exposure_i_result, meta_result)
 
-  return(output)
+  return(exposure_i_result)
 
 
 }

### 2021-08-24 R/functions.R
return proper output in `calc_binned_household_MR_het`

@@ -2313,7 +2313,7 @@ calc_binned_household_MR_het <- function(exposure_info, outcomes_to_run, househo
   colnames(meta_result)[-c(1:2)] <- paste0("bin_slope_", colnames(meta_result)[-c(1:2)])
   output <- left_join(exposure_i_result, meta_result)
 
-  return(exposure_i_result)
+  return(output)
 
 
 }

### 2021-08-24 analysis/update_meeting_25_08_2021.Rmd
intial update meeeting 25-08-21

---
title: "Update Meeting: 25-08-2021"
author: "jennysjaarda"
date: "2021-08-24"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
  toc: true
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
library(targets)
library(dplyr)
library(DT)
library(knitr)

```

## Last meeting summary.

## Model summary

```{r proxyMR_DAG, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/proxyMR_DAG.png", error = FALSE)
```

## Results

### Trends amongst bins

```{r MR_binned_het, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
tar_load(household_MR_binned_het)
DT::datatable(household_MR_binned_het, width = "100%")
```

### Standard MR

```{r standard_MR, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
tar_load(standard_MR_summary_BF_sig)
DT::datatable(standard_MR_summary_BF_sig, width = "100%")
```

### Paths from X-index to Y-partner

```{r proxyMR_products, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
tar_load(proxyMR_figure_data)
DT::datatable(proxyMR_figure_data, width = "100%")
```

#### Heterogeneity among paths

```{r proxyMR_figure1, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
tar_load(proxyMR_prod_comparison_figure)
proxyMR_prod_comparison_figure
```

#### Sex heterogeneity amongst paths

``````{r proxyMR_figure2, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
tar_load(proxyMR_sex_comparison_figure)
proxyMR_sex_comparison_figure
```

#### Hierarchical clustering

### 2021-08-24 R/functions.R
fix some problems with figure function

@@ -2773,6 +2773,12 @@ prep_proxyMR_figure_data <- function(proxyMR_comparison, traits_corr2_filled){
 
 create_proxy_prod_comparison_fig_ind <- function(data, exposure_sex, x, y, overlay_var, count){
 
+  custom_col <- c("#a6cee3",
+                  "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99",
+                  "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6",
+                  "#6a3d9a", "#ffff99", "#b15928")
+
+
   # exposure_sex <- "male"
   # data <- comparison_result
   # overlay_var <- "omega_vs_gam_BF_sig_sex_specific"
@@ -2844,6 +2850,9 @@ create_proxy_sex_comparison_fig_ind <- function(data, var, count){
 
   # data <- comparison_result
   # var <- "rho
+
+  num_result <- dim(data)[1]/2
+
 
   vars_to_select <- paste0(var, c("_beta", "_sex_het_p"))
   sex_het_p_var <- paste0(var, "_sex_het_p")

### 2021-08-24 R/functions.R
fix some problems with figures 2

@@ -2848,6 +2848,13 @@ create_proxy_prod_comparison_fig_ind <- function(data, exposure_sex, x, y, overl
 create_proxy_sex_comparison_fig_ind <- function(data, var, count){
 
 
+
+  custom_col <- c("#a6cee3",
+                  "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99",
+                  "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6",
+                  "#6a3d9a", "#ffff99", "#b15928")
+
+
   # data <- comparison_result
   # var <- "rho
 

### 2021-08-24 R/functions.R
move legends to bottom

@@ -2838,7 +2838,7 @@ create_proxy_prod_comparison_fig_ind <- function(data, exposure_sex, x, y, overl
   )
 
   fig_no_legend <- fig + theme(legend.position="none")
-  if(count==9){fig_no_legend <- fig_no_legend + theme(legend.position = c(0.5, 0.15))}
+  #if(count==9){fig_no_legend <- fig_no_legend + theme(legend.position = c(0.5, 0.15))}
 
 
   return(list(fig_no_legend = fig_no_legend, legend = legend))
@@ -2907,7 +2907,7 @@ create_proxy_sex_comparison_fig_ind <- function(data, var, count){
   )
 
   fig_no_legend <- fig + theme(legend.position="none")
-  if(count==3){fig_no_legend <- fig_no_legend + theme(legend.position = c(0.5, 0.08))}
+  #if(count==3){fig_no_legend <- fig_no_legend + theme(legend.position = c(0.5, 0.08))}
 
 
   return(list(fig_no_legend = fig_no_legend, legend = legend))
@@ -2954,7 +2954,10 @@ create_proxy_prod_comparison_fig <- function(proxyMR_figure_data){
 
   figures_grid <- plot_grid(plotlist = figures, nrow=3, ncol = 3, byrow = F, labels="AUTO", rel_heights = c(1, 1, 1))
 
-  return(figures_grid)
+  figures_grid_plus_legend <- plot_grid(figures_grid, legend, ncol = 1, nrow =2, rel_heights = c(1, 0.05))
+
+
+  return(figures_grid_plus_legend)
 }
 
 create_proxy_sex_comparison_fig <- function(proxyMR_figure_data){
@@ -2979,12 +2982,17 @@ create_proxy_sex_comparison_fig <- function(proxyMR_figure_data){
     fig_temp <- create_proxy_sex_comparison_fig_ind(proxyMR_figure_data, var = var, panel)
     sex_het_figures[[panel]] <- fig_temp[["fig_no_legend"]]
 
+
     if(panel==1) {legend <- fig_temp[["legend"]]}
     panel <- panel + 1
 
   }
 
   sex_het_figures_grid <- plot_grid(plotlist = sex_het_figures, nrow=1, ncol = 3, byrow = F, labels="AUTO", rel_heights = c(1, 1, 1))
+
[…]

### 2021-08-24 R/functions.R
add legend to sex het plot

@@ -2993,7 +2993,7 @@ create_proxy_sex_comparison_fig <- function(proxyMR_figure_data){
   figures_grid_plus_legend <- plot_grid(sex_het_figures_grid, legend, ncol = 1, nrow =2, rel_heights = c(1, 0.03))
 
 
-  return(sex_het_figures_grid)
+  return(figures_grid_plus_legend)
 }
 
 

### 2021-08-31 analysis/index.Rmd
clean up some website docs

@@
 output:
-  workflowr::wflow_html:
-    toc: false
+  bookdown::html_document2:
+    toc: true
 editor_options:
@@
 
-Welcome to my research website.
+This is the website for the research project "proxyMR".
 
-Check-out my latex tests [here](latex_test.html)
+If you have cloned the project to a local computer this website is rendered in the `docs` subdirectory of the project directory.
+
+If you are using `workflowr` to publish the research website it will also be rendered online to GitHub Pages.
+
+This page acts as a table of contents for the website. There are links to the web pages generated from the analysis notebooks and to the rendered versions of manuscripts/documents/presentations.
+
+------------------------------------------------------------------------
+
+# Overview documents {.unnumbered}
+
+## [About](about.html) {.unnumbered}
+
+This notebook explains the central ideas behind the project.
+
+## [Notes](notes.html) {.unnumbered}
+
+This notebook is for keeping notes of any points that may be useful for
+later project or manuscript development and which are not covered in the
+analysis notebooks or at risk of getting lost in the notebooks.
+
+## [Workflow management](workflow.html) {.unnumbered}
+
+This project uses the [`targets`](https://wlandau.github.io/targets/) and [`workflowr`](https://github.com/jdblischak/workflowr) packages for managing the workflow of the project (making sure that the dependencies between computational steps are satisfied).
+This notebook contains notes on the proposed workflow for
+using `targets` and `workflowr`.
+
+## [Meetings](meetings.html) {.unnumbered}
+
+## [Project Workflow Status](m_00_status.html) {.unnumbered}
+
+This notebook displays the computational status of the project, that is,
+whether everything is up to date. Ironically, this is currently the only
+notebook that must be run manually, so this notebook only displays the
+status **when the notebook was last executed** and there is no
+indication whether the project status has changed since then.
+
+------------------------------------------------------------------------
+
+# Publications and Presentations {.unnumbered}
+
+Links to rendered manuscripts and presentations will go here.
+
+------------------------------------------------------------------------
+
+# Analayses Notebooks {.unnumbered}
+
+These notebooks capture the analyses that were carried out to develop
+the code of the core processing pipeline. They are organised as
+side-chains to the core processing pipeline.
+
+Typically, a meta notebook will analyse the data available at one stage
+of the core pipeline, to guide the writing of the functions required to
+get to the next stage of the core pipeline. These meta notebooks
+generally conclude with the definition of a function that will be used
+in the core pipeline.
+
+There may be multiple notebooks all relating to different aspects of the
+same stage of the core pipeline.
+
+Sometimes the analyses are more diffuse - characterising the data in a
+way that may be helpful for guiding the development of future core
+stages, but not immediately resulting in the development of functions
+for the core pipeline.
+
+## 01 Read, check, and standardise the entity data {.unnumbered}
+
+Determine the initial data preparation of the imported entity records.
+
+### [m_01_1\_read_raw_entity_data](m_01_1_read_raw_entity_data.html) {.unnumbered}
+
+Read the raw entity data.
+
+### [m_01_2\_exclusions](m_01_2_exclusions.html) {.unnumbered}
+
+Apply the row exclusions (exclude test records and all records with
+status other than ACTIVE & VERIFIED).
+
+### [m_01_3\_drop_novar](m_01_3_drop_novar.html) {.unnumbered}
+
+Drop variables with no variation.
+
+### [m_01_4\_parse_dates](m_01_4_parse_dates.html) {.unnumbered}
+
+Parse the date columns.
+
+### [m_01_5\_check_admin](m_01_5_check_admin.html) {.unnumbered}
+
+Check the administrative variables (data relating to the administration
+of voter registration).
+
+### [m_01_6\_check_resid](m_01_6_check_resid.html) {.unnumbered}
+
+Check the residence variables - residential address and phone number.
+
+### [m_01_7\_check_demog](m_01_7_check_demog.html) {.unnumbered}
+
+Check the demographic variables - sex, age, and birth place.
+
+### [m_01_8\_check_name](m_01_8_check_name.html) {.unnumbered}
+
+Check the name variables.
+
+### [m_01_9\_clean_vars](m_01_9_clean_vars.html) {.unnumbered}
+
+Clean all the variables.
+
+------------------------------------------------------------------------
+
+# Acknowledgements
+
+Primary analysis performed by: [Jenny Sjaarda](jennysjaarda@gmail.com).
+
+Project supervisor: [Professor Zoltan Kutalik](https://wp.unil.ch/sgg/people/zoltan/).
+
+This work has been supported by the [Swiss National Science Foundation](http://www.snf.ch/en/funding/Pages/default.aspx).

### 2021-08-31 analysis/setup.Rmd
clean up some website docs

@@
 title: "Project setup"
-author: "jennysjaarda"
+author: "Jenny Sjaarda"
 date: "2021-06-14"
 output: workflowr::wflow_html
-editor_options:
-    toc: true
-    number_sections: true
 ---
@@
 *In general, [`targets`](https://books.ropensci.org/targets/dynamic.html) was used to manage long-running code and [`workflowr`](https://jdblischak.github.io/workflowr/index.html) was used to manage the [website](https://jennysjaarda.github.io/proxymr/index.html).*
-
-***
 
@@
 
-Open project in atom (or preferred text editor) and modify the following files:
+Open project in Rstudio (or preferred text editor) and modify the following files:
 
@@
   - `options("workflowr.view" = FALSE)`
-- To ensure git hub isn't manaaging large files, modify the `.gitignore` file, by adding the following lines:
-  - `analysis/*`
-  - `data/*`
-  - `!analysis/*.Rmd`
-  - `!data/*.md`
-  - `.git/`
-  - `.drake/`
+- To ensure git hub isn't managing large files, modify the `.gitignore` file, by adding the following lines:
+```
+analysis/*
+data/*
+!analysis/*.Rmd
+_targets
+*.out
+*.rds
+*.RData
+```
 - Save and push these changes to GitHub.
 - Pull to the server.
+
+# Create internal project folders.
+
+Return to sgg server and run the following:
+
+```{r, engine = 'bash', eval = FALSE}
+project_dir=/data/sgg2/jenny/projects/proxyMR
+mkdir $project_dir/data/raw
+mkdir $project_dir/data/processed
+mkdir $project_dir/docs/assets
+mkdir $project_dir/docs/generated_reports
+mkdir $project_dir/analysis/data_setup
+mkdir $project_dir/analysis/data_setup/IV_lists
+mkdir $project_dir/analysis/data_setup/IV_info
+mkdir $project_dir/analysis/data_setup/sex_heterogeneity
+mkdir $project_dir/analysis/traitMR
+mkdir $project_dir/analysis/traitMR/IVs
+mkdir $project_dir/analysis/traitMR/IVs/Neale
+mkdir $project_dir/analysis/traitMR/household_MR
+mkdir $project_dir/analysis/traitMR/household_GWAS
+mkdir $project_dir/analysis/traitMR/standard_GWAS
+mkdir $project_dir/analysis/traitMR/standard_MR
+mkdir $project_dir/analysis/traitMR/trait_info
+mkdir $project_dir/analysis/traitMR/pheno_files
+mkdir $project_dir/analysis/traitMR/pheno_files/phesant
+mkdir $project_dir/analysis/traitMR/pheno_files/raw_change
+mkdir $project_dir/output/figures
+mkdir $project_dir/output/tables
+mkdir $project_dir/output/figures/traitMR
+mkdir $project_dir/output/tables/traitMR
+
+```
+
+This will create the following directory structure in `proxyMR/`:
+
+```{r, engine = 'bash', eval = FALSE}
+proxyMR/
+├── .gitignore
+├── .gitattributes
+├── .Rprofile
+├── _workflowr.yml
+├── analysis/
+│   ├── data_setup/
+│   │   ├── sex_heterogeneity/
+│   │   ├── IV_info/
+│   │   └── IV_lists/
+│   ├── trait_MR/
+│   │   ├── IVs/
+│   │   │   └── Neale/
+│   │   ├── household_MR/
+│   │   ├── standard_MR/
+│   │   ├── household_GWAS/
+│   │   ├── standard_GWAS/
+│   │   ├── pheno_files/
+│   │   │   ├── phesant/
+│   │   │   └── raw_change/
+│   │   └── trait_info/
+│   ├── about.Rmd
+│   ├── index.Rmd
+│   ├── license.Rmd
+│   └── _site.yml
+├── code/
+│   ├── README.md
+├── data/
+│   ├── README.md
+│   ├── raw/
+│   └── processed/
+├── docs/
+│   ├── generated_reports/
+│   └── assets/
+├── proxyMR.Rproj
+├── output/
+│   ├── figures/
+│   │   └── trait_MR/
+│   ├── tables/
+│   │   └── trait_MR/
+│   └── README.md
+└── README.md
+```
+
+Note that after `renv` and `targets` is initialized below a few other directories and files will also be created, namely:
+
+```{r, engine = 'bash', eval = FALSE}
+proxyMR/
+├── renv/
+├── _targets/
+├── renv.lock
+└── .renvignore
+```
+
+# Initialize a renv directory
+
+[Renv](https://rstudio.github.io/renv/index.html) is a dependency management system for R. It is useful for making your project:
+
+Isolated: Installing a new or updated package for one project won’t break your other projects, and vice versa. That’s because `renv` gives each project its own private package library.
+
+Portable: Easily transport your projects from one computer to another, even across different platforms. `renv` makes it easy to install the packages your project depends on.
+
+Reproducible: `renv` records the exact package versions you depend on, and ensures those exact versions are the ones that get installed wherever you go.
+
+The renv package is a new effort to bring project-local R dependency management to your projects.
+The goal is for renv to be a robust, stable replacement for the Packrat package, with fewer surprises and better default behaviors.
+
+Initialize a `renv` project by simply running:
+
+```{r eval=FALSE}
+renv::init()
+tar_renv()
+```
+
+This creates a `renv` directory in your project folder.
+
+Now, every time you launch `R` from this directory or run `install.packages()`, `renv` will automatically keep track of your packages and versions.
+You are no longer in an ordinary R project; you’re in a `renv` project.
+The main difference is that a `renv` project has its own private package library.
+Any packages you install from inside a `renv` project are only available to that project; and packages you install outside of the project are not available to the project.
+
+Periodically call `renv::snapshot()` to save the state of the project library to the lockfile (called renv.lock), continue working on your project, installing and updating R packages as needed.
+Call `renv::restore()` to revert to the previous state as encoded in the lockfile if your attempts to update packages introduced some new problems.
+
+# Copy relevant files to project directory.
+
+Only do this once.
+The `Neale_SGG_directory.csv` file gives a summary of all SGG UKBB phenotypes that we have in our database and Neale phenotypes that have been downloaded.
+This file changes overtime as new phenotypes are requested from the UKBB and/or new Neale files are downloaded (i.e.
+more rows are added to the file).
+Because the `targets` package tracks input files and their changes, the current state of this file was copied to the project directory and this file was then used as an input file into the pipeline.
+This file is merely a reference file and changes to it don't correspond to changes in the underlying data used in the analyses.
+
+```{r eval = FALSE}
+Neale_summary_dir <- "/data/sgg2/jenny/data/Neale_UKBB_GWAS"
+Neale_SGG_dir_org <- paste0(Neale_summary_dir,"/Neale_SGG_directory.csv") 
+Neale_SGG_dir_cp <- paste0("data/Neale_SGG_directory_", format(Sys.Date(), "%d_%m_%Y"), ".csv")
+file.copy(Neale_SGG_dir_org, Neale_SGG_dir_cp)
+```
+
+# Run and summarize analyses.
+
+## Setup targets and execute plan.
+
+**Note that Part A and B are happening in parallel.**
+
+The [`targets`](https://books.ropensci.org/targets/index.html) package was used to manage teh pipeline and maintain a reproducible workflow.
+It integrates well with [HPC clusters](https://books.ropensci.org/targets/hpc.html) making it an ideal choice for this project.
+
+```{r eval=FALSE}
+library(targets)
+```
+
+For execution of `targets` plan: see the [`_targets.R`](https://github.com/jennysjaarda/proxyMR/blob/master/_targets.R) file.
+
+Configure a slurm template
+```{r eval=FALSE}
+options(clustermq.scheduler = "slurm", clustermq.template = "slurm_clustermq.tmpl")
+drake_hpc_template_file("slurm_clustermq.tmpl")
+
+# Write the file slurm_clustermq.tmpl. and edit manually
+```
+The file created using the `clustermq` template was edited manually to match [`slurm_clustermq.tmpl`](https://GitHub.com/jennysjaarda/proxyMR/blob/master/slurm_clustermq.tmpl)
+
+```{r comment=''}
+cat(readLines('slurm_clustermq.tmpl'), sep = '\n')
+```
+
+## Build and maintain website.
+
+Follow the general workflow outlined by [workflowr](https://jdblischak.GitHub.io/workflowr/articles/wflow-01-getting-started.html#the-workflow), with some minor revisions to accommodate workflow between personal computer and remote server:
+
+  1.
+Open a new or existing R Markdown file in `analysis/` (optionally using `wflow_open()`).
+(Usually created manually on personal computer and push to server to build later.) If creating manually, add the following to the top of the R Markdown file with an appropriate name for `Title`:
+
+```{r eval=FALSE}
+---
+title: "Title"
+site: workflowr::wflow_site
+output:
+  workflowr::wflow_html:
+    toc: true
+---
+```
+
+  2. Write documentation and perform analyses in the R Markdown file.
+
+  3. Run `commit` and `push` to upload revised R Markdown file to GitHub repository.
+
+  4. On server, pull changes using `wflow_git_pull()` (optionally using `git pull` from Terminal within cloned repository).
+
+  5.
+Within R console, run `wflow_build()`.
+This will create html files with `docs/` folder.
+These files cannot be viewed directly on server, but can be transferred and viewed via [FileZilla](https://filezilla-project.org/) or viewed directly by mounting the remote directory to your personal computer using [SSHFS](https://en.wikipedia.org/wiki/SSHFS) (recommended).
+Alternatively, Rmd files can be built within the targets pipeline using [`tar_render`](https://rdrr.io/github/wlandau/tarchetypes/man/tar_render.html), see examples in the `targets` manual [here](https://books.ropensci.org/targets/files.html#literate-programming).
+Perhaps the second option is preferred  as `tar_render` tracks all the Finds all the `tar_load()` / `tar_read() `dependencies in the report and inserts them into the target’s command.
+This enforces the proper dependency relationships.
+
+  6. Return to step 2 until satisfied with the result (optionally, edit Rmd file directly on server using `vi` if only small modifications are necessary).
+
+  7. Run `wflow_status()` to track repository.
+
+  8.
+Run `wflow_publish()` to commit the source files (R Markdown files or other files in `code/`, `data/`, and `output/`), build the HTML files, and commit the HTML files.
+If there are uncommited files in the directory that are not ".Rmd", `wflow_publish(all=T)` does not work.
+Alternatively, run the following with an informative `message`:
+
+```{r eval=FALSE}
+repo_status <- wflow_status()
+rmd_commit <- c(rownames(repo_status$status)[repo_status$status$modified],
+         rownames(repo_status$status)[repo_status$status$unpublished],
+         rownames(repo_status$status)[repo_status$status$scratch])
+
+wflow_publish(rmd_commit,
+              message="Updating website")
+
+```
+
+  9. Push the changes to GitHub with `wflow_git_push()` (or `git push` in the Terminal).

### 2021-08-31 master.R

@@ -1,42 +1,0 @@
-#master script
-
-library(workflowr)
-wflow_start("proxyMR")
-
-renv::init()
-tar_renv()
-
-install.packages("remotes")
-remotes::install_github("ropensci/targets")
-remotes::install_github("ropensci/tarchetypes")
-
-library(targets)
-
-#manually create R/ folder
-#create 'R/function.R' file
-#add '_drake.R' file
-
-wflow_use_github("jennysjaarda")
-
-## On terminal
-
-#git clone https://github.com/jennysjaarda/proxyMR.git
-
-# only do this once
-
-Neale_SGG_dir_org <- paste0(Neale_summary_dir,"/Neale_SGG_directory.csv") ## not tracking this as input because it changes overtime
-Neale_SGG_dir_cp <- paste0("data/Neale_SGG_directory_", format(Sys.Date(), "%d_%m_%Y"), ".csv")
-file.copy(Neale_SGG_dir_org, Neale_SGG_dir_cp)
-
-dir.create("output/figures")
-dir.create("output/tables")
-
-
-stats1 <- 2
-stats2 <- 2
-while ( (length(stats1) > 1 | length(stats2) > 1)){
-  stats1 <- suppressWarnings(system(paste("squeue -n", "process_Neale"), intern = TRUE))
-  stats2 <- suppressWarnings(system(paste("squeue -n", "clump_Neale_IVs"), intern = TRUE))
-  print("Still running...")
-  Sys.sleep(60)
-}

### 2021-09-06 R/functions.R
iterate figs as list, add xy binned plot (MR vs bin)

@@ -3553,7 +3553,7 @@ mr_scatter_plot_custom <-  function (mr_results, dat, mr_title, exposure_sex, ou
 }
 
 
-bin_plot_sex_specific <- function(harmonised_data, MR_binned, exposure_sex, group){
+mr_plot_binned_sex_specific <- function(harmonised_data, MR_binned, exposure_sex, group){
 
   if(exposure_sex=="male"){outcome_sex="female"}
   if(exposure_sex=="female"){outcome_sex="male"}
@@ -3574,7 +3574,7 @@ bin_plot_sex_specific <- function(harmonised_data, MR_binned, exposure_sex, grou
 
 
 
-  legend_title <- ifelse(group=="age_even_bins", "Median age \n of couples", "Time together \n in same household")
+  legend_title <- ifelse(group=="age_even_bins", "Median age \n of couples (years)", "Time together \n in same household (years)")
 
   index <- harmonise_dat_plot$beta.exposure < 0
   harmonise_dat_plot$beta.exposure[index] <- harmonise_dat_plot$beta.exposure[index] *-1
@@ -3616,9 +3616,6 @@ mr_plot_sex_specific <- function(harmonised_data, MR_binned, custom_col){
 
   outcome_levels <- levels(bin_summary$exposure_sex)
 
-
-  #legend_title <- ifelse(group=="age_even_bins", "Median age \n of couples", "Time together \n in same household")
-
   legend_title <- "Exposure sex"
 
   index <- harmonise_dat_plot$beta.exposure < 0
@@ -3648,7 +3645,7 @@ mr_plot_sex_specific <- function(harmonised_data, MR_binned, custom_col){
   return(plot)
 }
 
-forest_plot_sex_specific <- function(harmonised_data, MR_binned, group, custom_col){
+forest_plot_binned_sex_specific <- function(harmonised_data, MR_binned, group, custom_col){
 
   harmonised_data.sex_male <- harmonised_data[[paste0("exp_", "male", "_harmonised_data")]] %>% dplyr::filter(grouping_var == "age_even_bins") %>% dplyr::filter(bin== "all") %>% mutate_at('bin', as.factor)
   harmonised_data.sex_female <- harmonised_data[[paste0("exp_", "female", "_harmonised_data")]] %>% dplyr::filter(grouping_var == "age_even_bins") %>% dplyr::filter(bin== "all") %>% mutate_at('bin', as.factor)
[…]

### 2021-09-06 _targets.R

@@ -11,7 +11,7 @@ options(clustermq.scheduler = "slurm", clustermq.template = "slurm_clustermq.tmp
 tar_option_set(
   resources = tar_resources(
     clustermq = tar_resources_clustermq(template = list(num_cores = 1, account = "sgg",
-                                                        ntasks = 4, partition = "sgg",
+                                                        ntasks = 8, partition = "sgg",
                                                         log_file="/data/sgg2/jenny/projects/proxyMR/proxymr_%a_clustermq.out"))
   ),
   packages = c("tidyverse", "data.table", "cutr", "ukbtools", "rbgen", "bigsnpr", "TwoSampleMR",
@@ -465,7 +465,7 @@ list(
   tar_target(
     household_MR_binned_AM_figs,
     create_MR_binned_AM_figs(household_harmonised_data, household_MR_binned_meta, custom_col),
-    pattern = map(household_harmonised_data, household_MR_binned_meta)
+    pattern = map(household_harmonised_data, household_MR_binned_meta), iteration = "list"
   ),
 
   tar_target(

### 2021-09-06 R/functions.R
add arguements to `xy_plot_binned_meta`

@@ -3905,7 +3905,7 @@ xy_plot_binned_sex_specifc <- function(harmonised_data, MR_binned, group, custom
 
 }
 
-xy_plot_binned_meta <- function(){
+xy_plot_binned_meta <- function(harmonised_data, MR_binned, group, custom_col){
 
   harmonised_data.sex_male <- harmonised_data[[paste0("exp_", "male", "_harmonised_data")]] %>% dplyr::filter(grouping_var == "age_even_bins") %>% dplyr::filter(bin== "all") %>% mutate_at('bin', as.factor)
   harmonised_data.sex_female <- harmonised_data[[paste0("exp_", "female", "_harmonised_data")]] %>% dplyr::filter(grouping_var == "age_even_bins") %>% dplyr::filter(bin== "all") %>% mutate_at('bin', as.factor)
@@ -3950,6 +3950,7 @@ xy_plot_binned_meta <- function(){
   return(plot)
 
 }
+
 create_MR_binned_AM_figs <- function(household_harmonised_data, household_MR_binned_meta, custom_col){
 
   names_data <- names(household_harmonised_data)

### 2021-09-06 R/functions.R
change line to dashed in xy_plot

@@ -3881,7 +3881,7 @@ xy_plot_binned_sex_specifc <- function(harmonised_data, MR_binned, group, custom
 
   plot <- ggplot2::ggplot(data = fig_data, ggplot2::aes(x = bin_median_plot,
                                                         y = IVW_beta, colour = factor(exposure_sex))) +
-    geom_smooth(method="lm",formula=y~x, se = F, fullrange = T) +
+    geom_smooth(method="lm",formula=y~x, se = F, fullrange = T, linetype = "dashed") +
     ggplot2::geom_errorbar(ggplot2::aes(ymin = IVW_beta - IVW_se, ymax = IVW_beta + IVW_se),
                            colour = "grey", width = 0) +
     ggplot2::geom_point(ggplot2::aes(colour = factor(exposure_sex))) +
@@ -3932,7 +3932,7 @@ xy_plot_binned_meta <- function(harmonised_data, MR_binned, group, custom_col){
 
   plot <- ggplot2::ggplot(data = fig_data_meta, ggplot2::aes(x = bin_median,
                                                               y = IVW_meta_beta)) +
-    geom_smooth(method="lm",formula=y~x, se = F, fullrange = T, color = custom_col[2]) +
+    geom_smooth(method="lm",formula=y~x, se = F, fullrange = T, color = custom_col[2], linetype = "dashed") +
     ggplot2::geom_errorbar(ggplot2::aes(ymin = IVW_meta_beta - IVW_meta_se, ymax = IVW_meta_beta + IVW_meta_se),
                            colour = "grey", width = 0) +
     ggplot2::geom_point(color = custom_col[2]) +

### 2021-09-06 R/functions.R
add a sidebyside sex figure

@@ -3951,6 +3951,51 @@ xy_plot_binned_meta <- function(harmonised_data, MR_binned, group, custom_col){
 
 }
 
+
+xy_plot_binned_single_sex <- function(harmonised_data, MR_binned, exposure_sex, group, custom_col){
+
+  fig_data <- MR_binned %>% filter(bin!="all") %>% separate(bin, c("bin_start_temp", "bin_stop_temp"), ",", remove = F) %>% filter(grouping_var==!!group) %>% mutate(bin_start = substring(bin_start_temp, 2)) %>%
+    mutate(bin_stop = str_sub(bin_stop_temp,1,nchar(bin_stop_temp)-1)) %>% rowwise() %>%
+    mutate(bin_median = median(c(as.numeric(bin_start), as.numeric(bin_stop)))) %>%
+    mutate(bin_median_plot = case_when(exposure_sex=="female" ~ bin_median + 0.4,
+                                       exposure_sex=="male" ~ bin_median - 0.4))
+
+  harmonise_dat_plot <- rbind(harmonised_data.sex_male, harmonised_data.sex_female) %>% mutate_at('exposure_sex', as.factor)
+  trait_description <- as.character(harmonise_dat_plot[1, "exposure_description"])
+
+
+  if(exposure_sex=="male"){outcome_sex="female"}
+  if(exposure_sex=="female"){outcome_sex="male"}
+
+  fig_data_sex_spec <- fig_data_sex_spec <- fig_data %>% dplyr::filter(exposure_sex==!!exposure_sex)
+
+  plot <- ggplot2::ggplot(data = fig_data_sex_spec, ggplot2::aes(x = bin_median,
+                                                                 y = IVW_beta, colour = factor(exposure_sex))) +
+    geom_smooth(method="lm",formula=y~x, se = F, fullrange = T, linetype = "dashed", color = custom_col[2]) +
+    ggplot2::geom_errorbar(ggplot2::aes(ymin = IVW_beta - IVW_se, ymax = IVW_beta + IVW_se),
+                           colour = "grey", width = 0) +
+    ggplot2::geom_point(color = custom_col[2]) +
+
+
+    scale_x_continuous(breaks=x_ticks, limits = c(min(x_ticks)-2, max(x_ticks)+2),
+                       labels = x_labels) +
+
+    theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
+                       panel.grid.minor = element_blank(), axis.line=element_blank(),
[…]

### 2021-09-06 R/functions.R
fix small typo

@@ -4057,8 +4057,8 @@ create_MR_binned_AM_figs <- function(household_harmonised_data, household_MR_bin
 
   for(group in c("age_even_bins", "time_together_even_bins")){
 
-    xy_binM <- xy_plot_binned_single_sex(harmonised_data, MR_binned, "male", group, custom_col)
-    xy_binF <- xy_plot_binned_single_sex(harmonised_data, MR_binned, "female", group, custom_col)
+    xy_binM <- xy_plot_binned_single_sex(household_harmonised_data.AM, household_MR_binned_meta.AM, "male", group, custom_col)
+    xy_binF <- xy_plot_binned_single_sex(household_harmonised_data.AM, household_MR_binned_meta.AM, "female", group, custom_col)
 
     xy_bin_side_by_side <- plot_grid(xy_binM, xy_binF)
     assign(paste0("XY_sex_sidebyside_", group, "_fig"), p_combined)

### 2021-09-06 R/functions.R
fix typo

@@ -3954,6 +3954,9 @@ xy_plot_binned_meta <- function(harmonised_data, MR_binned, group, custom_col){
 
 xy_plot_binned_single_sex <- function(harmonised_data, MR_binned, exposure_sex, group, custom_col){
 
+  harmonised_data.sex_male <- harmonised_data[[paste0("exp_", "male", "_harmonised_data")]] %>% dplyr::filter(grouping_var == "age_even_bins") %>% dplyr::filter(bin== "all") %>% mutate_at('bin', as.factor)
+  harmonised_data.sex_female <- harmonised_data[[paste0("exp_", "female", "_harmonised_data")]] %>% dplyr::filter(grouping_var == "age_even_bins") %>% dplyr::filter(bin== "all") %>% mutate_at('bin', as.factor)
+
   fig_data <- MR_binned %>% filter(bin!="all") %>% separate(bin, c("bin_start_temp", "bin_stop_temp"), ",", remove = F) %>% filter(grouping_var==!!group) %>% mutate(bin_start = substring(bin_start_temp, 2)) %>%
     mutate(bin_stop = str_sub(bin_stop_temp,1,nchar(bin_stop_temp)-1)) %>% rowwise() %>%
     mutate(bin_median = median(c(as.numeric(bin_start), as.numeric(bin_stop)))) %>%

### 2021-09-06 R/functions.R
add xticks definition

@@ -3966,6 +3966,12 @@ xy_plot_binned_single_sex <- function(harmonised_data, MR_binned, exposure_sex, 
   harmonise_dat_plot <- rbind(harmonised_data.sex_male, harmonised_data.sex_female) %>% mutate_at('exposure_sex', as.factor)
   trait_description <- as.character(harmonise_dat_plot[1, "exposure_description"])
 
+  x_ticks <- unique(fig_data %>% pull(bin_median))
+  x_labels <- unique(fig_data %>% pull(bin))
+
+  xlab <- ifelse(group == "time_together_even_bins", "Time together in same household (years)", "Median age of couples (years)")
+
+  legend_title <- "Exposure sex"
 
   if(exposure_sex=="male"){outcome_sex="female"}
   if(exposure_sex=="female"){outcome_sex="male"}

### 2021-09-06 R/functions.R
saved wrong plot

@@ -1157,6 +1157,38 @@ extract_trait_info <- function(pheno_data){
 }
 
 
+calc_PC_traits <- function(exposure_info_list, sqc, relatives){
+
+  full_df <- numeric()
+
+  for(i in 1:length(exposure_info_list)){
+
+    exposure_info_i <- exposure_info_list[[i]]
+    phes_ID <- exposure_info_i %>% filter(Value=="phes_ID") %>% pull(Info)
+    phesant_file <- exposure_info_i %>% filter(Value=="phesant_file") %>% pull(Info)
+    tsv_data <- fread(phesant_file, header=TRUE, sep='\t',select=c("userId","sex","age", phes_ID))
+
+    if(i==1){
+      full_df <- tsv_data
+    } else full_df <- merge(full_df, tsv_data[,c(1, 4)])
+
+  }
+
+  IDs <- as.integer(full_df$userId)
+
+  related_IDs <- ukb_gen_samples_to_remove(relatives, ukb_with_data = IDs)
+  unrelated_data <- full_df[-which(full_df$userId %in% related_IDs ),]
+
+  sqc_head <- ukb_gen_sqc_names(sqc, col_names_only = FALSE)
+
+  calc_PC_data <- unrelated_data[,-c(1:3)]
+  res.pca <- prcomp(calc_PC_data, scale = TRUE)
+  fviz_eig(res.pca)
+
+
+
+}
+
 write_data_prep <- function(traits, traits_to_run, out1, out2){
 
   for(i in 1:dim(traits_to_run)[1]){
@@ -4070,7 +4102,7 @@ create_MR_binned_AM_figs <- function(household_harmonised_data, household_MR_bin
     xy_binF <- xy_plot_binned_single_sex(household_harmonised_data.AM, household_MR_binned_meta.AM, "female", group, custom_col)
 
     xy_bin_side_by_side <- plot_grid(xy_binM, xy_binF)
-    assign(paste0("XY_sex_sidebyside_", group, "_fig"), p_combined)
+    assign(paste0("XY_sex_sidebyside_", group, "_fig"), xy_bin_side_by_side)
   }
 
 

### 2021-09-06 _targets.R

@@ -378,6 +378,10 @@ list(
     pattern = map(exposures_to_run), iteration = "list"
   ),
 
+  tar_target(
+    PC_traits,
+    calc_PC_traits(exposure_info, data_sqc, data_relatives)
+  ),
 
   tar_target(
     IV_data_summary_run,

### 2021-09-06 R/functions.R
calc PC of traits

@@ -1157,7 +1157,7 @@ extract_trait_info <- function(pheno_data){
 }
 
 
-calc_PC_traits <- function(exposure_info_list, sqc, relatives){
+calc_PC_traits <- function(exposure_info_list, sqc, fam, relatives){
 
   full_df <- numeric()
 
@@ -1176,17 +1176,34 @@ calc_PC_traits <- function(exposure_info_list, sqc, relatives){
 
   IDs <- as.integer(full_df$userId)
 
-  related_IDs <- ukb_gen_samples_to_remove(relatives, ukb_with_data = IDs)
-  unrelated_data <- full_df[-which(full_df$userId %in% related_IDs ),]
 
-  sqc_head <- ukb_gen_sqc_names(sqc, col_names_only = FALSE)
+  # This option takes too long
+  # related_IDs <- ukb_gen_samples_to_remove(relatives, ukb_with_data = IDs)
+  # unrelated_data <- full_df[-which(full_df$userId %in% related_IDs ),]
+
+  ##fam file in same order as sample QC (sqc) file
+  sqc <- ukb_gen_sqc_names(sqc, col_names_only = FALSE)
+
+  sqc$ID <- fam[,1]
+  unrelated_IDs <- sqc[which(sqc$used_in_pca_calculation==1), "ID"]
+  unrelated_data <- full_df[which(full_df$userId %in% unrelated_IDs ),]
 
   calc_PC_data <- unrelated_data[,-c(1:3)]
-  res.pca <- prcomp(calc_PC_data, scale = TRUE)
-  fviz_eig(res.pca)
 
+  cor_matrix <- cor(calc_PC_data, use = "pairwise.complete.obs")
 
+  res.pca <- prcomp((cor_matrix), scale = TRUE)
+  # to visualize the PC results
+  # fviz_eig(res.pca)
 
+  return(res.pca)
+
+}
+
+calc_num_tests_by_PC <- function(prcomp_result, threshold){
+
+  num_tests <- which(summary(prcomp_result)$importance[3,] > threshold)[1]
+  return(num_tests)
 }
 
 write_data_prep <- function(traits, traits_to_run, out1, out2){
@@ -2827,7 +2844,7 @@ summarize_proxyMR_comparison <- function(proxyMR_comparison, traits_corr2_filled
                                                 TRUE ~ TRUE)) %>%
[…]

### 2021-09-06 _targets.R

@@ -15,7 +15,8 @@ tar_option_set(
                                                         log_file="/data/sgg2/jenny/projects/proxyMR/proxymr_%a_clustermq.out"))
   ),
   packages = c("tidyverse", "data.table", "cutr", "ukbtools", "rbgen", "bigsnpr", "TwoSampleMR",
-               "ggplot2", "purrr", "rmeta", "PASWR2", "cowplot", "meta", "strex", "RColorBrewer", "forestplot"),
+               "ggplot2", "purrr", "rmeta", "PASWR2", "cowplot", "meta", "strex", "RColorBrewer",
+               "forestplot"),
   error = "workspace",
   memory = "transient",
   garbage_collection = TRUE
@@ -380,7 +381,12 @@ list(
 
   tar_target(
     PC_traits,
-    calc_PC_traits(exposure_info, data_sqc, data_relatives)
+    calc_PC_traits(exposure_info, data_sqc, data_fam, data_relatives)
+  ),
+
+  tar_target(
+    num_tests_by_PCs,
+    calc_num_tests_by_PCs(PC_traits, 0.995)
   ),
 
   tar_target(

### 2021-09-06 R/functions.R
change function name calc_num_tests_by_PC

@@ -1200,7 +1200,7 @@ calc_PC_traits <- function(exposure_info_list, sqc, fam, relatives){
 
 }
 
-calc_num_tests_by_PC <- function(prcomp_result, threshold){
+calc_num_tests_by_PCs <- function(prcomp_result, threshold){
 
   num_tests <- which(summary(prcomp_result)$importance[3,] > threshold)[1]
   return(num_tests)

### 2021-09-06 R/functions.R
fix calc num tests

@@ -1203,7 +1203,7 @@ calc_PC_traits <- function(exposure_info_list, sqc, fam, relatives){
 calc_num_tests_by_PCs <- function(prcomp_result, threshold){
 
   num_tests <- which(summary(prcomp_result)$importance[3,] > threshold)[1]
-  return(num_tests)
+  return(num_tests[[1]])
 }
 
 write_data_prep <- function(traits, traits_to_run, out1, out2){

### 2021-09-06 analysis/update_meeting_06_09_2021.Rmd
update meeting 6/09/21

---
title: "Update Meeting: 06-09-2021"
author: "jennysjaarda"
date: "2021-09-06"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
library(targets)
library(dplyr)
library(DT)
library(knitr)

```

# Model summaries

```{r proxyMR_DAG, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/proxyMR_DAG.png", error = FALSE)
knitr::include_graphics("assets/proxyMR_DAG_MV.png", error = FALSE)

```

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

tar_load(household_MR_binned_het)
tar_load(num_tests_by_PCs)
tar_load(standard_MR_summary_BF_sig)
tar_load(household_MR_summary_BF_sig)
tar_load(proxyMR_comparison_summary)

```

# Questions

1. Impact of meta-analyzing after MR is calculated?

2. Trend across bins: should it be a weighted linear model?

3. Less stringent BF correction: 
    - Calculate correlation matrix of all 131 traits tested. 
    - Perform PCs on the resulting corr matrix. 
    - Number of tests = # of PCs needed until variance explained is > 99.5%.
    - Resulting # of tests: `r num_tests_by_PCs`.

4. Filter by same-person MR or proxyMR?

5. Summed `rho` and `gamma` and compared this to `omega`, for the variance - it's just the sum of these two variances?

6. MV MR: X -> Z -> Y
    - Drop Z's if not significant in MV MR (i.e. `Y ~ X + Z1 + Z2 + Z3 + ...`)
    - Do we drop from the model itself and recompute the beta-coefficients? Or simply not include them in the result sum?

# AM summary

```{r AM_summary, echo = TRUE, include = TRUE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

AM_result <- household_MR_binned_het %>% filter(same_trait)

AM_sig_sex_het <- AM_result %>% filter(sex_het_pval < 0.05/num_tests_by_PCs) %>% dplyr::select(exposure_ID, exposure_sex, IVW_beta, IVW_se, sex_het_pval) %>% unique()

AM_sig_age_slope <- AM_result %>% filter(grouping_var=="age_even_bins") %>% filter(bin_slope_pval_meta < 0.05/num_tests_by_PCs) %>%
  dplyr::select(exposure_ID, exposure_sex, IVW_beta, IVW_se, bin_slope_beta_meta, bin_slope_pval_meta) %>% unique()

AM_sig_tt_slope <- AM_result %>% filter(grouping_var=="time_together_even_bins") %>% filter(bin_slope_pval_meta < 0.05/num_tests_by_PCs) %>%
  dplyr::select(exposure_ID, exposure_sex, IVW_beta, IVW_se, bin_slope_beta_meta, bin_slope_pval_meta) %>% unique()

AM_sig_age_sex_het <- AM_result %>% filter(grouping_var=="age_even_bins") %>% filter(bin_slope_sex_het_pval < 0.05/num_tests_by_PCs) %>%
  dplyr::select(exposure_ID, exposure_sex, IVW_beta, IVW_se, bin_slope_beta, bin_slope_pval, bin_slope_beta_meta, bin_slope_pval_meta) %>% unique()

AM_sig_tt_sex_het <- AM_result %>% filter(grouping_var=="time_together_even_bins") %>% filter(bin_slope_sex_het_pval < 0.05/num_tests_by_PCs) %>%
  dplyr::select(exposure_ID, exposure_sex, IVW_beta, IVW_se, bin_slope_beta, bin_slope_pval, bin_slope_beta_meta, bin_slope_pval_meta) %>% unique()

```

# ProxyMR

Restricting XY pairs: 
  - If we base it on a significant MR in the same individuals (same MR: $X \rightarrow Y$), the number of tests is: `r dim(standard_MR_summary_BF_sig)[1]`.
  - If we base it on a significant **proxyMR** ($X_{i} \rightarrow Y_{p}$), the number of tests is: `r dim(household_MR_summary_BF_sig)[1]`.

Can we use the PC multiple-test correction we applied above here, or just use conservative raw BF correction?

Comparison of omega to rho + gamma instead of just rho or gamma alone. There are more significant results using both.

```{r proxy_MR_update, echo = TRUE, include = TRUE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

table(proxyMR_comparison_summary$omega_vs_gam_BF_sig_meta)

table(proxyMR_comparison_summary$omega_vs_rho_BF_sig_meta)

table(proxyMR_comparison_summary$omega_vs_gam_rho_BF_sig_meta)

#which(proxyMR_comparison_summary$omega_vs_gam_rho_BF_sig_meta & !proxyMR_comparison_summary$omega_vs_rho_BF_sig_meta)[1:10]

t(proxyMR_comparison_summary[205,c("exposure_ID", "outcome_ID", "exposure_description", "outcome_description", "gam_meta_beta", "rho_meta_beta", "omega_meta_beta", "gam_rho_meta_beta",
                                    "gam_meta_se", "rho_meta_se", "omega_meta_se", "gam_rho_meta_se",
                                   "omega_vs_gam_meta_p", "omega_vs_rho_meta_p", "omega_vs_gam_rho_meta_p")])

```

### 2021-09-14 analysis/couple_selection.Rmd
add a few Rmd notebooks

---
title: "Couple selection in UKBB"
author: "jennysjaarda"
date: "2021-09-08"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
library(targets)
library(dplyr)

```

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

tar_load(data_main_UKBB_raw_count)
tar_load(data_id_age)
tar_load(hh_pairs)
tar_load(hh_pairs_kin)
tar_load(data_household_relationships)

```

**Couples were defined and selected according to the following scheme**

# UKBB full sample

The initial UKBB sample comprised of N = **`r data_main_UKBB_raw_count`** individuals.

# European, genetic, consenting sample

Each phenotype in the SGG database was filtered for the following criteria based on the genetic qc file (found here `r files_ref %>% filter(custom_names == "sqc") %>% pull(files)`:

- `excess.relatives=0`
- `putative.sex.chromosome.aneuploidy=0`
- `in.white.british.ancestry.subset==1`

Redacted samples and participants that removed consent were also excluded. Specifically, redacted samples are any IDs in `c(-1, -2, -3)`, from the same genetic QC sample and the samples which removed consent, according to the exclusion file `w16389_20200204.csv`.

After filtering, N = **`r dim(data_id_age)[1]`** individuals remained.

***A few notes on the PHESANT processing:***

- Raw phenotypes in the SGG database were processed according the the `PHESANT` pipeline. 
- For details on the processing by PHESANT see: https://github.com/astheeggeggs/PHESANT. 
- The pipeline is not the simplest to follow and does a lot of referring to other variables and has been modified slightly based on the data we have. 
- The main thing to note is all continuous variables were transformed to a normal distribution using an inverse normal rank transformation.
- Each phenotype file contains ID, age, and sex columns (the first 3).
- Note that this filtering procedure does not necessarily remove related individuals. For this, please use the function 'ukb_gen_samples_to_remove()' from the [UKBB r-package](https://cran.r-project.org/web/packages/ukbtools/vignettes/explore-ukb-data.html).

# Houshold pairs

Within the European, genetic, consenting sample, we limited to only individuals which came from households with exactly two participants in the UKBB based on the file: `r files_ref %>% filter(custom_names == "household_info") %>% pull(files)`

```{r, hh_pairs_filtering1, include=FALSE}
## the `hh_pairs` target includes all UKBB participants so is not a direct subset of the filter above, so we need to first remove people that are not in the PHESANT files to know how many pairs are in this sample (PHESANT files have already been filtered as above)

pairs <- hh_pairs_kin
household_relationships <- data_household_relationships
field <- household_relationships_field

pairs$sex1 <- household_relationships[["sex"]][match(pairs[["HOUSEHOLD_MEMBER1"]], household_relationships$userId)]
pairs$sex2 <- household_relationships[["sex"]][match(pairs[["HOUSEHOLD_MEMBER2"]], household_relationships$userId)]

#merge(pairs, household_relationships, by.x=, by.y=)
pairs_full_sex <- pairs[which(!is.na(pairs$sex1) & !is.na(pairs$sex2)),]
pairs_not_full_sex <- pairs[which(!(!is.na(pairs$sex1) & !is.na(pairs$sex2))),]

```

After filtering, N = **`r dim(pairs_full_sex)[1]*2`** individuals and N = **`r dim(pairs_full_sex)[1]`** couples remained.

# Unrelated pairs

Household couples defined above were filtered based on kinship (estimate of the kinship coefficient for pair based on the set of markers used in the kinship inference) from the following file: `r tar_read(path_relatives)`. Couples were excluded if kinship was > `0.05`.

```{r, hh_pairs_filtering2, include=FALSE}

pairs_unrelated <- subset(pairs_full_sex, pairs_full_sex$kinship<0.05)

```

After filtering for relatedness, N = **`r dim(pairs_unrelated)[1]*2`** individuals and N = **`r dim(pairs_unrelated)[1]`** couples remained.

# Heterosexual pairs

Unrelated, household pairs were filtered to only include heterosexual couples.

```{r, hh_pairs_filtering3, include=FALSE}

pairs_unrelated_hetero <- pairs_unrelated[-which(pairs_unrelated$sex1==pairs_unrelated$sex2),]
pairs_unrelated_hetero2 <- pairs_unrelated_hetero[-which(is.na(pairs_unrelated_hetero$sex1) | is.na(pairs_unrelated_hetero$sex2)),]

temp1 <- pairs_unrelated_hetero[which(pairs_unrelated_hetero$sex1==0),c("HOUSEHOLD_MEMBER2", "HOUSEHOLD_MEMBER1", "HOUSE_ID", "kinship", "sex2", "sex1", "tar_group")]

colnames(temp1) <- colnames(pairs_unrelated_hetero)

temp2 <- pairs_unrelated_hetero[which(pairs_unrelated_hetero$sex1==1),c("HOUSEHOLD_MEMBER1", "HOUSEHOLD_MEMBER2", "HOUSE_ID", "kinship", "sex1", "sex2", "tar_group")]
pairs_filter <- rbind(temp2, temp1)
## 79150      6

```

After removing homosexual pairs, N = **`r dim(pairs_filter)[1]*2`** individuals and N = **`r dim(pairs_filter)[1]`** couples remained.

# Couple pairs

Finally, using the data at data-field `6141`, ["How are people in household related to participant"](https://biobank.ctsu.ox.ac.uk/showcase/field.cgi?id=6141), heterosexual, unrelated, household pairs were filtered to only include couples who both said: `Husband, wife, or partner`.

```{r hh_pairs_filtering3,  include = FALSE}

pairs_filter$house_rel_member1 <- household_relationships[[field]][match(pairs_filter[["HOUSEHOLD_MEMBER1"]], household_relationships$userId)]
pairs_filter$house_rel_member2 <- household_relationships[[field]][match(pairs_filter[["HOUSEHOLD_MEMBER2"]], household_relationships$userId)]
pairs_filter2 <- pairs_unrelated_hetero[which(pairs_filter$house_rel_member1==TRUE & pairs_filter$house_rel_member2==TRUE ) ,]
## 75048     6

colnames(pairs_filter2)[which(colnames(pairs_filter2)=="sex1")] <- "HOUSEHOLD_MEMBER1_sex"
colnames(pairs_filter2)[which(colnames(pairs_filter2)=="sex2")] <- "HOUSEHOLD_MEMBER2_sex"

```

After removing non-romantic pairs, N = **`r dim(pairs_filter2)[1]*2`** individuals and N = **`r dim(pairs_filter2)[1]`** couples remained.

### 2021-09-14 analysis/notes.Rmd
add a few Rmd notebooks

---
title: "Notes"
author: "jennysjaarda"
date: "2021-09-08"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Meta-analysis across sexes

There are two inputs into MR:

1. SNP/exposure effects. 
2. SNP/outcome effects.

If we want to perform a MR between exposure $X$ and outcome $Y$ comprising both sexes, we have three options:

1. Use joint GWAS statistics (i.e. $X \sim G$ is run in sexes combined, adjusted for sex).
2. Use sex-specific GWAS statistics and then meta-analyze **before** running the MR. 
3. Use sex-specific GWAS statistics and then meta-analyze the **after** running the MR.

In other words, there are three options for pooling the data:

1. At the selection of subgroup for GWAS analysis level. 
2. At the SNP/exposure and SNP/outcome level.
3. At the MR level.

### 2021-09-14 analysis/todo_list.Rmd
add a few Rmd notebooks

---
title: "To do"
author: "jennysjaarda"
date: "2021-09-08"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# 06/09/2021

- ~~Sum $\rho$ and $\gamma$.~~
- ~~Meta-analyzed binned results across sexes.~~ 
- ~~Fix variance calculation mistake in calculating $\rho$ and $\gamma$.~~
- ~~Write script/function which replaces `Neale_ID` with trait description.~~
- Clean up `index.Rmd`
- Clean up `about.Rmd`
- Add Rmd building to `targets` pipeline. 
- MVMR. 
- AM MR summary Rmd.
- PC correlation among couples summary Rmd. 
- Clean up update meeting Rmd.

# 08/09/2021

- ~~Weight trend line by inverse variance.~~ 
- Update AF in Neale summary stats. 
- Create couple selection Rmd. 
- Number of tests for filtering by $\omega$ should exclude those with only one trait. (i.e. number of tests would be $j^2 - j$).
- ~~Filter by $\omega$ instead of same-person MR.~~ 
- Force intercept through 0 in omega vs gamma/rho plot. 
- Explore impact of meta-anlaysis at SNP/trait level vs. MR-level.
  - Meta-analyze standard MR SNP/trait estimates across sexes. 
- Hierarchy figure (add labels).
- Make visio figures in LucidChart.
- Explore IV overlap between X and Y:
    - Determine percentage of IV overlap between X IVs and Y IVs. 
    - MV MR, add SNPs/trait effect from $X_{i} \rightarrow Y_{p}$ to $Y_{i} \rightarrow Y_{p}$ MR (i.e. run a MV MR of the form: $Y_{p} \sim Y_{i} + X_{i}$).
- Add script/function which replaces `Neale_ID` with trait description to main functions.

### 2021-09-14 _targets.R
count number of lines in one of the original raw UKBB files

@@ -16,7 +16,7 @@ tar_option_set(
   ),
   packages = c("tidyverse", "data.table", "cutr", "ukbtools", "rbgen", "bigsnpr", "TwoSampleMR",
                "ggplot2", "purrr", "rmeta", "PASWR2", "cowplot", "meta", "strex", "RColorBrewer",
-               "forestplot"),
+               "forestplot", "R.utils"),
   error = "workspace",
   memory = "transient",
   garbage_collection = TRUE
@@ -115,6 +115,13 @@ list(
   tar_target(
     data_UKBB_sample,
     fread(path_UKBB_sample, skip=2, header=F,data.table=F)
+  ),
+
+  ## COUNT RAW DATA
+
+  tar_target(
+    data_main_UKBB_raw_count,
+    countLines(path_main_UKBB_raw)[[1]]-1 # subtract 1 to account for header
   ),
 
   ##################################

### 2021-09-14 code/settings.R

@@ -39,22 +39,14 @@ household_correlation_threshold <-0.1
 irnt=TRUE
 
 
-#phesant_directory <- read.table(paste0(UKBB_processed,"/PHESANT/","PHESANT_file_directory.txt"), header=T)
-#phesant_file_list <- unique(phesant_directory$File)
-
 # Data-Field 6141
 # Description:	How are people in household related to participant
 # limit to only individuals who responded YES to husband, wife, or partner
 
 relatedness_field <- "6141_1"
-#relations_file <- as.character(phesant_directory[which(phesant_directory[,2]==relatedness_field),"File"])
 
 # name of first phesant file
-#first_phesant_file <- as.character(phesant_directory[1,"File"])
 
-
-#time_at_address_file <- "/data/sgg2/jenny/data/UKBB_processed/PHESANT/ukb31459/bin1/out_bin1..tsv"
-#time_at_address_raw_file <- "/data/sgg2/jenny/data/UKBB_raw/pheno/ukb31459.csv"
 time_at_address_field <- "699"
 time_at_address_raw_field <- "699-0.0"
 
@@ -88,10 +80,9 @@ time_at_address_field <- "699"
 time_at_address_raw_field <- "699-0.0"
 
 ## Input DATA
-
 files_custom_names <- c("household_info", "phesant_directory", "relatives", "fam", "sqc", "time_at_address",
                  "time_at_address_raw", "UKBB_directory", "Neale_SGG_dir", "Neale_manifest", "code_process_Neale", "Neale_variants",
-                 "UKBB_sample")
+                 "UKBB_sample", "main_UKBB_raw")
 
 files <- c(paste0(UKBB_dir,"/pheno/ukb6881.csv"), paste0(UKBB_processed_dir,"/PHESANT/","PHESANT_file_directory.txt"),
           paste0(UKBB_dir,"/geno/","ukb1638_rel_s488366.dat"),  paste0(UKBB_dir,"/plink/_001_ukb_cal_chr9_v2.fam"),
@@ -99,7 +90,8 @@ files <- c(paste0(UKBB_dir,"/pheno/ukb6881.csv"), paste0(UKBB_processed_dir,"/PH
           paste0(UKBB_processed_dir, "/PHESANT/ukb31459/bin1/out_bin1..tsv"), paste0(UKBB_dir, "/pheno/ukb31459.csv"),
[…]

### 2021-09-14 analysis/couple_selection.Rmd
update some docs

@@
 
-Each phenotype in the SGG database was filtered for the following criteria based on the genetic qc file (found here `r files_ref %>% filter(custom_names == "sqc") %>% pull(files)`:
+Each phenotype in the SGG database was filtered for the following criteria based on the genetic qc file (located here: ``r files_ref %>% filter(custom_names == "sqc") %>% pull(files)``):
 
@@
 
-Within the European, genetic, consenting sample, we limited to only individuals which came from households with exactly two participants in the UKBB based on the file: `r files_ref %>% filter(custom_names == "household_info") %>% pull(files)`
+Within the European, genetic, consenting sample, we limited to only individuals which came from households with exactly two participants in the UKBB based on the file: ``r files_ref %>% filter(custom_names == "household_info") %>% pull(files)``.
 
@@
 
-Household couples defined above were filtered based on kinship (estimate of the kinship coefficient for pair based on the set of markers used in the kinship inference) from the following file: `r tar_read(path_relatives)`.
+Household couples defined above were filtered based on kinship (estimate of the kinship coefficient for pair based on the set of markers used in the kinship inference) from the following file: ``r files_ref %>% filter(custom_names == "relatives") %>% pull(files)``.
 Couples were excluded if kinship was > `0.05`.
@@
 
-Finally, using the data at data-field `6141`, ["How are people in household related to participant"](https://biobank.ctsu.ox.ac.uk/showcase/field.cgi?id=6141), heterosexual, unrelated, household pairs were filtered to only include couples who both said: `Husband, wife, or partner`.
+Finally, using the data at data-field `6141`, ["How are people in household related to participant"](https://biobank.ctsu.ox.ac.uk/showcase/field.cgi?id=6141), the heterosexual, unrelated, household pairs were filtered to only include couples who both responded `Husband, wife, or partner`.
 
-```{r hh_pairs_filtering3,  include = FALSE}
+```{r hh_pairs_filtering4,  include = FALSE}
 

### 2021-09-14 analysis/about.Rmd
add some documents to the website

---
title: "About"
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

# Background

Sharing environment with others plays an important role on our behaviors and health outcomes.

```{r AM_vs_convergence, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/AM_vs_convergence.png", error = FALSE)
```

## Assortative mating

- Observation of increased phenotypic similarity between couples compared to random pairs.
- This is most prominent in anthropometric traits, such as BMI and height, measures of SES and various behavioral and lifestyle measures (diet, interests, hobbies, etc.).
- This is because people tend to choose partners more similar to themselves (i.e. opposites do not actually attract in reality).
- There is also genetic evidence of assortative mating, meaning couples are not only more phenotypically similar but also more genetically similar (see [Genetic evidence of assortative mating in humans](https://www.nature.com/articles/s41562-016-0016) [Matt Robinson] and [Genetic determination of height-mediated mate choice](https://genomebiology.biomedcentral.com/articles/10.1186/s13059-015-0833-8) [Albert Tenesa])

## Couple convergence

- Additionally, couples living together can become even more similar over the years.
- Unknown to what extent couples converge and influence each other over time.
- Complicated by the fact that people tend to choose partners that are also geographically closer to them, which induces both genetic adn phenotypic similarity.

## Horizontal indirect genetic effects

- **Indirect genetic effects (IGEs)**: the effect of a genotype of one individual on the phenotype of other individuals.
- **Vertical IGEs**: IGEs across generation.
  - For example, maternal genetic effects on their offspring phenotypes (such as birthweight).
  - Non-transmitted parental genotypes have also been described to influence offspring phenotypes (e.g. educational attainment).
  - Suggest vertical IGEs can be mediated through both *biological* and *social beahviors*.
- **Horizontal IGEs**: IGEs within generation. 
  - Reported for school friends in relation to educational attainment.
  - Unknown how much these observations are spread across traits, particularly in *partners* and *siblings*.
- [Tenesa et al.](https://www.nature.com/articles/s41562-020-00991-9) sought to identify evidence of IGEs in UKB couples.
  - Used linear mixed models to estimate partner indirect heritability and found evidence of partner heritability on ~50% of traits. 
  - Further analyses suggested that for at least ~25% of these traits, the partner heritability is consistent with the existence of indirect genetic effects.

# Objectives

1. Identify traits which are directly assorted for. In other words, which traits are responsible for the phenotypic similarity we observe amongst couples. In the case of similarity observed with respect to both diet and BMI, perhaps one phenotype is driving the similarity between couples and the other is either indirect or a confounder (as diet and BMI are themselves correlated).
2. Do these patterns differ amongst sex (i.e. are females or males more particular with respect to certain traits)?
3. Is there evidence for these patterns changing with age? 
4. Can we find evidence for convergence overtime? 
5. What are the traits that are responsible for the path from an exposure in an index case impacting an outcome in their partner?

```{r convergence_paths, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/convergence_schematic.png", error = FALSE)
```

### 2021-09-14 analysis/index.Rmd
add some documents to the website

--- 2021-07-22 analysis/index.Rmd
@@
 
-Welcome to my research website.
+This is the website for the research project "proxyMR".
 
-Check-out my latex tests [here](latex_test.html)
+If you have cloned the project to a local computer this website is rendered in the `docs` subdirectory of the project directory.
+
+If you are using `workflowr` to publish the research website it will also be rendered online to GitHub Pages.
+
+This page acts as a table of contents for the website. There are links to the web pages generated from the analysis notebooks and to the rendered versions of manuscripts/documents/presentations.
+
+------------------------------------------------------------------------
+
+# Overview documents {.unnumbered}
+
+## [About](about.html) {.unnumbered}
+
+This notebook explains the central ideas behind the project.
+
+## [Notes](notes.html) {.unnumbered}
+
+This notebook is for keeping notes of any points that may be useful for
+later project or manuscript development.
+
+## [To do](todo_list.html) {.unnumbered}
+
+Current tasks and completed tasks (roughly organized by dates/weeks).
+
+## [Workflow management](workflow.html) {.unnumbered}
+
+This project uses the [`targets`](https://wlandau.github.io/targets/) and [`workflowr`](https://github.com/jdblischak/workflowr) packages for managing the workflow of the project (making sure that the dependencies between computational steps are satisfied).
+This notebook contains notes on the proposed workflow for
+using `targets` and `workflowr`.
+
+## [Meetings](meetings.html) {.unnumbered}
+
+Weekly meeting were held roughly once per week or every other week.
+
+------------------------------------------------------------------------
+
+# Publications and Presentations {.unnumbered}
+
+- Group Meeting [07-09-2021](presentations/group_meeting_070921/group_meeting_proxyMR_070921.pdf)
+
+------------------------------------------------------------------------
+
+# Analyses Notebooks {.unnumbered}
+
+These notebooks capture the analyses that were carried out to develop
+the code of the core [processing pipeline](pipeline_overview.html).
+
+## 1. Prepare data {.unnumbered}
+
+Prepare phenotypes and couples for analysis in MR pipeline.
+
+### [Couple selection](couple_selection.html) {.unnumbered}
+
+Identify couples in the UKBB.
+
+### [Phenotype selection](phenotype_selection.html) {.unnumbered}
+
+Select phenotypes for analysis in pipeline.
+
+### [Final phenotype list](final_phenotype_list.html) {.unnumbered}
+
+Final list of phenotypes for analysis in pipeline and some basic characteristics (correlation among couples, number of independent IVs).
+
+## 2. Run MR pipeline {.unnumbered}
+
+### [MR between partners (partner vs index)]() {.unnumbered}
+
+- Identify SNP-exposure effects (GW-significance) for phenotypes (sex specific) from Neale database [$X_{i} \sim G_{i}$].
+- Test SNP-outcome effects between genetic instruments and partner outcome phenotypes (sex specific) [$Y_{p} \sim G_{i}$].
+- Combine in MR to assess causal relationships between index and partner [$Y_{p} \sim X_{i}$].
+
+### [Assortative mating MR]() {.unnumbered}
+
+- Restrict to only MRs where exposure is the same as outcome (for downstream analyses) [$X_{p} \sim X_{i}$].
+
+### [MR within same individual]() {.unnumbered}
+
+- Identify SNP-exposure effects (GW-significance) for phenotypes (sex specific) from Neale database [$X_{i} \sim G_{i}$].
+- Pull SNP-outcome effects  between genetic instruments and outcome phenotypes (sex specific) from Neale database [$Y_{i} \sim G_{i}$].
+- Combine in MR to assess causal relationship between all phenotypes [$Y_{i} \sim X_{i}$].
+
+### [Compare paths towards outcome in partner]() {.unnumbered}
+
+- Compare paths from index exposure phenotype to partner outcome phenotype [$X_{i} \rightarrow Y_{p}$].
+
+### [PC correlation among couples]() {.unnumbered}
+
+------------------------------------------------------------------------
+
+# Acknowledgements {.unnumbered}
+
+Primary analysis performed by: [Jenny Sjaarda](jennysjaarda@gmail.com).
+
+Project supervisor: [Professor Zoltan Kutalik](https://wp.unil.ch/sgg/people/zoltan/).
+
+This work has been supported by the [Swiss National Science Foundation](http://www.snf.ch/en/funding/Pages/default.aspx).

### 2021-09-14 analysis/final_phenotype_list.Rmd
update some docs

---
title: "Final phenotype list"
author: "Jenny Sjaarda"
date: "2021-09-02"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
library(targets)
library(dplyr)
library(DT)
library(knitr)
library(plotly)
library(ggplot2)

```

```{r load_targets, include = FALSE}

tar_load(traits_final)

```

After [phenotype selection](phenotype_selection.html), `r dim(traits_final)[1]` traits remained for downstream analyses.

The final set of traits and some basic properties of interest can be seen in the table below.

```{r final_phenotypes, echo=FALSE}

traits_final.DT <- traits_final %>% as_tibble() %>% mutate(r = sqrt(as.numeric(r2))) %>% dplyr::select(Neale_pheno_ID, description, variable_type, category, N_pairs, r2, r, num_IVs, num_IVs_pass_het)
DT::datatable(traits_final.DT, width = "100%") %>% formatRound(columns=c('r2', 'r'), digits = 3)

```

------------------------------------------------------------------------

A few summaries of various properties are show below.

```{r final_phenotypes_summaries, echo = FALSE}
variable_type_tab <- traits_final.DT %>% count(variable_type)
category_tab <- traits_final.DT %>% count(category)

knitr::kable(variable_type_tab, format = "pandoc")
knitr::kable(category_tab, format = "pandoc")

```

------------------------------------------------------------------------

```{r IV_freq, echo = FALSE}

ggplot(traits_final.DT, aes(as.numeric(num_IVs_pass_het))) + geom_histogram(binwidth = 10, alpha=0.5) + theme_classic()  + xlab("Number of IVs") + ylab("Frequency")

ggplot(traits_final.DT, aes(as.numeric(r))) + geom_histogram(binwidth = 0.01, alpha=0.5) + theme_classic()  + xlab("Correlation (r)") + ylab("Frequency")

```

### 2021-09-14 analysis/phenotype_selection.Rmd
update some docs

---
title: "Phenotype selection"
author: "Jenny Sjaarda"
date: "2021-09-02"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
library(targets)
library(dplyr)
library(DT)
library(knitr)

```

```{r load_targets, include = FALSE}

tar_load(data_phesant_directory)
tar_load(Neale_SGG_dir_filt)
tar_load(traits_corr)
tar_load(traits_corr2)
tar_load(traits_corr3)
tar_load(traits_corr4)
tar_load(traits_corr5)

```

**Phenotypes were selected according to the following scheme**

*Note that as none of the filtering criteria apply any BF correction for the number of phenotypes, the order of the steps does not matter.*

# UKBB Phenotypes in internal database processed by PHESANT.

All SGG phenotypes went internal processing according the PHESANT pipeline. Some traits were excluded for various reasons (genetic trait, insufficient sample size, etc.).

Number of traits initially = **`r length(unique(data_phesant_directory[,2]))`**.

Traits included and their correlations amongst couples can be seen below.

```{r initial_corr, echo=FALSE}

traits_corr.DT <- traits_corr %>% as_tibble() %>% mutate_all(parse_guess) %>% mutate(r = sqrt(r2))
DT::datatable(traits_corr.DT, width = "100%") %>% formatRound(columns=c('r2', 'r'), digits = 3)
```

# Merge with Neale summary stats.

Limited to only traits that were processed by Neale and had both sex-specific and joint summary statistics available (most traits that were excluded did not have all 3 subgroups were not available, some other traits were excluded by the Neale group).

Number remaining after filter = **`r length(unique(Neale_SGG_dir_filt$phenotype))`**.

The traits that *are* in the SGG database but were **not** included due to Neale filtering can be seen below. Can compare with the [Neale Manifest](https://docs.google.com/spreadsheets/d/1kvPoupSzsSFBNSztMzl04xMoSC3Kcx3CrjVf4yBmESU/edit?ts=5b5f17db#gid=178908679) (search by ID) to get a sense of why traits were excluded (likely there is not a GWAS file for `both_sexes`, `male` and `female`).

```{r SGG_Neale_non_link, echo=FALSE}

## traits corr is the same dimension as (1) above
Neale_SGG_dir_filt_unique <- Neale_SGG_dir_filt[,c("phenotype", "phenotype_sub", "sgg_phesant_name", "description", "SGG_available", "variable_type")] %>% unique()
non_merged <- anti_join(traits_corr, Neale_SGG_dir_filt_unique, by = c("ID"="sgg_phesant_name")) %>% as_tibble() %>% mutate_all(parse_guess) %>% mutate(r = sqrt(r2))
DT::datatable(non_merged, width = "100%") %>% formatRound(columns=c('r2', 'r'), digits = 3)

```

# Evidence of correlation among couples.

Traits were excluded if the correlation among couples was < `r household_correlation_threshold` (i.e. $\sqrt{(r^2)}$).

Number remaining after filter = **`r dim(traits_corr2)[1]/3`**.

The traits remaining after this filter can be seen below.

```{r corr_filter, echo=FALSE}

traits_corr2.DT <- traits_corr2 %>% as_tibble() %>% mutate(r = sqrt(r2)) %>% dplyr::select(Neale_pheno_ID, description, variable_type, category, N_pairs, r2, r) %>% unique()
DT::datatable(traits_corr2.DT, width = "100%") %>% formatRound(columns=c('r2', 'r'), digits = 3)
```

# Sufficient power for MR based on number of IVs.

Phenotypes were excluded if the number of valid IV was < `r num_IVs_threshold`. IVs were defined based on a p-value of p < `r IV_threshold` in `both_sexes` after clumping (performed in PLINK with the options `--clump-kb 10000` and `--clump-r2 0.001` using the 1000 Genomes European samples as a reference).

Number remaining after filter = **`r dim(traits_corr3$to_run)[1]`**.

The traits remaining after IV filtering can be seen below.

```{r IV_filter, echo=FALSE}

traits_corr3_filt <- traits_corr3$to_run 
traits_corr3_filt.DT <- traits_corr3_filt %>% as_tibble() %>% mutate(r = sqrt(as.numeric(r2))) %>% dplyr::select(Neale_pheno_ID, description, variable_type, category, N_pairs, r2, r, num_IVs)
DT::datatable(traits_corr3_filt.DT, width = "100%") %>% formatRound(columns=c('r2', 'r'), digits = 3)
```

# No evidence of sex-heterogeneity among valid IVs.

Using the sex-specific summary statistics, heterogeneity between sexes was calculated. IVs that showed significant evidence of heterogeneity between sexes were included (p < 0.05/[number of IVs]). After filtering IVs with significant heterogeneity, phenotypes were filtered to again have IVs < `r num_IVs_threshold`.

Number remaining after filter = **`r dim(traits_corr4$to_run)[1]`**.

The traits remaining after sex-heterogeneity filtering can be seen below.

```{r sex_het_filter, echo=FALSE}

traits_corr4_filt <- traits_corr4$to_run 
traits_corr4_filt.DT <- traits_corr4_filt %>% as_tibble() %>% mutate(r = sqrt(as.numeric(r2))) %>% dplyr::select(Neale_pheno_ID, description, variable_type, category, N_pairs, r2, r, num_IVs, num_IVs_pass_het)
DT::datatable(traits_corr4_filt.DT, width = "100%") %>% formatRound(columns=c('r2', 'r'), digits = 3)
```

# Non-binary.

Phenotypes were limited to `continuous_irnt` or `ordinal` (i.e. `binary` traits were removed).

Number remaining after filter = **`r dim(traits_corr5)[1]`**.

The traits remaining after sex-heterogeneity filtering can be seen below.

```{r continuous_filter, echo=FALSE}

traits_corr5.DT <- traits_corr5 %>% as_tibble() %>% mutate(r = sqrt(as.numeric(r2))) %>% dplyr::select(Neale_pheno_ID, description, variable_type, category, N_pairs, r2, r, num_IVs, num_IVs_pass_het)
DT::datatable(traits_corr5.DT, width = "100%") %>% formatRound(columns=c('r2', 'r'), digits = 3)

```

### 2021-09-14 analysis/todo_list.Rmd

@@
 - ~~Write script/function which replaces `Neale_ID` with trait description.~~
-- Clean up `index.Rmd`
-- Clean up `about.Rmd`
+- ~~Clean up `index.Rmd`~~
+- ~~Clean up `about.Rmd`~~
 - Add Rmd building to `targets` pipeline. 

### 2021-09-15 R/functions.R
calculate PC correlation in couples, but IVT first in European subset (previously had done it in entire sample)

@@ -351,10 +351,13 @@ ivt <- function(x){
 }
 
 
-compute_pc_corr <- function(sqc_munge, pairs_filter){
+compute_pc_corr <- function(sqc_munge, pairs_filter, data_id_sex){
 
-  sqc_munge <- sqc_munge %>% mutate_at(vars(starts_with("PC_")), ivt)
-  PCs <- colnames(sqc_munge)[-which(colnames(sqc_munge)=="ID")]
+
+  euro_phes_data <- merge(sqc_munge, data_id_sex, by.x = "ID", by.y = "userId")
+
+  sqc_munge <- euro_phes_data %>% mutate_at(vars(starts_with("PC_")), ivt)
+  PCs <- colnames(sqc_munge)[-which(colnames(sqc_munge) %in% c("ID", "sex"))]
   trait_corr <- numeric()
   ID_sub <- NA
   for (PC in PCs)

### 2021-09-15 _targets.R

@@ -200,7 +200,7 @@ list(
 
   tar_target(
     PCs_corr,
-    compute_pc_corr(sqc_munge, hh_pairs_filter)
+    compute_pc_corr(sqc_munge, hh_pairs_filter, data_id_sex)
   ),
   # write_traits_corr = write.csv(trait_corrs, file_out( "output/tables/1.household_correlations.csv"), row.names=F),
 

### 2021-09-15 analysis/geographical_impact.Rmd
adding Rmd of geographical impact

---
title: Assessing impact of geography as confounder in causal relationships between couples
author: "jennysjaarda"
date: "2021-09-14"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")

library(DT)
library(knitr)
library(targets)
library(tidyverse)
```

```{r load_targets, include = FALSE}

tar_load(data_id_sex)
tar_load(PCs_corr)

```

# Background

- Geographical location is clearly a strong indicator of couple choice. 
- Not only do couples tend to choose partners that are phenotypically similar to them (because of common interests, political views, etc.), but also because of location (i.e. convenience).
- Geographical similarity induces genetic and in turn phenotypic similarity. 
- We sought to explore the impact of genography on mate-choice by adding PCs to the pipeline.

# Analysis

## GWAS of PCs

- GWAS of first 40 PCs run using [bgenie](https://jmarchini.org/bgenie/).
- First, inverse normal rank transformation (INT) was applied within genetic, consenting, European sample (N = `r dim(data_id_sex)[1]`) before performing GWAS (this was done before dividing into sex-specific subgroups).
- After INT, PCs were residualized for `age` and `sex`.
- Run in:
  - Entire sample (residualized for both age and sex).  
  - In sex-specific subgroups (residualized for only age). 
- This was run in the entire genetic, consenting, European UKB sample (not just couples).
- GWAS was **not** adjusted for PCs (as is typical in GWAS) because PCs are linearly independent of each other (although the sample used for PC calculation - by the UKB investigators - would be slightly different that our GWAS sample, so the PCs would not necessarily be completely linearly independent in this sample).

## Correlation of PCs between couples

- Test raw "phenotypic" correlation of PCs between partners in UKB. 
- As above, PCs were transformed using INT within genetic, consenting, European sample (N = `r dim(data_id_sex)[1]`) and correlations were calculated within couples on the transformed PCs.

```{r pc_corr, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

PCs_corr.DT <- PCs_corr %>% mutate(r = sqrt(as.numeric(r2))) 
DT::datatable(PCs_corr, width = "100%")  %>% formatSignif(columns=c('r2', 'r'), digits=3)

```

### 2021-09-15 R/functions.R
add `replace_Neale_ID` function

@@ -2792,6 +2792,19 @@ pull_z_summ_stats <- function(){
 }
 
 
+## data -> data_set with a column that has Neale_ID that we want to replace with description (eg. "exposure_ID")
+## traits_corr2_filled -> reference data set which has both Neale_ID and descrition
+## data_Neale_ID_col -> the column name of the Neale_ID (eg. "exposure_ID")
+## output_column -> the output column name of the description (eg. "description")
+
+replace_Neale_ID <- function(data, traits_corr2_filled, data_Neale_ID_col, output_column){
+
+  descriptions <- traits_corr2_filled[match(data[[data_Neale_ID_col]], traits_corr2_filled$Neale_pheno_ID), "description"]
+  data[[data_Neale_ID_col]] <- descriptions
+  colnames(data)[which(colnames(data)==data_Neale_ID_col)] <- output_column
+  return(data)
+
+}
 
 find_sig_standard_MR_summary <- function(standard_MR_summary){
 
@@ -2803,7 +2816,9 @@ find_sig_standard_MR_summary <- function(standard_MR_summary){
 find_sig_household_MR_summary <- function(household_MR_summary){
 
   denom <- dim(household_MR_summary)[1]/2 #divide by 2 because each meta result is there twice (one row/sex)
-  sig_only <- household_MR_summary %>% filter(IVW_meta_pval < 0.05/denom)
+  num_traits <- sqrt(denom)
+  num_tests <- num_traits^2 - num_traits
+  sig_only <- household_MR_summary %>% filter(IVW_meta_pval < 0.05/num_tests)
   return(sig_only)
 }
 

### 2021-09-15 analysis/geographical_impact.Rmd

@@
 
-- GWAS of first 40 PCs run using [bgenie](https://jmarchini.org/bgenie/).
-- First, inverse normal rank transformation (INT) was applied within genetic, consenting, European sample (N = `r dim(data_id_sex)[1]`) before performing GWAS (this was done before dividing into sex-specific subgroups).
-- After INT, PCs were residualized for `age` and `sex`.
-- Run in:
+- Performed GWAS of first 40 PCs, run using [bgenie](https://jmarchini.org/bgenie/).
+- Sample selection: genetic, consenting, European sample (N = `r dim(data_id_sex)[1]`), in attempt to best replicate the GWAS used in other analyses from the Neale lab. 
+- Data transformation:
+    - First, inverse normal rank tansformation (INT) was applied within the selected (this was done before dividing into sex-specific subgroups).
+    - Next, PCs were residualized for `age` and `sex`.
+- GWAS was run in:
     - Entire sample (residualized for both age and sex).  
-  - In sex-specific subgroups (residualized for only age). 
-- This was run in the entire genetic, consenting, European UKB sample (not just couples).
+    - Sex-specific subgroups (residualized for only age). 
 - GWAS was **not** adjusted for PCs (as is typical in GWAS) because PCs are linearly independent of each other (although the sample used for PC calculation - by the UKB investigators - would be slightly different that our GWAS sample, so the PCs would not necessarily be completely linearly independent in this sample).
@@
 PCs_corr.DT <- PCs_corr %>% mutate(r = sqrt(as.numeric(r2))) 
-DT::datatable(PCs_corr, width = "100%")  %>% formatSignif(columns=c('r2', 'r'), digits=3)
+DT::datatable(PCs_corr.DT, width = "100%")  %>% formatSignif(columns=c('r2', 'r'), digits=3)
 

### 2021-09-15 R/functions.R
change colnames of heterogeneity results

@@ -2443,8 +2443,8 @@ calc_binned_household_MR_het <- function(exposure_info, outcomes_to_run, househo
 
         meta_summ <- c(Q_stat_meta, Q_pval_meta, lm_summary_meta, lm_summary_meta_weight)
         names(meta_summ) <- c("Q_stat_meta", "Q_pval_meta",
-                              "bin_slope_beta_meta", "bin_slope_se_meta", "bin_slope_pval_meta",
-                              "bin_slope_beta_meta_wt", "bin_slope_se_meta_wt", "bin_slope_pval_meta_wt")
+                              "bin_slope_meta_beta", "bin_slope_meta_se", "bin_slope_pval_meta",
+                              "bin_slope_meta_beta_wt", "bin_slope_meta_se_wt", "bin_slope_meta_pval_wt")
 
         out_temp <- as.data.frame(t(c(description_sum, all_sum, diff_sum, meta_summ))) %>%
           mutate_if(is.factor,as.character) %>%

### 2021-09-15 R/functions.R
fix one more column name in het results

@@ -2443,7 +2443,7 @@ calc_binned_household_MR_het <- function(exposure_info, outcomes_to_run, househo
 
         meta_summ <- c(Q_stat_meta, Q_pval_meta, lm_summary_meta, lm_summary_meta_weight)
         names(meta_summ) <- c("Q_stat_meta", "Q_pval_meta",
-                              "bin_slope_meta_beta", "bin_slope_meta_se", "bin_slope_pval_meta",
+                              "bin_slope_meta_beta", "bin_slope_meta_se", "bin_slope_meta_pval",
                               "bin_slope_meta_beta_wt", "bin_slope_meta_se_wt", "bin_slope_meta_pval_wt")
 
         out_temp <- as.data.frame(t(c(description_sum, all_sum, diff_sum, meta_summ))) %>%

### 2021-09-15 R/functions.R
calculate PC adjustment of filtered exposure list

@@ -2830,6 +2830,26 @@ AM_filter_household_MR_summary <- function(household_MR_summary){
 }
 
 
+find_AM_sig_exposure_info <- function(household_MR_summary_AM, exposure_info, num_tests_by_PCs){
+
+
+  AM_sig_traits <- household_MR_summary_AM %>% filter(IVW_meta_pval < 0.05/num_tests_by_PCs) %>% pull(exposure_ID) %>% unique()
+  exposure_info_sub <- list()
+  j <- 1
+  for(i in 1:length(exposure_info)){
+    exposure_info_i <- exposure_info[[i]]
+    exposure_ID <- exposure_info_i %>% filter(Value=="trait_ID") %>% pull(Info)
+
+    if(exposure_ID %in% AM_sig_traits){
+      exposure_info_sub[[j]] <- exposure_info_i
+      j <- j+1
+    }
+  }
+  return(exposure_info_sub)
+
+}
+
+
 z.test_p <- Vectorize(function(x, sigma.x, y, sigma.y) {z.test(x, sigma.x, y, sigma.y)$p},
                         vectorize.args = c("x", "sigma.x", "y", "sigma.y"))
 
@@ -2948,17 +2968,16 @@ run_proxyMR_comparison <- function(exposure_info, household_MR_summary_BF_sig, h
     prod_diff_list <- list()
 
     prod_diff_result <- summarized_result %>%
-      mutate(omega_vs_gam_p = z.test_p(gam_beta, gam_se, omega_beta, omega_se)) %>%
-      mutate(omega_vs_rho_p = z.test_p(rho_beta, rho_se, omega_beta, omega_se)) %>%
-      mutate(omega_vs_gam_rho_p = z.test_p(gam_rho_beta, gam_rho_se, omega_beta, omega_se)) %>%
-      mutate(gam_vs_rho_p = z.test_p(rho_beta, rho_se, gam_beta, gam_se)) %>%
+      mutate(omega_vs_gam_pval = z.test_p(gam_beta, gam_se, omega_beta, omega_se)) %>%
+      mutate(omega_vs_rho_pval = z.test_p(rho_beta, rho_se, omega_beta, omega_se)) %>%
+      mutate(omega_vs_gam_rho_pval = z.test_p(gam_rho_beta, gam_rho_se, omega_beta, omega_se)) %>%
+      mutate(gam_vs_rho_pval = z.test_p(rho_beta, rho_se, gam_beta, gam_se)) %>%
 
 
[…]

### 2021-09-15 _targets.R

@@ -507,7 +507,7 @@ list(
 
 
   tar_target(
-    household_MR, ## `household_MR` is run in full sample only, not binned by age / time-together categories
+    household_MR, ## `household_MR` is run in full sample only, not binned by age / time-together categories.
     run_household_MR_comprehensive(exposure_info, outcomes_to_run, household_harmonised_data, MR_method_list),
     pattern = map(exposure_info, household_harmonised_data), iteration = "list"
   ),
@@ -525,9 +525,25 @@ list(
   ),
 
   tar_target(
-    ##filter to only MR between same traits
+    ## filter to only MR between same traits.
     household_MR_summary_AM,
     AM_filter_household_MR_summary(household_MR_summary)
+  ),
+
+  tar_target(
+    ## after AM MR results is reduced to only those significant based on `num_tests_by_PCs`, reduce further for testing heterogeneity between sexes and amongst bins to reduce number of tests.
+    exposure_info_AM_sig,
+    find_AM_sig_exposure_info(household_MR_summary_AM, exposure_info, num_tests_by_PCs)
+  ),
+
+  tar_target(
+    PC_traits_AM_sig,
+    calc_PC_traits(exposure_info_AM_sig, data_sqc, data_fam, data_relatives)
+  ),
+
+  tar_target(
+    num_tests_by_PCs_AM_sig,
+    calc_num_tests_by_PCs(PC_traits_AM_sig, 0.995)
   ),
 
   tar_target(

### 2021-09-15 analysis/AM_MR_summary.Rmd
add AM MR summary

---
title: "Assortative mating MR summary"
author: "Jenny Sjaarda"
date: "2021-09-15"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
source("R/functions.R")
library(targets)
library(dplyr)
library(DT)
library(knitr)

```

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

tar_load(traits_final)
tar_load(household_MR_binned_het)
tar_load(household_MR_summary_AM)
tar_load(traits_corr2_filled)
tar_load(num_tests_by_PCs)
tar_load(num_tests_by_PCs_AM_sig)

```

# Background

- For certain analyses it is worth only looking at MR results between couples where the exposure trait is the same as the outcome trait. 
- These results we have termed, the assortative mating MR results.

```{r single_trait_fig, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/single_trait_MR.png", error = FALSE)
```

# Results

```{r AM_data_prep, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, include = FALSE}

AM_sig_traits <- household_MR_summary_AM %>% filter(IVW_meta_pval < 0.05/num_tests_by_PCs) %>% pull(exposure_ID) %>% unique()
AM_result_sig <- household_MR_summary_AM %>% filter(exposure_ID %in% AM_sig_traits)

household_MR_binned_het_temp <- replace_Neale_ID(household_MR_binned_het, traits_corr2_filled, "exposure_ID", "exposure_description")
household_MR_binned_het_temp2 <- replace_Neale_ID(household_MR_binned_het_temp, traits_corr2_filled, "outcome_ID", "outcome_description")
household_MR_binned_het_descript <- household_MR_binned_het_temp2

```

- The number of analyses corresponds to the number of traits = `r dim(traits_final)[1]`. 
- The significant results after adjusting for multiple hypothesis testing are shown below. 
- `IVW_meta_beta` and `IVW_meta_pval` corresponds to the beta and p-value of the meta-analyzed MR across sexes, respectively.  
- After adjusting for multiple tests (p < `0.05 / `r num_tests_by_PCs``), the number of significant AM MRs was `r length(AM_sig_traits)` (see below on details for how number of effective tests was obtained).

```{r AM_overview, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

AM_result_sig.DT <- AM_result_sig %>% dplyr::select(exposure_description, IVW_meta_beta, IVW_meta_pval) %>% unique()
DT::datatable(AM_result_sig.DT, width = "100%") %>% formatSignif(columns=c('IVW_meta_beta', 'IVW_meta_pval'), digits=3)

```

**A note on multiple testing adjustment:**
- Adjustment was based on PCs of correlation matrix.  
- Calculate correlation matrix of all `r dim(traits_final)[1]` traits tested. 
- Perform PCs on the resulting corr matrix. 
- Number of tests = # of PCs needed until variance explained is > 99.5%. 
- Resulting # of effective tests: `r num_tests_by_PCs`.

## Sex heterogeneity

- For each AM MR ($X_{p} \sim X_i$), analyses were run in each sex separately. 
- Of the `r length(AM_sig_traits)` significant results shown above, we tested to see if there was any difference in AM MR estimates between sexes. 
- A similar adjustment for multiple hypothesis testing was applied as above, resulting in the number of effective tests as `r num_tests_by_PCs_AM_sig`. 
- The table below displays the results that significantly differ among sexes (p < `0.05/`r num_tests_by_PCs_AM_sig``).

```{r AM_sex_het, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

AM_sig_sex_het <- AM_result_sig %>% filter(IVW_sex_het_pval < 0.05/num_tests_by_PCs_AM_sig)

AM_result_sig.DT <- AM_sig_sex_het %>% dplyr::select(exposure_description, IVW_beta, IVW_pval, IVW_meta_beta, IVW_meta_pval, IVW_sex_het_pval) %>% unique()
DT::datatable(AM_result_sig.DT, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'IVW_meta_beta', 'IVW_meta_pval', 'IVW_sex_het_pval'), digits=3)

```

- Only `r dim(AM_result_sig.DT)[1]/2` was significant after adjustment. 
- The table below shows the results significant at `p < 0.05`.

```{r AM_sex_het_nom, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

AM_sig_sex_het_nom <- AM_result_sig %>% filter(IVW_sex_het_pval < 0.05)
AM_result_sig_nom.DT <- AM_sig_sex_het_nom %>% dplyr::select(exposure_description, IVW_beta, IVW_pval, IVW_meta_beta, IVW_meta_pval, IVW_sex_het_pval) %>% unique()
DT::datatable(AM_result_sig_nom.DT, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'IVW_meta_beta', 'IVW_meta_pval', 'IVW_sex_het_pval'), digits=3)

```

## Binned results

- For each AM MR ($X_{p} \sim X_i$), MR analyses were run in the full sample as well as in 5 roughly equal sized bins according to time couple had been together (`time_together`) estimated using [time at household variable](https://biobank.ctsu.ox.ac.uk/showcase/field.cgi?id=699), and median age (`age`). 
- Of the `r length(AM_sig_traits)` significant results shown above, we tested to see if there was any significant difference in AM MR estimates amongst the two grouping variables. 
- Specifically, the outcome data set was split into 5 bins, and the SNP-outcome effect was estimated in each bin separately. 
- These outcome-SNP effects were then used to generate bin-specific MR estimates, using the same SNP-exposure effects from Neale. 
- To estimate the difference among bins, we used two approaches: 
  1. The Cochran's Q test to test for heterogeneity in meta-analyses.
  2. Tested the significance of the slope of linear model of median bin (either `age` or `time together`) versus the bin-specific MR estimate.  
- The two results are shown below. Note that each statistic was computed in each sex separately and meta-analyzed.

### Q-statistic across bins

```{r bin_q_het_setup, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

q_het_result_age <- household_MR_binned_het_descript %>% filter(same_trait==TRUE) %>% filter(grouping_var=="age_even_bins") %>% filter(IVW_meta_pval < 0.05/num_tests_by_PCs) %>%
  dplyr::select(exposure_description, exposure_sex, outcome_sex, grouping_var, IVW_beta, IVW_pval, Q_pval, Q_pval_meta) %>% arrange(Q_pval_meta)

q_het_result_tt <- household_MR_binned_het_descript %>% filter(same_trait==TRUE) %>% filter(grouping_var=="time_together_even_bins") %>% filter(IVW_meta_pval < 0.05/num_tests_by_PCs) %>%
  dplyr::select(exposure_description, exposure_sex, outcome_sex, grouping_var, IVW_beta, IVW_pval, Q_pval, Q_pval_meta) %>% arrange(Q_pval_meta)

```

The number of significant q-statistics significant after multiple hypothesis testing (p < `0.05/`r num_tests_by_PCs_AM_sig` = `r 0.05/num_tests_by_PCs_AM_sig```) in each group was:
  - Binned by median *age* of couples, number of significant results within meta = `r length(which(q_het_result_age$Q_pval_meta < 0.05/num_tests_by_PCs_AM_sig))/2` and sex-specific = `r length(which(q_het_result_age$Q_pval < 0.05/num_tests_by_PCs_AM_sig))/2`.
  - Binned by *time* couple had been together, number of significant results within meta = `r length(which(q_het_result_tt$Q_pval_meta < 0.05/num_tests_by_PCs_AM_sig))/2` and sex-specific = `r length(which(q_het_result_tt$Q_pval < 0.05/num_tests_by_PCs_AM_sig))/2`.

All results are displayed below (sort by Q_pval to view significant results):

```{r bin_q_het, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

DT::datatable(q_het_result_age, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'Q_pval', 'Q_pval_meta'), digits=3)
DT::datatable(q_het_result_tt, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'Q_pval', 'Q_pval_meta'), digits=3)
```

### Trend across bins

```{r bin_slope_setup, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

slope_result_age <- household_MR_binned_het_descript %>% filter(same_trait==TRUE) %>% filter(grouping_var=="age_even_bins") %>% filter(IVW_meta_pval < 0.05/num_tests_by_PCs) %>%
  dplyr::select(exposure_description, exposure_sex, grouping_var, bin_slope_beta, bin_slope_pval, bin_slope_meta_pval, bin_slope_meta_pval_wt, bin_slope_sex_het_pval) %>% arrange(bin_slope_meta_pval)

slope_result_tt <- household_MR_binned_het_descript %>% filter(same_trait==TRUE) %>% filter(grouping_var=="time_together_even_bins") %>% filter(IVW_meta_pval < 0.05/num_tests_by_PCs) %>%
  dplyr::select(exposure_description, exposure_sex, grouping_var, bin_slope_beta, bin_slope_pval, bin_slope_meta_pval, bin_slope_meta_pval_wt, bin_slope_sex_het_pval) %>% arrange(bin_slope_meta_pval)

```

Slopes were calculated by estimating the beta-coefficient of a linear regression between MR estimate within each bin (outcome) and median bin (either age or time at same household). Linear models were run both *unweighted* and *weighted* for the by the inverse of the SE of the MR estimate.

The number of significant slopes ($MR_{bin} \sim median_{bin}$) significant after multiple hypothesis testing (p < `0.05/`r num_tests_by_PCs_AM_sig` = `r 0.05/num_tests_by_PCs_AM_sig```) in each group was:
  - Binned by median *age* of couples, number of significant results within meta = `r length(which(slope_result_age$bin_slope_meta_pval < 0.05/num_tests_by_PCs_AM_sig))/2`, weighted by inverse of the SE = `r length(which(slope_result_age$bin_slope_meta_pval_wt < 0.05/num_tests_by_PCs_AM_sig))/2`, different between sexes = `r length(which(slope_result_age$bin_slope_sex_het_pval < 0.05/num_tests_by_PCs_AM_sig))/2`
  - Binned by *time* couple had been together, number of significant results within meta = `r length(which(slope_result_tt$bin_slope_meta_pval < 0.05/num_tests_by_PCs_AM_sig))/2` and weighted by inverse of the SE = `r length(which(slope_result_tt$bin_slope_meta_pval_wt < 0.05/num_tests_by_PCs_AM_sig))/2`, different between sexes = `r length(which(slope_result_tt$bin_slope_sex_het_pval < 0.05/num_tests_by_PCs_AM_sig))/2`
All results are displayed below (sort by Q_pval to view significant results):

All results are displayed below (sort by various p-values to view significant results):

```{r bin_q_het, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

DT::datatable(slope_result_age, width = "100%") %>% formatSignif(columns=c('bin_slope_beta', 'bin_slope_pval', 'bin_slope_meta_pval', 'bin_slope_meta_pval_wt', 'bin_slope_sex_het_pval'), digits=3)
DT::datatable(slope_result_tt, width = "100%") %>% formatSignif(columns=c('bin_slope_beta', 'bin_slope_pval', 'bin_slope_meta_pval', 'bin_slope_meta_pval_wt', 'bin_slope_sex_het_pval'), digits=3)
```

### 2021-09-15 analysis/update_meeting_05_07_2021.Rmd
update all update meetings Rmds

--- 2021-07-05 analysis/update_meeting_05_07_2021.Rmd
@@
 title: "Update Meeting: 05-07-2021"
-author: "jennysjaarda"
+author: "Jenny Sjaarda"
 date: "2021-07-02"
@@
 
-```{r, echo=FALSE,include = FALSE}
-# You need this code to conduct the magic dependences attaching...
-DT::datatable(matrix())
+```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
+
+tar_load(household_MR_summary)
+
 ```
 
-## Last week summary.
+# Last week summary.
 
@@
   - Going through paper in-depth. 
+- Next step: run MR between all traits within couples (previously had only looked at single trait at a time).
 
-## Results.
+```{r single_trait_MR, echo = FALSE}
+library(knitr)
+knitr::include_graphics("assets/single_trait_MR.png", error = FALSE)
+```
 
-- Univariate MR.
+# Results.
 
-```{r sex_het, echo = FALSE}
+Ran MR between all pairs of traits within couples.
+Number of tests = $j^2 * 2$, where $j$ is the number of traits.
+MR models were run in each sex sepeartely (male to female and female to male), which is why we multiply by 2.
 
-tar_load(household_MR_exhaustive_summary)
+```{r two_trait_MR_fig, echo = FALSE}
+library(knitr)
+knitr::include_graphics("assets/two_trait_model.png", error = FALSE)
+```
 
-household_MR_exhaustive_summary <- household_MR_exhaustive_summary %>% mutate(same_traits = case_when(exposure_ID == outcome_ID ~ TRUE, 
-                                                                   TRUE ~ FALSE))
+The results significant after multiple hypothesis testing can be seen below.
 
-DT::datatable(household_MR_exhaustive_summary, width = "100%")
+```{r two_trait_MR_results, echo = FALSE}
+
+household_MR_summary.DT <- household_MR_summary %>% dplyr::select(exposure_description, outcome_description, same_trait, exposure_sex, IVW_beta, IVW_pval, IVW_meta_beta, IVW_meta_pval)
+
+num_tests <- dim(household_MR_summary.DT)[1]/2
+household_MR_summary.DT.BF <- household_MR_summary.DT %>% filter(IVW_pval < 0.05/num_tests)
+
+DT::datatable(household_MR_summary.DT.BF, width = "100%") %>% formatRound(columns=c('IVW_beta', 'IVW_pval', 'IVW_meta_beta', 'IVW_meta_pval'), digits = 3)
 
@@
 
-## Questions.
+The total number of tests = `r num_tests`, the number significant after adjusting for BF correction = `r dim(household_MR_summary.DT.BF)[1]`.
+
+# Questions.
 
@@
     5. No evidence of sex heterogeneity among valid IVs. 
-  - Could add more phenotpyes as outcomes
+  - Could add more phenotpyes as outcomes?
 
-## Next steps.
+# Next steps.
 
--
+- Calculate MR in same person to finish all analyses in diagram above.

### 2021-09-15 analysis/update_meeting_06_09_2021.Rmd
update all update meetings Rmds

@@
 title: "Update Meeting: 06-09-2021"
-author: "jennysjaarda"
+author: "Jenny Sjaarda"
 date: "2021-09-06"
@@
 source("code/settings.R")
+source("R/functions.R")
 library(targets)
@@
 library(knitr)
-
-```
-
-# Model summaries
-
-```{r proxyMR_DAG, echo = FALSE}
-library(knitr)
-knitr::include_graphics("assets/proxyMR_DAG.png", error = FALSE)
-knitr::include_graphics("assets/proxyMR_DAG_MV.png", error = FALSE)
+library(kableExtra)
 
@@
 
-# Questions
+# Last meeting summary.
+
+- Need to sum $\rho$ and $\gamma$ together and calculate difference of the ***sum*** from $\omega$ (previously only calculated differences between each pair, not sum).
+- Use less stringent BF calculation for AM MR (for calculating trends across bins).
+
+# Model summaries.
+
+```{r proxyMR_DAG, echo = FALSE}
+library(knitr)
+knitr::include_graphics("assets/two_trait_model.png", error = FALSE)
+knitr::include_graphics("assets/MV_trait_model.png", error = FALSE)
+
+```
+
+# Questions.
 
 1. Impact of meta-analyzing after MR is calculated? 
-
 2. Trend across bins: should it be a weighted linear model? 
-
 3. Less stringent BF correction: 
@@
     - Resulting # of tests: `r num_tests_by_PCs`.  
-
 4. Filter by same-person MR or proxyMR?
-
-5. Summed `rho` and `gamma` and compared this to `omega`, for the variance - it's just the sum of these two variances?
-
-6. MV MR: X -> Z -> Y
-    - Drop Z's if not significant in MV MR (i.e. `Y ~ X + Z1 + Z2 + Z3 + ...`)
+5. Summed $\rho$ and $\gamma$ and compared this to `omega`, for the variance - it's just the sum of these two variances? 
+6. MV MR: $X \rightarrow  Z \rightarrow Y$
+    - Drop Z's if not significant in MV MR (i.e. $Y \sim X + Z_1 + Z_2 + Z_3...$)
     - Do we drop from the model itself and recompute the beta-coefficients? Or simply not include them in the result sum?
 
-# AM summary
+# AM summary.
 
-```{r AM_summary, echo = TRUE, include = TRUE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
-
-AM_result <- household_MR_binned_het %>% filter(same_trait)
-
-AM_sig_sex_het <- AM_result %>% filter(sex_het_pval < 0.05/num_tests_by_PCs) %>% dplyr::select(exposure_ID, exposure_sex, IVW_beta, IVW_se, sex_het_pval) %>% unique()
-
-AM_sig_age_slope <- AM_result %>% filter(grouping_var=="age_even_bins") %>% filter(bin_slope_pval_meta < 0.05/num_tests_by_PCs) %>%
-  dplyr::select(exposure_ID, exposure_sex, IVW_beta, IVW_se, bin_slope_beta_meta, bin_slope_pval_meta) %>% unique()
-
-AM_sig_tt_slope <- AM_result %>% filter(grouping_var=="time_together_even_bins") %>% filter(bin_slope_pval_meta < 0.05/num_tests_by_PCs) %>%
-  dplyr::select(exposure_ID, exposure_sex, IVW_beta, IVW_se, bin_slope_beta_meta, bin_slope_pval_meta) %>% unique()
-
-AM_sig_age_sex_het <- AM_result %>% filter(grouping_var=="age_even_bins") %>% filter(bin_slope_sex_het_pval < 0.05/num_tests_by_PCs) %>%
-  dplyr::select(exposure_ID, exposure_sex, IVW_beta, IVW_se, bin_slope_beta, bin_slope_pval, bin_slope_beta_meta, bin_slope_pval_meta) %>% unique()
-
-AM_sig_tt_sex_het <- AM_result %>% filter(grouping_var=="time_together_even_bins") %>% filter(bin_slope_sex_het_pval < 0.05/num_tests_by_PCs) %>%
-  dplyr::select(exposure_ID, exposure_sex, IVW_beta, IVW_se, bin_slope_beta, bin_slope_pval, bin_slope_beta_meta, bin_slope_pval_meta) %>% unique()
-
-```
+A summary of single trait MR within couples (AM MR) can be found [here](AM_MR_summary.html).
 
@@
 Restricting XY pairs:
+
   - If we base it on a significant MR in the same individuals (same MR: $X \rightarrow Y$), the number of tests is: `r dim(standard_MR_summary_BF_sig)[1]`.
@@
 
-Comparison of omega to rho + gamma instead of just rho or gamma alone. There are more significant results using both.
+Comparison of omega to rho + gamma instead of just rho or gamma alone.
+There are more significant results using both, would have expected there to be less.
+An example of one such result is shown below.
 
-```{r proxy_MR_update, echo = TRUE, include = TRUE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
+```{r proxy_MR_update, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
 
-table(proxyMR_comparison_summary$omega_vs_gam_BF_sig_meta)
+t1 <- as.data.frame(table(proxyMR_comparison_summary$omega_vs_gam_BF_sig_meta))
+colnames(t1) <- c("omega_vs_gam_BF_sig_meta", "Frq")
 
-table(proxyMR_comparison_summary$omega_vs_rho_BF_sig_meta)
+kbl(t1) %>% kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
 
-table(proxyMR_comparison_summary$omega_vs_gam_rho_BF_sig_meta)
+t2 <- as.data.frame(table(proxyMR_comparison_summary$omega_vs_rho_BF_sig_meta))
+colnames(t2) <- c("omega_vs_rho_BF_sig_meta", "Frq")
+
+kbl(t2) %>% kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
+
+t3 <- as.data.frame(table(proxyMR_comparison_summary$omega_vs_gam_rho_BF_sig_meta))
+colnames(t3) <- c("omega_vs_gam_rho_BF_sig_meta", "Frq")
+
+kbl(t3) %>% kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
 
@@
 
-t(proxyMR_comparison_summary[205,c("exposure_ID", "outcome_ID", "exposure_description", "outcome_description", "gam_meta_beta", "rho_meta_beta", "omega_meta_beta", "gam_rho_meta_beta",
+sample_dat <- t(proxyMR_comparison_summary[241,c("exposure_ID", "outcome_ID", "exposure_description", "outcome_description", "gam_meta_beta", "rho_meta_beta", "omega_meta_beta", "gam_rho_meta_beta",
                                     "gam_meta_se", "rho_meta_se", "omega_meta_se", "gam_rho_meta_se",
                                    "omega_vs_gam_meta_p", "omega_vs_rho_meta_p", "omega_vs_gam_rho_meta_p")])
+sample_dat <- as.data.frame(sample_dat)
+colnames(sample_dat) <- c("Info")
+
+kbl(sample_dat) %>% kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
 

### 2021-09-15 analysis/update_meeting_14_06_2021.Rmd
update all update meetings Rmds

@@
 title: "Update Meeting: 14-06-2021"
-author: "jennysjaarda"
+author: "Jenny Sjaarda"
 date: "2021-06-14"
@@
 source("code/settings.R")
-load("summary.Rdata")
+source("R/functions.R")
 library(DT)
 library(knitr)
+library(targets)
 
@@
 
-```{r, echo=FALSE,include = FALSE}
-# You need this code to conduct the magic dependencies attaching...
-DT::datatable(matrix())
+```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
+
+tar_load(traits_final)
+tar_load(household_MR_binned_het)
+tar_load(traits_corr2_filled)
+
 ```
 
-## Last week summary.
+# Last week summary.
 
-### Getting back up to speed.
+## Getting back up to speed.
 
@@
 
-### Next steps:
+## Next steps:
 
@@
 
-### Follow-up questions:
+## Follow-up questions:
 
@@
 
-### Responses:
+## Responses:
 
@@
 
-## Results.
+# Results.
 
-### Traits analyzed.
+## Traits analyzed.
 
@@
 
-DT::datatable(traits, width = "100%")
+traits_final.DT <- traits_final %>% as_tibble() %>% mutate(r = sqrt(as.numeric(r2))) %>% dplyr::select(Neale_pheno_ID, description, variable_type, category, N_pairs, r2, r, num_IVs, num_IVs_pass_het)
+DT::datatable(traits_final.DT, width = "100%") %>% formatRound(columns=c('r2', 'r'), digits = 3)
 
@@
 
-### Summary of results
+## Summary of AM MR results
 
@@
 
-reduced_mr_summary <- calc_mr_sex_het %>% dplyr::select(trait_ID, trait_description, 
-                                                        exposure_sex, N_snps,"IVW/Wald_summary","IVW/Wald_pval", 
-                                                        "MR_meta_est", "MR_meta_pval", "sex_het")
+AM_result <- household_MR_binned_het %>% filter(same_trait)
+reduced_mr_summary <- AM_result %>% dplyr::select("exposure_ID", "exposure_sex","IVW_beta","IVW_pval", 
+                                                        "IVW_meta_beta", "IVW_meta_pval", "sex_het_pval")
 
-DT::datatable(reduced_mr_summary, width = "100%")
+reduced_mr_summary.DT <- replace_Neale_ID(reduced_mr_summary, traits_corr2_filled, "exposure_ID", "description")
+
+DT::datatable(reduced_mr_summary.DT, width = "100%") %>% formatRound(columns=c('IVW_beta', 'IVW_pval', 'IVW_meta_beta', 'IVW_meta_pval', 'sex_het_pval'), digits = 3)
 
@@
 
-### Testing discrepancy amongst bins.
+## Testing discrepancy amongst bins.
 
@@
 
-mr_summary_age_reduced <- mr_summary_age %>% dplyr::select(trait_ID, exposure_sex, Q_stat, Q_pval)
-DT::datatable(mr_summary_age_reduced, width = "100%")
+mr_summary_age_reduced <- AM_result %>% dplyr::filter(grouping_var == "age_even_bins") %>% dplyr::select(exposure_ID, exposure_sex, grouping_var, Q_pval, Q_pval_meta)
+mr_summary_age_reduced.DT <- replace_Neale_ID(mr_summary_age_reduced, traits_corr2_filled, "exposure_ID", "description")
 
-mr_summary_tt_reduced <- mr_summary_tt %>% dplyr::select(trait_ID, exposure_sex, Q_stat, Q_pval)
-DT::datatable(mr_summary_tt_reduced, width = "100%")
+DT::datatable(mr_summary_age_reduced.DT, width = "100%") %>% formatRound(columns=c('Q_pval', 'Q_pval_meta'), digits = 3)
+
+mr_summary_tt_reduced <- AM_result %>% dplyr::filter(grouping_var == "time_together_even_bins") %>% dplyr::select(exposure_ID, exposure_sex, grouping_var, Q_pval, Q_pval_meta)
+mr_summary_tt_reduced.DT <- replace_Neale_ID(mr_summary_tt_reduced, traits_corr2_filled, "exposure_ID", "description")
+
+DT::datatable(mr_summary_tt_reduced.DT, width = "100%") %>% formatRound(columns=c('Q_pval', 'Q_pval_meta'), digits = 3)
 

### 2021-09-15 analysis/update_meeting_14_07_2021.Rmd
update all update meetings Rmds

--- 2021-07-14 analysis/update_meeting_14_07_2021.Rmd
@@
 title: "Update Meeting: 14-07-2021"
-author: "jennysjaarda"
+author: "Jenny Sjaarda"
 date: "2021-06-14"
@@
 source("code/settings.R")
-load("summary.Rdata")
 library(DT)
@@
 library(targets)
-
 ```
 
-```{r, echo=FALSE,include = FALSE}
-# You need this code to conduct the magic dependences attaching...
-DT::datatable(matrix())
-```
-
-## Last week summary.
+# Last week summary.
 
@@
 
-## Update.
+# Update.
 
-### Reprocessed PHESANT
+## Reprocessed PHESANT phenotypes.
 
-- For all new phenotype files since starting the project and added them to the pipeline.
+To include all possible phenotypes, reprocessed PHESANT for all new phenotype files that have been added to SGG databased since starting the project and added them to the pipeline.
+The list of phenotypes included can be seen below.
 
-```{r new_phenos, echo = FALSE}
+```{r new_phenos, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
 
@@
 
-### Identified a problem with the Neale summary stats.
+## Identified a problem with the Neale summary stats.
+
 - Only affecting the `low_confidence_variant` column. 
@@
 
-### Adding PCs to the pipeline.
+## Adding PCs to the pipeline.
 
 - Currently running GWAS of first 40 PCs, residualized for `age` and `sex`. Applied inverse normal rank transformation first.  
-- Test between partners.
+- Test correlation between partners.
 
-```{r pc_corr, echo = FALSE}
+```{r pc_corr, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
 
 tar_load(PCs_corr)
-
-DT::datatable(PCs_corr, width = "100%")
+PCs_corr.DT <- PCs_corr %>% mutate(r = sqrt(as.numeric(r2))) 
+DT::datatable(PCs_corr.DT, width = "100%")  %>% formatSignif(columns=c('r2', 'r'), digits=3)
 
@@
 
-## Next steps.
+# Next steps.
 
@@
 $$
-\begin{eqnarray*}
-	(1) \; X_{i} &=& G_{i}\cdot {\beta_{x, i\rightarrow i} } + \epsilon \\
-	(2) \; X_{p} &=& G_{i}\cdot {\beta_{x, i\rightarrow p} } + \epsilon \\
-	(3) \; Y_{i} &=& G_{i}\cdot {\beta_{y, i\rightarrow i} } + \epsilon \\
-	(4) \; Y_{p} &=& G_{i}\cdot {\beta_{y, i\rightarrow p} } + \epsilon \\
-	(5) \; Y_{i} &=& \alpha_j\cdot X_{i} +\epsilon \\
-	(6) \; X_{p} &=& \alpha_{x, am}\cdot X_{i{}} +\epsilon \\
-	(7) \; Y_{p} &=& \alpha_{y, am}\cdot Y_{i{}} +\epsilon \\
-	(8) \; Y_{p} &=& \alpha_{k}\cdot X_{i} +\epsilon
-
-\end{eqnarray*}
+\begin{align}
+(1) \; X_{i} = G_{i}\cdot {\beta_{x, i\rightarrow i} } + \epsilon \\
+(2) \; X_{p} = G_{p}\cdot {\beta_{x, p\rightarrow p} } + \epsilon \\
+(3) \; X_{p} = G_{i}\cdot {\beta_{x, i\rightarrow p} } + \epsilon \\
+(4) \; Y_{i} = G_{i}\cdot {\beta_{y, i\rightarrow i} } + \epsilon \\
+(5) \; Y_{p} = G_{i}\cdot {\beta_{y, i\rightarrow p} } + \epsilon \\
+(6) \; Y_{i} = \alpha_{x_{i} \rightarrow y_{i}}\cdot X_{i} +\epsilon \\
+(7) \; Y_{p} = \alpha_{x_{p} \rightarrow y_{p}}\cdot X_{p} +\epsilon \\
+(8) \; X_{p} = \alpha_{x_{i} \rightarrow x_{p}}\cdot X_{i{}} +\epsilon \\
+(9) \; Y_{p} = \alpha_{y_{i} \rightarrow y_{p}}\cdot Y_{i{}} +\epsilon \\
+(10) \; Y_{p} = \alpha_{x_{i} \rightarrow y_{p}}\cdot X_{i} +\epsilon
+\end{align}
 $$
 
-- Equation (1) and (3) are simply same-person GWAS result (obtained from Neale).
-- Equation (2) and (4) are household GWAS results (calculated in-house).
-- Equation (5), and ${\alpha_{j}}$, in the causal effect of a trait ${X}$ on ${Y}$ from a standard MR single-person MR.
-- Equation (6) and (7), represent assortative mating effects for traits ${X}$ and ${Y}$, respectively.
-- Equation (8) ${\alpha_{k}}$ is the result of phenotype ${X}$ in the index, having a causal effect on a phenotype ${Y}$ in the partner. 
-- This effect is either due to the product of one of the two:
-  - ${\alpha_{j}}$ with ${\alpha_{x, am}}$ 
-  - ${\alpha_{j}}$ with ${\alpha_{y, am}}$
+- Equation (1), (2), and (4) are simply same-person GWAS result (obtained from Neale).
+- Equation (3) and (5) are household GWAS results (calculated in-house).
+- Equation (6) and (7) (${\alpha_{x_{i} \rightarrow y_{i}}}$ and $\alpha_{x_{p} \rightarrow y_{p}}$), in the causal effect of a trait ${X}$ on ${Y}$ from a standard MR single-person MR, sex specific.
+- Equation (8) and (9), represent assortative mating effects for traits ${X}$ and ${Y}$, respectively ($\alpha_{x_{i} \rightarrow x_{p}}$ and $\alpha_{y_{i} \rightarrow y_{p}}$).
+- Equation (10) $\alpha_{x_{i} \rightarrow y_{p}}$ is the result of phenotype ${X}$ in the index, having a causal effect on a phenotype ${Y}$ in the partner. 
+    - This effect is either due to:
+        (1) The product of $\alpha_{x_{i} \rightarrow x_{p}}$ with $\alpha_{x_{p} \rightarrow y_{p}}$, or 
+        (2) The product of $\alpha_{x_{i} \rightarrow y_{i}}$ with $\alpha_{y_{i} \rightarrow y_{p}}$, or
+        (3) An additional 3rd (or 4th, etc.) variable which amplifies the effect from ${X_{i}}$ to ${Y_{p}}$ that doesn't simply act through a combination of assortative mating and a causal effect of $X$ on $Y$.

### 2021-09-15 analysis/update_meeting_25_08_2021.Rmd
update all update meetings Rmds

@@
 title: "Update Meeting: 25-08-2021"
-author: "jennysjaarda"
+author: "Jenny Sjaarda"
 date: "2021-08-24"
@@
 
-## Last meeting summary.
+```{r load_targets, include = FALSE}
 
-## Model summary
+tar_load(household_MR_binned_het)
+tar_load(traits_final)
+tar_load(standard_MR_summary_BF_sig)
+tar_load(traits_corr2_filled)
+tar_load(proxyMR_comparison_summary)
+tar_load(proxyMR_prod_comparison_figure)
+tar_load(proxyMR_sex_comparison_figure)
+```
+
+# Last meeting summary.
+
+- Discussed summary of pathways from $X$ in index ($X_i$) to $Y$ in the partner ($Y_p$).
+- Need for running the same individual MR (i.e. simply the standard MR) so we can compare the two pathways (see DAG below).
+- Multiply perpendicular paths together and to calculate the variance between them use the formula [here]( https://stats.stackexchange.com/questions/397839/varxy-if-x-and-y-are-independent-random-variables).
+
+# Model summary.
 
@@
 library(knitr)
-knitr::include_graphics("assets/proxyMR_DAG.png", error = FALSE)
+knitr::include_graphics("assets/two_trait_model.png", error = FALSE)
 ```
 
-## Results
+# Results.
 
-### Trends amongst bins
+## Heterogeneity amonst different groups.
 
-```{r MR_binned_het, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
-tar_load(household_MR_binned_het)
-DT::datatable(household_MR_binned_het, width = "100%")
+### Sex heterogeneity
+
+For each proxyMR ($Y_{p} \sim X_i$), analyses were run in each sex separately.
+The results below display the results that differ among sexes, after BF correction (i.e.
+`0.05 / (`r dim(traits_final)[1]` * `r dim(traits_final)[1]`)`)
+
+```{r sig_sex_het, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
+
+household_MR_binned_het_temp <- replace_Neale_ID(household_MR_binned_het, traits_corr2_filled, "exposure_ID", "exposure_description")
+household_MR_binned_het_temp2 <- replace_Neale_ID(household_MR_binned_het_temp, traits_corr2_filled, "outcome_ID", "outcome_description")
+household_MR_binned_het <- household_MR_binned_het_temp2
+
+num_tests <- dim(traits_final)[1]^2
+sex_het_BF_result <- household_MR_binned_het %>% filter(grouping_var=="age_even_bins") %>% 
+  dplyr::select(exposure_description, outcome_description, exposure_sex, outcome_sex, same_trait, IVW_beta, IVW_pval, sex_het_pval) %>% filter(as.numeric(sex_het_pval) < 0.05/num_tests)
+DT::datatable(sex_het_BF_result, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'sex_het_pval'), digits=3)
 ```
 
-### Standard MR
+However in this case, it may be more sensible to only consider same traits (i.e.
+testing for more of an "assortative mating" effect), reducing the number of tests to ``r dim(traits_final)[1]``.
+However, none pass BF-significance (p < `0.05/`r dim(traits_final)[1]` = `r 0.05/dim(traits_final)[1]``).
+The most significant results is `lifetime number of sexual partners`.
+
+```{r sex_het_same_trait, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
+
+sex_het_same_trait_result <- household_MR_binned_het %>% filter(grouping_var=="age_even_bins") %>% filter(same_trait==TRUE) %>%
+ dplyr::select(exposure_description, exposure_sex, outcome_sex, same_trait, IVW_beta, IVW_pval, sex_het_pval) %>% arrange(sex_het_pval)
+
+DT::datatable(sex_het_same_trait_result, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'sex_het_pval'), digits=3)
+```
+
+### Binned heterogeneity
+
+For each proxyMR ($Y_{p} \sim X_i$), MR analyses were run in the full sample as well as in 5 roughly equal sized bins according to `time_together` and `age`.
+Specifically, the outcome data set was split into 5 bins, and the SNP-outcome effect was estimated in each bin separately.
+These outcome-SNP effects were then used to generate bin-specific MR estimates, using the same SNP-exposure effects from Neale.
+
+To estimate the difference among bins, used two approaches:
+
+  1. The Cochran's Q test to test for heterogeneity in meta-analyses.
+  2. Tested the significance of the slope of linear model of median bin (either `age` or `time together`) versus the bin-specific MR estimate.
+
+The two results are shown below, restricted to only same_traits. Note that each statistic was computed in each sex separately and meta-analyzed.
+
+#### Q-statistic results
+
+```{r bin_q_het, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
+
+q_het_result_age <- household_MR_binned_het %>% filter(same_trait==TRUE) %>% filter(grouping_var=="age_even_bins") %>% 
+  dplyr::select(exposure_description, exposure_sex, outcome_sex, grouping_var, IVW_beta, IVW_pval, Q_pval, Q_pval_meta)
+
+q_het_result_tt <- household_MR_binned_het %>% filter(same_trait==TRUE) %>% filter(grouping_var=="time_together_even_bins") %>% 
+  dplyr::select(exposure_description, exposure_sex, outcome_sex, grouping_var, IVW_beta, IVW_pval, Q_pval, Q_pval_meta)
+
+DT::datatable(q_het_result_age, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'Q_pval', 'Q_pval_meta'), digits=3)
+DT::datatable(q_het_result_tt, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'Q_pval', 'Q_pval_meta'), digits=3)
+```
+
+#### Slope results
+
+```{r bin_slope_het, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
+
+slope_het_result_age <- household_MR_binned_het %>% filter(same_trait==TRUE) %>% filter(grouping_var=="age_even_bins") %>% 
+  dplyr::select(exposure_description, exposure_sex, outcome_sex, grouping_var, IVW_beta, IVW_pval, bin_slope_beta, bin_slope_pval, bin_slope_meta_beta, bin_slope_meta_pval, bin_slope_sex_het_pval)
+
+slope_het_result_tt <- household_MR_binned_het %>% filter(same_trait==TRUE) %>% filter(grouping_var=="time_together_even_bins") %>% 
+  dplyr::select(exposure_description, exposure_sex, outcome_sex, grouping_var, IVW_beta, IVW_pval, bin_slope_beta, bin_slope_pval, bin_slope_meta_beta, bin_slope_meta_pval, bin_slope_sex_het_pval)
+
+DT::datatable(slope_het_result_age, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'bin_slope_beta', 'bin_slope_pval', 'bin_slope_meta_beta', 'bin_slope_meta_pval', 'bin_slope_sex_het_pval'), digits=3)
+DT::datatable(slope_het_result_tt, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'bin_slope_beta', 'bin_slope_pval', 'bin_slope_meta_beta', 'bin_slope_meta_pval', 'bin_slope_sex_het_pval'), digits=3)
+```
+
+## Standard MR
+
+To complete all MR anlayses in the DAG above, needed to generate same-person MR estimates (i.e.
+standard MR).
+These were also run sex-specific, using Neale summary statistics.
+As expected, there are many BF-significant summary MR results.
+Significance was determined based on the meta-analyzed result across sexes.
+Thus the number of tests was equal to: ``r dim(traits_final)[1]`^2 - `r dim(traits_final)[1]` = `r dim(traits_final)[1]^2 - dim(traits_final)[1]` (number of traits = ``r dim(traits_final)[1]``).
+The BF-significant results can be seen below.
 
 ```{r standard_MR, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
-tar_load(standard_MR_summary_BF_sig)
-DT::datatable(standard_MR_summary_BF_sig, width = "100%")
+
+standard_MR_summary_BF_sig_sub <- standard_MR_summary_BF_sig %>% dplyr::select(exposure_description, outcome_description, N_snps, IVW_meta_beta, IVW_meta_pval) %>% unique()
+DT::datatable(standard_MR_summary_BF_sig_sub, width = "100%") %>% formatSignif(columns=c('IVW_meta_beta', 'IVW_meta_pval'), digits=3)
 ```
 
-### Paths from X-index to Y-partner
+Number of BF-signifcant MR pairs: `r dim(standard_MR_summary_BF_sig)[1]`.
 
-```{r proxyMR_products, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
-tar_load(proxyMR_figure_data)
-DT::datatable(proxyMR_figure_data, width = "100%")
+## Paths from X-index to Y-partner
+
+There are three distinct paths from $X_{index}$ to $Y_{partner}$ as illustrated in the DAG above:
+
+  1. Assortative mating through $X$: $\gamma$
+  2. Assortative mating through $Y$: $\rho$
+  3. Indirectly (i.e. through a latent variable which either increases or decreases the observed effect of $X_{index}$ to $Y_{partner}$): $\omega$
+
+The estimates of $\gamma$, $\rho$ and $\omega$ can be seen below, meta-analyzed across sex.
+
+```{r proxyMR_products, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), eval = TRUE}
+
+proxyMR_comparison_summary.DT <- proxyMR_comparison_summary %>% filter(exposure_sex=="male") %>% 
+  dplyr::select(exposure_ID, outcome_ID, exposure_description, outcome_description, 
+                gam_meta_beta, gam_meta_pval, gam_sex_het_pval, 
+                rho_meta_beta, rho_meta_pval, rho_sex_het_pval,
+                rho_meta_beta, rho_meta_pval, rho_sex_het_pval)
+
+DT::datatable(proxyMR_comparison_summary.DT, width = "100%") %>% formatSignif(columns=c('gam_meta_beta', 'gam_meta_pval', 'gam_sex_het_pval', 
+                                                                                        'rho_meta_beta', 'rho_meta_pval', 'rho_sex_het_pval',
+                                                                                        'rho_meta_beta', 'rho_meta_pval', 'rho_sex_het_pval'), digits=3)
 ```
 
-#### Heterogeneity among paths
+### Heterogeneity among paths
 
-```{r proxyMR_figure1, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
-tar_load(proxyMR_prod_comparison_figure)
+Next, we wanted to compare the three paths from $X_index$ to $Y_partner$. This was done using a standard heterogeneity test.
+
+```{r proxyMR_product_comparison, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), eval = TRUE}
+
+proxyMR_comparison_summary_prod <- proxyMR_comparison_summary %>%
+  dplyr::select(exposure_description, outcome_description, 
+                exposure_category, outcome_category, exposure_sex,
+                omega_vs_gam_p, omega_vs_gam_meta_p,
+                omega_vs_rho_p, omega_vs_rho_meta_p,
+                gam_vs_rho_p, gam_vs_rho_meta_p,
+                omega_vs_gam_rho_p, omega_vs_gam_rho_meta_p)
+
+DT::datatable(proxyMR_comparison_summary_prod, width = "100%") %>% formatSignif(
+  columns = c(
+    'omega_vs_gam_p',
+    'omega_vs_gam_meta_p',
+    'omega_vs_rho_p',
+    'omega_vs_rho_meta_p',
+    'gam_vs_rho_p',
+    'gam_vs_rho_meta_p',
+    'omega_vs_gam_rho_p',
+    'omega_vs_gam_rho_meta_p'
+  ),
+  digits = 3
+)
+```
+
+``````{r proxyMR_figure1, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600}
 proxyMR_prod_comparison_figure
@@
 
-#### Sex heterogeneity amongst paths
+### Sex heterogeneity amongst paths
 
-``````{r proxyMR_figure2, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
-tar_load(proxyMR_sex_comparison_figure)
+``````{r proxyMR_figure2, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600}
 proxyMR_sex_comparison_figure
@@
 
-#### Hierarchical clustering
+### Hierarchical clustering
+
+Working on a figure like the one [here](https://www.r-graph-gallery.com/310-custom-hierarchical-edge-bundling.html) and [here](https://www.r-graph-gallery.com/311-add-labels-to-hierarchical-edge-bundling.html).

### 2021-09-15 analysis/AM_MR_summary.Rmd
add AM MR summary

@@
 tar_load(num_tests_by_PCs_AM_sig)
+tar_load(exposures_to_run)
+tar_load(household_MR_AM_FvsM_fig)
 
@@
 
-- For certain analyses it is worth only looking at MR results between couples where the exposure trait is the same as the outcome trait. 
-- These results we have termed, the assortative mating MR results.
+For certain analyses it is worth only looking at MR results between couples where the exposure trait is the same as the outcome trait. We've called these analyses *assortative mating MR (AM MR)*.
 
@@
 
-- The number of analyses corresponds to the number of traits = `r dim(traits_final)[1]`. 
+- The number of analyses corresponds to the number of traits (`r dim(traits_final)[1]`). 
 - The significant results after adjusting for multiple hypothesis testing are shown below. 
-- `IVW_meta_beta` and `IVW_meta_pval` corresponds to the beta and p-value of the meta-analyzed MR across sexes, respectively.  
-- After adjusting for multiple tests (p < `0.05 / `r num_tests_by_PCs``), the number of significant AM MRs was `r length(AM_sig_traits)` (see below on details for how number of effective tests was obtained).
+- `IVW_meta_beta` and `IVW_meta_pval` corresponds to the beta and p-value of the meta-analyzed MR across sexes, respectively (i.e.
+MR estimates were computed in each sex-seperately using sex-specific SNP-exposure and SNP-outcome results and then meta-analyzed).
+- After adjusting for multiple tests (p < `0.05/`r num_tests_by_PCs``), identified `r length(AM_sig_traits)` significant AM MR relationships (see below on details for how number of effective tests was obtained).
 
@@
 **A note on multiple testing adjustment:**
+
   - Adjustment was based on PCs of correlation matrix.  
@@
 - Of the `r length(AM_sig_traits)` significant results shown above, we tested to see if there was any difference in AM MR estimates between sexes. 
-- A similar adjustment for multiple hypothesis testing was applied as above, resulting in the number of effective tests as `r num_tests_by_PCs_AM_sig`. 
+- A similar adjustment for multiple hypothesis testing was applied as above, resulting in `r num_tests_by_PCs_AM_sig` number of effective tests. 
 - The table below displays the results that significantly differ among sexes (p < `0.05/`r num_tests_by_PCs_AM_sig``).
@@
 
-- Only `r dim(AM_result_sig.DT)[1]/2` was significant after adjustment. 
-- The table below shows the results significant at `p < 0.05`.
+Only `r dim(AM_result_sig.DT)[1]/2` was significant after adjustment. The table below shows the results significant at `p < 0.05`.
 
@@
 
+In general, the male to female estimate is slightly larger than the female to male estimate, as shown in the figure below.
+
+```{r sex_het_fig, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
+household_MR_AM_FvsM_fig
+```
+
 ## Binned results
 
-- For each AM MR ($X_{p} \sim X_i$), MR analyses were run in the full sample as well as in 5 roughly equal sized bins according to time couple had been together (`time_together`) estimated using [time at household variable](https://biobank.ctsu.ox.ac.uk/showcase/field.cgi?id=699), and median age (`age`).
+- For each AM MR ($X_{p} \sim X_i$), analyses were run in the full sample as well as in 5 roughly equal sized bins according to time couple had been together (`time_together_even_bins`) estimated using [time at household variable](https://biobank.ctsu.ox.ac.uk/showcase/field.cgi?id=699), and median age (`age_even_bins`).
 - Of the `r length(AM_sig_traits)` significant results shown above, we tested to see if there was any significant difference in AM MR estimates amongst the two grouping variables. 
-- Specifically, the outcome data set was split into 5 bins, and the SNP-outcome effect was estimated in each bin separately. 
+- Specifically, the outcome data set was split into 5 bins, and the SNP-outcome effect was estimated in each bin separately (and each sex seperately). 
 - These outcome-SNP effects were then used to generate bin-specific MR estimates, using the same SNP-exposure effects from Neale. 
@@
     2. Tested the significance of the slope of linear model of median bin (either `age` or `time together`) versus the bin-specific MR estimate.  
-- The two results are shown below. Note that each statistic was computed in each sex separately and meta-analyzed.
+- The two results are shown below.
 
@@
 
-The number of significant q-statistics significant after multiple hypothesis testing (p < `0.05/`r num_tests_by_PCs_AM_sig` = `r 0.05/num_tests_by_PCs_AM_sig```) in each group was:
-  - Binned by median *age* of couples, number of significant results within meta = `r length(which(q_het_result_age$Q_pval_meta < 0.05/num_tests_by_PCs_AM_sig))/2` and sex-specific = `r length(which(q_het_result_age$Q_pval < 0.05/num_tests_by_PCs_AM_sig))/2`.
-  - Binned by *time* couple had been together, number of significant results within meta = `r length(which(q_het_result_tt$Q_pval_meta < 0.05/num_tests_by_PCs_AM_sig))/2` and sex-specific = `r length(which(q_het_result_tt$Q_pval < 0.05/num_tests_by_PCs_AM_sig))/2`.
+The number of significant q-statistics significant after multiple hypothesis testing (p < `0.05/`r num_tests_by_PCs_AM_sig` = `r 0.05/num_tests_by_PCs_AM_sig``) in each group was:
 
-All results are displayed below (sort by Q_pval to view significant results):
+  - Binned by median *age* of couples, number of significant results within meta = `r length(which(q_het_result_age$Q_pval_meta < 0.05/num_tests_by_PCs_AM_sig))/2` and sex-specific = `r length(which(q_het_result_age$Q_pval < 0.05/num_tests_by_PCs_AM_sig))`.
+  - Binned by *time* couple had been together, number of significant results within meta = `r length(which(q_het_result_tt$Q_pval_meta < 0.05/num_tests_by_PCs_AM_sig))/2` and sex-specific = `r length(which(q_het_result_tt$Q_pval < 0.05/num_tests_by_PCs_AM_sig))`.
+
+All results are displayed below (sort by Q_pval to view significant results).
 
@@
 
-Slopes were calculated by estimating the beta-coefficient of a linear regression between MR estimate within each bin (outcome) and median bin (either age or time at same household).
+Slopes were calculated by estimating the beta-coefficient of a linear regression between MR estimate within each bin (dependent variable) and median bin (independent variable, either age or time at same household).
 Linear models were run both *unweighted* and *weighted* for the by the inverse of the SE of the MR estimate.
 
-The number of significant slopes ($MR_{bin} \sim median_{bin}$) significant after multiple hypothesis testing (p < `0.05/`r num_tests_by_PCs_AM_sig` = `r 0.05/num_tests_by_PCs_AM_sig```) in each group was:
-  - Binned by median *age* of couples, number of significant results within meta = `r length(which(slope_result_age$bin_slope_meta_pval < 0.05/num_tests_by_PCs_AM_sig))/2`, weighted by inverse of the SE = `r length(which(slope_result_age$bin_slope_meta_pval_wt < 0.05/num_tests_by_PCs_AM_sig))/2`, different between sexes = `r length(which(slope_result_age$bin_slope_sex_het_pval < 0.05/num_tests_by_PCs_AM_sig))/2`
-  - Binned by *time* couple had been together, number of significant results within meta = `r length(which(slope_result_tt$bin_slope_meta_pval < 0.05/num_tests_by_PCs_AM_sig))/2` and weighted by inverse of the SE = `r length(which(slope_result_tt$bin_slope_meta_pval_wt < 0.05/num_tests_by_PCs_AM_sig))/2`, different between sexes = `r length(which(slope_result_tt$bin_slope_sex_het_pval < 0.05/num_tests_by_PCs_AM_sig))/2`
-All results are displayed below (sort by Q_pval to view significant results):
+The number of significant trends ($\beta$ estimates from the model: $\alpha_{bin} \sim median_{bin}$) significant after multiple hypothesis testing (p < `0.05/`r num_tests_by_PCs_AM_sig` = `r 0.05/num_tests_by_PCs_AM_sig``) in each group was:
 
-All results are displayed below (sort by various p-values to view significant results):
+  - Binned by median *age* of couples, number of significant results after adjusting for number of tests:
+      - $\beta$ unweighted, meta-anlayzed across sexes: `r length(which(slope_result_age$bin_slope_meta_pval < 0.05/num_tests_by_PCs_AM_sig))/2`
+      - $\beta$ weighted by inverse of the SE, meta-anlayzed across sexes: `r length(which(slope_result_age$bin_slope_meta_pval_wt < 0.05/num_tests_by_PCs_AM_sig))/2`
+      - $\beta$ significantly different between sexes: `r length(which(slope_result_age$bin_slope_sex_het_pval < 0.05/num_tests_by_PCs_AM_sig))/2`
+  - Binned by *time* couple had been together, number of significant results after adjusting for number of tests:
+      - $\beta$ unweighted, meta-anlayzed across sexes: `r length(which(slope_result_tt$bin_slope_meta_pval < 0.05/num_tests_by_PCs_AM_sig))/2`
+      - $\beta$ weighted by inverse of the SE, meta-anlayzed across sexes: `r length(which(slope_result_tt$bin_slope_meta_pval_wt < 0.05/num_tests_by_PCs_AM_sig))/2`
+      - $\beta$ significantly different between sexes: `r length(which(slope_result_tt$bin_slope_sex_het_pval < 0.05/num_tests_by_PCs_AM_sig))/2`.
 
-```{r bin_q_het, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
+All results are displayed below (sort by various p-values to view significant results).
+
+```{r bin_slope, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
 
@@
 ```
+
+### Figures
+
+A few figures of interest are shown below.
+
+```{r binned_figures, include = T, echo = FALSE}
+
+target1 <- which(exposures_to_run$Neale_pheno_ID=="1080")
+figs1 <- tar_read(household_MR_binned_AM_figs, branches = target1)[[1]]
+
+figs1[[5]]
+figs1[[7]]
+
+target2 <- which(exposures_to_run$Neale_pheno_ID=="1239")
+figs2 <- tar_read(household_MR_binned_AM_figs, branches = target2)[[1]]
+
+figs2[[5]]
+figs2[[7]]
+
+target3 <- which(exposures_to_run$Neale_pheno_ID=="20153_irnt")
+figs3 <- tar_read(household_MR_binned_AM_figs, branches = target3)[[1]]
+figs3[[5]]
+figs3[[9]]
+
+target4 <- which(exposures_to_run$Neale_pheno_ID=="1050")
+figs4 <- tar_read(household_MR_binned_AM_figs, branches = target4)[[1]]
+figs4[[4]]
+figs4[[8]]
+```

### 2021-09-15 analysis/index.Rmd
misc website updates

@@
 title: "Home"
+author: "Jenny Sjaarda"
 site: workflowr::wflow_site
@@
 
-## [Workflow management](workflow.html) {.unnumbered}
+## [Project Workflow Status](status.html) {.unnumbered}
+
+This notebook displays the computational status of the project, that is, whether everything is up to date.
+Ironically, this is currently the only notebook that must be run manually, so this notebook only displays the status **when the notebook was last executed** and there is no indication whether the project status has changed since then.
 
 This project uses the [`targets`](https://wlandau.github.io/targets/) and [`workflowr`](https://github.com/jdblischak/workflowr) packages for managing the workflow of the project (making sure that the dependencies between computational steps are satisfied).
-This notebook contains notes on the proposed workflow for
-using `targets` and `workflowr`.
 
@@
 
-### [Assortative mating MR]() {.unnumbered}
+### [Assortative mating MR](AM_MR_summary.html) {.unnumbered}
 
@@
 
-### [PC correlation among couples]() {.unnumbered}
+### [Assess impact of geography on causal relationships between couples](geographical_impact.html) {.unnumbered}
 

### 2021-09-15 R/functions.R
change some labels in figures

@@ -3876,7 +3876,7 @@ forest_plot_binned_sex_specific <- function(harmonised_data, MR_binned, group, c
 
   outcome_levels <- levels(bin_summary$exposure_sex)
 
-  xlab <- paste0("AM MR estimate for ", tolower(trait_description), " (sex-specific)")
+  xlab <- paste0("AM MR estimate for ", tolower(trait_description), "\n(sex-specific)")
 
   xmin <- min(bin_summary$IVW_beta - bin_summary$IVW_se)
   xmax <- max(bin_summary$IVW_beta + bin_summary$IVW_se)
@@ -4093,7 +4093,7 @@ xy_plot_binned_sex_specifc <- function(harmonised_data, MR_binned, group, custom
   x_ticks <- unique(fig_data %>% pull(bin_median))
   x_labels <- unique(fig_data %>% pull(bin))
 
-  xlab <- ifelse(group == "time_together_even_bins", "Time together in same household (years)", "Median age of couples (years)")
+  xlab <- ifelse(group == "time_together_even_bins", "Time together in\nsame household (years)", "Median age of\ncouples (years)")
 
   legend_title <- "Exposure sex"
 

### 2021-09-24 analysis/update_meeting_23_09_2021.Rmd
add update meeting 23-09-2021

---
title: "Update Meeting: 23-09-2021"
author: "Jenny Sjaarda"
date: "2021-09-20"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
library(targets)
library(dplyr)
library(DT)
library(knitr)
library(ggplot2)

```

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

tar_load(proxyyMR_IV_overlap)
tar_load(proxyMR_MR_paths_summary_yiyp_adj)
tar_load(proxyMR_comparison_summary_yiyp_adj)
```

# Last meeting summary.

- Because the results comparing $\omega$ to $\rho + \gamma$ were slightly confusing (see last meeting summary [here](https://jennysjaarda.github.io/proxyMR/update_meeting_06_09_2021.html)), decided we should explore the impact of $X_i$ on the $Y_i \rightarrow Y_p$ causal effects.

```{r two_trait_model, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/two_trait_model.png", error = FALSE)
```

- Two tasks to explore:
    1. Identify overlap between $X_i \rightarrow X_p$ IVs and $Y_i \rightarrow Y_p$. 
    2. Adjust the $Y_i \rightarrow Y_p$ MR results for effects of $X_i$ (as depicted below).

```{r YiYp_adj_model, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/two_trait_model_yiyp_adj.png", error = FALSE)
```

# Results.

## Overlap between IVs for $X$ and $Y$.

- Is simply calculating the SNP overlap good enough? 
- Better to calculate signal overlap using LD at some $r^2$ cutoff (what threshold would be useful?).
  - For each IV used to estimate the causal effect $Y_i \rightarrow Y_p$,
  - Test if they are in LD with any IV used to estimate the causal effect $X_i \rightarrow X_p$ at $r^2$ > `0.9`, 
  - Repeat for all IVs for $Y_i$ to calcualte the percentage overlap between the two sets of IVs. 
  - Define percentage overlap as: (# of $Y_i$ IVs are in LD with at least one $X_i$ IV at $r^2$ > `0.9`) / (total number of $Y_i$ IVs used).

The results can be seen below. The `YiXi_IV_exact_overlap` column indicates the percentage of $Y_i$ IVs that are an exact match with $X_i$ IVs (same `rs#`). The `YiXi_IV_sig_overlap` column denotes the percentage overlap of signals, as defined above.

```{r XiYi_IV_overlap, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
proxyyMR_IV_overlap.DT <- proxyyMR_IV_overlap %>% dplyr::select(exposure_description, outcome_description, YiXi_IV_exact_overlap, YiXi_IV_sig_overlap) %>% unique() %>% arrange(desc(YiXi_IV_sig_overlap))

DT::datatable(proxyyMR_IV_overlap.DT, width = "100%") %>% formatRound(columns=c('YiXi_IV_exact_overlap', 'YiXi_IV_sig_overlap'), digits = 3)
```

The results with the most overlap involve traits that are nearly identical.

## Adjust $Y_i \rightarrow Y_p$.

Performed MV MR $Y_i \rightarrow Y_p$ relationships: $Y_p \sim Y_i + X_i$. Where IVs were the same as the two trait MR ($Y_p \sim Y_i$) based on $Y$, but the effect of the IVs on $X_i$ were also included in the model.

```{r YiYp_adj, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
yiyp_adj.DT <- proxyMR_MR_paths_summary_yiyp_adj %>% dplyr::select(exposure_description, outcome_description, exposure_sex, starts_with("yiyp_IVW"))

DT::datatable(yiyp_adj.DT, width = "100%") %>% formatRound(columns=c('YiXi_IV_exact_overlap', 'YiXi_IV_sig_overlap'), digits = 3)

fig_dat <- full_join(proxyyMR_IV_overlap, proxyMR_MR_paths_summary_yiyp_adj)

create_fig <- function(fig_dat, custom_col){

  #AM_result <- household_MR_binned_het %>% filter(same_trait)

  plot_data <- fig_dat %>% #dplyr::select(exposure_ID, exposure_sex, IVW_beta, IVW_se, IVW_pval, sex_het_pval) %>% unique() %>%
    #pivot_wider(names_from = "exposure_sex", values_from = c("IVW_beta", "IVW_se", "IVW_pval")) %>%
    mutate(overlap_above_50 = case_when(YiXi_IV_sig_overlap > 0.5 ~ TRUE, TRUE ~ FALSE))

  legend_title <- "IV overlap"
  plot <- ggplot2::ggplot(data = plot_data, ggplot2::aes(x = yiyp_IVW_beta_adj,
                                                         y = yiyp_IVW_beta, color = factor(overlap_above_50))) +
    geom_point(alpha = 3/4) +

    geom_smooth(mapping = aes(x = yiyp_IVW_beta_adj, y = yiyp_IVW_beta), method = "lm", se=FALSE, formula = y~x+0, fullrange=TRUE, color = custom_col[4]) +
    geom_abline(slope=1, intercept=0, color = "black") +
    scale_color_manual(values = custom_col[c(1,2)], labels=c("overlap > 50%","overlap =< 50%")) +
    theme_minimal() +
    ggplot2::labs(colour = legend_title, x = paste0("MR estimate (Y-index to Y-partner)"),
                  y = paste("AM MR estimate adjusted for IV effect on relevant X"))
}

```

## Updated results with adjusted $Y_i$

```{r proxy_MR_update, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

t1 <- as.data.frame(table(proxyMR_comparison_summary_yiyp_adj$omega_vs_gam_BF_sig_meta))
colnames(t1) <- c("omega_vs_gam_BF_sig_meta", "Frq")

kbl(t1) %>% kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

t2 <- as.data.frame(table(proxyMR_comparison_summary_yiyp_adj$omega_vs_rho_BF_sig_meta))
colnames(t2) <- c("omega_vs_rho_BF_sig_meta", "Frq")

kbl(t2) %>% kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

t3 <- as.data.frame(table(proxyMR_comparison_summary_yiyp_adj$omega_vs_gam_rho_BF_sig_meta))
colnames(t3) <- c("omega_vs_gam_rho_BF_sig_meta", "Frq")

```

## MVMR

Run all MVMR models but need to sum up along all $\beta$ from the MV model to create the $\gamma_{mv}$.

MV MR: $X \rightarrow  Z \rightarrow Y$
  - Drop Z's if not significant in MV MR (i.e. $Y \sim X + Z_1 + Z_2 + Z_3...$)
  - Do we drop from the model itself and recompute the beta-coefficients? Or simply not include them in the result sum?

```{r MV_model, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/MV_trait_model_yiyp_adj.png", error = FALSE)
``

### 2021-09-24 analysis/update_meeting_23_09_2021.Rmd
select the right columns for DT

@@
 ```{r YiYp_adj, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
-yiyp_adj.DT <- proxyMR_MR_paths_summary_yiyp_adj %>% dplyr::select(exposure_description, outcome_description, exposure_sex, starts_with("yiyp_IVW"))
+yiyp_adj.DT <- proxyMR_MR_paths_summary_yiyp_adj %>% dplyr::select(exposure_description, outcome_description, exposure_sex, yiyp_IVW_beta, yiyp_IVW_se, yiyp_IVW_pval, yiyp_IVW_beta_adj, yiyp_IVW_se_adj, yiyp_IVW_pval_adj)
 
-DT::datatable(yiyp_adj.DT, width = "100%") %>% formatRound(columns=c('YiXi_IV_exact_overlap', 'YiXi_IV_sig_overlap'), digits = 3)
+format_round_cols <- colnames(dplyr::select_if(yiyp_adj.DT, is.numeric))
+DT::datatable(yiyp_adj.DT, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)
 
@@
 }
+
+create_fig(fig_dat, custom_col)
 

### 2021-09-24 R/functions.R
fix small typo in cat message `prune_z_summ_stats`

@@ -3576,7 +3576,7 @@ prune_z_summ_stats <- function(MV_z, z_summ_stats, prune_threshold){
     #mv_result <- mr_mvivw(mv_data_format) # see `str(mv_result)` to know how to access the results
 
     output_list[[paste0(outcome_ID, "_vs_", exposure_ID, "_", exposure_sex)]] <- (mv_data_format)
-    cat(paste0("Finished pruning summary statistics for ", i, " of ", dim(MV_z_data)[1], " traits.\n\n"))
+    cat(paste0("Finished pruning summary statistics for ", i, " of ", length(z_summ_stats), " traits.\n\n"))
 
 
 

### 2021-09-24 analysis/update_meeting_23_09_2021.Rmd
publish update meeting 24-09-21

@@
 library(ggplot2)
+library(kableExtra)
 
@@
 
-- Two tasks to explore:
+- Two things to explore:
     1. Identify overlap between $X_i \rightarrow X_p$ IVs and $Y_i \rightarrow Y_p$. 
@@
 
-- Is simply calculating the SNP overlap good enough? 
-- Better to calculate signal overlap using LD at some $r^2$ cutoff (what threshold would be useful?).
+Is simply calculating the SNP overlap good enough? Better to calculate signal overlap using LD at some $r^2$ cutoff (what threshold would be useful?).
+
   - For each IV used to estimate the causal effect $Y_i \rightarrow Y_p$,
@@
 The `YiXi_IV_exact_overlap` column indicates the percentage of $Y_i$ IVs that are an exact match with $X_i$ IVs (same `rs#`).
-The `YiXi_IV_sig_overlap` column denotes the percentage overlap of signals, as defined above.
+The `YiXi_IV_sig_overlap` column denotes the proportion overlap of signals, as defined above.
 
-```{r XiYi_IV_overlap, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
+```{r XiYi_IV_overlap, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
 proxyyMR_IV_overlap.DT <- proxyyMR_IV_overlap %>% dplyr::select(exposure_description, outcome_description, YiXi_IV_exact_overlap, YiXi_IV_sig_overlap) %>% unique() %>% arrange(desc(YiXi_IV_sig_overlap))
@@
 
-```{r YiYp_adj, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
-yiyp_adj.DT <- proxyMR_MR_paths_summary_yiyp_adj %>% dplyr::select(exposure_description, outcome_description, exposure_sex, yiyp_IVW_beta, yiyp_IVW_se, yiyp_IVW_pval, yiyp_IVW_beta_adj, yiyp_IVW_se_adj, yiyp_IVW_pval_adj)
+```{r YiYp_adj, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
+fig_dat <- full_join(proxyyMR_IV_overlap, proxyMR_MR_paths_summary_yiyp_adj)
+yiyp_adj.DT <- fig_dat %>% dplyr::select(exposure_description, outcome_description, exposure_sex, YiXi_IV_sig_overlap, yiyp_IVW_beta, yiyp_IVW_se, yiyp_IVW_pval, yiyp_IVW_beta_adj, yiyp_IVW_se_adj, yiyp_IVW_pval_adj)
 
@@
 
-fig_dat <- full_join(proxyyMR_IV_overlap, proxyMR_MR_paths_summary_yiyp_adj)
+```
+
+We can compare the $\alpha_{y_{i}\rightarrow y_{p}}$ before and after adjustment, and clearly see a band of estimates centered around $y = 0$, which matches nearly perfectly the instances where $Y$ IVs share more than 50% of $X$ IVs.
+There are several cases where the adjusted results caused the $abs(\alpha_{y_{i}\rightarrow y_{p}})$ to **increase**.
+
+```{r YiYp_adj_fig_comparison,  echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
+
+create_fig_meta <- function(fig_dat, custom_col){
+
+  #AM_result <- household_MR_binned_het %>% filter(same_trait)
+
+  plot_data <- fig_dat %>% #dplyr::select(exposure_ID, exposure_sex, IVW_beta, IVW_se, IVW_pval, sex_het_pval) %>% unique() %>%
+    #pivot_wider(names_from = "exposure_sex", values_from = c("IVW_beta", "IVW_se", "IVW_pval")) %>%
+    mutate(overlap_above_50 = case_when(YiXi_IV_sig_overlap > 0.5 ~ TRUE, TRUE ~ FALSE))
+
+  legend_title <- "IV overlap"
+  plot <- ggplot2::ggplot(data = plot_data, ggplot2::aes(x = yiyp_IVW_meta_beta,
+                                                         y = yiyp_IVW_meta_beta_adj, color = factor(overlap_above_50))) +
+    geom_point(alpha = 3/4) +
+
+    geom_smooth(mapping = aes(x = yiyp_IVW_meta_beta, y = yiyp_IVW_meta_beta_adj), method = "lm", se=FALSE, formula = y~x+0, fullrange=TRUE, color = custom_col[4]) +
+    geom_abline(slope=1, intercept=0, color = "black") +
+    scale_color_manual(values = custom_col[c(1,2)], labels=c("overlap =< 50%", "overlap > 50%")) +
+    theme_minimal() +
+    ggplot2::labs(colour = legend_title, x = paste0("MR estimate\n(Y-index to Y-partner)"),
+                  y = paste("AM MR estimate adjusted for\nIV effect on relevant X")) + ggtitle("Comparison of adjusted and unadjusted AM estimated\n(meta-analyzed across sexes)") +
+    theme(plot.title = element_text(hjust = 0.5))
+
+  return(plot)
+}
 
@@
   legend_title <- "IV overlap"
-  plot <- ggplot2::ggplot(data = plot_data, ggplot2::aes(x = yiyp_IVW_beta_adj,
-                                                         y = yiyp_IVW_beta, color = factor(overlap_above_50))) +
+  plot <- ggplot2::ggplot(data = plot_data, ggplot2::aes(x = yiyp_IVW_beta,
+                                                         y = yiyp_IVW_beta_adj, color = factor(overlap_above_50))) +
     geom_point(alpha = 3/4) +
 
-    geom_smooth(mapping = aes(x = yiyp_IVW_beta_adj, y = yiyp_IVW_beta), method = "lm", se=FALSE, formula = y~x+0, fullrange=TRUE, color = custom_col[4]) +
+    geom_smooth(mapping = aes(x = yiyp_IVW_beta, y = yiyp_IVW_beta_adj), method = "lm", se=FALSE, formula = y~x+0, fullrange=TRUE, color = custom_col[4]) +
     geom_abline(slope=1, intercept=0, color = "black") +
-    scale_color_manual(values = custom_col[c(1,2)], labels=c("overlap > 50%","overlap =< 50%")) +
+    scale_color_manual(values = custom_col[c(1,2)], labels=c("overlap =< 50%", "overlap > 50%")) +
     theme_minimal() +
-    ggplot2::labs(colour = legend_title, x = paste0("MR estimate (Y-index to Y-partner)"),
-                  y = paste("AM MR estimate adjusted for IV effect on relevant X"))
+    ggplot2::labs(colour = legend_title, x = paste0("MR estimate\n(Y-index to Y-partner)"),
+                  y = paste("AM MR estimate adjusted for\nIV effect on relevant X"), title = "Comparison of adjusted and unadjusted AM estimated \n (sex-specific results)") + 
+    theme(plot.title = element_text(hjust = 0.5))
+
+  return(plot)
 }
 
-create_fig(fig_dat, custom_col)
+plot_meta <- create_fig_meta(fig_dat, custom_col)
+plot <- create_fig(fig_dat, custom_col)
+
+plot_meta
+plot
 
@@
 
-## Updated results with adjusted $Y_i$
+## Updated two trait results with adjusted $Y_i$.
+
+The results below show the significant difference between $\omega$ vs $\rho$, $\gamma$ and their sum.
+Concerned that most of the significant results involve cases where the adjusted $abs(\alpha_{y_{i}\rightarrow y_{p}})$ is **greater** than the unadjusted case.
 
@@
 
-t1 <- as.data.frame(table(proxyMR_comparison_summary_yiyp_adj$omega_vs_gam_BF_sig_meta))
+yi_adj_increase <- which(abs(proxyMR_MR_paths_summary_yiyp_adj$yiyp_IVW_meta_beta) < abs(proxyMR_MR_paths_summary_yiyp_adj$yiyp_IVW_meta_beta_adj))
+
+proxyMR_inv <- proxyMR_comparison_summary_yiyp_adj[yi_adj_increase, ]
+
+t1 <- as.data.frame(table(proxyMR_comparison_summary_yiyp_adj$omega_vs_gam_BF_sig_meta)/2)
+t1.2 <- as.data.frame(table(proxyMR_inv$omega_vs_gam_BF_sig_meta)/2)
+
 colnames(t1) <- c("omega_vs_gam_BF_sig_meta", "Frq")
@@
 
-t2 <- as.data.frame(table(proxyMR_comparison_summary_yiyp_adj$omega_vs_rho_BF_sig_meta))
+t2 <- as.data.frame(table(proxyMR_comparison_summary_yiyp_adj$omega_vs_rho_BF_sig_meta)/2)
+t2.2 <- as.data.frame(table(proxyMR_inv$omega_vs_rho_BF_sig_meta)/2)
 colnames(t2) <- c("omega_vs_rho_BF_sig_meta", "Frq")
@@
 
-t3 <- as.data.frame(table(proxyMR_comparison_summary_yiyp_adj$omega_vs_gam_rho_BF_sig_meta))
+t3 <- as.data.frame(table(proxyMR_comparison_summary_yiyp_adj$omega_vs_gam_rho_BF_sig_meta)/2)
+t3.2 <- as.data.frame(table(proxyMR_inv$omega_vs_gam_rho_BF_sig_meta)/2)
+
 colnames(t3) <- c("omega_vs_gam_rho_BF_sig_meta", "Frq")
+kbl(t3) %>% kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
 
@@
 
-## MVMR
+## MVMR.
 
-Run all MVMR models but need to sum up along all $\beta$ from the MV model to create the $\gamma_{mv}$.
+Have so far run / am running all MVMR models but need to sum up along all $\beta$ from the MV model to create the $\gamma_{mv}$. (Having some computing issues with pruning.)
 
-MV MR: $X \rightarrow  Z \rightarrow Y$
+Should we use the adjusted $Y_i \rightarrow Y_p$ for calculating $\rho$?
+
+MV MR: $X \rightarrow  Z \rightarrow Y$.
+
   - Drop Z's if not significant in MV MR (i.e. $Y \sim X + Z_1 + Z_2 + Z_3...$)
@@
 ```{r MV_model, echo = FALSE}
-library(knitr)
-knitr::include_graphics("assets/MV_trait_model_yiyp_adj.png", error = FALSE)
-``
+
+knitr::include_graphics("assets/MV_trait_model.png", error = FALSE)
+```

### 2021-09-27 R/functions.R
remove dietary traits, re-add some binary trait remove left-side body traits

@@ -917,6 +917,27 @@ continuous_filter <- function(traits){
   return(output)
 }
 
+non_diet_filter <- function(traits){ # filter out only binary diet traits?
+  output <- traits %>% filter(!category=="diet") %>% filter(!grepl("intake", description))
+  return(output)
+}
+
+non_qual_filter <- function(traits){
+  output <- traits %>% filter(!grepl("Qualifications:", description))
+  return(output)
+}
+
+non_left_side_filter <- function(traits){
+  output <- traits %>% filter(!grepl("left", description))
+  return(output)
+}
+
+non_redundant_filter <- function(traits){
+  temp <- non_qual_filter(traits)
+  output <- non_left_side_filter(temp)
+  return(output)
+}
+
 write_final_filter <- function(traits_final, output_name){
   output_name <- ("output/tables/household_correlations.final_filter.csv")
   write.csv(traits_final, "output/tables/household_correlations.final_filter.csv", row.names=F)

### 2021-09-27 _targets.R

@@ -11,7 +11,7 @@ options(clustermq.scheduler = "slurm", clustermq.template = "slurm_clustermq.tmp
 tar_option_set(
   resources = tar_resources(
     clustermq = tar_resources_clustermq(template = list(num_cores = 1, account = "sgg",
-                                                        ntasks = 4, partition = "sgg",
+                                                        ntasks = 1, partition = "sgg",
                                                         log_file="/data/sgg2/jenny/projects/proxyMR/proxymr_%a_clustermq.out"))
   ),
   packages = c("tidyverse", "data.table", "cutr", "ukbtools", "rbgen", "bigsnpr", "TwoSampleMR",
@@ -327,10 +327,23 @@ list(
     continuous_filter(traits_corr4$to_run)
   ),
 
+  # filter for non-diet traits only
+  tar_target(
+    traits_corr5,
+    non_diet_filter(traits_corr4$to_run)
+  ),
+
+  # filter out left-side traits (highly correlated with right-side)
+  # filter out categorical 'Qualifications:', because data is roughly captured with continuous variables
+  tar_target(
+    traits_corr6,
+    non_redundant_filter(traits_corr5)
+  ),
+
   # final list of traits to run in pipeline, just a copy of above target
   tar_target(
     traits_final,
-    traits_corr5
+    traits_corr6
   ),
 
   tar_target(

### 2021-09-27 _targets.R
remove the continuous filter

@@ -321,11 +321,6 @@ list(
     traits_corr4,
     sex_het_filter(traits_corr3$to_run, IV_data_summary, num_IVs_threshold)
   ),
-  # filter for continuous traits only
-  tar_target(
-    traits_corr5,
-    continuous_filter(traits_corr4$to_run)
-  ),
 
   # filter for non-diet traits only
   tar_target(

### 2021-09-27 analysis/phenotype_selection.Rmd
add phenotype and couple selection schematic links
edit phenotype selection based on removing dietary traits and left-side etc

@@
 tar_load(traits_corr5)
+tar_load(traits_corr6)
 
@@
 *Note that as none of the filtering criteria apply any BF correction for the number of phenotypes, the order of the steps does not matter.*
+
+An overview of phenotype selection and exclusions can be seen [here](assets/phenotype_selection.png).
 
@@
 
-# Non-binary.
+# Non-diet.
 
-Phenotypes were limited to `continuous_irnt` or `ordinal` (i.e. `binary` traits were removed).
+Dietary phenotypes were removed due to high correlation amongst phenotypes, insufficient power and difficult interpretation.
 
@@
 
-The traits remaining after sex-heterogeneity filtering can be seen below.
+The traits remaining after removing dietary traits can be seen below.
 
-```{r continuous_filter, echo=FALSE}
+```{r non_diet_filter, echo=FALSE}
 
@@
 ```
+
+# Non-redundant phenotypes.
+
+Lastly, left-side body traits (highly correlated with right-side) were removed and all 'qualifications' data (corresponding to field [6138](https://biobank.ctsu.ox.ac.uk/showcase/field.cgi?id=6138)) because data is roughly captured with continuous variables (such as years of education).
+
+Number remaining after filter = **`r dim(traits_corr6)[1]`**.
+
+The traits remaining after removing dietary traits can be seen below.
+
+```{r non_redundant_filter, echo=FALSE}
+
+traits_corr6.DT <- traits_corr6 %>% as_tibble() %>% mutate(r = sqrt(as.numeric(r2))) %>% dplyr::select(Neale_pheno_ID, description, variable_type, category, N_pairs, r2, r, num_IVs, num_IVs_pass_het)
+DT::datatable(traits_corr6.DT, width = "100%") %>% formatRound(columns=c('r2', 'r'), digits = 3)
+
+```

### 2021-09-27 _targets.R
try a few markdown docs in the targets pipeline

@@ -396,6 +396,11 @@ list(
   ),
 
   tar_target(
+    corr_mat_traits,
+    calc_corr_mat_traits(exposure_info, data_sqc, data_fam, data_relatives)
+  ),
+
+  tar_target(
     PC_traits,
     calc_PC_traits(exposure_info, data_sqc, data_fam, data_relatives)
   ),
@@ -730,9 +735,15 @@ list(
     path_bgenie_pcs_unzip,
     unzip_bgenie(path_bgenie_pcs),
     pattern = map(path_bgenie_pcs), format = "file"
-  )
+  ),
 
 
+
+  ### markdown docs
+
+  tar_render(rmd_index, "analysis/index.Rmd"),
+  tar_render(rmd_final_phenotype_list, "analysis/final_phenotype_list.Rmd"),
+  tar_render(rmd_phenotype_selection, "analysis/index.Rmd") # Here is our call to tar_render().
 
 
 

### 2021-10-05 analysis/setup.Rmd
copy PHESANT and UKBB file directories to project directory in case they change with time

@@
 options("workflowr.view" = FALSE) # if using cluster
-options("workflowr.view" = TRUE) # defaule, if using Rstudio on personal computer
+options("workflowr.view" = TRUE) # default, if using Rstudio on personal computer
 
@@
 - Because Terminal cannot generate a preview and `workflowr` doesn't like the sysgit, to the `.Rprofile` file, add:
-  - `options(workflowr.sysgit = "")`
-  - `options("workflowr.view" = FALSE)`
+
+```{r, engine = 'bash', eval = FALSE}
+library("targets")
+
+if(Sys.info()["sysname"]=="Linux"){
+  options("workflowr.view" = FALSE)
+}
+```
+
 - To ensure git hub isn't managing large files, modify the `.gitignore` file, by adding the following lines:
@@
 Because the `targets` package tracks input files and their changes, the current state of this file was copied to the project directory and this file was then used as an input file into the pipeline.
-This file is merely a reference file and changes to it don't correspond to changes in the underlying data used in the analyses.
+This file is merely a reference file and changes to it doesn't correspond to changes in the underlying data used in the analyses.
+
+Similarly, `PHESANT_file_directory.txt` and `UKBB_pheno_directory.csv` were copied once to the project directory in case more phenotypes were added to the SGG UKBB database and changed these files, but do not necessarily need to be added to the proxyMR project.
 
@@
 Neale_summary_dir <- "/data/sgg2/jenny/data/Neale_UKBB_GWAS"
+UKBB_processed_dir <- "/data/sgg2/jenny/data/UKBB_processed"
+
 Neale_SGG_dir_org <- paste0(Neale_summary_dir,"/Neale_SGG_directory.csv") 
+PHESANT_dir_org <- paste0(UKBB_processed_dir,"/PHESANT/","PHESANT_file_directory.txt")
+UKBB_dir_org <- paste0(UKBB_processed_dir,"/UKBB_pheno_directory.csv")
+
 Neale_SGG_dir_cp <- paste0("data/Neale_SGG_directory_", format(Sys.Date(), "%d_%m_%Y"), ".csv")
+PHESANT_dir_cp <- paste0("data/PHESANT_file_directory_", format(Sys.Date(), "%d_%m_%Y"), ".txt")
+UKBB_dir_cp <- paste0("data/UKBB_pheno_directory_", format(Sys.Date(), "%d_%m_%Y"), ".csv")
+
 file.copy(Neale_SGG_dir_org, Neale_SGG_dir_cp)
+file.copy(PHESANT_dir_org, PHESANT_dir_cp)
+file.copy(UKBB_dir_org, UKBB_dir_cp)
+
 ```

### 2021-10-05 code/settings.R

@@ -70,7 +70,11 @@ Neale_summary_dir <- "/data/sgg2/jenny/data/Neale_UKBB_GWAS"
 Neale_output_dir <- "/data/sgg3/data/neale_files"
 
 ## files
+
 Neale_SGG_dir_file_cp <- "data/Neale_SGG_directory_15_07_2021.csv"
+PHESANT_file_dir_cp <- paste0("data/PHESANT_file_directory_05_10_2021.txt")
+UKBB_pheno_dir_cp <- paste0("data/UKBB_pheno_directory_05_10_2021.csv")
+
 Neale_manifest_file <- "UKBB GWAS Imputed v3 - File Manifest Release 20180731 - Manifest 201807.tsv"
 Neale_variant_file <- paste0(Neale_summary_dir, "/variants.tsv")
 
@@ -85,12 +89,12 @@ files_custom_names <- c("household_info", "phesant_directory", "relatives", "fam
                  "time_at_address_raw", "UKBB_directory", "Neale_SGG_dir", "Neale_manifest", "code_process_Neale", "Neale_variants",
                  "UKBB_sample", "main_UKBB_raw")
 
-files <- c(paste0(UKBB_dir,"/pheno/ukb6881.csv"), paste0(UKBB_processed_dir,"/PHESANT/","PHESANT_file_directory.txt"),
+files <- c(paste0(UKBB_dir,"/pheno/ukb6881.csv"), PHESANT_file_dir_cp,
           paste0(UKBB_dir,"/geno/","ukb1638_rel_s488366.dat"),  paste0(UKBB_dir,"/plink/_001_ukb_cal_chr9_v2.fam"),
           paste0(UKBB_dir,"/geno/ukb_sqc_v2.txt"),
           paste0(UKBB_processed_dir, "/PHESANT/ukb31459/bin1/out_bin1..tsv"), paste0(UKBB_dir, "/pheno/ukb31459.csv"),
-          paste0(UKBB_processed_dir,"/UKBB_pheno_directory.csv"),
-          Neale_SGG_dir_file_cp, paste0(Neale_output_dir,"/",Neale_manifest_file), "code/process_Neale.sh",
+          UKBB_pheno_dir_cp,
+          Neale_SGG_dir_file_cp, paste0(Neale_output_dir,"/", Neale_manifest_file), "code/process_Neale.sh",
[…]

### 2021-10-05 R/functions.R
fix problem selecting download link from Neale

@@ -58,7 +58,7 @@ download_neale_files  <-  function(phenotype_ids,
 
 
   to_download  =  read_tsv(reference_file) %>%
-    filter(str_detect("Phenotype Code", phenotype_ids)) %>%
+    filter(str_detect(`Phenotype Code`, phenotype_ids)) %>%
     dplyr::select(wget = 'wget command') %>%
     mutate(wget = str_replace(wget, '[.]bgz$', '.gz')) %>%
     filter(str_detect(.$wget, paste0('-O .*[.]', sex, '([.]v2)?[.]tsv[.]gz$'))) %>%
@@ -390,6 +390,46 @@ compute_pc_corr <- function(sqc_munge, pairs_filter, data_id_sex){
 
 
 }
+
+compute_pc_trait_corr <- function(sqc_munge, pairs_filter, Neale_pheno_ID, pheno_data){
+
+  pheno_data_full <- rbind(pheno_data[[1]], pheno_data[[2]])
+  pheno_data_ivs <- pheno_data_full %>% mutate_at(vars(starts_with("PC_")), ivt)
+
+  PCs <- colnames(pheno_data_full)[which(startsWith(colnames(pheno_data_full), "PC_"))]
+  trait_corr <- numeric()
+  ID_sub <- NA
+  for (PC in PCs)
+  {
+    id <- as.character(PC)
+    r2 <- NA
+    pairs_filter_copy <- pairs_filter
+    tsv_data <- sqc_munge %>% dplyr::select(ID, !!PC)
+    pairs_filter_copy$trait1 <- tsv_data[[PC]][match(pairs_filter_copy[["HOUSEHOLD_MEMBER1"]], tsv_data$ID)]
+    pairs_filter_copy$trait2 <- tsv_data[[PC]][match(pairs_filter_copy[["HOUSEHOLD_MEMBER2"]], tsv_data$ID)]
+    complete_pairs <- pairs_filter_copy[which(complete.cases(pairs_filter_copy$trait1 ,pairs_filter_copy$trait2 )),]
+    n_completed_pairs <- length(pairs_filter_copy[which(complete.cases(pairs_filter_copy$trait2, pairs_filter_copy$trait1)),"trait2"])
+    if(n_completed_pairs>1)
+    {
+      sd1 <- sd(complete_pairs$trait1,na.rm=TRUE)
+      sd2 <- sd(complete_pairs$trait2,na.rm=TRUE)
+      if(sd1!=0 & sd2!=0)
+      {
[…]

### 2021-10-05 R/functions.R
remove `ukb_gen_samples_to_remove` function use because PHESANT files are already filtered for non-relateds.

@@ -1148,19 +1148,9 @@ prep_pheno_data <- function(traits, Neale_pheno_ID, sqc, fam, relatives){
   pheno_male_full <- merge(pheno_male_comp, sqc_sub, by.x="IID", by.y="ID")
   pheno_female_full <- merge(pheno_female_comp, sqc_sub, by.x="IID", by.y="ID")
 
-  maleIDs <- as.integer(pheno_male_full$IID)
-  femaleIDs <- as.integer(pheno_female_full$IID)
-
-
-  related_males <- ukb_gen_samples_to_remove(relatives, ukb_with_data = maleIDs)
-  related_females <- ukb_gen_samples_to_remove(relatives, ukb_with_data = femaleIDs)
-
-  unrelated_male_data <- pheno_male_full[-which(pheno_male_full$IID %in% related_males ),]
-  unrelated_female_data <- pheno_female_full[-which(pheno_female_full$IID %in% related_females ),]
-
   cat(paste0("Successfully prepared phenotype for processing (created directories, phenotype file inputs, trait summary, etc).\n"))
 
-  pheno_data = list(unrelated_male_data = unrelated_male_data, unrelated_female_data = unrelated_female_data)
+  pheno_data = list(unrelated_male_data = pheno_male_full, unrelated_female_data = pheno_female_full)
   return(pheno_data)
 }
 

### 2021-10-18 R/functions.R
calc impact by PCs

@@ -1245,8 +1245,39 @@ calc_pc_trait_corr <- function(Neale_pheno_ID, pheno_data){
     result <- rbind(result, output_row)
   }
 
-  colnames(result) <- c("Neale_pheno_ID", "PC", "cor_both_sexes", "cor_male", "cor_female")
+  colnames(result) <- c("Neale_pheno_ID", "PC", "corr_both_sexes", "corr_male", "corr_female")
+
+  result <- result %>% as_tibble() %>% type_convert() %>% mutate_at(c("Neale_pheno_ID"), as.character)
   return(result)
+}
+
+calc_corr_impact_by_PCs <- function(traits_corr, PCs_corr, PC_trait_corr){
+
+  result <- numeric()
+
+  Neale_pheno_ID <- PC_trait_corr$Neale_pheno_ID[1]
+
+  trait_couple_r2 <- as.numeric(traits_corr[which(traits_corr$ID==Neale_pheno_ID),"r2"])
+  trait_couple_r <- sqrt(trait_couples_r2)
+
+  PCs <- PCs_corr$description
+
+  for(PC in PCs){
+
+    trait_PC_r <- PC_trait_corr[which(PC_trait_corr$PC==PC),"corr_both_sexes"][[1]]
+    PC_couple_r2 <- as.numeric(PCs_corr[which(PCs_corr$description==PC),"r2"])
+    PC_couple_r <- sqrt(PC_couple_r2)
+
+    correlation_due_to_confounding <- trait_PC_r^2 *PC_couple_r
+
+    temp_row <- cbind(Neale_pheno_ID, PC, trait_couple_r, PC_couple_r, trait_PC_r, correlation_due_to_confounding)
+    result <- rbind(result, temp_row)
+  }
+
+  colnames(result) <- c("Neale_pheno_ID", "PC", "trait_couple_corr", "PC_couple_corr", "trait_PC_corr", "corr_due_to_confounding")
+
+  result <- result %>% as_tibble() %>% type_convert() %>% mutate_at(c("Neale_pheno_ID"), as.character)
+
 }
 
 write_data_prep <- function(traits, traits_to_run, out1, out2){

### 2021-10-18 _targets.R

@@ -756,6 +756,11 @@ list(
     pattern = map(path_bgenie_pcs), format = "file"
   ),
 
+  tar_target(
+    corr_impact_by_PCs,
+    calc_corr_impact_by_PCs(traits_corr, PCs_corr, PC_trait_corr),
+    pattern = map(PC_trait_corr)
+  ),
 
 
   ### markdown docs

### 2021-10-18 R/functions.R
correct for phes_ID vs Neale_pheno_ID

@@ -1256,8 +1256,9 @@ calc_corr_impact_by_PCs <- function(traits_corr, PCs_corr, PC_trait_corr){
   result <- numeric()
 
   Neale_pheno_ID <- PC_trait_corr$Neale_pheno_ID[1]
+  phes_ID <- gsub("_irnt", "", Neale_pheno_ID)
 
-  trait_couple_r2 <- as.numeric(traits_corr[which(traits_corr$ID==Neale_pheno_ID),"r2"])
+  trait_couple_r2 <- as.numeric(traits_corr[which(traits_corr$ID==phes_ID),"r2"])
   trait_couple_r <- sqrt(trait_couples_r2)
 
   PCs <- PCs_corr$description

### 2021-10-18 analysis/update_meeting_15_09_2021.Rmd
add updated meeting 15-09-21

---
title: "Update Meeting: 15-10-2021"
author: "jennysjaarda"
date: "2021-10-12"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
library(targets)
library(dplyr)
library(DT)
library(knitr)
library(ggplot2)
library(kableExtra)

```

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
tar_load(household_MR_summary)
tar_load(PCs_corr)
tar_load(PC_trait_corr)
tar_load(corr_impact_by_PCs)
```

# Sample definition issue.

## Update.

Noticed that the sample sizes we were working with were quite a bit larger than the Neale database. After a bit of digging, I think I figured out the main reason(s): it seems to be a result of the way we filtered related individuals.

- First of all, because we’ve done all analyses in each sex separately, I also have performed the related filters in each sex separately. Meaning there is surely related individuals between sexes that remain (of course each individual couple has been filtered separately for their own relatedness). 
- Second, I am using a function to retain the maximal set of unrelated individuals which is relatively computation heavy and I believe that Neale would have simply filtered based on `used_in_pca_calculation` from the QC file, which results in <<N.

I assume Neale also applied related filtering *before* PHESANT transformations, while I applied it after (I was thinking that if someone was performing an analysis on a subset of UKB, they would have unnecessarily lost individuals even though the two related individuals weren’t in their subgroup). The problem is, these differences result in slightly different phenotypes in some cases I’ve realized.

For example, phenotype 1050 (amount of time spent outside, hours) is converted by the PHESANT pipeline to an ordinal variable with only 3 categories because the upper categories are so small. Where to split the categories directly depends on the data, with the goal to have roughly equal sized groups. Anyways, in Neale it’s split in different place than in our data.

### Discussion points.

- Should we apply the same filter as Neale to get roughly the same sample size and data (particularly for these categorical phenotypes)? 
    - Regarding the category definitions: best to have something as similar as possible to Neale's, because the instrument selection is based on his summary stats, which used his definition. But if it is hard to replicate his definition, it's not a big problem, it is much more important that we move forward fast

- And/or is it a problem that there we would have related individuals across sexes? 
    - For the relateds: unfortunately, they have to be excluded (male-female relatives too, not only same-sex relatives), because when we meta-analyse male->female and female->male results, we assume inependence in the estimates, which is not true if you don't exclude all relatives.

## Update 2.0.

I was able to nearly mimic the sample size of the round 1 GWAS, but not the round 2 GWAS. In the blog post describing the [first release](http://www.nealelab.is/blog/2017/9/11/details-and-considerations-of-the-uk-biobank-gwas), it is mentioned that N = 337,199 individuals are used in the GWAS. I can nearly replicate this number exactly with my data by removing un consenting, non-white and only retaining those that were `used_in_pca_calculation` (from the genetic QC file). However in the [round 2 blog post](http://www.nealelab.is/blog/2019/9/16/biomarkers-gwas-results), the sample size increases to N = 361,194 individuals.

The details compairing the two samples are found on the [github page](https://github.com/Nealelab/UK_Biobank_GWAS/blob/master/README.md), specifically:

  * __imputed-v3 parameters__
	* Used.in.pca.calculation filter (unrelated samples)
	* sex chromosome aneuploidy filter
	* Use provided PCs for European sample selection to determine British ancestry
	  * Use 7 standard deviations away from the 1st 6 PCs
	  * Further Filter to self-reported 'white-British' / 'Irish' / 'White'
	* **QCed sample count: 361194 samples** 
  * __imputed-v2 parameters__
    * Used.in.pca.calculation filter (unrelated samples)
    * sex chromosome aneuploidy filter
    * White.british.ancestry filter
    * **QCed sample count: 337199 samples**

### Discussion points.

Our current filtering matches closely the v2 filtering. If we tried to mimic the v3 filtering, we could increase our sample size. What do we think about this discrepency? And more importantly, the decrease in power due to smaller sample size (about 25K less than the v3 parameters)?

# Progress update.

## Phenotype selection.

- Added back binary phenotypes. 
- Removed all dietary phenotypes (binary, ordinal and continuous).
- Removed some redundant phenotypes:
  - Left-side body traits (highly correlated with right-side) were removed
  - All 'qualifications' data (corresponding to field [6138](https://biobank.ctsu.ox.ac.uk/showcase/field.cgi?id=6138)) because data is roughly captured with continuous variables (such as years of education).
- See [phenotype selection](phenotpye_selection.html) document for complete summary of selection and remaining phenotypes.

## Two-trait MR filtering.

Restrict to only XY pairs that are have correlation < `0.8`. Correlations were calculated in the full biobank (unrelated, British, consenting etc.), not just within couples, in both sexes combined. Trait pairs and their correlations are summarized in the table below.

```{r corr_traits_summary, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

household_MR_summary.DT <- household_MR_summary %>% filter(same_trait==FALSE) %>% dplyr::select(exposure_ID, outcome_ID, exposure_description, outcome_description, corr_traits) %>% unique()

format_round_cols <- colnames(dplyr::select_if(household_MR_summary.DT, is.numeric))
DT::datatable(household_MR_summary.DT, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)

```

In total, `#r length(which(household_MR_summary.DT$corr_traits > 0.8))` trait pairs were removed due to correlation > `0.8`.

**Note**: after filtering out pairs with high correlation, a bonferroni correction applied (p < 0.05 / [num remaining pairs after correaltion filter]) before examining paths from $X_i \rightarrow Y_p$ (i.e we only examined pairs that were significant along the diagonal).

```{r two_trait_model, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/two_trait_model.png", error = FALSE)
```

## Adjust $Y_i \rightarrow Y_p$ for effect of $X_i \rightarrow Y_p$ in MVMR.

Previously, I had adjusted the $Y_i \rightarrow Y_p$ MR results for effects of $X_i$ (as depicted below), but had only used $Y$ IVs.

```{r YiYp_adj_model, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/two_trait_model_yiyp_adj.png", error = FALSE)
```

Decided best to also include all $X$ IVs in the model as well. MRMR model is: $Y_p \sim Y_i + X_i$.

To ensure a set of independent SNPs, $X$ and $Y$ IVs were clumped using at a threshold of $r^2$ < `r prune_threshold` using the 1000 Genomes European data set. Clumping p-values for priorization of SNPs, were based on the Neale summary statistics with $Y_i$ in both sexes, ensuring the same set of SNPs were selected in the male and female models.

## MVMR

For each $X_i \rightarrow  Y_p$ MR, we included relevant covariates (Z's) in a MVMR framework in the $X_p \rightarrow Y_p$ MR. Z's were selected based on having BF significance in **both** of the MRs:

1. $X \rightarrow  Z$
2. $Z \rightarrow  Y$

After Z's were selected, they were further pruned based on phenotypic correlation (same as above). First, any Z with phenotypic correlation with $X$ or $Y$ > `r z_prune_threshold` was removed. Second, beginning with the most significant $Z_j$ association with $Y$ (i.e. $Z \rightarrow  Y$), all reamining Z's with phenoytpic correlation > `r z_prune_threshold` were removed. This process was repeated until all remaining Z's had phenotypic correlation < `r z_prune_threshold` with each other and $X$ and $Y$.

The resulting MVMR model is: $Y \sim X + Z_1 + ... + Z_k$ run in same individual (see figure).

```{r MV_model, echo = FALSE}

knitr::include_graphics("assets/MV_trait_model.png", error = FALSE)
```

Finally, IVs for all $Z$ and $X$ were clumped as above (using at a threshold of $r^2$ < `r prune_threshold` using the 1000 Genomes European data set. Clumping p-values for priorization of SNPs, were based on the Neale summary statistics with $X_i$ in both sexes, ensuring the same set of SNPs were selected in the male and female models).

## PC impact

Tested the following:
- cor(PC_i,PC_p) for each PC (i.e. within couple PC correlation).
- cor(X,PC) for all PCs.

PC correlations among couples is shown below.

```{r PC_corr_couples, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

PCs_corr.DT <- PCs_corr %>% mutate(r = sqrt(as.numeric(r2))) 
DT::datatable(PCs_corr.DT, width = "100%")  %>% formatSignif(columns=c('r2', 'r'), digits=3)

```

Correlation of PCs with traits in pipeline is shown below.

```{r PC_traits_corr, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

## merge the list of dataframes 
PCs_trait_corr.DT <- PC_trait_corr %>% mutate(r = sqrt(as.numeric(r2))) 
DT::datatable(PCs_corr.DT, width = "100%")  %>% formatSignif(columns=c('r2', 'r'), digits=3)

```

Next, I calculated the correlation due to counfounding as `cor(X,PC)^2*cor(PC_i,PC_p)` values, and plotted them against the raw `cor(X_i,X_p)` values.

```{r PC_trait_plot, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

plot <- ggplot2::ggplot(data = corr_impact_by_PCs, ggplot2::aes(x = trait_couple_corr,
                                                                y = corr_due_to_confounding, color = factor(overlap_above_50))) +
  geom_point(alpha = 3/4) +

  geom_smooth(mapping = aes(x = trait_couple_corr, y = corr_due_to_confounding), method = "lm", se=FALSE, formula = y~x+0, fullrange=TRUE, color = custom_col[4]) +
  geom_abline(slope=1, intercept=0, color = "black") +
  scale_color_manual(values = custom_col[c(1,2)], labels=c("overlap =< 50%", "overlap > 50%")) +
  theme_minimal() +
  ggplot2::labs(colour = legend_title, x = paste0("MR estimate\n(Y-index to Y-partner)"),
                y = paste("AM MR estimate adjusted for\nIV effect on relevant X"), title = "Comparison of adjusted and unadjusted AM estimated \n (sex-specific results)") + 
  theme(plot.title = element_text(hjust = 0.5))

```

### 2021-10-19 R/functions.R
calc `corr_impact_by_coords`

@@ -1281,6 +1281,42 @@ calc_corr_impact_by_PCs <- function(traits_corr, PCs_corr, PC_trait_corr){
 
 }
 
+calc_corr_impact_by_coords <- function(outcomes_to_run, traits_corr, corr_mat_traits){
+
+  result <- numeric()
+
+  for(i in 1:dim(outcomes_to_run)[1]){
+
+    Neale_pheno_ID <- outcomes_to_run$Neale_pheno_ID[i]
+    phes_ID <- gsub("_irnt", "", Neale_pheno_ID)
+    trait_couple_r2 <- as.numeric(traits_corr[which(traits_corr$ID==phes_ID),"r2"])
+    trait_couple_r <- sqrt(trait_couple_r2)
+
+    trait_NC_r <- corr_mat_traits[phes_ID, "129"]
+    trait_EC_r <- corr_mat_traits[phes_ID, "130"]
+
+    NC_couple_r2 <- as.numeric(traits_corr[which(traits_corr$ID=="129"),"r2"])
+    EC_couple_r2 <- as.numeric(traits_corr[which(traits_corr$ID=="130"),"r2"])
+
+    NC_couple_r <- sqrt(NC_couple_r2)
+    EC_couple_r <- sqrt(EC_couple_r2)
+
+    correlation_due_to_confounding_NC <- trait_NC_r^2 *NC_couple_r
+    correlation_due_to_confounding_EC <- trait_EC_r^2 *EC_couple_r
+
+    temp_row_NC <- cbind(Neale_pheno_ID, "129", trait_couple_r, NC_couple_r, trait_NC_r, correlation_due_to_confounding_NC)
+    temp_row_EC <- cbind(Neale_pheno_ID, "130", trait_couple_r, EC_couple_r, trait_EC_r, correlation_due_to_confounding_EC)
+
+    result_temp <- rbind(temp_row_NC, temp_row_EC)
+    result <- rbind(result, result_temp)
+  }
+
+  colnames(result) <- c("Neale_pheno_ID", "coordinate", "trait_couple_corr", "coordinate_couple_corr", "trait_coordiante_corr", "corr_due_to_confounding")
+  result <- result %>% as_tibble() %>% type_convert() %>% mutate_at(c("Neale_pheno_ID"), as.character)
+
+
+}
+
 write_data_prep <- function(traits, traits_to_run, out1, out2){
 
   for(i in 1:dim(traits_to_run)[1]){

### 2021-10-19 _targets.R

@@ -763,6 +763,13 @@ list(
   ),
 
 
+  tar_target(
+    corr_impact_by_coords,
+    calc_corr_impact_by_coords(outcomes_to_run, traits_corr, corr_mat_traits),
+    pattern = map(PC_trait_corr)
+  ),
+
+
   ### markdown docs
 
   tar_render(rmd_index, "analysis/index.Rmd"),

### 2021-10-19 _targets.R
don't map over anything in `corr_impact_by_coords`

@@ -765,8 +765,7 @@ list(
 
   tar_target(
     corr_impact_by_coords,
-    calc_corr_impact_by_coords(outcomes_to_run, traits_corr, corr_mat_traits),
-    pattern = map(PC_trait_corr)
+    calc_corr_impact_by_coords(outcomes_to_run, traits_corr, corr_mat_traits)
   ),
 
 

### 2021-10-19 _targets.R
select correct `z_summ_stats`

@@ -668,7 +668,7 @@ list(
     map(MV_z)
   ),
 
-  # filter to only z's that have correlation < 0.7 (play with this a bit).
+  # filter to only z's that have correlation < 0.8 (play with this a bit).
 
   tar_target(
     z_summ_stats,

### 2021-10-19 analysis/update_meeting_15_10_2021.Rmd
fix date in update meeting
add results to 15-10-21 meeting

--- 2021-10-18 analysis/update_meeting_15_09_2021.Rmd
@@
 tar_load(corr_impact_by_PCs)
+tar_load(corr_impact_by_coords)
+tar_load(proxyMR_MR_paths_summary_yiyp_adj)
+tar_load(MV_z_corr_filter)
+tar_load(traits_corr2_filled)
 ```
@@
 If we tried to mimic the v3 filtering, we could increase our sample size.
-What do we think about this discrepency?
+What do we think about this discrepancy?
 And more importantly, the decrease in power due to smaller sample size (about 25K less than the v3 parameters)?
+*This is just to keep in mind, if we want to increase our sample in the future if some results are nearly significant - could try adjusting our inclusion criteria.*
 
@@
   - All 'qualifications' data (corresponding to field [6138](https://biobank.ctsu.ox.ac.uk/showcase/field.cgi?id=6138)) because data is roughly captured with continuous variables (such as years of education).
-- See [phenotype selection](phenotpye_selection.html) document for complete summary of selection and remaining phenotypes.
+- See [phenotype selection](phenotype_selection.html) document for complete summary of selection and remaining phenotypes.
 
@@
 
-Restrict to only XY pairs that are have correlation < `0.8`.
+For the two trait MR, restrict to only XY pairs that are have correlation < `0.8`.
 Correlations were calculated in the full biobank (unrelated, British, consenting etc.), not just within couples, in both sexes combined.
@@
 
-```{r corr_traits_summary, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
+```{r corr_traits_summary, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
 
@@
 
-In total, `#r length(which(household_MR_summary.DT$corr_traits > 0.8))` trait pairs were removed due to correlation > `0.8`.
+In total, `r length(which(household_MR_summary.DT$corr_traits > 0.8))` trait pairs were removed due to correlation > `0.8` from `r dim(household_MR_summary.DT)[1]` pairs.
 
-**Note**: after filtering out pairs with high correlation, a bonferroni correction applied (p < 0.05 / [num remaining pairs after correaltion filter]) before examining paths from $X_i \rightarrow Y_p$ (i.e we only examined pairs that were significant along the diagonal).
+**Note**: after filtering out pairs with high correlation, a Bonferroni correction applied (p < 0.05 / [num remaining pairs after correlation filter]) before examining paths from $X_i \rightarrow Y_p$ (i.e we only examined pairs that were significant along the diagonal).
 
@@
 
-## MVMR
+The results before and after adjustment are shown below:
 
-For each $X_i \rightarrow  Y_p$ MR, we included relevant covariates (Z's) in a MVMR framework in the $X_p \rightarrow Y_p$ MR. Z's were selected based on having BF significance in **both** of the MRs:
+```{r YiYp_adj, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
+yiyp_adj.DT <- proxyMR_MR_paths_summary_yiyp_adj %>% dplyr::select(exposure_description, outcome_description, exposure_sex, yiyp_IVW_beta, yiyp_IVW_se, yiyp_IVW_pval, yiyp_IVW_beta_adj, yiyp_IVW_se_adj, yiyp_IVW_pval_adj)
+
+format_round_cols <- colnames(dplyr::select_if(yiyp_adj.DT, is.numeric))
+DT::datatable(yiyp_adj.DT, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)
+
+```
+
+## MVMR.
+
+For each $X_i \rightarrow  Y_p$ MR, we included relevant co-variates (Z's) in a MVMR framework in the $X_p \rightarrow Y_p$ MR.
+Z's were selected based on having BF significance in **both** of the MRs:
 
@@
 
-## PC impact
+```{r MV_z, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
+
+ z_list <- MV_z_corr_filter$MV_z_prune
+average_num_z <- mean((sapply(z_list, length)))
+
+```
+
+On average, the number of Z's to include was: **`r average_num_z`**, which seems quite high.
+
+See an example of Z's included for one trait pair below (first trait is exposure, second is outcome).
+
+```{r MV_z_example, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
+
+i <- 1
+exposure_description <- MV_z_corr_filter$exposure_description[i]
+outcome_description <- MV_z_corr_filter$outcome_description[i]
+
+exposure_description
+outcome_description
+
+z_list_ex <- tibble("z" = MV_z_corr_filter$MV_z_prune[i][[1]])
+
+z_list_ex.DT <- replace_Neale_ID(z_list_ex, traits_corr2_filled, "z", "outcome_description")
+
+DT::datatable(z_list_ex.DT, width = "100%")
+```
+
+## Impact of geography.
+
+Wanted to assess impact of geography on couple trait correaltion. Investigated this by using both genetic PCs and birth coordinates.
+
+### PC impact.
 
@@
 
-```{r PC_corr_couples, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
+```{r PC_corr_couples, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
 
 PCs_corr.DT <- PCs_corr %>% mutate(r = sqrt(as.numeric(r2))) 
-DT::datatable(PCs_corr.DT, width = "100%")  %>% formatSignif(columns=c('r2', 'r'), digits=3)
+format_round_cols <- colnames(dplyr::select_if(PCs_corr.DT, is.numeric))
+DT::datatable(PCs_corr.DT, width = "100%")  %>% formatRound(columns=format_round_cols, digits = 3)
 
@@
 
-```{r PC_traits_corr, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
+```{r PC_traits_corr, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
 
 ## merge the list of dataframes 
-PCs_trait_corr.DT <- PC_trait_corr %>% mutate(r = sqrt(as.numeric(r2))) 
-DT::datatable(PCs_corr.DT, width = "100%")  %>% formatSignif(columns=c('r2', 'r'), digits=3)
+PCs_trait_corr.DT <- PC_trait_corr %>% bind_rows()
+format_round_cols <- colnames(dplyr::select_if(PCs_trait_corr.DT, is.numeric))
+DT::datatable(PCs_trait_corr.DT, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)
 
@@
 
-```{r PC_trait_plot, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
+```{r PC_trait_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
 
 plot <- ggplot2::ggplot(data = corr_impact_by_PCs, ggplot2::aes(x = trait_couple_corr,
-                                                                y = corr_due_to_confounding, color = factor(overlap_above_50))) +
+                                                                y = corr_due_to_confounding)) +
   geom_point(alpha = 3/4) +
@@
   geom_smooth(mapping = aes(x = trait_couple_corr, y = corr_due_to_confounding), method = "lm", se=FALSE, formula = y~x+0, fullrange=TRUE, color = custom_col[4]) +
-  geom_abline(slope=1, intercept=0, color = "black") +
-  scale_color_manual(values = custom_col[c(1,2)], labels=c("overlap =< 50%", "overlap > 50%")) +
   theme_minimal() +
-  ggplot2::labs(colour = legend_title, x = paste0("MR estimate\n(Y-index to Y-partner)"),
-                y = paste("AM MR estimate adjusted for\nIV effect on relevant X"), title = "Comparison of adjusted and unadjusted AM estimated \n (sex-specific results)") + 
-  theme(plot.title = element_text(hjust = 0.5))
+  ggplot2::labs(x = paste0("Couple trait correlation"),
+                y = paste("Couple correlation due to confounding"))
+
+plot
 
 ```
+
+```{r PC_traits_confounding, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
+
+## merge the list of dataframes 
+corr_impact_by_PCs.DT <- replace_Neale_ID(corr_impact_by_PCs, traits_corr2_filled, "Neale_pheno_ID", "outcome_description")
+format_round_cols <- colnames(dplyr::select_if(corr_impact_by_PCs.DT, is.numeric))
+DT::datatable(corr_impact_by_PCs.DT, width = "100%")  %>% formatRound(columns=format_round_cols, digits = 3)
+
+```
+
+### Coordiante impact.
+
+Performed the same analysis as above but replacing PCs with North and East birth co-ordinates (data field [129](https://biobank.ctsu.ox.ac.uk/crystal/field.cgi?id=129) and [130](https://biobank.ctsu.ox.ac.uk/crystal/field.cgi?id=130), respectively).
+
+```{r coord_trait_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
+
+legend_title <- "Co-ordinate" 
+plot <- ggplot2::ggplot(data = corr_impact_by_coords, ggplot2::aes(x = trait_couple_corr,
+                                                                y = corr_due_to_confounding, color = factor(coordinate))) +
+  geom_point(alpha = 3/4) +
+
+  geom_smooth(mapping = aes(x = trait_couple_corr, y = corr_due_to_confounding), method = "lm", se=FALSE, formula = y~x+0, fullrange=TRUE, color = custom_col[4]) +
+  scale_color_manual(values = custom_col[c(1,2)], labels=c("North co-ordinate", "East co-ordinate")) +
+  theme_minimal() +
+  ggplot2::labs(x = paste0("Couple trait correlation"),
+                y = paste("Couple correlation due to confounding"))
+
+plot
+
+```
+
+```{r coordin_traits_confounding, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
+
+## merge the list of dataframes 
+corr_impact_by_coords.DT <- replace_Neale_ID(corr_impact_by_coords, traits_corr2_filled, "Neale_pheno_ID", "outcome_description")
+format_round_cols <- colnames(dplyr::select_if(corr_impact_by_coords.DT, is.numeric))
+DT::datatable(corr_impact_by_coords.DT, width = "100%")  %>% formatRound(columns=format_round_cols, digits = 3)
+
+```

### 2021-10-19 analysis/update_meeting_15_10_2021.Rmd
add table description of coordinate impact

@@
 
+The table below gives the data in the plot above:
+
+- `outcome_description` corresponds to the trait. 
+- `PC` is the genetic PC. 
+- `trait_couple_corr` is the correlation of `outcome_description` in couples. 
+- `PC_couple_corr` is the correlation of `PC` in couples. 
+- `trait_PC_corr` is the correlation of `outcome_description` with `PC`.
+-  `corr_due_to_confounding` corresponds to the formula: `cor(X,PC)^2*cor(PC_i,PC_p)`.
+
 ```{r PC_traits_confounding, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
@@
 
+The table below gives the data in the plot above:
+
+- `outcome_description` corresponds to the trait. 
+- `coordinate` is the coordiante (North: 129 or East: 130). 
+- `trait_couple_corr` is the correlation of `outcome_description` in couples. 
+- `coordinate_couple_corr` is the correlation of `coordinate` in couples. 
+- `trait_coordinate_corr` is the correlation of `outcome_description` with `coordinate`.
+-  `corr_due_to_confounding` corresponds to the formula: `cor(X,coord)^2*cor(coord_i,coord_p)`.
+
 ```{r coordin_traits_confounding, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

### 2021-10-20 _targets.R
create figures for path summary

@@ -706,6 +706,16 @@ list(
   ),
 
   tar_target(
+    proxyMR_prod_comparison_fig_yiyp_adj,
+    create_proxy_prod_comparison_fig(proxyMR_comparison_summary_yiyp_adj)
+  ),
+
+  tar_target(
+    proxyMR_sex_comparison_fig_yiyp_adj,
+    create_proxy_sex_comparison_fig(proxyMR_comparison_summary_yiyp_adj)
+  ),
+
+  tar_target(
     PC_gwas_input,
     prep_PC_GWAS(data_id_age, data_id_sex, sqc_munge, data_UKBB_sample)
   ),

### 2021-10-20 R/functions.R
select proper columns in `create_proxy_sex_comparison_fig_ind`

@@ -3473,8 +3473,8 @@ create_proxy_sex_comparison_fig_ind <- function(data, var, count){
   num_result <- dim(data)[1]/2
 
 
-  vars_to_select <- paste0(var, c("_beta", "_sex_het_p"))
-  sex_het_p_var <- paste0(var, "_sex_het_p")
+  vars_to_select <- paste0(var, c("_beta", "_sex_het_pval"))
+  sex_het_p_var <- paste0(var, "_sex_het_pval")
   beta_var <- paste0(var, "_beta")
 
   fig_data <- data %>% dplyr::select("exposure_ID", "outcome_ID", "exposure_sex", vars_to_select) %>%

### 2021-10-20 analysis/update_meeting_28_10_2021.Rmd
add update meeting for 28-10-21

---
title: "update_meeting_28_10_2021"
author: "jennysjaarda"
date: "2021-10-20"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
source("R/functions.R")
library(targets)
library(dplyr)
library(DT)
library(knitr)
library(ggplot2)
library(kableExtra)

```

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
tar_load(household_MR_summary)
tar_load(PCs_corr)
tar_load(PC_trait_corr)
tar_load(corr_impact_by_PCs)
tar_load(corr_impact_by_coords)
tar_load(proxyMR_MR_paths_summary_yiyp_adj)
tar_load(MV_z_corr_filter)
tar_load(traits_corr2_filled)
tar_load(proxyMR_prod_comparison_fig_yiyp_adj)
tar_load(proxyMR_prod_comparison_figure)

```

# Progress update.

## Compairing paths from $X_i \rightarrow Y_p$.

Without adjustment.

``````{r proxyMR_figure1, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600}
proxyMR_prod_comparison_figure
```

With adjustment.

``````{r proxyMR_figure2, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600}
proxyMR_prod_comparison_fig_yiyp_adj
```

### 2021-10-26 analysis/update_meeting_25_10_2021.Rmd
wflow_rename("analysis/update_meeting_28_10_2021.Rmd", "analysis/update_meeting_25_10_2021.Rmd")

---
title: "Update Meeting: 25-10-2021"
author: "jennysjaarda"
date: "2021-10-20"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
source("R/functions.R")
library(targets)
library(dplyr)
library(DT)
library(knitr)
library(ggplot2)
library(kableExtra)
library(cowplot)

```

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
tar_load(household_MR_summary)
tar_load(PCs_corr)
tar_load(PC_trait_corr)
tar_load(corr_impact_by_PCs)
tar_load(corr_impact_by_coords)
tar_load(proxyMR_MR_paths_summary_yiyp_adj)
tar_load(proxyMR_comparison_summary_yiyp_adj)
tar_load(MV_z_corr_filter)
tar_load(traits_corr2_filled)
tar_load(proxyMR_prod_comparison_fig_yiyp_adj)
tar_load(proxyMR_prod_comparison_fig)

```

# Progress update.

## Impact of geography.

Wanted to assess impact of geography on couple trait correlation. Investigated this by using both genetic PCs and birth coordinates.

### PC impact.

Tested the following:

- cor(PC_i,PC_p) for each PC (i.e. within couple PC correlation).
- cor(X,PC) for all PCs.

Next, I calculated the correlation due to counfounding as `cor(X,PC)^2*cor(PC_i,PC_p)` values, and plotted them against the raw `cor(X_i,X_p)` values.

```{r PC_trait_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

plot_data <- corr_impact_by_PCs %>% dplyr::select(Neale_pheno_ID, trait_couple_corr, corr_due_to_confounding_all) %>% unique()
plot <- ggplot2::ggplot(data = plot_data, ggplot2::aes(x = trait_couple_corr,
                                                       y = corr_due_to_confounding_all)) +
  geom_point(alpha = 3/4) +

  geom_smooth(mapping = aes(x = trait_couple_corr, y = corr_due_to_confounding_all), method = "lm", se=FALSE, formula = y~x+0, fullrange=TRUE, color = custom_col[4]) +
  theme_minimal() +
  ggplot2::labs(x = paste0("Couple trait correlation"),
                y = paste("Couple correlation due to confounding"))

plot

```

The table below gives the data in the plot above:

- `outcome_description` corresponds to the trait. 
- `trait_couple_corr` is the correlation of `outcome_description` in couples. 
-  `corr_due_to_confounding_all` corresponds to the formula: `cor(X,PC)^2*cor(PC_i,PC_p)`, summed across all PCs.

```{r PC_traits_confounding, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

corr_impact_by_PCs.DT <- replace_Neale_ID(plot_data, traits_corr2_filled, "Neale_pheno_ID", "outcome_description") %>% ungroup()
format_round_cols <- colnames(dplyr::select_if(corr_impact_by_PCs.DT, is.numeric))
DT::datatable(corr_impact_by_PCs.DT, width = "100%")  %>% formatRound(columns=format_round_cols, digits = 3)

```

### Coordiante impact.

Performed the same analysis as above but replacing PCs with North and East birth co-ordinates (data field [129](https://biobank.ctsu.ox.ac.uk/crystal/field.cgi?id=129) and [130](https://biobank.ctsu.ox.ac.uk/crystal/field.cgi?id=130), respectively).

```{r coord_trait_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

plot_data <- corr_impact_by_coords %>% dplyr::select(Neale_pheno_ID, trait_couple_corr, corr_due_to_confounding_all) %>% unique()
plot <- ggplot2::ggplot(data = plot_data, ggplot2::aes(x = trait_couple_corr,
                                                       y = corr_due_to_confounding_all)) +
  geom_point(alpha = 3/4) +

  geom_smooth(mapping = aes(x = trait_couple_corr, y = corr_due_to_confounding_all), method = "lm", se=FALSE, formula = y~x+0, fullrange=TRUE, color = custom_col[4]) +
  theme_minimal() +
  ggplot2::labs(x = paste0("Couple trait correlation"),
                y = paste("Couple correlation due to confounding"))

plot

```

The table below gives the data in the plot above:

- `outcome_description` corresponds to the trait. 
- `trait_couple_corr` is the correlation of `outcome_description` in couples. 
-  `corr_due_to_confounding_all` corresponds to the formula: `cor(X,PC)^2*cor(PC_i,PC_p)`, summed across both coorindates.

```{r coord_traits_confounding, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

corr_impact_by_coords.DT <- replace_Neale_ID(plot_data, traits_corr2_filled, "Neale_pheno_ID", "outcome_description") %>% ungroup()
format_round_cols <- colnames(dplyr::select_if(corr_impact_by_coords.DT, is.numeric))
DT::datatable(corr_impact_by_coords.DT, width = "100%")  %>% formatRound(columns=format_round_cols, digits = 3)

```

## Compairing paths from $X_i \rightarrow Y_p$.

We performed two analyses to compare $\rho$, $\gamma$ and $\omega$:

- No adjustment in $Y_i \rightarrow Y_p$ MR. 
- Adjustment for effects on $X_i$ in $Y_i \rightarrow Y_p$ MR,  MVMR model as follows: $Y_p \sim Y_i + X_i$ (IVs were a combination of all $X$ and $Y$ IVs, pruned at standard parameters using 1000G European data).

```{r YiYp_adj_model, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/two_trait_model_yiyp_adj.png", error = FALSE)
```

Broad overview of results are shown in the figure below.

- Without adjustment.

```{r proxyMR_figure_no_adj, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600}
proxyMR_prod_comparison_fig
```

- With adjustment.

```{r proxyMR_figure_adj, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600}
proxyMR_prod_comparison_fig_yiyp_adj
```

Going forward, we will just use the adjusted results.

### Exploring the difference between $\rho$ and $\gamma$.

In general, $\rho$ is significantly larger that $\gamma$, meaning that $X_i \rightarrow Y_p$ favors paths where assortative mating is through $X$ rather than $Y$. This makes a lot of biological and intuitive sense. In other words, exposures are passed from index to partner, rather than outcomes.

A summary of the linear model of $\rho$ vs $\gamma$, forced through the intercept, is below.

```{r proxyMR_summary_rho_vs_gam, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

data <- proxyMR_comparison_summary_yiyp_adj
exposure_sex <- "meta"

fig_data <- data[which(data[["exposure_sex"]]=="male"),]

x_start <- "gam"
y_start <- "rho"
overlay_var_start <- "gam_vs_rho_BF_sig"
overlay_var = paste0(overlay_var_start, "_meta")

x = paste0(x_start, "_beta")
y = paste0(y_start, "_beta")

count <- 0

fig_data <- fig_data %>% rename(x := !!x) %>% rename(y := !!y) %>% rename(overlay_var := !!overlay_var) %>%
    mutate(x_plot = case_when(x < 0 ~ abs(x), TRUE ~ x)) %>%
    mutate(y_plot = case_when(x < 0 ~ -1*y, TRUE ~ y))

summary(lm(y ~ x + 0, data = fig_data))

```

The corresponding figure is below. The blue line includes only BF-significant $\rho$ and $\gamma$, where the green line includes all points (analagous to the linear model above).

```{r proxyMR_fig_rho_vs_gam, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600, message = FALSE, warning = FALSE}

plot <- create_proxy_prod_comparison_fig_ind(data, exposure_sex, x, y, overlay_var, count)

plot[[1]]

```

### Exploring the difference between $\omega$ and [$\rho$ and $\gamma$].

A few observations:

- In general, $\omega$ is significantly larger than both $\rho$ and $\gamma$, as shown in the figures below. 
- On the other hand, when we sum $\rho$ and $\gamma$, $\omega$ is in general lower than the sum. 
- There are a number of instances where $\omega$ is significantly **smaller** than the sum &rightarrow not as interesting? *But what does this mean?* That we didn't have power in the $\omega$ (i.e diagonal) MR to capture the relationship that is likely only going through the either $\rho$ or $\gamma$ (or a sum of the two)? 
- There are a few cases where $\omega$ is significantly **larger** than the sum. See below for details.

```{r proxyMR_fig_omega_vs_gam, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600, message = FALSE, warning = FALSE}
data <- proxyMR_comparison_summary_yiyp_adj
exposure_sex <- "meta"

fig_data <- data[which(data[["exposure_sex"]]=="male"),]

x_start <- "gam"
y_start <- "omega"
overlay_var_start <- "omega_vs_gam_BF_sig"
overlay_var = paste0(overlay_var_start, "_meta")

x = paste0(x_start, "_beta")
y = paste0(y_start, "_beta")

count <- 0

plot <- create_proxy_prod_comparison_fig_ind(data, exposure_sex, x, y, overlay_var, count)

plot[[1]]

```

```{r proxyMR_fig_omega_vs_rho, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600, message = FALSE, warning = FALSE}
data <- proxyMR_comparison_summary_yiyp_adj
exposure_sex <- "meta"

fig_data <- data[which(data[["exposure_sex"]]=="male"),]

x_start <- "rho"
y_start <- "omega"
overlay_var_start <- "omega_vs_rho_BF_sig"
overlay_var = paste0(overlay_var_start, "_meta")

x = paste0(x_start, "_beta")
y = paste0(y_start, "_beta")

count <- 0

plot <- create_proxy_prod_comparison_fig_ind(data, exposure_sex, x, y, overlay_var, count)

plot[[1]]

```

```{r proxyMR_fig_omega_vs_gam_rho, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600, message = FALSE, warning = FALSE}
data <- proxyMR_comparison_summary_yiyp_adj
exposure_sex <- "meta"

fig_data <- data[which(data[["exposure_sex"]]=="male"),]

x_start <- "gam_rho"
y_start <- "omega"
overlay_var_start <- "omega_vs_gam_rho_BF_sig"
overlay_var = paste0(overlay_var_start, "_meta")

x = paste0(x_start, "_beta")
y = paste0(y_start, "_beta")

count <- 0

plot <- create_proxy_prod_comparison_fig_ind(data, exposure_sex, x, y, overlay_var, count)

plot[[1]]

 fig_data <- fig_data %>% rename(x := !!x) %>% rename(y := !!y) %>% rename(overlay_var := !!overlay_var) %>%
    mutate(x_plot = case_when(x < 0 ~ abs(x), TRUE ~ x)) %>%
    mutate(y_plot = case_when(x < 0 ~ -1*y, TRUE ~ y))

 summary(lm(y ~ x + 0, data = fig_data))

```

A summary of the cases where $\omega$ is significantly larger or smaller than the sum of $\rho$ and $\gamma$ are shown in the two tables below (taking absolute values of each).

```{r proxyMR_summary_omega_vs_gam_rho, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600, message = FALSE, warning = FALSE}

omega_larger.DT <- proxyMR_comparison_summary_yiyp_adj %>% filter(omega_vs_gam_rho_BF_sig_meta) %>% filter(abs(omega_meta_beta) > abs(gam_rho_meta_beta)) %>% dplyr::select(exposure_description, outcome_description, gam_rho_meta_beta, omega_meta_beta) %>% unique() 
format_round_cols <- colnames(dplyr::select_if(omega_larger.DT, is.numeric))
DT::datatable(omega_larger.DT, width = "100%")  %>% formatRound(columns=format_round_cols, digits = 3)

omega_smaller.DT <- proxyMR_comparison_summary_yiyp_adj %>% filter(omega_vs_gam_rho_BF_sig_meta) %>% filter(abs(gam_rho_meta_beta) > abs(omega_meta_beta)) %>% dplyr::select(exposure_description, outcome_description, gam_rho_meta_beta, omega_meta_beta) %>% unique() 
format_round_cols <- colnames(dplyr::select_if(omega_smaller.DT, is.numeric))
DT::datatable(omega_smaller.DT, width = "100%")  %>% formatRound(columns=format_round_cols, digits = 3)

#fig_data %>% filter(omega_meta_beta < 0 & gam_rho_meta_beta > 0) %>% dplyr::select(x,y, exposure_description, outcome_description, overlay_var)

```

## Next steps.

- In the $Y_i \rightarrow Y_p$ MR, rather than performing a MVMR ($Y_p \sim Y_i + X_i$), try in a two step process as follows: 
    - Calculate the effect of $Y_p$ without effects of $X_i$ on $Y_p$ as: $Y_{resid} = Y_p - \omega*X_I$.
    - Run MR with the residualized $Y$ as: $Y_{resid} \sim Y_i$

**Question**: What is the corresponding SEs on $Y_{resid}$?

- Check for cases of reverse causation in the same-person MR. 
- Remove SNPs were the effect is significantly larger on the outcome than the exposure. 
- In the results including both sexes, meta-analyze at the SNP-level rather than the MR level (note that right now the two horizontal, same person MRs in the DAG above are different because they are sex specific. I then meta-analyze $\omega$, $\gamma$ and $\rho$ as the *final* step).
- Calculate the variance of $\rho + \gamma$ as: $2 * cor(\rho, \gamma) * SE_{\gamma} * SE{\rho}$, i.e. include the extra-term of 2*correlation.

# PolyMR.

- Finished discussion points. 
- Waiting on Jonathan to add bibtex file before adding my references. 
- Haven't done a summary on CATE - who was going to do that? 
- *Question:* Why is there often an inflection point around the (0,0) point? Is this an effect of standardization?

### 2021-11-02 R/functions.R
summarize household MVMR and meta-analyze.

@@ -2983,6 +2983,58 @@ summarize_household_MR_comprehensive <- function(household_MR, corr_mat_traits){
 
 }
 
+summarize_household_MVMR <- function(household_MVMR, traits_corr2_filled, corr_mat_traits){
+
+  outcome_traits <- traits_corr2_filled[which(traits_corr2_filled[["Neale_file_sex"]]=="both"),]
+  exposure_ID <- household_MVMR[["exposure_ID"]][1]
+  exposure_description <- as.character(outcome_traits[which(outcome_traits[["Neale_pheno_ID"]]==exposure_ID), "description"])
+
+  cat(paste0("Summarizing and meta-analyzing household MMVR results across sexes for all outcomes with phenotype `", exposure_ID, "` as exposure.\n\n"))
+
+  meta_list <- list()
+  for(var in c("yi", "xi", "xp")){
+
+    meta_temp <- household_MVMR %>% dplyr::select(exposure_ID, outcome_ID, starts_with(var)) %>% rename_all(~stringr::str_replace(., paste0("^", var, "_"),"")) %>%
+      dplyr::group_by(exposure_ID, outcome_ID) %>%
+      # note can't easily use the `z.test_p` function because it isn't set up for vector of beta and se's, but rather one argument for each
+      dplyr::group_by(exposure_ID, outcome_ID) %>% group_modify(~ summarize_sex_specific_results(.x$beta, .x$se))
+
+    colnames(meta_temp)[-c(1:2)] <- paste0(var, "_", colnames(meta_temp)[-c(1:2)])
+    meta_list[[var]] <- meta_temp
+
+  }
+
+  meta_result <- reduce(meta_list, full_join)
+
+  summarized_result <- full_join(household_MVMR, meta_result)
+  summarized_result$exposure_description <- NA
+  summarized_result$outcome_description <- NA
+  summarized_result$corr_traits <- NA
+
+  for(i in 1:dim(summarized_result)[1]){
+    outcome_ID <- household_MVMR[["outcome_ID"]][i]
+    exposure_IDX <- exposure_ID
+    outcome_IDX <- outcome_ID
+    if(endsWith(exposure_ID, "_irnt")) {
+      exposure_IDX <- gsub("_irnt", "", exposure_ID)
+    }
+    if(endsWith(outcome_ID, "_irnt")) {
+      outcome_IDX <- gsub("_irnt", "", outcome_ID)
+    }
+    corr_traits <- corr_mat_traits[exposure_IDX, outcome_IDX]
[…]

### 2021-11-02 _targets.R

@@ -532,13 +532,19 @@ list(
   ),
 
   tar_target(
-    household_MVMR, ## `household_MR` is run in full sample only, not binned by age / time-together categories.
+    household_MVMR, # to run model: $Y_p \sim X_i + Y_i + X_p$
     {
       path_household_GWAS
       path_outcome_stats
       run_household_MVMR(exposure_info, outcomes_to_run)
     },
     pattern = map(exposure_info), iteration = "list"
+  ),
+
+  tar_target(
+    household_MVMR_meta,
+    summarize_household_MVMR(household_MVMR, traits_corr2_filled, corr_mat_traits),
+    pattern = map(household_MVMR)
   ),
 
   tar_target(

### 2021-11-02 _targets.R
change name of meta MVMR target

@@ -542,16 +542,16 @@ list(
   ),
 
   tar_target(
-    household_MVMR_meta,
-    summarize_household_MVMR(household_MVMR, traits_corr2_filled, corr_mat_traits),
-    pattern = map(household_MVMR)
-  ),
-
-  tar_target(
     ## summarize into one table, ignore leave-1-out analyses for now, meta-analyze across sexes and calculate heterogeneity statistic between sexes.
     household_MR_summary,
     summarize_household_MR_comprehensive(household_MR, corr_mat_traits),
     pattern = map(household_MR)
+  ),
+
+  tar_target(
+    household_MVMR_summary,
+    summarize_household_MVMR(household_MVMR, traits_corr2_filled, corr_mat_traits),
+    pattern = map(household_MVMR)
   ),
 
   tar_target(

### 2021-11-02 analysis/update_meetings_03_11_2021.Rmd
add update meeting for 03-11-21

---
title: "Update Meeting: 03-11-2021"
author: "Jenny Sjaarda"
date: "2021-11-01"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
source("R/functions.R")
library(targets)
library(dplyr)
library(DT)
library(knitr)
library(ggplot2)
library(kableExtra)

```

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
tar_load(household_MVMR_summary)
tar_load(household_MR_summary)

```

# Progress update

Decided to run MVMR as follows: $Y_p \sim X_i + Y_i + X_p$, using instruments for ${X_i, Y_i, X_p}$. In this MR, the coefficient for $X_i$ would represent the *direct* $X_i \rightarrow Y_p$ causal effect and hopefully this would be close to zero in most cases. Running simply the $Y_p \sim X_i$ MR (with only $X_i$ instruments), would give you the $X_i \rightarrow Y_p$ *total* effect. Then we can compare the direct and total effect results.

A summary of the MVMR results are below, filtered to only traits with abs(correlation) < 0.8, and then BF significant with $X_i$.

The columns `IVW_meta_beta` and `IVW_meta_pval` correspond to the meta-analyzed across sexes univariate results from the $Y_p ~ X_i$ MR. The subsequent columns correspond to the MVMR MR results.

```{r MVMR_summary, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

household_MVMR_summary.DT <- household_MVMR_summary %>% filter(same_trait==FALSE) %>% filter(abs(corr_traits) < 0.8)

num_tests <- dim(household_MVMR_summary.DT)[1]
household_MVMR_BF.DT <- household_MVMR_summary.DT %>% filter(xi_meta_pval < 0.05/num_tests) %>% dplyr::select(exposure_ID, outcome_ID, exposure_description, outcome_description, corr_traits, yi_meta_beta, yi_meta_pval, xi_meta_beta, xi_meta_pval, xp_meta_beta, xp_meta_pval) %>% unique()

household_MR_summary_sub <- household_MR_summary %>% dplyr::select(exposure_ID, outcome_ID, IVW_meta_beta, IVW_meta_pval) %>% unique()
MVMR_uniMR_comparison.DT <- inner_join(household_MR_summary_sub, household_MVMR_BF.DT)
format_round_cols <- colnames(dplyr::select_if(MVMR_uniMR_comparison.DT, is.numeric))

DT::datatable(MVMR_uniMR_comparison.DT, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3) %>% dplyr::select(exposure_ID, outcome_ID, exposure_description, outcome_description, corr_traits, everything())

```

### 2021-11-02 analysis/update_meeting_03_11_2021.Rmd
wflow_publish("analysis/update_meeting_03_11_2021.Rmd")

--- 2021-11-02 analysis/update_meetings_03_11_2021.Rmd
@@
 
-The columns `IVW_meta_beta` and `IVW_meta_pval` correspond to the meta-analyzed across sexes univariate results from the $Y_p ~ X_i$ MR. The subsequent columns correspond to the MVMR MR results.
+The column `corr_traits` corresponds to the raw correlation between traits in the biobank.
+The columns `IVW_meta_beta` and `IVW_meta_pval` correspond to the meta-analyzed across sexes univariate results from the $Y_p ~ X_i$ MR.
+The subsequent columns correspond to the MVMR MR results.
 
-```{r MVMR_summary, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
+```{r MVMR_summary, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
 
@@
 household_MR_summary_sub <- household_MR_summary %>% dplyr::select(exposure_ID, outcome_ID, IVW_meta_beta, IVW_meta_pval) %>% unique()
-MVMR_uniMR_comparison.DT <- inner_join(household_MR_summary_sub, household_MVMR_BF.DT)
+MVMR_uniMR_comparison.DT <- inner_join(household_MR_summary_sub, household_MVMR_BF.DT)  %>% dplyr::select(exposure_ID, outcome_ID, exposure_description, outcome_description, corr_traits, everything())
 format_round_cols <- colnames(dplyr::select_if(MVMR_uniMR_comparison.DT, is.numeric))
 
-DT::datatable(MVMR_uniMR_comparison.DT, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3) %>% dplyr::select(exposure_ID, outcome_ID, exposure_description, outcome_description, corr_traits, everything())
+DT::datatable(MVMR_uniMR_comparison.DT, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)
 

### 2021-11-02 R/functions.R
use abs value of correlation as filtering

@@ -3143,7 +3143,8 @@ find_sig_standard_MR_summary <- function(standard_MR_summary){
 find_non_corr_household_MR_summary <- function(household_MR_summary){
 
   # filter to only pairs that have correlation < 0.8
-  output <- household_MR_summary[which(household_MR_summary$corr_traits<0.8),]
+  output <- household_MR_summary[which(abs(household_MR_summary$corr_traits)<0.8),]
+
   return(output)
 }
 

### 2021-11-03 analysis/update_meeting_03_11_2021.Rmd
wflow_publish("analysis/update_meeting_03_11_2021.Rmd")

@@
 
-household_MR_summary_sub <- household_MR_summary %>% dplyr::select(exposure_ID, outcome_ID, IVW_meta_beta, IVW_meta_pval) %>% unique()
+household_MR_summary_sub <- household_MR_summary %>% dplyr::select(exposure_ID, outcome_ID, IVW_meta_beta, IVW_meta_se, IVW_meta_pval) %>% unique()
 MVMR_uniMR_comparison.DT <- inner_join(household_MR_summary_sub, household_MVMR_BF.DT)  %>% dplyr::select(exposure_ID, outcome_ID, exposure_description, outcome_description, corr_traits, everything())
@@
 ```
+
+```{r direct_vs_indirect_effects_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
+
+plot_data <-  inner_join(household_MR_summary_sub, household_MVMR_summary.DT)  %>% dplyr::select(exposure_ID, outcome_ID, exposure_description, outcome_description, corr_traits, IVW_meta_beta, IVW_meta_se, yi_meta_beta, yi_meta_se, xi_meta_beta, xi_meta_se, xp_meta_beta, xp_meta_se) %>% unique()
+
+#bquote('x axis'~(Å^2))
+#expression(Sepal~Length[cm])
+
+#Indirect~effect~of~X[i]~on~Y[p]~(from~univariate~MR)
+#xlab <- bquote('Indirect effect of x_i on y_p (from univariate MR)'~(Å^2))
+
+plot <-
+  ggplot2::ggplot(data = plot_data, ggplot2::aes(x = IVW_meta_beta,
+                                                 y = xi_meta_beta)) +
+  geom_point(alpha = 3 / 4) +
+  geom_smooth(
+    mapping = aes(x = IVW_meta_beta, y = xi_meta_beta),
+    method = "lm",
+    se = FALSE,
+    formula = y ~ x + 0,
+    fullrange = TRUE,
+    color = custom_col[4]
+  ) +
+  ggplot2::geom_errorbar(
+    ggplot2::aes(ymin = xi_meta_beta - xi_meta_se, ymax = xi_meta_beta + xi_meta_se),
+    colour = "grey",
+    width = 0
+  ) +
+  ggplot2::geom_errorbarh(
+    ggplot2::aes(xmin = IVW_meta_beta - IVW_meta_se, xmax = IVW_meta_beta + IVW_meta_se),
+    colour = "grey",
+    height = 0
+  ) +
+  theme_minimal() +
+  ggplot2::labs(
+    x = expression(Indirect~effect~of~X[i]~on~Y[p]~(from~univariate~MR)),
+    y = expression(Direct~effect~of~X[i]~on~Y[p]~(from~MVMR))
+  ) + 
+  geom_abline(slope=1, intercept=0)
+
+plot
+
+```

### 2021-11-11 R/functions.R
summarize `proxyMR_comparison_yiyp_adj`

@@ -3355,6 +3355,166 @@ run_household_MVMR <- function(exposure_info, outcomes_to_run){
 }
 
 
+run_household_MVMR_joint <- function(exposure_info, outcomes_to_run){
+
+  output_list <- list()
+  exposure_ID <- exposure_info %>% filter(Value=="trait_ID") %>% pull(Info)
+
+  cat(paste0("\nRunning MVMR analyses for all outcomes with phenotype `", exposure_ID, "` as exposure.\n\n"))
+
+  result <- numeric()
+
+  exposure_ID_prev <- ""
+  outcome_ID_prev <- ""
+
+  for(i in 1:dim(outcomes_to_run)[1]){
+
+    outcome_ID <- outcomes_to_run$Neale_pheno_ID[[i]]
+    MR_complete_i <- list()
+
+    #for(exposure_sex in c("male", "female")){
+
+
+      #######################################
+      ## Standard GWAS results with Y IVs ###
+      #######################################
+
+      GWAS_file_yIVs_1 <- paste0("analysis/traitMR/standard_GWAS_rev_filter/", exposure_ID, "/", exposure_ID, "_vs_", outcome_ID, "_GWAS.csv")
+      xi_YIV_GWAS_results <- fread(GWAS_file_yIVs_1, data.table = F) %>% filter(sex == "meta")
+      #xi_YIV_GWAS_results_both <- fread(GWAS_file_yIVs_1, data.table = F) %>% filter(sex == "both_sexes")
+      ## this will pull the standard GWAS results for phenotype `exposure_ID` for IVs from `outcome_ID`, i.e. the effect of G on Xi for only Y IVs.
+
+      GWAS_file_yIVs_2 <- paste0("analysis/traitMR/standard_GWAS_rev_filter/", outcome_ID, "/", outcome_ID, "_vs_", outcome_ID, "_GWAS.csv")
+      yi_YIV_GWAS_results <- fread(GWAS_file_yIVs_2, data.table = F) %>% filter(sex == "meta")
+      ## this will pull the standard GWAS results for phenotype `exposure_ID` for IVs from `outcome_ID`, i.e. the effect of G on Yi for only Y IVs.
+
+      #######################################
+      ## Standard GWAS results with X IVs ###
+      #######################################
+
+      GWAS_file_xIVs_1 <- paste0("analysis/traitMR/standard_GWAS_rev_filter/", exposure_ID, "/", exposure_ID, "_vs_", exposure_ID, "_GWAS.csv")
[…]

### 2021-11-11 _targets.R

@@ -689,6 +689,8 @@ list(
   ),
 
   ## Household MVMR:  $Y_p \sim X_i + Y_i + X_p$
+  ## Decided that this won't work afterall since we don't have additional instruments for X_p, beyond those that are already instruments for X_i. The only way around would be to run a GWAS on X_p, but it's not worth the effort.
+  ## If we change our minds the `run_household_MVMR_joint` needs some minor attention before running.
 
   tar_target(
     household_MVMR,
@@ -696,6 +698,17 @@ list(
       path_household_GWAS
       path_outcome_stats
       run_household_MVMR(exposure_info, outcomes_to_run)
+    },
+    pattern = map(exposure_info), iteration = "list"
+  ),
+
+
+  tar_target(
+    household_MVMR_joint,
+    {
+      path_household_GWAS_filter
+      path_outcome_stats_filter
+      run_household_MVMR_joint(exposure_info, outcomes_to_run)
     },
     pattern = map(exposure_info), iteration = "list"
   ),
@@ -808,13 +821,13 @@ list(
   ),
 
   tar_target(
-    proxyMR_MR_paths_summary_yiyp_adj, ## summarize the different MR paths in each Xi -> Yp
-    summarize_proxyMR_paths(proxyMR_comparison_yiyp_adj)
+    proxyMR_MR_paths_summary_yiyp_adj_joint, ## summarize the different MR paths in each Xi -> Yp
+    summarize_proxyMR_paths(proxyMR_comparison_yiyp_adj_joint)
   ),
 
   tar_target(
-    proxyMR_comparison_summary_yiyp_adj,
-    summarize_proxyMR_comparison(proxyMR_comparison_yiyp_adj, traits_corr2_filled)
+    proxyMR_comparison_summary_yiyp_adj_joint,
+    summarize_proxyMR_comparison_joint(proxyMR_comparison_yiyp_adj_joint, traits_corr2_filled)
   ),
 
   ## Add z's to proxyMR

### 2021-11-11 analysis/update_meeting_11_11_2021.Rmd
add update meeting 11-11-21

---
title: "Update Meeting: 11-11-2021"
author: "Jenny Sjaarda"
date: "2021-11-11"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
source("R/functions.R")
library(targets)
library(dplyr)
library(DT)
library(knitr)
library(ggplot2)
library(kableExtra)

```

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
tar_load(proxyMR_comparison_summary_yiyp_adj_joint)
tar_load(proxyMR_MR_paths_summary_yiyp_adj_joint)

```

# Progress update.

- Last time, we ran a MVMR as follows: $Y_p \sim X_i + Y_i + X_p$, using instruments for ${X_i, Y_i, X_p}$. Unfortunately, realized this won't work after all because there is too much co-linearity between $X_i$ and $X_p$. We also don't have additional instruments for $X_p$, beyond those that are already instruments for $X_i$. The only way around would be to run a GWAS on $X_p$, but probably not worth that effort (unlikely to find any additional instrument anyways).
- Updated pipeline to meta-analyze across sexes at the SNP level rather than the MR level (minimize weak-instrument bias). 
- In all two-trait MRs, I’ve filtered out SNPs that show evidence of reverse causation based on same-person statistics (i.e. from Neale, meta-analyzed across sexes as per update above) and then use those same, filtered IVs in the couple MRs. IVs were filtered based on the code below (for reference), essentially removing any SNP that is more strongly associated with the outcome than the expsoure at a threshold of p < `# r reverse_MR_threshold`, corresponding to the `reverse_MR_threshold` in the code below.

```{r eval=FALSE}

reverse_t_threshold  =  stats::qnorm(reverse_MR_threshold)

dat_filter <- dat %>%
  mutate(z_score.exposure = beta.exposure / se.exposure) %>%
  mutate(z_score.outcome = beta.outcome / se.outcome) %>%
  mutate(std_beta.exposure = z_score.exposure / sqrt(samplesize.exposure)) %>%
  mutate(std_beta.outcome = z_score.outcome / sqrt(samplesize.outcome)) %>%
  mutate(std_se.exposure = 1 / sqrt(samplesize.exposure)) %>%
  mutate(std_se.outcome = 1 / sqrt(samplesize.outcome)) %>%
  mutate(std_t_stat = (abs(std_beta.exposure) - abs(std_beta.outcome)) /
           sqrt(std_se.exposure^2 + std_se.outcome^2)) %>%
  filter(std_t_stat > reverse_t_threshold)
```

# Results.

As a recap, the model is below.

```{r proxyMR_figure_adj, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600}
proxyMR_prod_comparison_fig_yiyp_adj
```

A summary of the regressions between $\omega$, $\rho$ and $\gamma$ is below.

```{r regression_summary, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600}
data <- proxyMR_comparison_summary_yiyp_adj_joint
summary(lm(rho_beta ~ omega_beta + 0, data = data))
summary(lm(gam_beta ~ omega_beta + 0, data = data))
summary(lm(gam_rho_beta ~ omega_beta + 0, data = data))
```

```{r resid_summary, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600}
rho_resid <- resid(lm(rho_beta ~ gam_beta, data = data))
data$gam_rho_resid <- rho_resid + data$gam_beta
summary(lm(gam_rho_resid ~ omega_beta + 0, data = data))
```

```{r proxyMR_fsig_omega_vs_gam, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600, message = FALSE, warning = FALSE}

data$exposure_sex <- "joint"

x_start <- "gam"
y_start <- "omega"
overlay_var_start <- "omega_vs_gam_BF_sig"
overlay_var = paste0(overlay_var_start)

x = paste0(x_start, "_beta")
y = paste0(y_start, "_beta")

count <- 0

plot <- create_proxy_prod_comparison_fig_ind(data, exposure_sex = "joint", x, y, overlay_var, count)

plot[[1]]

```

### 2021-11-22 analysis/update_meeting_03_11_2021.Rmd
wflow_publish("analysis/update_meeting_03_11_2021.Rmd")

@@
 Decided to run MVMR as follows: $Y_p \sim X_i + Y_i + X_p$, using instruments for ${X_i, Y_i, X_p}$.
-In this MR, the coefficient for $X_i$ would represent the *direct* $X_i \rightarrow Y_p$ causal effect and hopefully this would be close to zero in most cases.
-Running simply the $Y_p \sim X_i$ MR (with only $X_i$ instruments), would give you the $X_i \rightarrow Y_p$ *total* effect.
+In this MR, the coefficient for $X_i$ would represent the **direct** $X_i \rightarrow Y_p$ causal effect and hopefully this would be close to zero in most cases.
+Running simply the $Y_p \sim X_i$ MR (with only $X_i$ instruments), would give you the $X_i \rightarrow Y_p$ **total** effect.
 Then we can compare the direct and total effect results.
 
-A summary of the MVMR results are below, filtered to only traits with abs(correlation) < 0.8, and then BF significant with $X_i$.
+A summary of the MVMR results are below, filtered to only traits with `abs(correlation) < 0.8`, and then BF significant with $X_i$.
 
@@
 num_tests <- dim(household_MVMR_summary.DT)[1]
-household_MVMR_BF.DT <- household_MVMR_summary.DT %>% filter(xi_meta_pval < 0.05/num_tests) %>% dplyr::select(exposure_ID, outcome_ID, exposure_description, outcome_description, corr_traits, yi_meta_beta, yi_meta_pval, xi_meta_beta, xi_meta_pval, xp_meta_beta, xp_meta_pval) %>% unique()
+household_MVMR_BF.DT <- household_MVMR_summary.DT %>% filter(xi_meta_pval < 0.05/num_tests*2) %>% dplyr::select(exposure_ID, outcome_ID, exposure_description, outcome_description, corr_traits, yi_meta_beta, yi_meta_pval, xi_meta_beta, xi_meta_pval, xp_meta_beta, xp_meta_pval) %>% unique()
 
@@
 
+The plot below displays the direct effect (from MVMR) vs indirect effects (from univariate MR) of $X_i$ on $Y_p$ for all combinations of traits with `abs(correlation) < 0.8` (i.e.
+the data above plus non-significant results).
+Total number of tests is `r dim(household_MVMR_summary.DT)[1]/2`.
+The second plot shows only BF-significant results according to the MVMR.
+
 ```{r direct_vs_indirect_effects_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
 
-plot_data <-  inner_join(household_MR_summary_sub, household_MVMR_summary.DT)  %>% dplyr::select(exposure_ID, outcome_ID, exposure_description, outcome_description, corr_traits, IVW_meta_beta, IVW_meta_se, yi_meta_beta, yi_meta_se, xi_meta_beta, xi_meta_se, xp_meta_beta, xp_meta_se) %>% unique()
+  plot_data <-  inner_join(household_MR_summary_sub, household_MVMR_summary.DT)  %>% dplyr::select(exposure_ID, outcome_ID, exposure_description, outcome_description, corr_traits, IVW_meta_beta, IVW_meta_se, yi_meta_beta, yi_meta_se, xi_meta_beta, xi_meta_se, xi_meta_pval, xp_meta_beta, xp_meta_se) %>% unique()
+
+sig_plot_data <- plot_data %>% filter(xi_meta_pval < 0.05/num_tests*2)
 
@@
 
+plot_sig <-
+  ggplot2::ggplot(data = sig_plot_data, ggplot2::aes(x = IVW_meta_beta,
+                                                 y = xi_meta_beta)) +
+  geom_point(alpha = 3 / 4) +
+  geom_smooth(
+    mapping = aes(x = IVW_meta_beta, y = xi_meta_beta),
+    method = "lm",
+    se = FALSE,
+    formula = y ~ x + 0,
+    fullrange = TRUE,
+    color = custom_col[4]
+  ) +
+  ggplot2::geom_errorbar(
+    ggplot2::aes(ymin = xi_meta_beta - xi_meta_se, ymax = xi_meta_beta + xi_meta_se),
+    colour = "grey",
+    width = 0
+  ) +
+  ggplot2::geom_errorbarh(
+    ggplot2::aes(xmin = IVW_meta_beta - IVW_meta_se, xmax = IVW_meta_beta + IVW_meta_se),
+    colour = "grey",
+    height = 0
+  ) +
+  theme_minimal() +
+  ggplot2::labs(
+    x = expression(Indirect~effect~of~X[i]~on~Y[p]~(from~univariate~MR)),
+    y = expression(Direct~effect~of~X[i]~on~Y[p]~(from~MVMR))
+  ) + 
+  geom_abline(slope=1, intercept=0)
+
+plot_sig
 ```

### 2021-11-22 analysis/update_meeting_11_11_2021.Rmd
wflow_publish("analysis/update_meeting_11_11_2021.Rmd")

@@
 library(kableExtra)
+library(cowplot)
 
@@
 
-```{r proxyMR_figure_adj, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600}
-proxyMR_prod_comparison_fig_yiyp_adj
+```{r YiYp_adj_model, echo = FALSE}
+library(knitr)
+knitr::include_graphics("assets/two_trait_model_yiyp_adj.png", error = FALSE)
 ```
@@
 
-```{r regression_summary, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600}
+```{r regression_summary, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
 data <- proxyMR_comparison_summary_yiyp_adj_joint
@@
 
-```{r resid_summary, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600}
+```{r resid_summary, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
 rho_resid <- resid(lm(rho_beta ~ gam_beta, data = data))
@@
 
-```{r proxyMR_fsig_omega_vs_gam, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600, message = FALSE, warning = FALSE}
+```{r proxyMR_sig_gam_vs_omega, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600, message = FALSE, warning = FALSE}
 
@@
 ```
+
+```{r proxyMR_sig_rho_vs_omega, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600, message = FALSE, warning = FALSE}
+
+data$exposure_sex <- "joint"
+
+x_start <- "rho"
+y_start <- "omega"
+overlay_var_start <- "omega_vs_rho_BF_sig"
+overlay_var = paste0(overlay_var_start)
+
+x = paste0(x_start, "_beta")
+y = paste0(y_start, "_beta")
+
+count <- 0
+
+plot <- create_proxy_prod_comparison_fig_ind(data, exposure_sex = "joint", x, y, overlay_var, count)
+
+plot[[1]]
+
+```

### 2021-11-22 R/functions.R
Run `household_MR_binned_joint_std`

@@ -2817,6 +2817,73 @@ binned_household_MR_ind_joint <- function(exposure_info, outcome_ID, household_h
 }
 
 
+binned_household_MR_ind_joint_std <- function(exposure_info, outcome_ID, household_harmonised_data, grouping_var, MR_method_list) {
+
+  exposure_ID <- exposure_info %>% filter(Value=="trait_ID") %>% pull(Info)
+  exposure_description <- exposure_info %>% filter(Value=="description") %>% pull(Info)
+
+  pheno_dir <- paste0("analysis/traitMR/")
+
+  household_harmonised_data_group <- household_harmonised_data %>% filter(grouping_var == !!grouping_var)
+
+  household_intervals <- levels(as.factor(household_harmonised_data_group$bin))
+  household_intervals_num <- strex::str_first_number(household_intervals)
+  household_intervals_num[which(is.na(household_intervals_num))] <- 1e3
+
+  full_MR_summary <-  numeric()
+  bin_summary <- numeric()
+
+  for(bin in household_intervals[order(household_intervals_num)]){
+
+    harmonise_sub <- household_harmonised_data_group[which(household_harmonised_data_group[["bin"]]==bin),]
+
+    harmonise_sub_std <- harmonise_sub %>%
+      mutate(z_score.exposure = beta.exposure / se.exposure) %>%
+      mutate(z_score.outcome = beta.outcome / se.outcome) %>%
+      mutate(std_beta.exposure = z_score.exposure / sqrt(samplesize.exposure)) %>%
+      mutate(std_beta.outcome = z_score.outcome / sqrt(samplesize.outcome)) %>%
+      mutate(std_se.exposure = 1 / sqrt(samplesize.exposure)) %>%
+      mutate(std_se.outcome = 1 / sqrt(samplesize.outcome)) %>%
+      mutate(std_t_stat = (abs(std_beta.exposure) - abs(std_beta.outcome)) /
+               sqrt(std_se.exposure^2 + std_se.outcome^2))
+
+    harmonise_sub_std <- harmonise_sub_std %>% dplyr::select(-beta.exposure, -se.exposure, -beta.outcome, -se.outcome) %>%
+      rename(beta.exposure = std_beta.exposure) %>% rename(se.exposure = std_se.exposure) %>%
+      rename(beta.outcome = std_beta.outcome) %>% rename(se.outcome = std_se.outcome)
+
[…]

### 2021-11-22 _targets.R

@@ -600,6 +600,12 @@ list(
   ),
 
   tar_target(
+    household_MR_binned_joint_std, # MR results run in data meta-analyzed at SNP-level. SNP effects were standardized first.
+    run_binned_household_MR_joint(exposure_info, outcomes_to_run, household_harmonised_data_meta_reverse_filter, grouping_var, MR_method_list = MR_method_list),
+    pattern = map(exposure_info, household_harmonised_data_meta_reverse_filter), iteration = "list"
+  ),
+
+  tar_target(
     path_household_MR_binned,
     {
       path_MR_dirs
@@ -881,56 +887,58 @@ list(
     create_proxy_sex_comparison_fig(proxyMR_comparison_summary_yiyp_adj)
   ),
 
-  tar_target(
-    PC_gwas_input,
-    prep_PC_GWAS(data_id_age, data_id_sex, sqc_munge, data_UKBB_sample)
-  ),
+  # Run a GWAS of PCs - based on results of PC correlation decided not to pursue this analysis. If we come back to it, it will need to be rerun in the proper UKB subgroup.
 
-  tar_target(
-    path_PC_gwas_input,
-    write_PC_gwas_input(PC_gwas_input),
-    format = "file"
-  ),
-
-  tar_target(
-    path_v2_snps,
-    create_UKBB_v2_snp_file_list(UKBB_processed_dir),
-    format = "file"
-  ),
-
-  tar_target(
-    file_v2_snps,
-    path_v2_snps
-  ),
-
-  tar_target(
-    bgenie_ukbb_chunks,
-      make_ukbb_chunks(file_v2_snps, chunk_size=5e4), pattern = map(file_v2_snps)
-  ),
-
-  tar_target(
-    bgenie_ukbb_chunks_run,
-    as_tibble(bgenie_ukbb_chunks)
-  ),
-
-  tar_target(
-    path_bgenie_GWAS_dir,
-    create_bgenie_GWAS_dir()
-  ),
-
-  tar_target(
-    path_bgenie_pcs,
-    launch_bgenie(bgenie_ukbb_chunks_run$chr, phenofile = path_PC_gwas_input, UKBB_dir,
[…]

### 2021-11-23 _targets.R
finish replacting `exposures_to_run` with `outcomes_to_run`

@@ -392,7 +392,7 @@ list(
     exposure_info,
     get_trait_info(traits_final,outcomes_to_run$Neale_pheno_ID,
                    data_Neale_manifest, Neale_summary_dir, Neale_output_dir),
-    pattern = map(exposures_to_run), iteration = "list"
+    pattern = map(outcomes_to_run), iteration = "list"
   ),
 
   tar_target(
@@ -426,8 +426,8 @@ list(
 
   tar_target(
     summ_stats,
-    create_summary_stats(exposures_to_run$Neale_pheno_ID, exposure_info, IV_data_summary_run),
-    pattern = map(exposures_to_run, exposure_info, IV_data_summary_run),
+    create_summary_stats(outcomes_to_run$Neale_pheno_ID, exposure_info, IV_data_summary_run),
+    pattern = map(outcomes_to_run, exposure_info, IV_data_summary_run),
     iteration = "list"
   ),
 
@@ -435,9 +435,9 @@ list(
     path_summ_stats,
     {
       create_trait_dirs
-      write_summ_stats(exposures_to_run$Neale_pheno_ID, summ_stats)
+      write_summ_stats(outcomes_to_run$Neale_pheno_ID, summ_stats)
     },
-    pattern = map(exposures_to_run, summ_stats), format = "file"
+    pattern = map(outcomes_to_run, summ_stats), format = "file"
   ),
 
   tar_target(

### 2021-11-23 analysis/couple_selection.Rmd
add phenotype and couple selection schematic links
wflow_publish("analysis/couple_selection.Rmd")

@@
 
-# UKBB full sample
+An overview of couple selection and exclusions can be seen [here](assets/couple_selection.png).
+
+# UKBB full sample.
 
@@
 
-# European, genetic, consenting sample
+# European, genetic, consenting sample.
 
@@
 - `in.white.british.ancestry.subset==1`
+- `used.in.pca.calculation==1`
 
@@
 - Note that this filtering procedure does not necessarily remove related individuals.
-For this, please use the function 'ukb_gen_samples_to_remove()' from the [UKBB r-package](https://cran.r-project.org/web/packages/ukbtools/vignettes/explore-ukb-data.html).
+For this, please use the function `ukb_gen_samples_to_remove()` from the [UKBB r-package](https://cran.r-project.org/web/packages/ukbtools/vignettes/explore-ukb-data.html).
 
-# Houshold pairs
+# Houshold pairs.
 
@@
 
-# Unrelated pairs
+# Unrelated pairs.
 
@@
 
-# Heterosexual pairs
+# Heterosexual pairs.
 
@@
 
-# Couple pairs
+# Couple pairs.
 

### 2021-11-25 R/functions.R
test correlation matrix of all traits (even those that don't pass all filters) to be able to run confounder analyses

@@ -1186,8 +1186,6 @@ extract_trait_info <- function(pheno_data){
 
 calc_corr_mat_traits <- function(outcomes_to_run, path_pheno_data){
 
-
-
   full_df <- numeric()
 
   for(i in 1:dim(outcomes_to_run)[1]){
@@ -1216,6 +1214,35 @@ calc_corr_mat_traits <- function(outcomes_to_run, path_pheno_data){
 
   return(cor_matrix)
 
+
+}
+
+calc_corr_mat_traits_all <- function(traits_all, path_pheno_data, sqc, fam, relatives){
+
+  full_df <- numeric()
+  traits <- traits_all %>% filter(Neale_file_sex=="both")
+
+  for(i in 1:dim(traits)[1]){
+
+    category <- as.character(traits[i,"category"])
+    Neale_pheno_ID <- as.character(traits[i,"Neale_pheno_ID"]) ## this is the Neale_id, used to be pheno_description
+
+    pheno_data <- prep_pheno_data(traits_all, Neale_pheno_ID, sqc, fam, relatives)
+
+    joint_data <- rbind(pheno_data[[1]], pheno_data[[2]]) %>% dplyr::select(-starts_with("PC_")) # remove PC columns
+
+    if(i==1){
+      full_df <- joint_data
+    } else full_df <- merge(full_df, joint_data[,c(1, 4)], all = TRUE)
+
+    print(i)
+  }
+
+  calc_PC_data <- full_df[,-c(1:3)]
+
+  cor_matrix <- cor(calc_PC_data, use = "pairwise.complete.obs")
+
+  return(cor_matrix)
 
 }
 
@@ -1345,6 +1372,51 @@ calc_corr_impact_by_coords <- function(outcomes_to_run, traits_corr, corr_mat_tr
 
   return(temp)
 }
+
+
+calc_corr_impact_by_traits <- function(outcomes_to_run, traits_corr, corr_mat_traits, trait){
+
+  result <- numeric()
+
+  for(i in 1:dim(outcomes_to_run)[1]){
+
+    Neale_pheno_ID <- outcomes_to_run$Neale_pheno_ID[i]
+    phes_ID <- gsub("_irnt", "", Neale_pheno_ID)
+    trait_couple_r2 <- as.numeric(traits_corr[which(traits_corr$ID==phes_ID),"r2"])
+    trait_couple_r <- sqrt(trait_couple_r2)
+
[…]

### 2021-11-25 _targets.R

@@ -401,6 +401,11 @@ list(
   ),
 
   tar_target(
+    corr_mat_traits_all,
+    calc_corr_mat_traits_all(traits_corr2_filled, path_pheno_data)
+  ),
+
+  tar_target(
     PC_trait_corr,
     calc_pc_trait_corr(outcomes_to_run$Neale_pheno_ID, pheno_data),
     pattern = map(outcomes_to_run, pheno_data), iteration = "list"
@@ -941,13 +946,21 @@ list(
     pattern = map(PC_trait_corr)
   ),
 
-
   tar_target(
     corr_impact_by_coords,
     calc_corr_impact_by_coords(outcomes_to_run, traits_corr, corr_mat_traits)
   ),
 
+  tar_target(
+    confounder_traits,
+    c("738", "845", "189_irnt") # household income, age completed education, townsend deprivation index
+  ),
 
+  tar_target(
+    corr_impact_by_traits,
+    calc_corr_impact_by_traits(outcomes_to_run, traits_corr, corr_mat_traits),
+    pattern = map(confounder_traits)
+  ),
   ### markdown docs
 
   tar_render(rmd_index, "analysis/index.Rmd"),

### 2021-11-25 analysis/phenotype_selection.Rmd
use `traits_corr2_filled` in couple correlation Rmd so all traits have a category.

@@
 tar_load(traits_corr2)
+tar_load(traits_corr2_filled)
 tar_load(traits_corr3)
@@
 
-traits_corr2.DT <- traits_corr2 %>% as_tibble() %>% mutate(r = sqrt(r2)) %>% dplyr::select(Neale_pheno_ID, description, variable_type, category, N_pairs, r2, r) %>% unique()
+traits_corr2.DT <- traits_corr2_filled %>% as_tibble() %>% mutate(r = sqrt(r2)) %>% dplyr::select(Neale_pheno_ID, description, variable_type, category, N_pairs, r2, r) %>% unique()
 DT::datatable(traits_corr2.DT, width = "100%") %>% formatRound(columns=c('r2', 'r'), digits = 3)

### 2021-11-25 _targets.R
add missing arguement to calculating cor_mat_traits_all

@@ -402,7 +402,7 @@ list(
 
   tar_target(
     corr_mat_traits_all,
-    calc_corr_mat_traits_all(traits_corr2_filled, path_pheno_data)
+    calc_corr_mat_traits_all(traits_corr2_filled, path_pheno_data, data_sqc, data_fam, data_relatives)
   ),
 
   tar_target(
@@ -953,12 +953,12 @@ list(
 
   tar_target(
     confounder_traits,
-    c("738", "845", "189_irnt") # household income, age completed education, townsend deprivation index
+    c("738", "845", "189_irnt", "20016_irnt") # household income, age completed education, townsend deprivation index, Fluid intelligence score
   ),
 
   tar_target(
     corr_impact_by_traits,
-    calc_corr_impact_by_traits(outcomes_to_run, traits_corr, corr_mat_traits),
+    calc_corr_impact_by_traits(outcomes_to_run, traits_corr, corr_mat_traits_all),
     pattern = map(confounder_traits)
   ),
   ### markdown docs

### 2021-11-28 analysis/update_meeting_29_11_2021.Rmd
add udpate meeting for 29/11/21

---
title: "Update Meeting: 29-11-2021"
author: "Jenny Sjaarda"
date: "2021-11-01"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
source("R/functions.R")
library(targets)
library(dplyr)
library(DT)
library(knitr)
library(ggplot2)
library(kableExtra)
library(cowplot)
library(plotly)
```

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
tar_load(couple_MR_vs_trait_corr)
tar_load(corr_impact_by_traits)

```

# Analyses update.

## MR effects vs. raw correlation.

```{r corr_vs_MR_effects, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}

plot_data <- couple_MR_vs_trait_corr %>% dplyr::select(IVW_beta_std, couple_r)

plot <- ggplot2::ggplot(data = plot_data, ggplot2::aes(x = couple_r,
                                                       y = IVW_beta_std)) +
  geom_point(alpha = 3/4) +

  geom_abline(slope=1, intercept=0) +

  theme_minimal() +
  ggplot2::labs(x = paste0("Couple correlation"),
                y = paste("Standardized MR effect"))

plot

ggplotly(plot)
```

## Re-calculate number of significance in same trait analyses.

## Compare male and female effects.

## Confounder analysis with added variables.

## Repeat hierarchical plot.

### 2021-11-28 R/functions.R
rerun sex specific household MR results at filtered level

@@ -2788,7 +2788,7 @@ binned_household_MR_ind <- function(exposure_info, outcome_ID, household_harmoni
     if(exposure_sex=="female"){index="HOUSEHOLD_MEMBER2"}
     opp_index <- ifelse(index=="HOUSEHOLD_MEMBER1", "HOUSEHOLD_MEMBER2", "HOUSEHOLD_MEMBER1")
 
-    household_harmonised_data_sex <- household_harmonised_data[[paste0("exp_", exposure_sex, "_harmonised_data")]] %>% filter(grouping_var == !!grouping_var)
+    household_harmonised_data_sex <- household_harmonised_data[[paste0("exp_", exposure_sex, "_harmonised_data_filter")]] %>% filter(grouping_var == !!grouping_var)
 
     household_intervals <- levels(as.factor(household_harmonised_data_sex$bin))
     household_intervals_num <- strex::str_first_number(household_intervals)

### 2021-11-28 _targets.R

@@ -583,8 +583,8 @@ list(
   ### NEED TO FIRST FILTER HARMONISED DATA RESULTS
   tar_target(
     household_MR_binned_sex_specific, # MR results are given binned in full sample and binned by time-together and mean age
-    run_binned_household_MR(exposure_info, outcomes_to_run, household_harmonised_data, grouping_var, MR_method_list = MR_method_list),
-    pattern = map(exposure_info, household_harmonised_data), iteration = "list"
+    run_binned_household_MR(exposure_info, outcomes_to_run, household_harmonised_data_reverse_filter, grouping_var, MR_method_list = MR_method_list),
+    pattern = map(exposure_info, household_harmonised_data_reverse_filter), iteration = "list"
   ),
 
   tar_target(

### 2021-11-29 analysis/update_meeting_29_11_2021.Rmd
wflow_publish("analysis/update_meeting_29_11_2021.Rmd")

@@
 library(targets)
+library(kableExtra)
 library(dplyr)
@@
 library(plotly)
+library(plyr)
 ```
@@
 tar_load(corr_impact_by_traits)
+tar_load(traits_corr2_filled)
+tar_load(household_MR_summary_AM)
+tar_load(num_tests_by_PCs)
+tar_load(num_tests_by_PCs_AM_sig)
+tar_load(household_MR_binned_het_sex_spec)
+tar_load(household_MR_binned_het_joint)
+tar_load(traits_final)
+tar_load(household_MR_AM_FvsM_fig)
+tar_load(confounder_traits)
 
@@
 
+Sought to compare the raw correlation amongst couples vs.
+the standardized MR effects.
+There are many traits where the `correlation > MR effects` including some traits of interest such as standing height, place of birth, among many others.
+A plot comparing the two and a summary table of the effects is shown below.
+
 ```{r corr_vs_MR_effects, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
 
-plot_data <- couple_MR_vs_trait_corr %>% dplyr::select(IVW_beta_std, couple_r)
+plot_data <- replace_Neale_ID(couple_MR_vs_trait_corr, traits_corr2_filled, "exposure_ID", "outcome_description") %>% 
+  dplyr::rename(Phenotype = outcome_description) %>% 
+  dplyr::rename(Correlation = couple_r) %>% dplyr::rename(MR_est = IVW_beta_std)
 
-plot <- ggplot2::ggplot(data = plot_data, ggplot2::aes(x = couple_r,
-                                                       y = IVW_beta_std)) +
+plot <- ggplot2::ggplot(data = plot_data, ggplot2::aes(x = Correlation,
+                                                       y = MR_est, Phenotype = Phenotype)) +
   geom_point(alpha = 3/4) +
@@
   geom_abline(slope=1, intercept=0) +
+
+  ggplot2::geom_errorbar(
+    ggplot2::aes(ymin = MR_est - IVW_se_std, ymax = MR_est + IVW_se_std),
+    colour = "grey",
+    width = 0
+  ) +
 
@@
 
-plot
+ggplotly(plot, tooltip = c("x", "y", "Phenotype"))
 
-ggplotly(plot)
 ```
 
-## Re-calculate number of significance in same trait analyses.
+```{r corr_vs_MR_effects_table, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
 
-## Compare male and female effects.
+plot_data.DT <- plot_data %>% dplyr::rename(IVW_beta_std = MR_est) %>% 
+  mutate(mr_corr_diff = Correlation - IVW_beta_std) %>% 
+  mutate(correlation_larger = ifelse(Correlation > IVW_beta_std, TRUE, FALSE)) %>%
+  dplyr::select(Phenotype, Correlation, IVW_beta_std, IVW_pval_std, mr_corr_diff, correlation_larger)
+
+format_round_cols <- colnames(dplyr::select_if(plot_data.DT, is.numeric))
+
+DT::datatable(plot_data.DT, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)
+```
+
+## Same-trait analyses.
+
+**A few key changes to the pipeline since last time we looked at these results:**
+
+1. List of traits of interest changed to exclude dietary traits and add binary traits (among a few other minor revisions), total number of traits went down. 
+2.
+Sample changed to ensure both males and females were not related (previously had only removed unrelated within males and females separately), our resulting samples closely matched Neale round 2 GWAS which resulted in much fewer couples (~50K instead of ~65K).
+3. Filtered SNPs for evidence of reverse causality (should not have any bearing on these results since these involve only one trait).
+4. Meta-analyses are performed at SNP-level rather than MR-level.
+
+### Re-calculate number of significant results in same trait analyses.
+
+```{r AM_data_prep, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, include = FALSE}
+
+AM_result_sig <- household_MR_summary_AM %>% filter(IVW_pval < 0.05/num_tests_by_PCs) 
+AM_sig_traits <-household_MR_summary_AM %>% filter(IVW_pval < 0.05/num_tests_by_PCs) %>% pull(exposure_ID) %>% unique()
+AM_sex_spec_result_sig <- household_MR_binned_het_sex_spec %>% filter(same_trait ==TRUE) %>% filter(exposure_ID %in% AM_sig_traits) %>% filter(sex_het_pval < 0.05/num_tests_by_PCs_AM_sig)
+
+```
+- The number of analyses corresponds to the number of traits (`r dim(traits_final)[1]`). 
+- The significant results after adjusting for multiple hypothesis testing are shown below. 
+- `IVW_meta_beta` and `IVW_meta_pval` corresponds to the beta and p-value of the meta-analyzed MR across sexes, respectively (i.e.
+MR estimates were computed in each sex-seperately using sex-specific SNP-exposure and SNP-outcome results and then meta-analyzed).
+- After adjusting for multiple tests (p < `0.05/`r num_tests_by_PCs``), identified `r dim(AM_result_sig)[1]` significant assortative mating MR results (corresponding to the table below).
+
+```{r, AM_sig, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, include = TRUE}
+
+AM_result_sig.DT <- AM_result_sig %>% dplyr::select(exposure_description, IVW_beta, IVW_pval) %>% unique()
+DT::datatable(AM_result_sig.DT, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval'), digits=3)
+
+```
+
+### Sex heterogeneity.
+
+- For each AM MR ($X_{p} \sim X_i$), analyses were run in each sex separately. 
+- Of the `r dim(AM_sig_traits)[1]` significant results shown above, we tested to see if there was any difference in AM MR estimates between sexes. 
+- A similar adjustment for multiple hypothesis testing was applied as above, resulting in `r num_tests_by_PCs_AM_sig` number of effective tests. 
+- After adjusting for multiple hypothesis testing (p < `0.05/`r num_tests_by_PCs_AM_sig``), `r dim(AM_sex_spec_result_sig)[1]` traits showed significant differences amongst sexes (this changed from before).
+- The table below shows the *nominally significant* results at `p < 0.05`.
+
+```{r AM_sex_het_nom, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
+
+AM_sig_sex_het_nom <- household_MR_binned_het_sex_spec %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% filter(sex_het_pval < 0.05)
+AM_sig_sex_het_nom <- replace_Neale_ID(AM_sig_sex_het_nom, traits_corr2_filled,"exposure_ID", "exposure_description")
+AM_result_sig_nom.DT <- AM_sig_sex_het_nom %>% dplyr::select(exposure_description, exposure_sex, IVW_beta, IVW_pval, sex_het_pval) %>% unique()
+DT::datatable(AM_result_sig_nom.DT, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'sex_het_pval'), digits=3)
+
+```
+
+Figures compared the male to female MR results, and vice versa are shown below. *Suggests evidence of regression dilution bias?*
+
+```{r sex_het_fig, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
+household_MR_AM_FvsM_fig[[1]]
+household_MR_AM_FvsM_fig[[2]]
+```
+
+### Binned results.
+
+```{r binned_results_prep, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, include = FALSE}
+
+slope_sig_age <- household_MR_binned_het_joint %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
+  filter(grouping_var == "age_even_bins") %>% 
+  filter(bin_slope_pval < 0.05/num_tests_by_PCs_AM_sig)
+
+slope_sig_tt <- household_MR_binned_het_joint %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
+  filter(grouping_var == "time_together_even_bins") %>% 
+  filter(bin_slope_pval < 0.05/num_tests_by_PCs_AM_sig)
+
+slope_wt_sig_age <- household_MR_binned_het_joint %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
+  filter(grouping_var == "age_even_bins") %>% 
+  filter(bin_slope_pval_wt < 0.05/num_tests_by_PCs_AM_sig)
+
+slope_wt_sig_tt <- household_MR_binned_het_joint %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
+  filter(grouping_var == "time_together_even_bins") %>% 
+  filter(bin_slope_pval_wt < 0.05/num_tests_by_PCs_AM_sig)
+
+sex_het_sig_age <- household_MR_binned_het_sex_spec %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
+  filter(grouping_var == "age_even_bins") %>% 
+  filter(bin_slope_sex_het_pval < 0.05/num_tests_by_PCs_AM_sig)
+
+sex_het_sig_tt <- household_MR_binned_het_sex_spec %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
+  filter(grouping_var == "time_together_even_bins") %>% 
+  filter(bin_slope_sex_het_pval < 0.05/num_tests_by_PCs_AM_sig)
+
+sex_het_wt_sig_age <- household_MR_binned_het_sex_spec %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
+  filter(grouping_var == "age_even_bins") %>% 
+  filter(bin_slope_sex_het_pval_wt < 0.05/num_tests_by_PCs_AM_sig)
+
+sex_het_wt_sig_tt <- household_MR_binned_het_sex_spec %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
+  filter(grouping_var == "time_together_even_bins") %>% 
+  filter(bin_slope_sex_het_pval_wt < 0.05/num_tests_by_PCs_AM_sig)
+
+sex_het_wt_sig_age_traits <- replace_Neale_ID(sex_het_wt_sig_age, traits_corr2_filled, "exposure_ID", "outcome_description") %>% pull(outcome_description) %>% unique()
+```
+
+- For each AM MR ($X_{p} \sim X_i$), analyses were run in the full sample as well as in 5 roughly equal sized bins according to time couple had been together (`time_together_even_bins`) estimated using [time at household variable](https://biobank.ctsu.ox.ac.uk/showcase/field.cgi?id=699), and median age (`age_even_bins`).
+- Of the `r length(AM_sig_traits)` significant results shown above, we tested to see if there was any significant difference in AM MR estimates amongst the two grouping variables. 
+- Specifically, the outcome data set was split into 5 bins, and the SNP-outcome effect was estimated in each bin separately (and each sex seperately). 
+- These outcome-SNP effects were then used to generate bin-specific MR estimates, using the same SNP-exposure effects from Neale. 
+- To estimate the difference among bins, we used two approaches: 
+    1. The Cochran's Q test to test for heterogeneity in meta-analyses *(of which we obtained no significant results, so not presenting in paper)*.
+    2. Tested the significance of the slope of linear model of median bin (either `age` or `time together`) versus the bin-specific MR estimate.
+
+------------------------------------------------------------------------
+
+Slopes were calculated by estimating the beta-coefficient of a linear regression between MR estimate within each bin (dependent variable) and median bin (independent variable, either age or time at same household).
+Linear models were run both *unweighted* and *weighted* for the by the inverse of the SE of the MR estimate.
+
+The number of significant trends ($\beta$ estimates from the model: $\alpha_{bin} \sim median_{bin}$) significant after multiple hypothesis testing (p < `0.05/`r num_tests_by_PCs_AM_sig` = `r 0.05/num_tests_by_PCs_AM_sig``) in each group was:
+
+  - Binned by median *age* of couples, number of significant results after adjusting for number of tests:
+      - $\beta$ unweighted, meta-anlayzed across sexes: `r dim(slope_sig_age)[1]`
+      - $\beta$ weighted by inverse of the SE, meta-anlayzed across sexes: `r dim(slope_wt_sig_age)[1]`
+      - $\beta$ unweighted, significantly different between sexes: `r dim(sex_het_sig_age)[1]/2`
+      - $\beta$ weighted by inverse of the SE, significantly different between sexes: `r dim(sex_het_wt_sig_age)[1]/2`
+
+  - Binned by *time* couple had been together, number of significant results after adjusting for number of tests:
+      - $\beta$ unweighted, meta-anlayzed across sexes: `r dim(slope_sig_tt)[1]`
+      - $\beta$ weighted by inverse of the SE, meta-anlayzed across sexes: `r dim(slope_wt_sig_tt)[1]`
+      - $\beta$ unweighted, significantly different between sexes: `r dim(sex_het_sig_tt)[1]/2`.
+      - $\beta$ weighted by inverse of the SE, significantly different between sexes: `r dim(sex_het_wt_sig_tt)[1]/2`
+
+```{r binned_results, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
+
+full_binn <- rbind.fill(household_MR_binned_het_sex_spec, household_MR_binned_het_joint) %>% filter(same_trait)
+full_binn.DT <- full_binn %>% dplyr::select(exposure_ID, exposure_sex, grouping_var, IVW_beta, IVW_pval, sex_het_pval, bin_slope_beta,
+                                            bin_slope_pval,bin_slope_sex_het_pval, bin_slope_sex_het_pval_wt) %>% unique() %>% arrange(exposure_ID)
+
+full_binn.DT <- replace_Neale_ID(full_binn.DT, traits_corr2_filled, "exposure_ID", "outcome_description")
+
+format_round_cols <- colnames(dplyr::select_if(full_binn.DT, is.numeric))
+
+DT::datatable(full_binn.DT, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)
+
+```
+
+### Summary.
+
+- From the same-trait analyses we unforutantely are not left with many significant results after adjusting for multiple tests. 
+- The plot comparing male to female estimates does seem to be an effect of regression dilution bias, since the inverse plot suggests the inverse conclusion? Am I interpretting this right? 
+- Our most significant results are the binned results by age, where we have 4 traits that show significant differences according to the slope (the result is the same regardless of using the weighted or unweighted beta estimates), but none of the joint results are BF significant.
+The four traits are: `r sex_het_wt_sig_age_traits`.
 
@@
 
-## Repeat hierarchical plot.
+Repeated the confounder analysis with the following 4 traits:
+
+```{r confounder_traits, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
+confounder_trait_tibble <- as_tibble(confounder_traits)
+confounder_trait_tibble_descript <- replace_Neale_ID(confounder_trait_tibble, traits_corr2_filled, "value", "outcome_description")
+confounder_trait_tibble_descript %>% 
+  kbl() %>%
+  kable_styling()
+```
+
+------------------------------------------------------------------------
+
+Below are the corresponding figures:
+
+```{r confounder_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
+
+plots <- list()
+for(i in 1:length(confounder_traits)){
+
+  trait <- confounder_traits[i]
+  trait_description <- confounder_trait_tibble_descript[[1]][i]
+
+  plot_data <- corr_impact_by_traits %>% filter(trait_interest==trait) %>% filter(trait_interest!=trait_j)
+  plot_data <- replace_Neale_ID(plot_data, traits_corr2_filled, "trait_j", "Phenotype")
+  plot <- ggplot2::ggplot(data = plot_data, ggplot2::aes(x = trait_j_couple_corr,
+                                                         y = corr_due_to_confounding, Phenotype = Phenotype)) +
+    geom_point(alpha = 3/4) +
+
+    geom_smooth(mapping = aes(x = trait_j_couple_corr, y = corr_due_to_confounding), method = "lm", se=FALSE, formula = y~x+0, fullrange=TRUE, color = custom_col[4]) +
+    theme_minimal() +
+    ggplot2::labs(x = paste0("Couple trait correlation"),
+                  y = paste0("Couple correlation due to confounding"),
+                  title = paste0("Impact of ", trait_description, " on couple correlation"))
+
+  plots[[i]] <- plot 
+}
+
+ggplotly(plots[[1]], tooltip = c("x", "y", "Phenotype"))
+ggplotly(plots[[2]], tooltip = c("x", "y", "Phenotype"))
+ggplotly(plots[[3]], tooltip = c("x", "y", "Phenotype"))
+ggplotly(plots[[4]], tooltip = c("x", "y", "Phenotype"))
+
+```

### 2022-01-12 R/functions.R
fix a few typos in `calc_corr_impact_by_coords`

@@ -1392,9 +1392,7 @@ calc_corr_impact_by_coords <- function(outcomes_to_run, traits_corr, corr_mat_tr
   EC_pheno_data <- prep_pheno_data(traits_all, "130_irnt", sqc, fam, relatives)
   EC_joint_data <- rbind(EC_pheno_data[[1]], EC_pheno_data[[2]]) %>% dplyr::select(-starts_with("PC_")) # remove PC columns
 
-
   for(i in 1:dim(outcomes_to_run)[1]){
-
 
     Neale_pheno_ID <- outcomes_to_run$Neale_pheno_ID[i]
     phes_ID <- gsub("_irnt", "", Neale_pheno_ID)
@@ -1416,13 +1414,13 @@ calc_corr_impact_by_coords <- function(outcomes_to_run, traits_corr, corr_mat_tr
     ###########################################################
 
     trait_pheno_data <- prep_pheno_data(traits_all, Neale_pheno_ID, sqc, fam, relatives)
-    trait_joint_data <- rbind(i_pheno_data[[1]], i_pheno_data[[2]]) %>% dplyr::select(-starts_with("PC_")) # remove PC columns
+    trait_joint_data <- rbind(trait_pheno_data[[1]], trait_pheno_data[[2]]) %>% dplyr::select(-starts_with("PC_")) # remove PC columns
 
-    full_df_NC <- merge(trait_interest_joint_data, NC_joint_data[,c(1, 4)], all = TRUE)
-    full_df_EC <- merge(trait_interest_joint_data, EC_joint_data[,c(1, 4)], all = TRUE)
+    full_df_NC <- merge(NC_joint_data, trait_joint_data[,c(1, 4)], all = TRUE)
+    full_df_EC <- merge(EC_joint_data, trait_joint_data[,c(1, 4)], all = TRUE)
 
-    full_df_complete_NC <- full_df[complete.cases(full_df_NC), ]
-    full_df_complete_EC <- full_df[complete.cases(full_df_EC), ]
+    full_df_complete_NC <- full_df_NC[complete.cases(full_df_NC), ]
+    full_df_complete_EC <- full_df_EC[complete.cases(full_df_EC), ]
 
     trait_NC_n_pairs <- dim(full_df_complete_NC)[1] # represents number of pairs in the correlation between the NC and trait `i`
     trait_EC_n_pairs <- dim(full_df_complete_EC)[1] # represents number of pairs in the correlation between the EC and trait `i`

### 2022-01-12 analysis/manuscript.Rmd
create a manuscript Rmd with figures

--- 2021-11-28 analysis/update_meeting_29_11_2021.Rmd
@@
 ---
-title: "Update Meeting: 29-11-2021"
+title: "Manuscript"
 author: "Jenny Sjaarda"
-date: "2021-11-01"
+date: "2022-01-04"
 output: workflowr::wflow_html
-editor_options:
-  chunk_output_type: console
 ---
@@
 
-# Analyses update.
+# A few notes.
 
-## MR effects vs. raw correlation.
+- The number of effective tests can be found at the target `num_tests_by_PCs`. 
+- The number of effective tests among those significant in the MR can be found at the target `num_tests_by_PCs_AM_sig`.
+
+# Relationship between causal effects and raw phenotypic correlation in couples
+
+Sought to compare the raw correlation amongst couples vs.
+the standardized MR effects.
+There are many traits where the `correlation > MR effects` including some traits of interest such as standing height, place of birth, among many others.
+A plot comparing the two and a summary table of the significant effects is shown below.
 
@@
 
-plot_data <- couple_MR_vs_trait_corr %>% dplyr::select(IVW_beta_std, couple_r)
+plot_data <- replace_Neale_ID(couple_MR_vs_trait_corr, traits_corr2_filled, "exposure_ID", "outcome_description") %>% 
+  dplyr::rename(Phenotype = outcome_description) %>% 
+  dplyr::rename(Correlation = couple_r) %>% dplyr::rename(MR_est = IVW_beta)
 
-plot <- ggplot2::ggplot(data = plot_data, ggplot2::aes(x = couple_r,
-                                                       y = IVW_beta_std)) +
+plot <- ggplot2::ggplot(data = plot_data, ggplot2::aes(x = Correlation,
+                                                       y = MR_est, Phenotype = Phenotype)) +
   geom_point(alpha = 3/4) +
@@
   geom_abline(slope=1, intercept=0) +
+
+  ggplot2::geom_errorbar(
+    ggplot2::aes(ymin = MR_est - IVW_se, ymax = MR_est + IVW_se),
+    colour = "grey",
+    width = 0
+  ) +
+  ggplot2::geom_errorbar(
+    ggplot2::aes(xmin = Correlation - couple_r_se, xmax = Correlation + couple_r_se),
+    colour = "grey",
+    width = 0
+  ) +
 
@@
 
-plot
+ggplotly(plot, tooltip = c("x", "y", "Phenotype"))
 
-ggplotly(plot)
 ```
 
-## Re-calculate number of significance in same trait analyses.
+```{r corr_vs_MR_effects_table, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
 
-## Compare male and female effects.
+plot_data.DT <- plot_data %>% dplyr::rename(IVW_beta = MR_est) %>% 
+  filter(diff_p < 0.05/num_tests_by_PCs) %>%
+  mutate(mr_corr_diff = Correlation - IVW_beta) %>% 
+  mutate(correlation_larger = ifelse(Correlation > IVW_beta, TRUE, FALSE)) %>%
+  dplyr::select(Phenotype, Correlation, IVW_beta, IVW_pval, mr_corr_diff, correlation_larger)
 
-## Confounder analysis with added variables.
+format_round_cols <- colnames(dplyr::select_if(plot_data.DT, is.numeric))
 
-## Repeat hierarchical plot.
+DT::datatable(plot_data.DT, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)
+```
+
+# Effect of sex, age and time together on causal effects in couples
+
+```{r AM_data_prep, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, include = FALSE}
+
+AM_result_sig <- household_MR_summary_AM %>% filter(IVW_pval < 0.05/num_tests_by_PCs) 
+AM_sig_traits <-household_MR_summary_AM %>% filter(IVW_pval < 0.05/num_tests_by_PCs) %>% pull(exposure_ID) %>% unique()
+AM_sex_spec_result_sig <- household_MR_binned_het_sex_spec %>% filter(same_trait ==TRUE) %>% filter(exposure_ID %in% AM_sig_traits) %>% filter(sex_het_pval < 0.05/num_tests_by_PCs_AM_sig)
+
+```
+
+- The number of analyses corresponds to the number of traits (`r dim(traits_final)[1]`). 
+- The significant results after adjusting for multiple hypothesis testing are shown below. 
+- `IVW_meta_beta` and `IVW_meta_pval` corresponds to the beta and p-value of the meta-analyzed MR across sexes, respectively (i.e.
+MR estimates were computed in each sex-seperately using sex-specific SNP-exposure and SNP-outcome results and then meta-analyzed).
+- After adjusting for multiple tests (p < `0.05/`r num_tests_by_PCs``), identified `r dim(AM_result_sig)[1]` significant assortative mating MR results (corresponding to the table below).
+
+# Impact of confounders on couple correlation
+
+We tested the impact of confounding on genetic PCs, N/E coordinates and the following 4 traits:
+
+```{r confounder_traits, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
+confounder_trait_tibble <- as_tibble(confounder_traits)
+confounder_trait_tibble_descript <- replace_Neale_ID(confounder_trait_tibble, traits_corr2_filled, "value", "outcome_description")
+confounder_trait_tibble_descript %>% 
+  kbl() %>%
+  kable_styling()
+```
+
+```{r confounder_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
+
+plots <- list()
+for(i in 1:length(confounder_traits)){
+
+  trait <- confounder_traits[i]
+  trait_description <- confounder_trait_tibble_descript[[1]][i]
+
+  plot_data <- corr_impact_by_traits %>% filter(trait_interest==trait) %>% filter(trait_interest!=trait_j)
+  plot_data <- replace_Neale_ID(plot_data, traits_corr2_filled, "trait_j", "Phenotype")
+  plot <-
+    ggplot2::ggplot(
+      data = plot_data,
+      ggplot2::aes(x = trait_j_couple_corr,
+                   y = corr_due_to_confounding, Phenotype = Phenotype)
+    ) +
+    geom_point(alpha = 3 / 4) +
+
+    ggplot2::geom_errorbar(
+      ggplot2::aes(ymin = corr_due_to_confounding - corr_due_to_confounding_se, ymax = corr_due_to_confounding + corr_due_to_confounding_se),
+      colour = "grey",
+      width = 0
+    ) +
+    ggplot2::geom_errorbar(
+      ggplot2::aes(xmin = trait_j_couple_corr - trait_j_couple_corr_se, xmax = trait_j_couple_corr + trait_j_couple_corr_se),
+      colour = "grey",
+      width = 0
+    ) +
+
+    theme_minimal() +
+    ggplot2::labs(
+      x = paste0("Couple trait correlation"),
+      y = paste0("Couple correlation due to confounding"),
+      title = paste0("Impact of ", trait_description, "\n on couple correlation")
+    )
+
+  plots[[i]] <- plot 
+}
+
+figures_grid <- plot_grid(plotlist = plots, nrow=2, ncol = 2, byrow = F, labels="AUTO")
+figures_grid
+```

### 2022-02-03 analysis/manuscript.Rmd
wflow_publish("analysis/manuscript.Rmd")

--- 2021-11-29 analysis/update_meeting_29_11_2021.Rmd
@@
 library(kableExtra)
-library(dplyr)
 library(DT)
@@
 library(plyr)
+library(dplyr)
+library(ggrepel)
+library(viridis)
+library(hrbrthemes)
+library(ggpubr)
+library(broom)
 ```
@@
 tar_load(couple_MR_vs_trait_corr)
+
+tar_load(corr_impact_by_PCs)
 tar_load(corr_impact_by_traits)
+tar_load(corr_impact_by_coords)
+tar_load(corr_potential_trait_confounders)
+tar_load(traits_corr)
+tar_load(PCs_corr)
 tar_load(traits_corr2_filled)
@@
 tar_load(household_MR_binned_het_sex_spec)
-tar_load(household_MR_binned_het_joint)
+tar_load(household_MR_binned_het_SNPmeta)
 tar_load(traits_final)
@@
 tar_load(confounder_traits)
+tar_load(proxyMR_prod_comparison_fig_yiyp_adj)
+tar_load(proxyMR_comparison_summary_yiyp_adj_SNPmeta)
+tar_load(household_MR_summary_AM)
+tar_load(household_MR_binned_het_SNPmeta)
 
@@
 
-# Analyses update.
+# A few notes (for myself).
 
-## MR effects vs. raw correlation.
+- The number of effective tests can be found at the target `num_tests_by_PCs`. 
+- The number of effective tests among those significant in the MR can be found at the target `num_tests_by_PCs_AM_sig`.
+
+# Relationship between causal effects and raw phenotypic correlation in couples.
 
@@
 There are many traits where the `correlation > MR effects` including some traits of interest such as standing height, place of birth, among many others.
-A plot comparing the two and a summary table of the effects is shown below.
+A plot comparing the two and a summary table of the significant effects is shown below.
 
@@
 
+interesting_phenotypes <- c("Basal metabolic rate", "Place of birth in UK - north co-ordinate", 
+                            "Systolic blood pressure, automated reading", "Standing height", 
+                            "Body fat percentage", "Forced vital capacity (FVC), Best measure","Red blood cell (erythrocyte) count")
+
 plot_data <- replace_Neale_ID(couple_MR_vs_trait_corr, traits_corr2_filled, "exposure_ID", "outcome_description") %>% 
-  dplyr::rename(Phenotype = outcome_description) %>% 
-  dplyr::rename(Correlation = couple_r) %>% dplyr::rename(MR_est = IVW_beta_std)
+  mutate(Phenotype = outcome_description) %>% 
+  mutate(Phenotype = case_when(Phenotype == "Basal metabolic rate" ~ "BMR",
+                               Phenotype == "Systolic blood pressure, automated reading" ~ "SBP",
+                               Phenotype == "Body fat percentage" ~ "BFP",
+                               Phenotype == "Forced vital capacity (FVC), Best measure" ~ "FVC",
+                               Phenotype == "Red blood cell (erythrocyte) count" ~ "RBC count",
+                               Phenotype == "Place of birth in UK - north co-ordinate" ~ "Place of birth, NC",
+                               TRUE ~ Phenotype)) %>%
+  dplyr::rename(Correlation = couple_r) %>% dplyr::rename(MR_est = IVW_beta) %>% 
+  mutate(diff_p_sig = case_when(diff_p < 0.05/num_tests_by_PCs ~ TRUE, TRUE ~ FALSE)) %>%
+  mutate(label_phenos = case_when(outcome_description %in% interesting_phenotypes ~ TRUE, TRUE ~ FALSE))
+
+  # TRUE - BF sig difference 
+  # FALSE - non-sig 
+  #mutate(diff_p_sig = ifelse(diff_p < 0.05/num_tests_by_PCs, TRUE, FALSE))
 
 plot <- ggplot2::ggplot(data = plot_data, ggplot2::aes(x = Correlation,
-                                                       y = MR_est, Phenotype = Phenotype)) +
-  geom_point(alpha = 3/4) +
-
+                                                       y = MR_est, 
+                                                       label = Phenotype, 
+                                                       color = factor(diff_p_sig)),) +
+  geom_point(alpha = 1.2) +
+  scale_color_manual(values = custom_col[c(1,2)], labels=c("Non-significant","BF-signifncant difference")) +
+  #geom_text(show.legend = FALSE) +
   geom_abline(slope=1, intercept=0) +
@@
   ggplot2::geom_errorbar(
-    ggplot2::aes(ymin = MR_est - IVW_se_std, ymax = MR_est + IVW_se_std),
+    ggplot2::aes(ymin = MR_est - IVW_se, ymax = MR_est + IVW_se),
+    colour = "grey",
+    width = 0
+  ) +
+  ggplot2::geom_errorbar(
+    ggplot2::aes(xmin = Correlation - couple_r_se, xmax = Correlation + couple_r_se),
     colour = "grey",
@@
   ggplot2::labs(x = paste0("Couple correlation"),
-                y = paste("Standardized MR effect"))
+                y = paste("Standardized MR effect")) +
+  theme(legend.position="bottom") + theme(legend.title=element_blank()) +
+  geom_text_repel(data = subset(plot_data, label_phenos==TRUE),
+                  nudge_y = -0.25 - subset(plot_data, label_phenos==TRUE)$MR_est, size = 2,
+                  direction = "x")
 
-ggplotly(plot, tooltip = c("x", "y", "Phenotype"))
+plot
+
+plot_data.DT <- plot_data %>% dplyr::rename(IVW_beta = MR_est) %>% 
+  filter(diff_p < 0.05/num_tests_by_PCs) %>%
+  mutate(mr_corr_diff = Correlation - IVW_beta) %>% 
+  mutate(correlation_larger = ifelse(Correlation > IVW_beta, TRUE, FALSE)) %>%
+  dplyr::select(outcome_description, Correlation, IVW_beta, IVW_pval, diff_p, mr_corr_diff)
 
@@
 
+## Summary of cases where correlation > MR estimate.
+
+The table below displays the `r dim(plot_data.DT)[1]` traits that have correlation significantly > MR-estimate.
+
 ```{r corr_vs_MR_effects_table, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
-
-plot_data.DT <- plot_data %>% dplyr::rename(IVW_beta_std = MR_est) %>% 
-  mutate(mr_corr_diff = Correlation - IVW_beta_std) %>% 
-  mutate(correlation_larger = ifelse(Correlation > IVW_beta_std, TRUE, FALSE)) %>%
-  dplyr::select(Phenotype, Correlation, IVW_beta_std, IVW_pval_std, mr_corr_diff, correlation_larger)
 
@@
 
-## Same-trait analyses.
+## Search for potential confounders to explain cases where correlation > MR estimate.
 
-**A few key changes to the pipeline since last time we looked at these results:**
+1. Restrict to only traits that have correlation significantly > MR-estimate (`r dim(plot_data.DT)[1]` traits have correlation > MR_estimate).
+2. For each of these traits, find the traits which have a causal effect on this trait (MR p-val < 0.05/[number of effective tests]).
+3. Calculate the correlation due to confounding as: $C = (\alpha_{z\rightarrow x})^2 * \alpha_{z_i\rightarrow z_p}$.
+4. Calculate the ratio of this correlation ($C$) and the correlation of $X$ in partners ($C/r_x$).
 
-1. List of traits of interest changed to exclude dietary traits and add binary traits (among a few other minor revisions), total number of traits went down. 
-2.
-Sample changed to ensure both males and females were not related (previously had only removed unrelated within males and females separately), our resulting samples closely matched Neale round 2 GWAS which resulted in much fewer couples (~50K instead of ~65K).
-3. Filtered SNPs for evidence of reverse causality (should not have any bearing on these results since these involve only one trait).
-4. Meta-analyses are performed at SNP-level rather than MR-level.
+A summary of the these results can be seen below:
 
-### Re-calculate number of significant results in same trait analyses.
+```{r confounder_summary, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
+
+number_of_potential_confounders <- dim(corr_potential_trait_confounders)[1]/dim(plot_data.DT)[1]
+
+non_sig_confounders <- corr_potential_trait_confounders %>% filter(!sig_confounder) %>% group_by(outcome_ID) %>%
+  dplyr::summarize(description = outcome_description[1], n = n()) %>% 
+  filter(n==number_of_potential_confounders) %>%
+  mutate(n = n-number_of_potential_confounders) %>% mutate(potential_confounders = NA) %>%
+  mutate(mean_C_ratio = NA) %>%
+  dplyr::rename(trait_ID = outcome_ID)
+
+sig_confounders <- corr_potential_trait_confounders %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
+  #mutate(potential_confounders = paste0(exposure_description, collapse = "; ")) %>%
+  dplyr::summarize(description = outcome_description[1], n = n(), 
+                   potential_confounders = paste0(exposure_description, collapse = "; "), 
+                   mean_C_ratio = mean(corr_due_to_confounding_ratio)) %>%
+  dplyr::rename(trait_ID = outcome_ID)
+
+confounder_summary <- rbind(non_sig_confounders, sig_confounders)
+
+format_round_cols <- colnames(dplyr::select_if(confounder_summary, is.numeric))
+
+DT::datatable(confounder_summary, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)
+
+```
+
+# Effect of sex, age and time together on causal effects in couples.
 
@@
 ```
-- The number of analyses corresponds to the number of traits (`r dim(traits_final)[1]`). 
+
+- For each trait in the pipeline, we tested the causal effect within couples (i.e. the number of analyses corresponds to the number of traits (`r dim(traits_final)[1]`)). 
 - The significant results after adjusting for multiple hypothesis testing are shown below. 
@@
 MR estimates were computed in each sex-seperately using sex-specific SNP-exposure and SNP-outcome results and then meta-analyzed).
-- After adjusting for multiple tests (p < `0.05/`r num_tests_by_PCs``), identified `r dim(AM_result_sig)[1]` significant assortative mating MR results (corresponding to the table below).
+- After adjusting for multiple tests (p < `0.05/`r num_tests_by_PCs``), identified `r dim(AM_result_sig)[1]` significant assortative mating (same-trait) MR results (corresponding to the table below)..
 
@@
 
-### Sex heterogeneity.
+## Sex heterogeneity.
 
@@
 
-Figures compared the male to female MR results, and vice versa are shown below. *Suggests evidence of regression dilution bias?*
-
-```{r sex_het_fig, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
-household_MR_AM_FvsM_fig[[1]]
-household_MR_AM_FvsM_fig[[2]]
-```
-
-### Binned results.
+## Binned results (exploring effect of age and time-together).
 
@@
 
-slope_sig_age <- household_MR_binned_het_joint %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
+slope_sig_age <- household_MR_binned_het_SNPmeta %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
   filter(grouping_var == "age_even_bins") %>% 
@@
 
-slope_sig_tt <- household_MR_binned_het_joint %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
+slope_sig_tt <- household_MR_binned_het_SNPmeta %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
   filter(grouping_var == "time_together_even_bins") %>% 
@@
 
-slope_wt_sig_age <- household_MR_binned_het_joint %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
+slope_wt_sig_age <- household_MR_binned_het_SNPmeta %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
   filter(grouping_var == "age_even_bins") %>% 
@@
 
-slope_wt_sig_tt <- household_MR_binned_het_joint %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
+slope_wt_sig_tt <- household_MR_binned_het_SNPmeta %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
   filter(grouping_var == "time_together_even_bins") %>% 
@@
 
-```{r binned_results, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
+Given that we have no significant results, how should we report these analyses?
 
-full_binn <- rbind.fill(household_MR_binned_het_sex_spec, household_MR_binned_het_joint) %>% filter(same_trait)
-full_binn.DT <- full_binn %>% dplyr::select(exposure_ID, exposure_sex, grouping_var, IVW_beta, IVW_pval, sex_het_pval, bin_slope_beta,
-                                            bin_slope_pval,bin_slope_sex_het_pval, bin_slope_sex_het_pval_wt) %>% unique() %>% arrange(exposure_ID)
+A few potential figures we could include for a given trait are shown below. 
+```{r binned_figures, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, include = TRUE}
 
-full_binn.DT <- replace_Neale_ID(full_binn.DT, traits_corr2_filled, "exposure_ID", "outcome_description")
+tar_load(exposure_info)
+for(i in 1:length(exposure_info)){
+  temp <- exposure_info[[i]]
+  if(temp[which(temp$Value=="trait_ID"), "Info"] == "130_irnt") {break}
+}
 
-format_round_cols <- colnames(dplyr::select_if(full_binn.DT, is.numeric))
+figs <- tar_read(household_MR_binned_AM_figs, branches = i)[[1]]
 
-DT::datatable(full_binn.DT, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)
+figs[[1]]
+figs[[2]]
+figs[[3]]
+figs[[4]]
+figs[[5]]
+figs[[6]]
+figs[[7]]
+figs[[8]]
+figs[[9]]
 
@@
 
-### Summary.
+# Heterogeneity stats
 
-- From the same-trait analyses we unforutantely are not left with many significant results after adjusting for multiple tests. 
-- The plot comparing male to female estimates does seem to be an effect of regression dilution bias, since the inverse plot suggests the inverse conclusion? Am I interpretting this right? 
-- Our most significant results are the binned results by age, where we have 4 traits that show significant differences according to the slope (the result is the same regardless of using the weighted or unweighted beta estimates), but none of the joint results are BF significant.
-The four traits are: `r sex_het_wt_sig_age_traits`.
+- In the single trait analysis, we also decided to investigate Cochran’s heterogeneity Q-stat to identify traits with high heterogeneity.
+- This would indicate that associations between the index genotype and partner’s phenotype may not only act indirectly through causal relationship between the traits, but some direct effect is present.
+- Ideally, for the majority of the traits, the heterogeneity would be low, confirming that (despite what Tenesa suggests) there are no direct index genome to partner phenome effects.
+- We tested heterogeneity statistics using the function [mr_heterogeneity](https://mrcieu.github.io/TwoSampleMR/reference/mr_heterogeneity.html) from the [TwoSampleMR](https://mrcieu.github.io/TwoSampleMR/index.html) package.
 
-## Confounder analysis with added variables.
+```{r, het_stots, include = FALSE}
+sig_het <- household_MR_summary_AM %>% filter(Het_IVW_Qpval_round < 0.05/num_tests_by_PCs)
+```
 
-Repeated the confounder analysis with the following 4 traits:
+The number of traits with significant heterogeneity in the MR is: `r dim(sig_het)[1]`.
+
+# Impact of confounders on couple correlation.
+
+We tested the impact of confounding on genetic PCs, N/E coordinates and the following 4 traits:
 
@@
 confounder_trait_tibble_descript %>% 
-  kbl() %>%
+  kbl(col.names = NULL) %>%
   kable_styling()
@@
 
-------------------------------------------------------------------------
+## Ratio of confounding to correlation.
 
-Below are the corresponding figures:
+We tested the ratio of confounding to correlation for each confounder trait tested.
+
+```{r corr_due_to_confounding_ratio, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
+
+corr_impact_by_coords_ratio <- corr_impact_by_coords %>% 
+  dplyr::select(Neale_pheno_ID, trait_couple_corr, trait_couple_corr_se, corr_due_to_confounding_all, corr_due_to_confounding_all_se) %>%
+  unique() %>%
+  mutate(trait_interest = "household_coords") %>%
+  mutate(trait_ratio = corr_due_to_confounding_all/trait_couple_corr)
+
+corr_impact_by_PCs_ratio <- corr_impact_by_PCs %>% 
+  dplyr::select(Neale_pheno_ID, trait_couple_corr, trait_couple_corr_se, corr_due_to_confounding_all, corr_due_to_confounding_all_se) %>%
+  unique() %>%
+  mutate(trait_interest = "PCs") %>%
+  mutate(trait_ratio = corr_due_to_confounding_all/trait_couple_corr)
+
+corr_impact_by_traits_ratio <- corr_impact_by_traits %>% group_by(trait_interest) %>% 
+  dplyr::select(trait_j, trait_interest, trait_j_couple_corr, trait_j_couple_corr_se, 
+                corr_due_to_confounding, corr_due_to_confounding_se) %>% 
+  mutate(trait_ratio = corr_due_to_confounding/trait_j_couple_corr)
+
+corr_impact_by_traits_ratio <- replace_Neale_ID(corr_impact_by_traits_ratio, traits_corr2_filled, "trait_interest", "outcome_description") %>%
+  dplyr::rename(trait_interest = outcome_description)
+
+corr_impact_ratios <- rbind(corr_impact_by_PCs_ratio, corr_impact_by_coords_ratio, corr_impact_by_traits_ratio)
+corr_impact_ratios_summary <- corr_impact_ratios %>%
+  group_by(trait_interest) %>%
+  dplyr::summarise(mean_ratio = mean(trait_ratio), median_ratio = median(trait_ratio), sd_ratio = sd(trait_ratio), number_of_traits = n())
+
+corr_impact_ratios_summary %>%
+  kbl() %>%
+  kable_styling()
+
+```
+
+*A note on the number of traits:*
+
+We tested the impact of `trait_interest` on couple correlation amongst all traits in our couple MR framework (number of traits = `r dim(household_MR_summary_AM)[1]`) with the exception of home location and birth place coordinates (2 each [N and E] * 2), we also removed the trait itself from the confounder analysis where applicable (in the case of traits `20016_irnt`, `738` and `845`).
+[Trait `189_irnt` was not included in our final couple MR analysis but could still be evaluated here because no MR-estimates were used.]
+
+## Plots.
+
+The figure below shows the 4 trait which showed the greatest impact on couple correlation (mean confounding ratio > 0.01).
 
 ```{r confounder_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
+
+geography_plots <- list()
 
@@
 
+  if(trait=="738"){trait_description <- "Average total household \nincome before tax"}
+  if(trait=="189_irnt"){trait_description <- "Townsend deprivation index \nat recruitment"}
+
   plot_data <- corr_impact_by_traits %>% filter(trait_interest==trait) %>% filter(trait_interest!=trait_j)
   plot_data <- replace_Neale_ID(plot_data, traits_corr2_filled, "trait_j", "Phenotype")
-  plot <- ggplot2::ggplot(data = plot_data, ggplot2::aes(x = trait_j_couple_corr,
-                                                         y = corr_due_to_confounding, Phenotype = Phenotype)) +
-    geom_point(alpha = 3/4) +
+  plot <-
+    ggplot2::ggplot(
+      data = plot_data,
+      ggplot2::aes(x = trait_j_couple_corr,
+                   y = corr_due_to_confounding, Phenotype = Phenotype)
+    ) +
+    theme(plot.title = element_text(hjust = 0.5)) +
+    geom_point(alpha = 3 / 4) +
 
-    geom_smooth(mapping = aes(x = trait_j_couple_corr, y = corr_due_to_confounding), method = "lm", se=FALSE, formula = y~x+0, fullrange=TRUE, color = custom_col[4]) +
+    ggplot2::geom_errorbar(
+      ggplot2::aes(ymin = corr_due_to_confounding - corr_due_to_confounding_se, ymax = corr_due_to_confounding + corr_due_to_confounding_se),
+      colour = "grey",
+      width = 0
+    ) +
+    ggplot2::geom_errorbar(
+      ggplot2::aes(xmin = trait_j_couple_corr - trait_j_couple_corr_se, xmax = trait_j_couple_corr + trait_j_couple_corr_se),
+      colour = "grey",
+      width = 0
+    ) +
+
     theme_minimal() +
-    ggplot2::labs(x = paste0("Couple trait correlation"),
-                  y = paste0("Couple correlation due to confounding"),
-                  title = paste0("Impact of ", trait_description, " on couple correlation"))
+    ggplot2::labs(
+      x = paste0("Couple trait correlation"),
+      y = paste0("Couple correlation \ndue to confounding"),
+      title = paste0(trait_description)
+      #title = paste0("Impact of ", trait_description, "\n on couple correlation")
+    ) +
+    theme(plot.title = element_text(size=11, hjust = 0.5)) + theme(axis.title = element_text(size = 10))
 
@@
 
-ggplotly(plots[[1]], tooltip = c("x", "y", "Phenotype"))
-ggplotly(plots[[2]], tooltip = c("x", "y", "Phenotype"))
-ggplotly(plots[[3]], tooltip = c("x", "y", "Phenotype"))
-ggplotly(plots[[4]], tooltip = c("x", "y", "Phenotype"))
+figures_grid <- plot_grid(plotlist = plots, nrow=2, ncol = 2, byrow = F, labels="AUTO")
+figures_grid
+```
+
+# Comparison of paths from index to partner.
+
+As a recap, the model is below.
+
+```{r YiYp_adj_model, echo = FALSE}
+library(knitr)
+knitr::include_graphics("assets/two_trait_model_yiyp_adj.png", error = FALSE)
+```
+
+We sought to compare the different paths from $X_i$ to $Y_p$ using the estimates computed as $\omega$, $\rho$ and $\gamma$.
+
+```{r proxyMR_figure2, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600}
+proxyMR_prod_comparison_fig_yiyp_adj
+```
+
+We considered the regression of the following estimates vs $\omega$ with an intercept passing through the origin (i.e. $estimate \sim \omega + 0$:
+
+1. $\rho$
+2. $\gamma$
+3. $\rho$ + $\gamma$
+4. $\rho_{resid} + \gamma$, where $\rho$  is residualized for $\gamma$ (i.e. $\rho_{resid} = resid(lm(\rho \sim \gamma))$)
+
+The results are summarized below:
+
+```{r regression_summary, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
+
+data <- proxyMR_comparison_summary_yiyp_adj_SNPmeta
+data$rho_resid <- resid(lm(rho_beta ~ gam_beta, data = data))
+data$gam_rho_resid <- data$rho_resid + data$gam_beta
+
+result <- numeric()
+for(outcome in c("rho", "gam", "gam_rho", "gam_rho_resid")){
+
+  col_to_rename <- paste0(outcome, "_beta")
+  if(outcome=="gam_rho_resid") {col_to_rename <- outcome}
+  data_model <- data %>% dplyr::rename(outcome = !!col_to_rename)
+  model <- summary(lm(outcome ~ omega_beta + 0, data = data_model)) 
+  r2 <- model %>% glance %>% pull(r.squared)
+  model_summary <- model %>% tidy() %>% 
+    mutate(term = "omega") %>% mutate(outcome = outcome) %>% mutate(r.squared = r2)
+
+  result <- rbind(result, model_summary)
+
+}
+
+result %>% dplyr::select(outcome, everything()) %>%
+  kbl() %>%
+  kable_styling()
 
 ```
+
+```{r remove_opp_sign, include = FALSE}
+
+opp_sign_rows <- numeric()
+for(col1 in c("gam_beta", "rho_beta", "gam_rho_beta", "gam_rho_resid", "omega_beta")){
+
+  for(col2 in c("gam_beta", "rho_beta", "gam_rho_beta", "gam_rho_resid", "omega_beta")){
+
+      temp <- which(sign(data[[col1]])!=sign(data[[col2]]))
+      opp_sign_rows <- c(opp_sign_rows, temp)
+  }
+
+}
+
+opp_sign_rows <- unique(opp_sign_rows)
+
+```
+
+To further explore these estimates, performed paired t-tests to find the significant differences between pairs of estimates.
+We first removed all instances where the sign did not match between any two pairings.
+In total, of the `r dim(data)[1]` trait pairs under consideration, `r length(opp_sign_rows)` were removed.
+
+```{r boxplot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
+
+df <- data[-opp_sign_rows,] %>% dplyr::select(exposure_ID, outcome_ID, gam_beta, rho_beta, gam_rho_beta, gam_rho_resid, omega_beta)
+
+data_boxplot  <- df %>%
+  pivot_longer(cols = !c("exposure_ID", "outcome_ID"), names_to = "data", values_to = "beta") %>%
+  mutate_at("data", as.factor) %>%
+  mutate(
+    data = fct_recode(as.factor(data),
+                      Gamma = "gam_beta",
+                      Rho = "rho_beta",
+                      Gamma_Rho = "gam_rho_beta",
+                      Gamma_Rho_resid = "gam_rho_resid",
+                      Omega = "omega_beta")
+  )
+
+pairwise.t.test(data_boxplot$beta, data_boxplot$data, p.adjust = "none")
+
+cmpr <- list(c("Omega", "Gamma_Rho"), c("Gamma","Gamma_Rho"), c("Gamma_Rho","Gamma_Rho_resid"), c("Rho", "Gamma_Rho"))
+
+position <- data_boxplot %>% group_by(data) %>% dplyr::summarise(median=median(beta)) %>% 
+  arrange(-desc(median)) %>% pull(data)
+
+data_boxplot$data <- factor(data_boxplot$data,levels=position)
+
+boxplot <- data_boxplot %>%
+  ggplot( aes(x=beta, y=data)) +
+  geom_boxplot() +
+  scale_fill_viridis(discrete = TRUE, alpha=0.6, option="A") +
+  theme_ipsum() +
+  theme(
+    legend.position="none",
+    plot.title = element_text(size=11)
+  ) +
+  labs(x ="Estimate", y = "") +
+  scale_y_discrete(labels=c("Gamma_Rho" = "\u03B3 + \u03C1", "Omega" = "\u03C9",
+                            "Gamma" = "\u03B3", "Gamma_Rho_resid" = expression( paste(rho ["resid"], " + ", gamma)), "Rho" = "\u03C1")) +
+  stat_compare_means(comparisons = cmpr, tip.length=0.01,
+                     label = "p.signif",
+                     symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1),
+                                        symbols = c("****", "***", "**", "*", "ns")))
+
+boxplot
+
+```

### 2022-02-03 analysis/manuscript.Rmd
wflow_publish("analysis/manuscript.Rmd")

@@
 
-# Heterogeneity stats
+# Heterogeneity stats.
 
@@
 ```
+
+## Paths summary.
+
+A summary of the $\omega$, $\rho$, $\gamma$ and $\rho_{resid} + \gamma$ estimates are summarized in the table below.
+Those with BF-significant differences between the estimates (according to a Z-test, p < 0.05/`r dim(data)[1]`) are given as T/F in the last three columns.
+
+```{r paths_summary, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
+
+paths_summary <- data %>% dplyr::select(exposure_description, outcome_description, gam_beta, rho_beta, omega_beta, gam_rho_resid, gam_vs_rho_BF_sig, omega_vs_rho_BF_sig, omega_vs_gam_BF_sig)
+
+format_round_cols <- colnames(dplyr::select_if(paths_summary, is.numeric))
+
+DT::datatable(paths_summary, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)
+
+```

### 2022-02-03 analysis/manuscript.Rmd
wflow_publish("analysis/manuscript.Rmd")

@@
 
-paths_summary <- data %>% dplyr::select(exposure_description, outcome_description, gam_beta, rho_beta, omega_beta, gam_rho_resid, gam_vs_rho_BF_sig, omega_vs_rho_BF_sig, omega_vs_gam_BF_sig)
+paths_summary <- data %>% dplyr::select(exposure_description, outcome_description, gam_beta, rho_beta, omega_beta, gam_rho_resid, gam_vs_rho_pval, omega_vs_rho_pval, omega_vs_gam_pval)
 

### 2022-02-05 R/functions.R
prune potential trait confounders list and proxyMR summary list

@@ -3240,9 +3240,12 @@ compare_mr_raw_corr <- function(exposure_info, household_MR_binned_joint_std, tr
 
 }
 
-find_potential_trait_confounders <- function(Neale_pheno_ID, Neale_pheno_ID_corr, standard_MR_summary_SNPmeta, household_MR_summary_SNPmeta, traits_corr, num_tests_by_PCs){
+find_potential_trait_confounders <- function(Neale_pheno_ID, Neale_pheno_ID_corr, standard_MR_summary_SNPmeta,
+                                             household_MR_summary_SNPmeta, traits_corr, num_tests_by_PCs, corr_mat_traits, corr_trait_threshold){
 
   # here `Neale_pheno_ID` is used as the outcome_ID, because we are trying to find potential confounders that have an impact on this `Neale_pheno_ID`.
+  # `Neale_pheno_ID_corr` refers to the couple correlation for `Neale_pheno_ID`.
+
   phes_ID <- gsub("_irnt", "", Neale_pheno_ID)
 
   pheno_i_MR <- standard_MR_summary_SNPmeta %>% filter(outcome_ID==Neale_pheno_ID)
@@ -3251,14 +3254,49 @@ find_potential_trait_confounders <- function(Neale_pheno_ID, Neale_pheno_ID_corr
     rename(exposure_ID_AM_IVW_beta = IVW_beta) %>%
     rename(exposure_ID_AM_IVW_se = IVW_se)
 
-  ## correlation due to confounding
+  corr_mat_traits_sub <- as.data.frame(corr_mat_traits) %>% dplyr::select(!!phes_ID) %>% rownames_to_column(var = "exposure_ID") %>% as_tibble() %>%
+    dplyr::rename(between_trait_correlation = !!phes_ID)
+
+  ## calculate correlation due to confounding
+  ## filter to only confounders with correlation with index trait < `corr_trait_threshold`
+  ## filter to only confounders with
   output <- pheno_i_MR %>% dplyr::select(exposure_ID, outcome_ID, exposure_description, outcome_description, IVW_beta, IVW_se, IVW_pval) %>%
     left_join(AM_MR, by = c("exposure_ID" = "exposure_ID")) %>%
[…]

### 2022-02-05 _targets.R

@@ -16,7 +16,7 @@ tar_option_set(
   ),
   packages = c("tidyverse", "data.table", "cutr", "ukbtools", "rbgen", "bigsnpr", "TwoSampleMR",
                "ggplot2", "purrr", "rmeta", "PASWR2", "cowplot", "meta", "strex", "RColorBrewer",
-               "forestplot", "R.utils", "MendelianRandomization"),
+               "forestplot", "R.utils", "MendelianRandomization", "rstatix"),
   error = "workspace",
   memory = "transient",
   storage = "worker",
@@ -628,7 +628,7 @@ list(
   tar_target(
     corr_potential_trait_confounders,
     find_potential_trait_confounders(Neale_pheno_ID=couple_MR_vs_trait_corr_sig$exposure_ID, Neale_pheno_ID_corr=couple_MR_vs_trait_corr_sig$couple_r,
-                                     standard_MR_summary_SNPmeta, household_MR_summary_SNPmeta, traits_corr, num_tests_by_PCs),
+                                     standard_MR_summary_SNPmeta, household_MR_summary_SNPmeta, traits_corr, num_tests_by_PCs, corr_mat_traits),
     pattern = map(couple_MR_vs_trait_corr_sig)
   ),
 
@@ -883,6 +883,11 @@ list(
   tar_target(
     proxyMR_comparison_summary_yiyp_adj_SNPmeta, ## compare the effects
     summarize_proxyMR_comparison_SNPmeta(proxyMR_comparison_yiyp_adj_SNPmeta, traits_corr2_filled)
+  ),
+
+  tar_target(
+    proxyMR_comparison_summary_yiyp_adj_SNPmeta_prune, ##  if there is a pair A-B and C-D and if max(corr(A,C)*corr(B,D),corr(A,D)*corr(B,C)) > `corr_trait_threshold` [0.8]
+    prune_proxyMR_comparison_SNPmeta(proxyMR_comparison_summary_yiyp_adj_SNPmeta, household_MR_summary_BF_sig, corr_mat_traits, corr_trait_threshold)
   ),
 
   ## Add z's to proxyMR

### 2022-02-05 analysis/manuscript.Rmd

@@
 boxplot
-
 ```
@@
 
-DT::datatable(paths_summary, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)
+format_round_cols <- c(colnames(dplyr::select(paths_summary, ends_with("_beta"))), "gam_rho_resid")
+format_signif_cols <- colnames(dplyr::select(paths_summary, ends_with("_pval")))
+
+DT::datatable(paths_summary, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3) %>%
+  formatSignif(columns = format_signif_cols, digits = 3)
 

### 2022-02-05 _targets.R
fix minor typo

@@ -628,7 +628,7 @@ list(
   tar_target(
     corr_potential_trait_confounders,
     find_potential_trait_confounders(Neale_pheno_ID=couple_MR_vs_trait_corr_sig$exposure_ID, Neale_pheno_ID_corr=couple_MR_vs_trait_corr_sig$couple_r,
-                                     standard_MR_summary_SNPmeta, household_MR_summary_SNPmeta, traits_corr, num_tests_by_PCs, corr_mat_traits),
+                                     standard_MR_summary_SNPmeta, household_MR_summary_SNPmeta, traits_corr, num_tests_by_PCs, corr_mat_traits, corr_trait_threshold),
     pattern = map(couple_MR_vs_trait_corr_sig)
   ),
 

### 2022-02-08 R/functions.R
calc gam_rho_resid in `summarize_proxyMR_comparison_SNPmeta` and update figure to include this estimate vs omega

@@ -5213,7 +5213,32 @@ summarize_proxyMR_comparison_SNPmeta <- function(proxyMR_comparison, traits_corr
   comparison_result_plus_cat <- left_join(comparison_result, traits_corr2_filled %>% dplyr::select(Neale_pheno_ID, category) %>% rename(exposure_category = category) %>% unique(), by = c("exposure_ID" = "Neale_pheno_ID")) %>%
     left_join(traits_corr2_filled %>% dplyr::select(Neale_pheno_ID, category) %>% rename(outcome_category = category) %>% unique(), by = c("outcome_ID" = "Neale_pheno_ID"))
 
-  return(comparison_result_plus_cat)
+  data <- comparison_result_plus_cat
+
+  model <- lm(rho_beta ~ gam_beta, data = data)
+
+  data$rho_resid <- resid(model)
+
+  a <- summary(model)$coefficients[2,1]
+  b <- summary(model)$coefficients[1,1]
+  # below is the same thing as `data$rho_resid`
+  # rho_resid <- data$rho_beta - data$gam_beta*a - b
+  data$gam_rho_resid <- data$rho_resid + data$gam_beta
+
+  cor_rho_gamma <- cor(data$rho_beta,data$gam_beta)
+
+  data$rho_var <- data$rho_se^2
+  data$gam_var <- data$gam_se^2
+
+  data$gam_rho_resid <- data$rho_resid + data$gam_beta
+  data$gam_rho_resid_var <- data$rho_var + data$gam_var *(1-a)^2 + 2*data$rho_se*(1-a)*data$gam_se*cor_rho_gamma
+  data$gam_rho_resid_se <- sqrt(data$gam_rho_resid_var)
+
+  output <- data %>% mutate(omega_vs_gam_rho_resid_pval = z.test_p(gam_rho_resid, gam_rho_resid_se, omega_beta, omega_se)) %>%
+    mutate(omega_vs_gam_rho_resid_BF_sig = case_when(TRUE ~ omega_vs_gam_rho_resid_pval < 0.05/num_result,
+                                           TRUE ~ TRUE))
+
+  return(output)
 
 }
 
@@ -5308,6 +5333,7 @@ create_proxy_prod_comparison_fig_ind <- function(data, exposure_sex, x, y, overl
   if(x == "gam_beta" | x == "gam_meta_beta"){x_label <- "\u03B3"}
   if(x == "rho_beta" | x == "rho_meta_beta"){x_label <- "\u03C1"}
[…]

### 2022-02-08 analysis/manuscript.Rmd

@@
 ```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
+
 tar_load(couple_MR_vs_trait_corr)
-
 tar_load(corr_impact_by_PCs)
@@
 4. Calculate the ratio of this correlation ($C$) and the correlation of $X$ in partners ($C/r_x$). 
+5. Remove any `potential_counders` with correlation with `trait_ID` > `r corr_trait_threshold`.
+6. For each `trait_ID` prune remaining `potential_counders` so that all have correlations <  `r corr_trait_threshold`, prioritized by their `corr_due_to_confounding_ratio`.
 
-A summary of the these results can be seen below:
+A summary of the top 5 confounders for each `trait_ID` can be seen below:
 
@@
 sig_confounders <- corr_potential_trait_confounders %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
+  arrange(-corr_due_to_confounding_ratio) %>%
+  mutate(total_potential_confounders = n()) %>%
+  slice_head(n = 5) %>%
   #mutate(potential_confounders = paste0(exposure_description, collapse = "; ")) %>%
-  dplyr::summarize(description = outcome_description[1], n = n(), 
+  dplyr::summarize(description = outcome_description[1],
+                   total_potential_confounders = total_potential_confounders[1],
                    potential_confounders = paste0(exposure_description, collapse = "; "), 
-                   mean_C_ratio = mean(corr_due_to_confounding_ratio)) %>%
+                   max_C_ratio = max(corr_due_to_confounding_ratio), 
+                   min_C_ratio = min(corr_due_to_confounding_ratio)) %>%
   dplyr::rename(trait_ID = outcome_ID)
 
-confounder_summary <- rbind(non_sig_confounders, sig_confounders)
+#confounder_summary <- rbind(non_sig_confounders, sig_confounders)
 
-format_round_cols <- colnames(dplyr::select_if(confounder_summary, is.numeric))
+format_round_cols <- colnames(dplyr::select_if(sig_confounders, is.numeric))
 
-DT::datatable(confounder_summary, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)
+DT::datatable(sig_confounders, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)
 
@@
 
+```{r confounder_effect_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
+
+plot_data <- corr_potential_trait_confounders %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
+  arrange(-corr_due_to_confounding_ratio) %>%
+  mutate(total_potential_confounders = n()) %>%
+  slice_head(n = 1) %>%
+  mutate(corr_minus_MR_est = abs(between_trait_correlation)- abs(IVW_beta))
+
+# plot <- ggplot2::ggplot(data = plot_data, ggplot2::aes(x = Correlation,
+#                                                        y = MR_est, 
+#                                                        label = Phenotype, 
+#                                                        color = factor(diff_p_sig)),) +
+#   geom_point(alpha = 1.2) +
+#   scale_color_manual(values = custom_col[c(1,2)], labels=c("Non-significant","BF-signifncant difference")) +
+#   #geom_text(show.legend = FALSE) +
+#   geom_abline(slope=1, intercept=0) +
+#   
+#   ggplot2::geom_errorbar(
+#     ggplot2::aes(ymin = MR_est - IVW_se, ymax = MR_est + IVW_se),
+#     colour = "grey",
+#     width = 0
+#   ) +
+#   ggplot2::geom_errorbar(
+#     ggplot2::aes(xmin = Correlation - couple_r_se, xmax = Correlation + couple_r_se),
+#     colour = "grey",
+#     width = 0
+#   ) +
+#   
+#   theme_minimal() +
+#   ggplot2::labs(x = paste0("Couple correlation"),
+#                 y = paste("Standardized MR effect")) +
+#   theme(legend.position="bottom") + theme(legend.title=element_blank()) +
+#   geom_text_repel(data = subset(plot_data, label_phenos==TRUE),
+#                   nudge_y = -0.25 - subset(plot_data, label_phenos==TRUE)$MR_est, size = 2,
+#                   direction = "x")
+
+```
 # Effect of sex, age and time together on causal effects in couples.
@@
 A summary of the $\omega$, $\rho$, $\gamma$ and $\rho_{resid} + \gamma$ estimates are summarized in the table below.
-Those with BF-significant differences between the estimates (according to a Z-test, p < 0.05/`r dim(data)[1]`) are given as T/F in the last three columns.
+The differences between the estimates (according to a Z-test) are given in the last three columns.
+Trait pairs were pruned such that:  if there is a pair A-B and C-D and if [max(corr(A,C)*corr(B,D),corr(A,D)*corr(B,C))] > `r corr_trait_threshold` then one pair is dropped.
+Pairs were prioritized based on the $\omega$ p-value.
 
 ```{r paths_summary, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
+
+tar_load(proxyMR_comparison_summary_yiyp_adj_SNPmeta_prune)
+
+data <- proxyMR_comparison_summary_yiyp_adj_SNPmeta_prune
+data$rho_resid <- resid(lm(rho_beta ~ gam_beta, data = data))
+data$gam_rho_resid <- data$rho_resid + data$gam_beta
 

### 2022-02-08 R/functions.R
fix minor typo in `create_proxy_prod_comparison_fig`

@@ -5502,7 +5502,7 @@ create_proxy_prod_comparison_fig <- function(proxyMR_figure_data){
 
   figures <- list()
 
-  proxyMR_figure_data %>% rename(gam_rho_resid_beta = gam_rho_resid)
+  proxyMR_figure_data <- proxyMR_figure_data %>% rename(gam_rho_resid_beta = gam_rho_resid)
   count <- 1
   for(panel in 1:4){
 

### 2022-02-08 R/functions.R
fix x_label in `create_proxy_prod_comparison_fig_ind`

@@ -5333,7 +5333,7 @@ create_proxy_prod_comparison_fig_ind <- function(data, exposure_sex, x, y, overl
   if(x == "gam_beta" | x == "gam_meta_beta"){x_label <- "\u03B3"}
   if(x == "rho_beta" | x == "rho_meta_beta"){x_label <- "\u03C1"}
   if(x == "gam_rho_beta" | x == "gam_rho_meta_beta"){x_label <- "\u03B3 + \u03C1"}
-  if(x == "gam_rho_resid_beta"){x_label <- expression( paste(rho ["resid"], " + ", gamma))}
+  if(x == "gam_rho_resid_beta"){x_label <- expression(paste(rho ["resid"], " + ", gamma))}
 
   if(y == "omega_beta" | y == "omega_meta_beta"){y_label <- "\u03C9"}
   if(y == "rho_beta" | y == "rho_meta_beta"){y_label <- "\u03C1"}
@@ -5362,7 +5362,7 @@ create_proxy_prod_comparison_fig_ind <- function(data, exposure_sex, x, y, overl
     geom_point(data = overlay_data, aes(x=x_plot, y=y_plot), color = custom_col[2]) +
     theme(legend.position="top") + geom_abline(slope=1, intercept=0) + theme(legend.title = element_blank()) +
     #xlab(paste0(x_label, " ", exposure_sex)) + ylab(paste0(y_label, " ", exposure_sex)) +
-    xlab(paste0(x_label)) + ylab(paste0(y_label)) +
+    xlab(x_label) + ylab(paste0(y_label)) +
 
     scale_x_continuous(breaks = seq(0, xmax, by = 0.2), limits = c(0, xmax)) +
     scale_y_continuous(breaks = seq(-2, ymax, by = 0.5)[seq(-2, ymax, by = 0.5) > ylim], limits = c(ylim, ymax))

### 2022-02-11 analysis/setup.Rmd
update Neale directory with newly downloaded files

@@
 mkdir $project_dir/analysis/traitMR/standard_GWAS
+mkdir $project_dir/analysis/traitMR/household_GWAS_rev_filter
+mkdir $project_dir/analysis/traitMR/standard_GWAS_rev_filter
 mkdir $project_dir/analysis/traitMR/standard_MR

### 2022-02-11 code/settings.R

@@ -73,7 +73,7 @@ Neale_output_dir <- "/data/sgg3/data/neale_files"
 
 ## files
 
-Neale_SGG_dir_file_cp <- "data/Neale_SGG_directory_15_07_2021.csv"
+Neale_SGG_dir_file_cp <- "data/Neale_SGG_directory_11_02_2021.csv"
 PHESANT_file_dir_cp <- paste0("data/PHESANT_file_directory_05_10_2021.txt")
 UKBB_pheno_dir_cp <- paste0("data/UKBB_pheno_directory_05_10_2021.csv")
 

### 2022-02-11 code/settings.R
change typo in date of new Neale_SGG_dir_file_cp

@@ -73,7 +73,7 @@ Neale_output_dir <- "/data/sgg3/data/neale_files"
 
 ## files
 
-Neale_SGG_dir_file_cp <- "data/Neale_SGG_directory_11_02_2021.csv"
+Neale_SGG_dir_file_cp <- "data/Neale_SGG_directory_11_02_2022.csv"
 PHESANT_file_dir_cp <- paste0("data/PHESANT_file_directory_05_10_2021.txt")
 UKBB_pheno_dir_cp <- paste0("data/UKBB_pheno_directory_05_10_2021.csv")
 

### 2022-02-11 R/functions.R
change genetic_IDs to numeric for proper merge in run_household_GWAS

@@ -2435,12 +2435,13 @@ household_GWAS_bin <- function(exposure_info, summ_stats, pheno_data, outcome_ID
 
     household_intervals <- levels(household_time[[grouping_var]])
 
-    genetic_IDs <- tibble(IID = as.character(rownames(IV_geno)))
+    #genetic_IDs <- tibble(IID = as.character(rownames(IV_geno)))
 
     ### SHOULD REALLY USE THIS FORM OF GENETIC_IDs where IIDs are numeric.
     ### male ID 4e6 does not merge with pheno_cov when IID is a character.
     ## see dim(pheno_cov) vs. dim(temp1)
-    #genetic_IDs <- tibble(IID = as.numeric(rownames(IV_geno)))
+
+    genetic_IDs <- tibble(IID = as.numeric(rownames(IV_geno)))
 
     # to reduce to only genetic IDs with good genetic data
     temp1 <- merge(pheno_cov,genetic_IDs, by.x=index, by.y="IID")

### 2022-02-11 _targets.R

@@ -11,7 +11,7 @@ options(clustermq.scheduler = "slurm", clustermq.template = "slurm_clustermq.tmp
 tar_option_set(
   resources = tar_resources(
     clustermq = tar_resources_clustermq(template = list(num_cores = 1, account = "sgg",
-                                                        ntasks = 1, partition = "sgg",
+                                                        ntasks = 4, partition = "sgg",
                                                         log_file="/data/sgg2/jenny/projects/proxyMR/proxymr_%a_clustermq.out"))
   ),
   packages = c("tidyverse", "data.table", "cutr", "ukbtools", "rbgen", "bigsnpr", "TwoSampleMR",

### 2022-02-24 analysis/phenotype_selection.Rmd
udpate phenotpye selection

@@
 Using the sex-specific summary statistics, heterogeneity between sexes was calculated.
-IVs that showed significant evidence of heterogeneity between sexes were included (p < 0.05/[number of IVs]).
+IVs that showed significant evidence of heterogeneity between sexes were excluded (p < 0.05/[number of IVs]).
 After filtering IVs with significant heterogeneity, phenotypes were filtered to again have IVs < `r num_IVs_threshold`.
@@
 
-Lastly, left-side body traits (highly correlated with right-side) were removed and all 'qualifications' data (corresponding to field [6138](https://biobank.ctsu.ox.ac.uk/showcase/field.cgi?id=6138)) because data is roughly captured with continuous variables (such as years of education).
+Lastly, we removed several duplicated and redundant phenotypes.
+Specifically, all left-side body traits (highly correlated with right-side) were removed and we also retained only one of the duplicated phenotypes for BMI and weight (retaining data fields 21001 and 21002, respectively).
+Additionally, all “qualifications” data was removed (corresponding to field 6138) due to the availability of finer-scale correlated variables, such as “age completed full time education” (data field 845).
 

### 2022-03-07 analysis/manuscript.Rmd
wflow_publish("analysis/manuscript.Rmd")

--- 2022-02-05 analysis/manuscript.Rmd
@@
 library(broom)
+library(data.table)
+library(xlsx)
 ```
@@
 tar_load(household_MR_binned_het_SNPmeta)
+tar_load(potential_trait_confounders_MVMR)
+tar_load(corr_mat_traits)
 
@@
 
+# Pearson vs. Spearman correlation.
+
+We wanted to ensure that INQT was not significantly impacting the correaltions of each trait, so we also calculated a non-parametric based correaltion (Spearman) and found consistent correlations.
+
+```{r pearson_vs_spearman, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE, warning = FALSE}
+tar_load(Neale_SGG_dir_filt)
+tar_load(traits_corr_spearman)
+tar_load(traits_corr)
+
+Neale_SGG_dir_filt_unique <- Neale_SGG_dir_filt[,c("phenotype", "phenotype_sub", "sgg_phesant_name", "description", "SGG_available", "variable_type")] %>% unique()
+
+pearson_vs_spearman_dat <- inner_join(traits_corr, unique(Neale_SGG_dir_filt_unique[,c("sgg_phesant_name", "variable_type")]), by = c("ID"="sgg_phesant_name")) %>% dplyr::rename(r2_pearson = r2) %>%
+  inner_join(traits_corr_spearman[,c("ID", "r2")]) %>% dplyr::rename(r2_spearman = r2) %>% 
+  as_tibble() %>% type_convert() %>%
+  mutate(r_pearson = sqrt(r2_pearson)) %>% 
+  mutate(r_spearman = sqrt(r2_spearman))
+
+pearson_vs_spearman_plot <- ggplot2::ggplot(data = pearson_vs_spearman_dat, 
+                                            ggplot2::aes(x = r_pearson,
+                                                         y = r_spearman, 
+                                                         color = factor(variable_type))) + 
+  geom_point() +
+  theme_minimal() +
+  ggplot2::labs(x = paste0("Pearson correlation"),
+                y = paste("Spearman correlation")) + 
+  scale_color_manual(name = "Variable type", values = custom_col[c(1,2,3)], labels=c("Binary","Continuous (IRNT)", "Ordinal"))
+
+pearson_vs_spearman_plot
+```
+
 # Relationship between causal effects and raw phenotypic correlation in couples.
@@
 There are many traits where the `correlation > MR effects` including some traits of interest such as standing height, place of birth, among many others.
-A plot comparing the two and a summary table of the significant effects is shown below.
+
+For example, if we take a trait such as place of birth (NC), this is very highly correlated among couples.
+However, the within-couple MR-estimate is smaller than the observed phenotypic correaltion.
+This is the case for many such traits.
+
+```{r pheno_corr_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE, warning = FALSE}
+
+### phenotypic correlation plot for 129_irnt (place of birth)
+
+tar_load(path_pheno_data)
+tar_load(hh_pairs_filter)
+Neale_pheno_ID <- "129_irnt" ## this is the Neale_id, used to be pheno_description
+phes_ID <- gsub("_irnt", "", Neale_pheno_ID)
+
+male_pheno_file <- path_pheno_data[which(endsWith(path_pheno_data, paste0("/phesant/", Neale_pheno_ID, "_male.txt")))]
+female_pheno_file <- path_pheno_data[which(endsWith(path_pheno_data, paste0("/phesant/", Neale_pheno_ID, "_female.txt")))]
+male_pheno_data <- fread(male_pheno_file, header=TRUE, select=c("IID","sex","age", phes_ID))
+female_pheno_data <- fread(female_pheno_file, header=TRUE, select=c("IID","sex","age", phes_ID))
+
+joint_data <- rbind(male_pheno_data, female_pheno_data)
+
+pairs_filter_copy <- hh_pairs_filter
+    pairs_filter_copy <- pairs_filter_copy %>%
+      mutate(HOUSEHOLD_MEMBER1_M = case_when(HOUSEHOLD_MEMBER1_sex == 1 ~ HOUSEHOLD_MEMBER1,
+                                             HOUSEHOLD_MEMBER2_sex == 1 ~ HOUSEHOLD_MEMBER2)) %>%
+
+      mutate(HOUSEHOLD_MEMBER2_F = case_when(HOUSEHOLD_MEMBER1_sex == 0 ~ HOUSEHOLD_MEMBER1,
+                                             HOUSEHOLD_MEMBER2_sex == 0 ~ HOUSEHOLD_MEMBER2)) %>%
+
+      dplyr::select(-HOUSEHOLD_MEMBER1, -HOUSEHOLD_MEMBER2) %>%
+      dplyr::rename(HOUSEHOLD_MEMBER1 = HOUSEHOLD_MEMBER1_M) %>%
+      dplyr::rename(HOUSEHOLD_MEMBER2 = HOUSEHOLD_MEMBER2_F)
+
+pairs_filter_copy$trait1 <- joint_data[[phes_ID]][match(pairs_filter_copy[["HOUSEHOLD_MEMBER1"]], joint_data$IID)]
+pairs_filter_copy$trait2 <- joint_data[[phes_ID]][match(pairs_filter_copy[["HOUSEHOLD_MEMBER2"]], joint_data$IID)]
+
+pheno_corr_dat <- pairs_filter_copy
+
+pheno_corr_plot <- ggplot2::ggplot(data = pheno_corr_dat, ggplot2::aes(x = trait1,
+                                                       y = trait2)) +
+  geom_point(alpha = 1.2, color = "gray") +
+  #geom_text(show.legend = FALSE) +
+  #geom_abline(slope=1, intercept=0) +
+
+  theme_minimal() +
+
+  #geom_text(aes(x=1,y=2,label=""),show.legend=F) +
+  ggplot2::labs(x = paste0("Male place of birth (NC)"),
+                y = paste("Female place of birth (NC)")) + 
+  geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
+  theme_ipsum()
+
+pheno_corr_plot
+
+```
+
+A plot comparing the two effects, and a summary table of the significant effects is shown below.
 
@@
                             "Systolic blood pressure, automated reading", "Standing height", 
-                            "Body fat percentage", "Forced vital capacity (FVC), Best measure","Red blood cell (erythrocyte) count")
+                            "Weight", "Forced vital capacity (FVC), Best measure",
+                            "Basophill count", "Pulse rate, automated reading")
 
-plot_data <- replace_Neale_ID(couple_MR_vs_trait_corr, traits_corr2_filled, "exposure_ID", "outcome_description") %>% 
+corr_vs_MR_effects_dat <- replace_Neale_ID(couple_MR_vs_trait_corr, traits_corr2_filled, "exposure_ID", "outcome_description") %>% 
   mutate(Phenotype = outcome_description) %>% 
@@
                                Phenotype == "Systolic blood pressure, automated reading" ~ "SBP",
+                               Phenotype == "Pulse rate, automated reading" ~ "Pulse rate",
                                Phenotype == "Body fat percentage" ~ "BFP",
+                               Phenotype == "Standing height" ~ "Height",
+                               Phenotype == "Reticulocyte percentage" ~ "Reticulocyte %",
+
+                               #Phenotype == "Weight" & outcome_ID == "21002_irnt" ~ "Weight",
                                Phenotype == "Forced vital capacity (FVC), Best measure" ~ "FVC",
@@
                                TRUE ~ Phenotype)) %>%
+
   dplyr::rename(Correlation = couple_r) %>% dplyr::rename(MR_est = IVW_beta) %>% 
   mutate(diff_p_sig = case_when(diff_p < 0.05/num_tests_by_PCs ~ TRUE, TRUE ~ FALSE)) %>%
-  mutate(label_phenos = case_when(outcome_description %in% interesting_phenotypes ~ TRUE, TRUE ~ FALSE))
+  mutate(outcome_description2 = case_when(outcome_description=="Weight" & outcome_ID == "23098_irnt" ~ "Weight2",
+                                          TRUE ~ outcome_description)) %>%
+  mutate(label_phenos = case_when(outcome_description2 %in% interesting_phenotypes ~ TRUE, TRUE ~ FALSE))
 
@@
 
-plot <- ggplot2::ggplot(data = plot_data, ggplot2::aes(x = Correlation,
+corr_vs_MR_effects_plot <- ggplot2::ggplot(data = corr_vs_MR_effects_dat, ggplot2::aes(x = Correlation,
                                                        y = MR_est, 
@@
   theme_minimal() +
+  theme(legend.text=element_text(size=11)) +
+  #geom_text(aes(x=1,y=2,label=""),show.legend=F) +
   ggplot2::labs(x = paste0("Couple correlation"),
@@
   theme(legend.position="bottom") + theme(legend.title=element_blank()) +
-  geom_text_repel(data = subset(plot_data, label_phenos==TRUE),
-                  nudge_y = -0.25 - subset(plot_data, label_phenos==TRUE)$MR_est, size = 2,
-                  direction = "x")
+  geom_text_repel(data = subset(corr_vs_MR_effects_dat, label_phenos==TRUE),
+                  nudge_y = -0.25 - subset(corr_vs_MR_effects_dat, label_phenos==TRUE)$MR_est, size = 3.25,
+                  direction = "x", show.legend = FALSE )
 
-plot
+corr_vs_MR_effects_plot
 
-plot_data.DT <- plot_data %>% dplyr::rename(IVW_beta = MR_est) %>% 
+corr_vs_MR_effects_sig_dat <- corr_vs_MR_effects_dat %>% dplyr::rename(IVW_beta = MR_est) %>% 
   filter(diff_p < 0.05/num_tests_by_PCs) %>%
@@
   mutate(correlation_larger = ifelse(Correlation > IVW_beta, TRUE, FALSE)) %>% 
-  dplyr::select(outcome_description, Correlation, IVW_beta, IVW_pval, diff_p, mr_corr_diff)
+  arrange(diff_p) %>% dplyr::select(-exposure_ID, -Phenotype, -outcome_description2, -label_phenos, -mr_corr_diff, -diff_p_sig) %>% 
+  dplyr::rename(trait_description = outcome_description) %>%
+  dplyr::rename(trait_ID = outcome_ID) %>%
+  dplyr::rename(couple_r = Correlation) %>%
+  dplyr::select(trait_description, trait_ID, everything())
+
+## prune for correlations < 0.8
+
+corr_vs_MR_effects_sig_dat.DT <- corr_vs_MR_effects_sig_dat %>% 
+  dplyr::select(trait_description, trait_ID, couple_r, IVW_beta, IVW_pval, diff_p, correlation_larger)
+
+#prune_pheno_table(corr_vs_MR_effects_sig_dat, "trait_ID", corr_mat_traits, corr_trait_threshold)
+
+# Could prune the results, but only one trait is removed so seems pointless:
+
+# Result are pruned to only show traits with correlation < `r corr_trait_threshold`.
+(The number of traits before pruning was `r dim(corr_vs_MR_effects_sig_dat)[1]` vs `r dim(corr_vs_MR_effects_sig_dat_pruned.DT)[1]` after pruning).
+```
+
+## Summary of cases where correlation is different from the MR estimate.
+
+The table below displays the `r dim(corr_vs_MR_effects_sig_dat.DT)[1]` traits that have correlation significantly different from the MR-estimate.
+Of the significant difference `r table(corr_vs_MR_effects_sig_dat$correlation_larger)[[1]]` have MR-estimates larger than than correlation, suggesting the presence of a negative confounder.
+
+```{r corr_vs_MR_effects_table, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
+
+format_round_cols <- colnames(dplyr::select_if(corr_vs_MR_effects_sig_dat.DT, is.numeric))
+
+DT::datatable(corr_vs_MR_effects_sig_dat.DT %>% dplyr::select(-trait_ID), width = "100%") %>% formatSignif(columns=format_round_cols, digits = 3)
 
@@
 
-## Summary of cases where correlation > MR estimate.
+## Search for potential confounders to explain cases where correlation differs from MR estimate.
 
-The table below displays the `r dim(plot_data.DT)[1]` traits that have correlation significantly > MR-estimate.
-
-```{r corr_vs_MR_effects_table, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
-
-format_round_cols <- colnames(dplyr::select_if(plot_data.DT, is.numeric))
-
-DT::datatable(plot_data.DT, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)
-```
-
-## Search for potential confounders to explain cases where correlation > MR estimate.
-
-1. Restrict to only traits that have correlation significantly > MR-estimate (`r dim(plot_data.DT)[1]` traits have correlation > MR_estimate).
+1. Restrict to only traits that have correlation significantly different from MR-estimate (`r dim(corr_vs_MR_effects_sig_dat)[1]` traits have correlation different from MR_estimate).
 2. For each of these traits, find the traits which have a causal effect on this trait (MR p-val < 0.05/[number of effective tests]).
@@
 4. Calculate the ratio of this correlation ($C$) and the correlation of $X$ in partners ($C/r_x$). 
+5. Remove any `potential_counders` with correlation with `trait_ID` > `r corr_trait_threshold`.
+6. For each `trait_ID` prune remaining `potential_counders` so that all have correlations <  `r corr_trait_threshold`, prioritized by their `corr_due_to_confounding_ratio`.
 
-A summary of the these results can be seen below:
+A summary of the top 5 confounders for each `trait_ID` can be seen below:
+
+```{r simple_C_model, echo = FALSE}
+library(knitr)
+knitr::include_graphics("assets/simple_confounder_model.png", error = FALSE)
+
+```
 
@@
 
-number_of_potential_confounders <- dim(corr_potential_trait_confounders)[1]/dim(plot_data.DT)[1]
+number_of_potential_confounders <- dim(corr_potential_trait_confounders)[1]/dim(corr_vs_MR_effects_sig_dat)[1]
 
@@
 sig_confounders <- corr_potential_trait_confounders %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
+  arrange(-corr_due_to_confounding_ratio) %>%
+  mutate(total_potential_confounders = n()) %>%
+  dplyr::summarize(description = outcome_description[1],
+                   total_potential_confounders = total_potential_confounders[1],
+                   potential_confounders = paste0(exposure_description, collapse = "; "), 
+                   max_C = max(corr_due_to_confounding), 
+                   min_C = min(corr_due_to_confounding)) %>%
+  dplyr::rename(trait_ID = outcome_ID) %>%
+  dplyr::rename(trait_description = description) %>% 
+  dplyr::select(trait_description, trait_ID, everything())
+
+sig_confounders_top5 <- corr_potential_trait_confounders %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
+  arrange(-corr_due_to_confounding_ratio) %>%
+  mutate(total_potential_confounders = n()) %>%
+  slice_head(n = 5) %>%
   #mutate(potential_confounders = paste0(exposure_description, collapse = "; ")) %>%
-  dplyr::summarize(description = outcome_description[1], n = n(), 
-                   potential_confounders = paste0(exposure_description, collapse = "; "), 
-                   mean_C_ratio = mean(corr_due_to_confounding_ratio)) %>%
-  dplyr::rename(trait_ID = outcome_ID)
+  dplyr::summarize(trait_description = outcome_description[1],
+                   total_potential_confounders = total_potential_confounders[1],
+                   top5_potential_confounders = paste0(exposure_description, collapse = "; "), 
+                   max_C = max(corr_due_to_confounding), 
+                   min_C = min(corr_due_to_confounding)) %>%
+  dplyr::rename(trait_ID = outcome_ID) %>%
+  left_join(corr_vs_MR_effects_sig_dat.DT, by = c("trait_description", "trait_ID")) %>%
+  dplyr::select(trait_description, trait_ID, couple_r, IVW_beta, IVW_pval,everything())
 
-confounder_summary <- rbind(non_sig_confounders, sig_confounders)
+format_round_cols <- colnames(dplyr::select_if(sig_confounders_top5, is.numeric))
 
-format_round_cols <- colnames(dplyr::select_if(confounder_summary, is.numeric))
+DT::datatable(sig_confounders_top5, width = "100%") %>% formatSignif(columns=format_round_cols, digits = 3)
 
-DT::datatable(confounder_summary, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)
+```
+
+We compared the difference in phenotypic estimate and MR-estimates to the maximum C for each trait, as shown in the figure below:
+
+```{r max_C_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
+
+max_C_data <- corr_potential_trait_confounders %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
+  arrange(-corr_due_to_confounding_ratio) %>%
+  mutate(total_potential_confounders = n()) %>%
+  slice_head(n = 1) %>%
+  left_join(couple_MR_vs_trait_corr %>% 
+              dplyr::rename(outcome_ID_IVW_beta = IVW_beta) %>%
+              dplyr::rename(outcome_ID_IVW_se = IVW_se) %>%
+              dplyr::rename(outcome_ID_couple_r = couple_r) %>% 
+              dplyr::rename(outcome_ID_couple_r_se = couple_r_se) %>% 
+              dplyr::select(outcome_ID, outcome_ID_IVW_beta, outcome_ID_IVW_se, outcome_ID_couple_r, outcome_ID_couple_r_se)) %>%
+  mutate(same_trait_corr_minus_MR_est = abs(outcome_ID_couple_r)- abs(outcome_ID_IVW_beta)) %>%
+
+  dplyr::rename(Z_beta = IVW_beta) %>%
+  dplyr::rename(Z_se = IVW_se) %>%
+  dplyr::rename(Z_pval = IVW_pval) %>%
+  mutate(Z_sq_beta = Z_beta^2) %>%
+  mutate(Z_sq_se = sqrt(4*Z_beta^2*Z_se^2+2*Z_se^4)) %>%
+  dplyr::rename(Z_AM_beta = exposure_ID_AM_IVW_beta) %>%
+  dplyr::rename(Z_AM_se = exposure_ID_AM_IVW_se) %>%
+  mutate(Z_prod = Z_sq_beta*Z_AM_beta) %>%
+  mutate(Z_prod_var = variance_of_product(Z_sq_beta, Z_sq_se, Z_AM_beta, Z_AM_se)) %>%
+  mutate(Z_prod_se = sqrt(Z_prod_var)) %>%
+  mutate(same_trait_corr_minus_MR_est_se = variance_of_sum(outcome_ID_IVW_se, outcome_ID_couple_r_se))
+
+max_C_plot <- ggplot2::ggplot(data = max_C_data, ggplot2::aes(x = corr_due_to_confounding,
+                                                       y = same_trait_corr_minus_MR_est)) +
+  geom_point(alpha = 1.2) +
+  #scale_color_manual(values = custom_col[c(1,2)], labels=c("Non-significant","BF-signifncant difference")) +
+  #geom_text(show.legend = FALSE) +
+  geom_abline(slope=1, intercept=0) +
+
+  ggplot2::geom_errorbar(
+    ggplot2::aes(ymin = same_trait_corr_minus_MR_est - same_trait_corr_minus_MR_est_se, ymax = same_trait_corr_minus_MR_est + same_trait_corr_minus_MR_est_se),
+    colour = "grey",
+    width = 0
+  ) +
+  ggplot2::geom_errorbar(
+    ggplot2::aes(xmin = corr_due_to_confounding - Z_prod_se, xmax = corr_due_to_confounding + Z_prod_se),
+    colour = "grey",
+    width = 0
+  ) +
+
+  theme_minimal() +
+  ggplot2::labs(x = paste0("max C"),
+                y = paste("abs(correlation) - abs(MR-estimate)")) +
+  theme(legend.title=element_blank()) +
+  #geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
+  theme_ipsum()
+
+max_C_plot
+
+```
+
+Finally, we computed a MVMR to sum up the effect of all potential confounders as shown in the DAG below and compared these C-statistics to the difference between phenotypic correlation and MR-estimates.
+
+```{r MVMR_C_model, echo = FALSE}
+library(knitr)
+knitr::include_graphics("assets/MVMR_confounder_model.png", error = FALSE)
+```
+
+```{r MVMR_C_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
+
+discrepentant_traits <- unique(corr_potential_trait_confounders$outcome_ID)
+
+output <- numeric()
+for(i in 1:length(potential_trait_confounders_MVMR)){
+
+  outcome_ID <- discrepentant_traits[i]
+
+  corr_potential_trait_confounders_i <- corr_potential_trait_confounders %>% filter(outcome_ID == !!outcome_ID)
+
+  MVMR_result_org <- potential_trait_confounders_MVMR[[i]]
+
+  MVMR_result_org <- MVMR_result_org %>% mutate(beta_2_var = 4*beta^2*se^2+2*se^4) %>%
+    mutate(beta_2_se = sqrt(beta_2_var)) %>% 
+    left_join(corr_potential_trait_confounders_i %>% dplyr::select(exposure_ID, exposure_ID_AM_IVW_beta, exposure_ID_AM_IVW_se), by = c("exposure" = "exposure_ID")) %>%
+    dplyr::rename(MVMR_Z_beta = beta) %>%
+    dplyr::rename(MVMR_Z_se = se) %>%
+    dplyr::rename(MVMR_Z_pval = pval) %>%
+    mutate(MVMR_Z_sq_beta = MVMR_Z_beta^2) %>%
+    dplyr::rename(MVMR_Z_sq_se = beta_2_se) %>% 
+    dplyr::rename(Z_AM_beta = exposure_ID_AM_IVW_beta) %>%
+    dplyr::rename(Z_AM_se = exposure_ID_AM_IVW_se) %>%
+    mutate(Z_prod = MVMR_Z_sq_beta*Z_AM_beta) %>%
+    mutate(Z_prod_var = variance_of_product(MVMR_Z_sq_beta, MVMR_Z_sq_se, Z_AM_beta, Z_AM_se))
+
+  MVMR_result_nom <- MVMR_result_org %>% filter(MVMR_Z_pval < 0.05)
+  MVMR_result_BF <- MVMR_result_org %>% filter(MVMR_Z_pval < 0.05/dim(MVMR_result_org)[1])
+
+  num_confounders_nom <- dim(MVMR_result_org)[1]
+  num_confounders_BF <- dim(MVMR_result_BF)[1]
+
+  row <- cbind(outcome_ID)
+  for(filter_type in c("nom", "BF")){
+
+    if(filter_type == "nom"){
+      t <- MVMR_result_org %>% filter(MVMR_Z_pval < 0.05)
+    }
+    if(filter_type == "BF"){
+      t <- MVMR_result_org %>% filter(MVMR_Z_pval < 0.05/dim(MVMR_result_org)[1])
+    }
+
+    num_confounders <- dim(t)[1]
+    C_sum <- sum(t$Z_prod)
+    C_sum_var <- sum(t$Z_prod_var)
+    C_sum_se <- sqrt(C_sum_var)
+
+    temp_row <- cbind(num_confounders, C_sum, C_sum_var, C_sum_se)
+    colnames(temp_row) <- paste0(c("num_confounders", "C_sum", "C_sum_var", "C_sum_se"), "_", filter_type)
+    row <- cbind(row, temp_row)
+
+  }
+
+  output <- rbind(output, row)
+
+}
+
+output <- output %>% as_tibble() %>% type_convert() %>% mutate(C_sum_pval = 2*pnorm(-abs(C_sum_nom/C_sum_se_nom)))
+
+MVMR_C_data <- max_C_data %>% left_join(output)
+
+MVMR_C_plot <- ggplot2::ggplot(data = MVMR_C_data, ggplot2::aes(x = C_sum_nom ,
+                                                       y =  same_trait_corr_minus_MR_est)) +
+  geom_point(alpha = 1.2) +
+  #scale_color_manual(values = custom_col[c(1,2)], labels=c("Non-significant","BF-signifncant difference")) +
+  #geom_text(show.legend = FALSE) +
+  #geom_abline(slope=1, intercept=0) +
+
+  # ggplot2::geom_errorbar(
+  #   ggplot2::aes(ymin = MR_est - IVW_se, ymax = MR_est + IVW_se),
+  #   colour = "grey",
+  #   width = 0
+  # ) +
+  # ggplot2::geom_errorbar(
+  #   ggplot2::aes(xmin = Correlation - couple_r_se, xmax = Correlation + couple_r_se),
+  #   colour = "grey",
+  #   width = 0
+  # ) +
+
+  theme_minimal() +
+  ggplot2::labs(x = paste0("MVMR C-sum (p < 0.05)"),
+                y = paste("abs(correlation) - abs(MR-estimate)")) +
+  theme(legend.title=element_blank()) +
+  geom_abline(slope=1, intercept=0)
+
+MVMR_C_plot
 
@@
 
-AM_result_sig.DT <- AM_result_sig %>% dplyr::select(exposure_description, IVW_beta, IVW_pval) %>% unique()
-DT::datatable(AM_result_sig.DT, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval'), digits=3)
+AM_result_sig.DT <- AM_result_sig %>% dplyr::select(exposure_description, exposure_ID, IVW_beta, IVW_se, IVW_pval) %>% unique() %>% 
+  dplyr::rename(trait_description = exposure_description) %>%
+  dplyr::rename(trait_ID = exposure_ID)
 
+DT::datatable(AM_result_sig.DT, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_se', 'IVW_pval'), digits=3)
+
+AM_sig_sex_het_nom <- household_MR_binned_het_sex_spec %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% filter(sex_het_pval < 0.05)
+AM_sig_sex_het_nom <- replace_Neale_ID(AM_sig_sex_het_nom, traits_corr2_filled,"exposure_ID", "exposure_description") %>%
+  dplyr::rename(trait_description = exposure_description) %>%
+  dplyr::rename(trait_ID = exposure_ID) %>%
+  dplyr::select(trait_description, trait_ID, everything())
+
+AM_result_sig_nom.DT <- AM_sig_sex_het_nom %>% dplyr::select(trait_description, trait_ID, exposure_sex, IVW_beta, IVW_se, IVW_pval, sex_het_pval) %>% unique()
 ```
@@
 - A similar adjustment for multiple hypothesis testing was applied as above, resulting in `r num_tests_by_PCs_AM_sig` number of effective tests. 
-- After adjusting for multiple hypothesis testing (p < `0.05/`r num_tests_by_PCs_AM_sig``), `r dim(AM_sex_spec_result_sig)[1]` traits showed significant differences amongst sexes (this changed from before).
-- The table below shows the *nominally significant* results at `p < 0.05`.
+- After adjusting for multiple hypothesis testing (p < `0.05/`r num_tests_by_PCs_AM_sig``), `r dim(AM_sex_spec_result_sig)[1]` traits showed significant differences amongst sexes.
+- The table below shows the *nominally significant* results at `p < 0.05` (of which there are `r dim(AM_result_sig_nom.DT)[1]/2`).
 
@@
 
-AM_sig_sex_het_nom <- household_MR_binned_het_sex_spec %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% filter(sex_het_pval < 0.05)
-AM_sig_sex_het_nom <- replace_Neale_ID(AM_sig_sex_het_nom, traits_corr2_filled,"exposure_ID", "exposure_description")
-AM_result_sig_nom.DT <- AM_sig_sex_het_nom %>% dplyr::select(exposure_description, exposure_sex, IVW_beta, IVW_pval, sex_het_pval) %>% unique()
 DT::datatable(AM_result_sig_nom.DT, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'sex_het_pval'), digits=3)
 
+nominally_sig_sex_differences <- AM_result_sig_nom.DT %>% pivot_wider(names_from = exposure_sex, values_from = c(IVW_beta, IVW_se, IVW_pval)) 
+# write.table(nominally_sig_sex_differences, "docs/manuscript_tables/nominally_sig_sex_differences.txt", row.names = F, sep=",", quote = F)
 ```
 
-## Binned results (exploring effect of age and time-together).
+## MR binned results (exploring effect of age and time-together).
 
@@
 
-sex_het_wt_sig_age_traits <- replace_Neale_ID(sex_het_wt_sig_age, traits_corr2_filled, "exposure_ID", "outcome_description") %>% pull(outcome_description) %>% unique()
 ```
@@
 
-Given that we have no significant results, how should we report these analyses?
+The significant traits correspond to the following traits:
+
+```{r sig_sex_diff, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, include = TRUE}
+
+sex_het_wt_sig_age_traits <- replace_Neale_ID(sex_het_wt_sig_age, traits_corr2_filled, "exposure_ID", "outcome_description") %>% pull(outcome_description) %>% unique()
+
+sex_het_sig_age_sub <- sex_het_sig_age %>% dplyr::select(exposure_ID, exposure_sex, grouping_var, bin_slope_beta, bin_slope_se, bin_slope_pval, bin_slope_sex_het_pval)
+sex_het_sig_tt_sub <- sex_het_sig_tt %>% dplyr::select(exposure_ID, exposure_sex, grouping_var, bin_slope_beta, bin_slope_se, bin_slope_pval, bin_slope_sex_het_pval)
+
+sig_sex_diff_dat <- replace_Neale_ID(rbind(sex_het_sig_age_sub, sex_het_sig_tt_sub), traits_corr2_filled, "exposure_ID", "exposure_description")
+
+DT::datatable(sig_sex_diff_dat, width = "100%") %>% formatSignif(columns=c('bin_slope_beta', 'bin_slope_se', 'bin_slope_pval', 'bin_slope_sex_het_pval'), digits=3)
+
+```
 
 A few potential figures we could include for a given trait are shown below.
+
 ```{r binned_figures, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, include = TRUE}
@@
 tar_load(exposure_info)
+
+branches_to_pull <- numeric()
+for(descriptions in unique(sig_sex_diff_dat$exposure_description))
+{
+
   for(i in 1:length(exposure_info)){
   temp <- exposure_info[[i]]
-  if(temp[which(temp$Value=="trait_ID"), "Info"] == "130_irnt") {break}
+  if(temp[which(temp$Value=="description"), "Info"] == descriptions) {break}
+  }
+  branches_to_pull <- c(i, branches_to_pull)
 }
 
-figs <- tar_read(household_MR_binned_AM_figs, branches = i)[[1]]
+figs <- list()
 
-figs[[1]]
-figs[[2]]
-figs[[3]]
-figs[[4]]
-figs[[5]]
-figs[[6]]
-figs[[7]]
-figs[[8]]
-figs[[9]]
+for(i in branches_to_pull){
+
+  name <- paste0("exposure_", i)
+  figs[[name]] <- tar_read(household_MR_binned_AM_figs, branches = i)[[1]]
+
+}
+
+figs[[1]][[4]]
+figs[[1]][[5]]
+figs[[2]][[4]]
+figs[[2]][[5]]
+
+```
+
+## Phenotypic binned results (exploring effect of age and time-together).
+
+We also tested the phenotypic correlation among the different bins. Below are the traits that show a significant trend either with Pearson's or Spearman's correlation (p < `r 0.05/num_tests_by_PCs`).
+
+```{r, pheno_binned_result, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
+
+tar_load(binned_pheno_corrs)
+
+binned_pheno_corr_summary <- bind_rows(lapply(binned_pheno_corrs, function(x) {x[[2]]}))
+
+binned_pheno_corr_sig <- binned_pheno_corr_summary %>% filter(bin_slope_pval_spearman < 0.05/num_tests_by_PCs | bin_slope_pval_pearson < 0.05/num_tests_by_PCs)
+
+binned_pheno_corr_sig.DT <- replace_Neale_ID(binned_pheno_corr_sig, traits_corr2_filled, "trait", "trait_description") %>%
+  dplyr::rename(trait_ID = trait) %>%
+  dplyr::select(trait_description, trait_ID, everything())
+
+format_round_cols <- colnames(dplyr::select_if(binned_pheno_corr_sig.DT, is.numeric))
+
+DT::datatable(binned_pheno_corr_sig.DT, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)
+
+# write.table(binned_pheno_corr_sig.DT, "docs/manuscript_tables/binned_phenotypic_correlations.txt", row.names = F, sep=",", quote = F)
 
@@
 
-```{r, het_stots, include = FALSE}
+```{r, het_stats, include = FALSE}
 sig_het <- household_MR_summary_AM %>% filter(Het_IVW_Qpval_round < 0.05/num_tests_by_PCs)
@@
   unique() %>%
-  mutate(trait_interest = "household_coords") %>%
+  mutate(confounder_trait = "household_coords") %>%
   mutate(trait_ratio = corr_due_to_confounding_all/trait_couple_corr)
@@
   unique() %>%
-  mutate(trait_interest = "PCs") %>%
+  mutate(confounder_trait = "PCs") %>%
   mutate(trait_ratio = corr_due_to_confounding_all/trait_couple_corr)
 
-corr_impact_by_traits_ratio <- corr_impact_by_traits %>% group_by(trait_interest) %>% 
-  dplyr::select(trait_j, trait_interest, trait_j_couple_corr, trait_j_couple_corr_se, 
+corr_impact_by_traits_ratio <- corr_impact_by_traits %>% dplyr::rename(confounder_trait = trait_interest) %>% 
+  group_by(confounder_trait) %>% 
+  dplyr::select(trait_j, confounder_trait, trait_j_couple_corr, trait_j_couple_corr_se, 
                 corr_due_to_confounding, corr_due_to_confounding_se) %>% 
@@
 
-corr_impact_by_traits_ratio <- replace_Neale_ID(corr_impact_by_traits_ratio, traits_corr2_filled, "trait_interest", "outcome_description") %>%
-  dplyr::rename(trait_interest = outcome_description)
+corr_impact_by_traits_ratio <- replace_Neale_ID(corr_impact_by_traits_ratio, traits_corr2_filled, "confounder_trait", "confounder_trait_description") %>% ungroup() %>%
+  dplyr::select(-confounder_trait) %>% dplyr::rename(confounder_trait = confounder_trait_description)
 
 corr_impact_ratios <- rbind(corr_impact_by_PCs_ratio, corr_impact_by_coords_ratio, corr_impact_by_traits_ratio)
+
 corr_impact_ratios_summary <- corr_impact_ratios %>%
-  group_by(trait_interest) %>%
+  group_by(confounder_trait) %>%
   dplyr::summarise(mean_ratio = mean(trait_ratio), median_ratio = median(trait_ratio), sd_ratio = sd(trait_ratio), number_of_traits = n())
@@
 
+# write.table(corr_impact_ratios_summary, "docs/manuscript_tables/corr_impact_ratios_summary.txt", row.names = F, sep=",", quote = F)
 ```
@@
 
-The figure below shows the 4 trait which showed the greatest impact on couple correlation (mean confounding ratio > 0.01).
+The figure below shows the 4 traits which showed the greatest impact on couple correlation (mean confounding ratio > 0.01).
 
@@
 
+confounder_traits_sub <- confounder_traits[c(2,3,4,5)]
+
 plots <- list()
-for(i in 1:length(confounder_traits)){
+for(i in 1:length(confounder_traits_sub)){
 
-  trait <- confounder_traits[i]
-  trait_description <- confounder_trait_tibble_descript[[1]][i]
+  trait <- confounder_traits_sub[i]
+  trait_description <- confounder_trait_tibble_descript[which(confounder_trait_tibble_descript[["value"]]==trait),"outcome_description"][[1]]
 
@@
   if(trait=="189_irnt"){trait_description <- "Townsend deprivation index \nat recruitment"}
+  if(trait=="6160_1"){trait_description <- "Sports club or gym user"}
 
@@
     ) +
+    xlim(0, 1.0) +
     theme(plot.title = element_text(size=11, hjust = 0.5)) + theme(axis.title = element_text(size = 10))
@@
 
+i <- 5
+
 figures_grid <- plot_grid(plotlist = plots, nrow=2, ncol = 2, byrow = F, labels="AUTO")
 figures_grid
+```
+
+```{r PC_confounder_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
+
+plot_data <- corr_impact_by_PCs %>% dplyr::select(Neale_pheno_ID, trait_couple_corr, corr_due_to_confounding_all, corr_due_to_confounding_all_se, trait_couple_corr_se, trait_couple_corr_se) %>% unique()
+
+plot_PCs <-
+  ggplot2::ggplot(
+    data = plot_data,
+    ggplot2::aes(x = trait_couple_corr,
+                 y = corr_due_to_confounding_all)
+  ) +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  geom_point(alpha = 3 / 4) +
+
+  ggplot2::geom_errorbar(
+    ggplot2::aes(ymin = corr_due_to_confounding_all - corr_due_to_confounding_all_se, ymax = corr_due_to_confounding_all + corr_due_to_confounding_all_se),
+    colour = "grey",
+    width = 0
+  ) +
+  ggplot2::geom_errorbar(
+    ggplot2::aes(xmin = trait_couple_corr - trait_couple_corr_se, xmax = trait_couple_corr + trait_couple_corr_se),
+    colour = "grey",
+    width = 0
+  ) +
+
+  theme_minimal() +
+  ggplot2::labs(
+    x = paste0("Couple trait correlation"),
+    y = paste0("Couple correlation \ndue to confounding"),
+    title = paste0("Principal components")
+    #title = paste0("Impact of ", trait_description, "\n on couple correlation")
+  ) +
+
+  xlim(0, 1.0) +
+  theme(plot.title = element_text(size=11, hjust = 0.5)) + theme(axis.title = element_text(size = 10))
+
+i <- 5
+plots[[i]] <- plot_PCs
+
+```
+
+```{r coord_confounder_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE, out.width = "200%", dpi=600}
+
+plot_data <- corr_impact_by_coords %>% dplyr::select(Neale_pheno_ID, trait_couple_corr, corr_due_to_confounding_all, corr_due_to_confounding_all_se, trait_couple_corr_se, trait_couple_corr_se) %>% unique()
+
+plot_coords <-
+  ggplot2::ggplot(
+    data = plot_data,
+    ggplot2::aes(x = trait_couple_corr,
+                 y = corr_due_to_confounding_all)
+  ) +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  geom_point(alpha = 3 / 4) +
+
+  ggplot2::geom_errorbar(
+    ggplot2::aes(ymin = corr_due_to_confounding_all - corr_due_to_confounding_all_se, ymax = corr_due_to_confounding_all + corr_due_to_confounding_all_se),
+    colour = "grey",
+    width = 0
+  ) +
+  ggplot2::geom_errorbar(
+    ggplot2::aes(xmin = trait_couple_corr - trait_couple_corr_se, xmax = trait_couple_corr + trait_couple_corr_se),
+    colour = "grey",
+    width = 0
+  ) +
+
+  theme_minimal() +
+  ggplot2::labs(
+    x = paste0("Couple trait correlation"),
+    y = paste0("Couple correlation \ndue to confounding"),
+    title = paste0("Place of birth \ncoordinates")
+    #title = paste0("Impact of ", trait_description, "\n on couple correlation")
+  ) +
+  xlim(0, 1.0) +
+  theme(plot.title = element_text(size=11, hjust = 0.5)) + theme(axis.title = element_text(size = 10))
+
+i <- 6
+plots[[i]] <- plot_coords
+
+figures_grid <- plot_grid(plotlist = plots, nrow=3, ncol = 2, byrow = T, labels="AUTO", label_size = 12)
+figures_grid
+
 ```
@@
 
-result %>% dplyr::select(outcome, everything()) %>%
+omega_regression_estimates <- result %>% dplyr::select(outcome, everything())
+
+omega_regression_estimates %>%
   kbl() %>%
@@
 opp_sign_rows <- numeric()
-for(col1 in c("gam_beta", "rho_beta", "gam_rho_beta", "gam_rho_resid", "omega_beta")){
+for(col1 in c("gam_beta", "rho_beta", "gam_rho_resid", "omega_beta")){
 
-  for(col2 in c("gam_beta", "rho_beta", "gam_rho_beta", "gam_rho_resid", "omega_beta")){
+  for(col2 in c("gam_beta", "rho_beta", "gam_rho_resid", "omega_beta")){
 
@@
 
-```{r boxplot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
+```{r boxplot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE, warning = FALSE}
 
-df <- data[-opp_sign_rows,] %>% dplyr::select(exposure_ID, outcome_ID, gam_beta, rho_beta, gam_rho_beta, gam_rho_resid, omega_beta)
+df <- data[-opp_sign_rows,] %>% dplyr::select(exposure_ID, outcome_ID, gam_beta, rho_beta, gam_rho_resid, omega_beta)
 
@@
                       Rho = "rho_beta",
-                      Gamma_Rho = "gam_rho_beta",
                       Gamma_Rho_resid = "gam_rho_resid",
@@
 
-cmpr <- list(c("Omega", "Gamma_Rho"), c("Gamma","Gamma_Rho"), c("Gamma_Rho","Gamma_Rho_resid"), c("Rho", "Gamma_Rho"))
+#cmpr <- list(c("Omega", "Gamma_Rho"), c("Gamma","Gamma_Rho"), c("Gamma_Rho","Gamma_Rho_resid"), c("Rho", "Gamma_Rho"))
 
@@
 
+data_boxplot <- data_boxplot %>% filter(data != "Gamma_Rho")
+
 boxplot <- data_boxplot %>%
-  ggplot( aes(x=beta, y=data)) +
-  geom_boxplot() +
+  ggplot( aes(x=data, y=beta)) +
+  geom_boxplot(outlier.shape = NA) +
+  coord_cartesian(ylim=c(-0.5, 0.5)) +
   scale_fill_viridis(discrete = TRUE, alpha=0.6, option="A") +
@@
   ) +
-  labs(x ="Estimate", y = "") +
-  scale_y_discrete(labels=c("Gamma_Rho" = "\u03B3 + \u03C1", "Omega" = "\u03C9",
+  #labs(x ="Estimate", y = "") +
+  scale_x_discrete(labels=c("Omega" = "\u03C9", 
                             "Gamma" = "\u03B3", "Gamma_Rho_resid" = expression( paste(rho ["resid"], " + ", gamma)), "Rho" = "\u03C1")) +
-  stat_compare_means(comparisons = cmpr, tip.length=0.01,
-                     label = "p.signif",
-                     symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1),
-                                        symbols = c("****", "***", "**", "*", "ns")))
+  theme(axis.title.x=element_blank(),
+        axis.title.y=element_blank())
+
+#+ #"Gamma_Rho" = "\u03B3 + \u03C1",
+
+  # stat_compare_means(comparisons = cmpr, tip.length=0.01,
+  #                    label = "p.signif",
+  #                    symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1),
+  #                                       symbols = c("****", "***", "**", "*", "ns"))) 
+  #
 
 boxplot
+
+figures_grid <- plot_grid(plotlist = list(proxyMR_prod_comparison_fig_yiyp_adj, boxplot), nrow=1, ncol = 2, byrow = F)
+figures_grid
+
 ```
@@
 A summary of the $\omega$, $\rho$, $\gamma$ and $\rho_{resid} + \gamma$ estimates are summarized in the table below.
-Those with BF-significant differences between the estimates (according to a Z-test, p < 0.05/`r dim(data)[1]`) are given as T/F in the last three columns.
+The differences between the estimates (according to a Z-test) are given in the last three columns.
+Trait pairs were pruned such that:  if there is a pair A-B and C-D and if [`max(corr(A,C)*corr(B,D),corr(A,D)*corr(B,C))`] > `r corr_trait_threshold` then one pair is dropped.
+Pairs were prioritized based on the $\omega$ p-value.
 
@@
 
-paths_summary <- data %>% dplyr::select(exposure_description, outcome_description, gam_beta, rho_beta, omega_beta, gam_rho_resid, gam_vs_rho_pval, omega_vs_rho_pval, omega_vs_gam_pval)
+tar_load(proxyMR_comparison_summary_yiyp_adj_SNPmeta_prune)
 
-format_round_cols <- colnames(dplyr::select_if(paths_summary, is.numeric))
+data <- proxyMR_comparison_summary_yiyp_adj_SNPmeta_prune
+data$rho_resid <- resid(lm(rho_beta ~ gam_beta, data = data))
+data$gam_rho_resid <- data$rho_resid + data$gam_beta
+
+paths_summary <- data %>% dplyr::select(exposure_description, outcome_description, gam_beta, rho_beta, omega_beta, gam_rho_resid, gam_vs_rho_pval, omega_vs_rho_pval, omega_vs_gam_pval, omega_vs_gam_rho_resid_pval)
 
@@
 ```
+
+```{r write_tables, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
+
+################################# 
+## Tables to include in paper ###
+#################################
+
+## Cases where correlation is > MR effects: `corr_vs_MR_effects_sig_dat`
+
+## Potential confounders for all cases where corr is > MR effects: `sig_confounders`
+
+## Significant same-trait couple MR effects: `AM_result_sig.DT`
+
+## Nominally significant same-trait sex-difference: `nominally_sig_sex_differences`
+
+## Phenotypic correlations by bin: `binned_pheno_corr_sig.DT`
+
+## Ratio of confounding to correlation: `corr_impact_ratios_summary`
+
+## Omega regression estimates: `omega_regression_estimates`
+
+## Pruend paths summary: `paths_summary`
+
+file_name <- "docs/manuscript_tables/manuscript_tables.xlsx"
+
+write.xlsx_wrapper <- function(dat, ...){
+
+  write.xlsx(as.data.frame(dat), ...)
+}
+
+write.xlsx_wrapper(corr_vs_MR_effects_sig_dat, file=file_name, sheetName="sig_corr_vs_MR", row.names=FALSE)
+write.xlsx_wrapper(sig_confounders, file=file_name, sheetName="sig_confounders", append=TRUE, row.names=FALSE)
+write.xlsx_wrapper(AM_result_sig.DT, file=file_name, sheetName="BF_sig_same_trait_MR", append=TRUE, row.names=FALSE)
+write.xlsx_wrapper(nominally_sig_sex_differences, file=file_name, sheetName="nominally_sig_sex_diff", append=TRUE, row.names=FALSE)
+write.xlsx_wrapper(binned_pheno_corr_sig.DT, file=file_name, sheetName="pheno_binned_corrs", append=TRUE, row.names=FALSE)
+write.xlsx_wrapper(corr_impact_ratios_summary, file=file_name, sheetName="corr_impact_ratios_summary", append=TRUE, row.names=FALSE)
+write.xlsx_wrapper(omega_regression_estimates, file=file_name, sheetName="omega_regression_estimates", append=TRUE, row.names=FALSE)
+write.xlsx_wrapper(paths_summary, file=file_name, sheetName="paths_summary", append=TRUE, row.names=FALSE)
+
+```

### 2022-03-07 analysis/index.Rmd
wflow_publish("analysis/index.Rmd")

@@
 
-## [Project Workflow Status](status.html) {.unnumbered}
+## [Project Workflow Status](project_status.html) {.unnumbered}
 
@@
 Weekly meeting were held roughly once per week or every other week.
+
+## [Manuscript](manuscript.html) {.unnumbered}
+
+Document with a final summary of the analyses and tables for the manuscript.
 
@@
 
-Primary analysis performed by: [Jenny Sjaarda](jennysjaarda@gmail.com).
+Primary analysis performed by: [Jenny Sjaarda](https://wp.unil.ch/sgg/people/jenny-sjaarda/).
 

### 2022-03-07 analysis/todo_list.Rmd
wflow_publish("analysis/todo_list.Rmd")

@@
 title: "To do"
-author: "jennysjaarda"
+author: "Jenny Sjaarda"
 date: "2021-09-08"
@@
 - MVMR. 
-- AM MR summary Rmd.
-- PC correlation among couples summary Rmd. 
-- Clean up update meeting Rmd.
+- ~~AM MR summary Rmd (and publish).~~
+- ~~PC correlation among couples summary Rmd.~~ 
+- ~~Clean up update meeting Rmd.~~
+  - ~~14-06-21~~
+  - ~~05-07-21~~
+  - ~~14-07-21~~
+  - ~~25-08-21~~
+  - ~~06-09-21~~
+- ~~Delete latex_test docs.~~
 
@@
 - ~~Weight trend line by inverse variance.~~ 
-- Update AF in Neale summary stats. 
-- Create couple selection Rmd. 
-- Number of tests for filtering by $\omega$ should exclude those with only one trait. (i.e. number of tests would be $j^2 - j$).
+- ~~Update AF in Neale summary stats.~~
+- ~~Create couple selection Rmd.~~ 
+- ~~Number of tests for filtering by $\omega$ should exclude those with only one trait. (i.e. number of tests would be $j^2 - j$).~~
 - ~~Filter by $\omega$ instead of same-person MR.~~ 
@@
 - Hierarchy figure (add labels).
-- Make visio figures in LucidChart.
-- Explore IV overlap between X and Y:
-    - Determine percentage of IV overlap between X IVs and Y IVs. 
-    - MV MR, add SNPs/trait effect from $X_{i} \rightarrow Y_{p}$ to $Y_{i} \rightarrow Y_{p}$ MR (i.e. run a MV MR of the form: $Y_{p} \sim Y_{i} + X_{i}$).
-- Add script/function which replaces `Neale_ID` with trait description to main functions.
+- ~~Make visio figures in LucidChart.~~
+- ~~Explore IV overlap between X and Y:~~
+    - ~~Determine percentage of IV overlap between X IVs and Y IVs.~~ 
+    - ~~MV MR, add SNPs/trait effect from $X_{i} \rightarrow Y_{p}$ to $Y_{i} \rightarrow Y_{p}$ MR (i.e. run a MV MR of the form: $Y_{p} \sim Y_{i} + X_{i}$).~~
+- ~~Add script/function which replaces `Neale_ID` with trait description to main functions.~~
+- Add status notebook, largely copied from here (https://github.com/rgayler/fa_sim_cal/blob/master/analysis/m_00_status.Rmd).
+
+# 15/09/2021
+
+- ~~Clean up figures for group meeting presentation Rmd.~~
+- ~~Delete RData and rds files.~~
+- MR between partners (partner vs index) Rmd (can take a lot from early July update meeting).
+- MR within same individual Rmd (can take a lot from August update meeting).
+- Compare paths towards outcome in partner Rmd  (can take a lot from August update meeting).
+- ~~Rerun het figures and AM MR summary.~~
+- ~~Add pipeline_overview.Rmd.~~
+
+## 24/09/2021
+
+- PC analysis
+    - Draw a DAG for this, confirm with Zoltan. 
+- ~~Delete `binary_added.csv` file.~~
+- ~~Remove left body trait, dietary traits, re-add binary traits.~~
+- ~~In 2 trait model, restrict to only variables with correlation < 0.8 (play with this a bit until we see that traits are really different).~~
+- ~~Also include X IVs in adjYiYp models in addition to Y IVs.~~
+- Check which unrelated group was using in PC GWAS.

### 2022-03-28 analysis/manuscript.Rmd
wflow_publish("analysis/manuscript.Rmd")

@@
 
+tar_load(Neale_SGG_dir_filt)
+tar_load(traits_corr_spearman)
 tar_load(couple_MR_vs_trait_corr)
@@
 tar_load(corr_impact_by_coords)
-tar_load(corr_potential_trait_confounders)
+tar_load(corr_potential_trait_confounders_pos)
 tar_load(traits_corr)
@@
 tar_load(household_MR_binned_het_SNPmeta)
-tar_load(potential_trait_confounders_MVMR)
+#tar_load(potential_trait_confounders_MVMR)
 tar_load(corr_mat_traits)
+tar_load(corr_impact_by_traits_causal)
+tar_load(household_MR_SNPmeta_sensitivity)
+tar_load(binned_pheno_corrs)
 
@@
 ```{r pearson_vs_spearman, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE, warning = FALSE}
-tar_load(Neale_SGG_dir_filt)
-tar_load(traits_corr_spearman)
-tar_load(traits_corr)
 
@@
 
-pearson_vs_spearman_dat <- inner_join(traits_corr, unique(Neale_SGG_dir_filt_unique[,c("sgg_phesant_name", "variable_type")]), by = c("ID"="sgg_phesant_name")) %>% dplyr::rename(r2_pearson = r2) %>%
-  inner_join(traits_corr_spearman[,c("ID", "r2")]) %>% dplyr::rename(r2_spearman = r2) %>% 
+## looks at all traits *before* the correlation filter
+
+pearson_vs_spearman_dat <-
+  inner_join(traits_corr,
+             unique(Neale_SGG_dir_filt_unique[, c("sgg_phesant_name", "variable_type")]),
+             by = c("ID" = "sgg_phesant_name")) %>%
+  dplyr::rename(r2_pearson = r2) %>%
+  inner_join(traits_corr_spearman[, c("ID", "r2")]) %>% dplyr::rename(r2_spearman = r2) %>%
   as_tibble() %>% type_convert() %>%
@@
 
-pearson_vs_spearman_plot <- ggplot2::ggplot(data = pearson_vs_spearman_dat, 
-                                            ggplot2::aes(x = r_pearson,
+pearson_vs_spearman_plot <-
+  ggplot2::ggplot(data = pearson_vs_spearman_dat,
+                  ggplot2::aes(
+                    x = r_pearson,
                     y = r_spearman,
-                                                         color = factor(variable_type))) + 
+                    color = factor(variable_type)
+                  )) +
   geom_point() +
@@
                 y = paste("Spearman correlation")) +
-  scale_color_manual(name = "Variable type", values = custom_col[c(1,2,3)], labels=c("Binary","Continuous (IRNT)", "Ordinal"))
+  scale_color_manual(
+    name = "Variable type",
+    values = custom_col[c(1, 2, 3)],
+    labels = c("Binary", "Continuous (IRNT)", "Ordinal")
+  )
 
@@
 
-Sought to compare the raw correlation amongst couples vs.
+We sought to compare the raw correlation amongst couples vs.
 the standardized MR effects.
@@
 ```
+
+## Summary of cases where correlation is different from the MR estimate.
 
@@
 
-## Summary of cases where correlation is different from the MR estimate.
-
 The table below displays the `r dim(corr_vs_MR_effects_sig_dat.DT)[1]` traits that have correlation significantly different from the MR-estimate.
-Of the significant difference `r table(corr_vs_MR_effects_sig_dat$correlation_larger)[[1]]` have MR-estimates larger than than correlation, suggesting the presence of a negative confounder.
+Of the significant difference `r table(corr_vs_MR_effects_sig_dat$correlation_larger)[[1]]` have MR-estimates larger than than correlation, suggesting the presence of a negative confounder (this table corresponds to "sig_corr_vs_MR" in the excel document).
 
@@
 
+Sought to identify potential confounders (from our list of traits in the pipeline) which may, in part, explain the discrepant estimates, as follows:
+
 1. Restrict to only traits that have correlation significantly different from MR-estimate (`r dim(corr_vs_MR_effects_sig_dat)[1]` traits have correlation different from MR_estimate).
-2. For each of these traits, find the traits which have a causal effect on this trait (MR p-val < 0.05/[number of effective tests]).
-3. Calculate the correlation due to confounding as: $C = (\alpha_{z\rightarrow x})^2 * \alpha_{z_i\rightarrow z_p}$.
-4. Calculate the ratio of this correlation ($C$) and the correlation of $X$ in partners ($C/r_x$). 
-5. Remove any `potential_counders` with correlation with `trait_ID` > `r corr_trait_threshold`.
-6. For each `trait_ID` prune remaining `potential_counders` so that all have correlations <  `r corr_trait_threshold`, prioritized by their `corr_due_to_confounding_ratio`.
-
-A summary of the top 5 confounders for each `trait_ID` can be seen below:
+1. Remove any `potential_counders` with correlation with `trait_ID` > `r corr_trait_threshold`.
+1.
+For each of the remaining $Z$ `potential_counders` for each trait $X$, find the potential confounders which have a causal effect on trait $X$ ($\alpha_{z\rightarrow x}$ p-val < 0.05/[number of effective tests]).
+1. Split results into those cases where MR-estimate is **greater** than the phenotypic correlation and those that are **less** than the correlation. 
+1.
+For traits where correlation was less than MR-estimate, we restricted to negative confounders (i.e.
+negative $\alpha_{z_i\rightarrow z_p}$), and conversely for traits where the correlation was greater than the MR-estimate we restricted to positive confounders (i.e.
+positive $\alpha_{z_i\rightarrow z_p}$) [note that only one significant MR-effect is negative].
+1. Of the remaining $Z$, we further filtered to those with a significant within couple MR effect ($\alpha_{z_i\rightarrow z_p}$ p < 0.05/[number of remaining $Z$]). 
+1. Calculate the correlation due to confounding as: $C = (\alpha_{z\rightarrow x})^2 * \alpha_{z_i\rightarrow z_p}$.
+1. Calculate the ratio of this correlation ($C$) and the correlation of $X$ in partners ($C/r_x$). 
+1. For each `trait_ID` prune remaining `potential_counders` so that all have correlations <  `r corr_trait_threshold`, prioritized by their `corr_due_to_confounding_ratio`.
 
@@
 knitr::include_graphics("assets/simple_confounder_model.png", error = FALSE)
-
-```
+```
+
+A summary of the top 5 confounders for each `trait_ID` can be seen below (a variation of this table, include *all* potential confounders is included in the excel sheet titled "sig_confounders"):
 
@@
 
-number_of_potential_confounders <- dim(corr_potential_trait_confounders)[1]/dim(corr_vs_MR_effects_sig_dat)[1]
-
-non_sig_confounders <- corr_potential_trait_confounders %>% filter(!sig_confounder) %>% group_by(outcome_ID) %>%
+number_of_potential_confounders <- dim(corr_potential_trait_confounders_pos)[1]/dim(corr_vs_MR_effects_sig_dat)[1]
+
+non_sig_confounders <- corr_potential_trait_confounders_pos %>% filter(!sig_confounder) %>% group_by(outcome_ID) %>%
   dplyr::summarize(description = outcome_description[1], n = n()) %>% 
@@
 
-sig_confounders <- corr_potential_trait_confounders %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
+sig_confounders <- corr_potential_trait_confounders_pos %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
   arrange(-corr_due_to_confounding_ratio) %>%
@@
 
-sig_confounders_top5 <- corr_potential_trait_confounders %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
+sig_confounders_top5 <- corr_potential_trait_confounders_pos %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
   arrange(-corr_due_to_confounding_ratio) %>%
@@
 
-max_C_data <- corr_potential_trait_confounders %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
+max_C_data <- corr_potential_trait_confounders_pos %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
   arrange(-corr_due_to_confounding_ratio) %>%
@@
   mutate(Z_prod_se = sqrt(Z_prod_var)) %>%
-  mutate(same_trait_corr_minus_MR_est_se = variance_of_sum(outcome_ID_IVW_se, outcome_ID_couple_r_se))
+  mutate(same_trait_corr_minus_MR_est_se = sqrt(variance_of_sum(outcome_ID_IVW_se, outcome_ID_couple_r_se)))
 
@@
   ggplot2::labs(x = paste0("max C"),
-                y = paste("abs(correlation) - abs(MR-estimate)")) +
+                y = paste("Correlation - MR-estimate")) +
   theme(legend.title=element_blank()) +
@@
 
-Finally, we computed a MVMR to sum up the effect of all potential confounders as shown in the DAG below and compared these C-statistics to the difference between phenotypic correlation and MR-estimates.
+Lastly, we attempted to compute a MVMR sum of the effects of all potential confounders as shown in the DAG below and compared these C-statistics to the difference between phenotypic correlation and MR-estimates, however the estimates did not make sense so we decided not to pursue this analysis.
 
@@
 
-```{r MVMR_C_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
-
-discrepentant_traits <- unique(corr_potential_trait_confounders$outcome_ID)
+```{r MVMR_C_plot, echo = FALSE, eval = FALSE, message=FALSE}
+
+discrepentant_traits <- unique(corr_potential_trait_confounders_pos$outcome_ID)
 
@@
 
-  corr_potential_trait_confounders_i <- corr_potential_trait_confounders %>% filter(outcome_ID == !!outcome_ID)
+  corr_potential_trait_confounders_pos_i <- corr_potential_trait_confounders_pos %>% filter(outcome_ID == !!outcome_ID)
 
@@
     mutate(beta_2_se = sqrt(beta_2_var)) %>% 
-    left_join(corr_potential_trait_confounders_i %>% dplyr::select(exposure_ID, exposure_ID_AM_IVW_beta, exposure_ID_AM_IVW_se), by = c("exposure" = "exposure_ID")) %>%
+    left_join(corr_potential_trait_confounders_pos_i %>% dplyr::select(exposure_ID, exposure_ID_AM_IVW_beta, exposure_ID_AM_IVW_se), by = c("exposure" = "exposure_ID")) %>%
     dplyr::rename(MVMR_Z_beta = beta) %>%
@@
 
-# Effect of sex, age and time together on causal effects in couples.
+# Single-trait causal effects in couples.
 
@@
 
-AM_result_sig <- household_MR_summary_AM %>% filter(IVW_pval < 0.05/num_tests_by_PCs) 
-AM_sig_traits <-household_MR_summary_AM %>% filter(IVW_pval < 0.05/num_tests_by_PCs) %>% pull(exposure_ID) %>% unique()
-AM_sex_spec_result_sig <- household_MR_binned_het_sex_spec %>% filter(same_trait ==TRUE) %>% filter(exposure_ID %in% AM_sig_traits) %>% filter(sex_het_pval < 0.05/num_tests_by_PCs_AM_sig)
+AM_result_sig <-
+  household_MR_summary_AM %>% filter(IVW_pval < 0.05 / num_tests_by_PCs)
+AM_sig_traits <-
+  household_MR_summary_AM %>% filter(IVW_pval < 0.05 / num_tests_by_PCs) %>% pull(exposure_ID) %>% unique()
+AM_sex_spec_result_sig <-
+  household_MR_binned_het_sex_spec %>% filter(same_trait == TRUE) %>% filter(exposure_ID %in% AM_sig_traits) %>% filter(sex_het_pval < 0.05 /
+                                                                                                                          num_tests_by_PCs_AM_sig)
 
@@
 MR estimates were computed in each sex-seperately using sex-specific SNP-exposure and SNP-outcome results and then meta-analyzed).
-- After adjusting for multiple tests (p < `0.05/`r num_tests_by_PCs``), identified `r dim(AM_result_sig)[1]` significant assortative mating (same-trait) MR results (corresponding to the table below)..
-
-```{r, AM_sig, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, include = TRUE}
+- After adjusting for multiple tests (p < `0.05/`r num_tests_by_PCs``), identified `r dim(AM_result_sig)[1]` significant assortative mating (same-trait) MR results (corresponding to the table below and table "BF_sig_same_trait_MR" in the excel document).
+
+```{r, AM_sig, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, include = TRUE, message = FALSE}
 
@@
 AM_result_sig_nom.DT <- AM_sig_sex_het_nom %>% dplyr::select(trait_description, trait_ID, exposure_sex, IVW_beta, IVW_se, IVW_pval, sex_het_pval) %>% unique()
-```
+
+AM_sensitivity <- bind_rows(household_MR_SNPmeta_sensitivity) %>% filter(same_trait)
+AM_joint.DT <- left_join(AM_result_sig, AM_sensitivity) %>% 
+  dplyr::rename(trait_description = exposure_description) %>%
+  dplyr::rename(trait_ID = exposure_ID) %>%
+  dplyr::select(trait_description, trait_ID, wt_median_summary, wt_median_pval, wt_mode_summary, wt_mode_pval)
+
+non_sig_MR_mode <- AM_joint.DT %>% filter(wt_mode_pval > 0.05)
+non_sig_MR_median <- AM_joint.DT %>% filter(wt_median_pval > 0.05)
+
+```
+
+## Heterogeneity stats.
+
+- In the single trait analysis, we also decided to investigate Cochran’s heterogeneity Q-stat to identify traits with high heterogeneity.
+- This would indicate that associations between the index genotype and partner’s phenotype may not only act indirectly through causal relationship between the traits, but some direct effect is present.
+- Ideally, for the majority of the traits, the heterogeneity would be low, confirming that (despite what Tenesa suggests) there are no direct index genome to partner phenome effects.
+- We tested heterogeneity statistics using the function [mr_heterogeneity](https://mrcieu.github.io/TwoSampleMR/reference/mr_heterogeneity.html) from the [TwoSampleMR](https://mrcieu.github.io/TwoSampleMR/index.html) package.
+
+```{r, het_stats, include = FALSE}
+sig_het <- household_MR_summary_AM %>% filter(Het_IVW_Qpval_round < 0.05/num_tests_by_PCs)
+```
+
+The number of traits with significant heterogeneity in the MR is: `r dim(sig_het)[1]`.
+
+## Sensitivity follow-up.
+
+Of those traits that had a significant causal-effect among couples by IVW, we tested the MR weighted-mode and MR weighted-median method for consistent findings.
+
+- We found `r dim(non_sig_MR_mode)[1]` trait(s) which was/were not significant using the weighted-mode approach (p > 0.05). 
+- We found `r dim(non_sig_MR_median)[1]` trait(s) which was/were not significant using the weighted-mode approach (p > 0.05).
+
+Results can be seen in the table below.
+
+```{r, AM_sensitivity, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, include = TRUE}
+
+DT::datatable(AM_joint.DT, width = "100%") %>% formatSignif(columns=c('wt_median_pval', 'wt_mode_pval'), digits=3)
+
+```
+
+# Effect of sex, age and time together on causal effects in couples.
 
@@
 - After adjusting for multiple hypothesis testing (p < `0.05/`r num_tests_by_PCs_AM_sig``), `r dim(AM_sex_spec_result_sig)[1]` traits showed significant differences amongst sexes.
-- The table below shows the *nominally significant* results at `p < 0.05` (of which there are `r dim(AM_result_sig_nom.DT)[1]/2`).
+- The table below shows the *nominally significant* results at `p < 0.05` (of which there are `r dim(AM_result_sig_nom.DT)[1]/2`) (a slight variation of this table corresponds to "nominally_sig_sex_diff" in the excel document).
+- The $p_{binomial}$ can be calculated as `1 – pbinom(15, 64, 0.05)`, and the fold enrichment as `15/(64*0.05)`.
+- Found that female estimates are on average larger than males (result of the paired t-test shown below).
 
@@
 # write.table(nominally_sig_sex_differences, "docs/manuscript_tables/nominally_sig_sex_differences.txt", row.names = F, sep=",", quote = F)
+
+t.test(nominally_sig_sex_differences$IVW_beta_male, nominally_sig_sex_differences$IVW_beta_female, paired = T)
+
 ```
@@
 - To estimate the difference among bins, we used two approaches: 
-    1. The Cochran's Q test to test for heterogeneity in meta-analyses *(of which we obtained no significant results, so not presenting in paper)*.
+    1. The Cochran's Q test to test for heterogeneity in meta-analyses *(of which we obtained no significant results, so not presented in paper)*.
     2. Tested the significance of the slope of linear model of median bin (either `age` or `time together`) versus the bin-specific MR estimate.
@@
 
-We also tested the phenotypic correlation among the different bins. Below are the traits that show a significant trend either with Pearson's or Spearman's correlation (p < `r 0.05/num_tests_by_PCs`).
+We also tested the phenotypic correlation among the different bins.
+Below are the traits that show a significant trend either with Pearson correlation (p < `r 0.05/num_tests_by_PCs`) (the table below corresponds to "pheno_binned_corrs" in the excel document).
 
@@
 
-tar_load(binned_pheno_corrs)
-
 binned_pheno_corr_summary <- bind_rows(lapply(binned_pheno_corrs, function(x) {x[[2]]}))
 
-binned_pheno_corr_sig <- binned_pheno_corr_summary %>% filter(bin_slope_pval_spearman < 0.05/num_tests_by_PCs | bin_slope_pval_pearson < 0.05/num_tests_by_PCs)
-
-binned_pheno_corr_sig.DT <- replace_Neale_ID(binned_pheno_corr_sig, traits_corr2_filled, "trait", "trait_description") %>%
+pheno_corr_pearson_sig <- binned_pheno_corr_summary %>% filter(bin_slope_pval_pearson < 0.05/num_tests_by_PCs) %>% arrange(group)
+
+pheno_corr_pearson_sig <- replace_Neale_ID(pheno_corr_pearson_sig, traits_corr2_filled, "trait", "trait_description") %>%
   dplyr::rename(trait_ID = trait) %>%
@@
 
-format_round_cols <- colnames(dplyr::select_if(binned_pheno_corr_sig.DT, is.numeric))
-
-DT::datatable(binned_pheno_corr_sig.DT, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)
-
-# write.table(binned_pheno_corr_sig.DT, "docs/manuscript_tables/binned_phenotypic_correlations.txt", row.names = F, sep=",", quote = F)
-
-```
-
-# Heterogeneity stats.
-
-- In the single trait analysis, we also decided to investigate Cochran’s heterogeneity Q-stat to identify traits with high heterogeneity.
-- This would indicate that associations between the index genotype and partner’s phenotype may not only act indirectly through causal relationship between the traits, but some direct effect is present.
-- Ideally, for the majority of the traits, the heterogeneity would be low, confirming that (despite what Tenesa suggests) there are no direct index genome to partner phenome effects.
-- We tested heterogeneity statistics using the function [mr_heterogeneity](https://mrcieu.github.io/TwoSampleMR/reference/mr_heterogeneity.html) from the [TwoSampleMR](https://mrcieu.github.io/TwoSampleMR/index.html) package.
-
-```{r, het_stats, include = FALSE}
-sig_het <- household_MR_summary_AM %>% filter(Het_IVW_Qpval_round < 0.05/num_tests_by_PCs)
-```
-
-The number of traits with significant heterogeneity in the MR is: `r dim(sig_het)[1]`.
+## create panel figure 
+pheno_corr_pearson_sig$branches_to_pull <- NA
+
+for(i in 1:dim(pheno_corr_pearson_sig)[1]){
+
+  trait <- pheno_corr_pearson_sig[i, "trait_ID"][[1]]
+  for(k in 1:length(exposure_info)){
+  temp <- exposure_info[[k]]
+  if(temp[which(temp$Value=="trait_ID"), "Info"] == trait) {break}
+  }
+
+  pheno_corr_pearson_sig$branches_to_pull[i] <- k
+}
+
+figs <- list()
+
+for(i in 1:dim(pheno_corr_pearson_sig)[1]){
+
+  k <- pheno_corr_pearson_sig$branches_to_pull[i] 
+  binned_pheno_corrs_i <- tar_read(binned_pheno_corrs, branches = k)[[1]]
+  exposure_info_i <- tar_read(exposure_info, branches = k)[[1]]
+  pheno_figs <- create_binned_pheno_figs (binned_pheno_corrs_i, exposure_info_i, grouping_var)
+
+  name <- paste0("exposure_", i)
+  group <- pheno_corr_pearson_sig[i, "group"][[1]]
+  fig_num <- ifelse(group == "age_even_bins", 1, 2)
+  figs[[name]] <- pheno_figs[[fig_num]] #tar_read(binned_pheno_figs, branches = i)[[1]][[fig_num]]
+
+}
+
+pheno_binned_figures <- plot_grid(plotlist = figs, nrow=2, ncol = 2, byrow = T, labels="AUTO")
+
+## format table 
+pheno_corr_pearson_sig.DT <- pheno_corr_pearson_sig %>% dplyr::select(-branches_to_pull) %>%
+  mutate(bin_type = case_when(group == "age_even_bins" ~ "age", 
+                              group == "time_together_even_bins" ~ "time_together")) %>%
+  dplyr::select(-group) %>% dplyr::select(trait_description, trait_ID, bin_type, everything())
+
+format_round_cols <- colnames(dplyr::select_if(pheno_corr_pearson_sig.DT, is.numeric))
+
+DT::datatable(pheno_corr_pearson_sig.DT, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)
+
+```
+
+```{r, pheno_binned_figs, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, out.width = "200%", dpi=600}
+
+pheno_binned_figures
+
+```
 
@@
 
-We tested the impact of confounding on genetic PCs, N/E coordinates and the following 4 traits:
+We tested the impact of confounding on genetic PCs, N/E coordinates and the following traits (however only presenting a selection in paper):
 
@@
 
-```{r corr_due_to_confounding_ratio, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
+```{r corr_due_to_confounding_ratio, eval = FALSE, echo = FALSE}
 
@@
 
+```{r corr_due_to_confounding_ratio_causal, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
+
+confounder_traits_sub <- confounder_traits[c(1,2,4,5)]
+
+corr_impact_by_traits_causal_summary <- corr_impact_by_traits_causal %>% group_by(exposure_ID) %>%
+   dplyr::summarise(mean_ratio = mean(corr_due_to_confounding_ratio), median_ratio = median(corr_due_to_confounding_ratio), sd_ratio = sd(corr_due_to_confounding_ratio), number_of_traits = n())
+
+corr_impact_by_traits_causal_summary <- replace_Neale_ID(corr_impact_by_traits_causal_summary, traits_corr2_filled, "exposure_ID", "outcome_description")
+
+corr_impact_by_traits_causal_summary <- corr_impact_by_traits_causal_summary[which(corr_impact_by_traits_causal_summary$exposure_ID %in% confounder_traits_sub),] %>% 
+  dplyr::rename(confounder_trait =  outcome_description) %>% 
+  dplyr::rename(confounder_trait_ID =  exposure_ID) %>% 
+  dplyr::select(confounder_trait, confounder_trait_ID, everything())
+
+corr_impact_by_traits_causal_summary %>%
+  kbl() %>%
+  kable_styling()
+
+```
+
 *A note on the number of traits:*
 
-We tested the impact of `trait_interest` on couple correlation amongst all traits in our couple MR framework (number of traits = `r dim(household_MR_summary_AM)[1]`) with the exception of home location and birth place coordinates (2 each [N and E] * 2), we also removed the trait itself from the confounder analysis where applicable (in the case of traits `20016_irnt`, `738` and `845`).
+We tested the impact of `trait_interest` on couple correlation amongst all traits in our couple MR framework (number of traits = `r dim(household_MR_summary_AM)[1]`) with the exception of birth place coordinates (2 each [N and E]), we also removed the trait itself from the confounder analysis where applicable (in the case of traits `20016_irnt`, `738` and `845`).
 [Trait `189_irnt` was not included in our final couple MR analysis but could still be evaluated here because no MR-estimates were used.]
@@
 
-```{r confounder_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
+```{r confounder_plot, echo = FALSE, eval = FALSE, message=FALSE}
+
+confounder_traits_sub <- confounder_traits[c(1,2,4,5)]
 
 geography_plots <- list()
-
-confounder_traits_sub <- confounder_traits[c(2,3,4,5)]
 
@@
 
+    geom_abline(slope=1, intercept=0) +
+    coord_cartesian(ylim = c(0, NA)) +
     theme_minimal() +
@@
     xlim(0, 1.0) +
+    xlim(0, 1.0) +
+
     theme(plot.title = element_text(size=11, hjust = 0.5)) + theme(axis.title = element_text(size = 10))
@@
 
-```{r PC_confounder_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
+```{r PC_confounder_plot, echo = FALSE, eval = FALSE, message=FALSE}
 
@@
 
-```{r coord_confounder_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE, out.width = "200%", dpi=600}
+```{r coord_confounder_plot, echo = FALSE, eval = FALSE, message=FALSE, out.width = "200%", dpi=600}
 
@@
 
+```{r causal_confounder_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE, dpi=600, fig.height=7}
+
+confounder_traits_sub <- confounder_traits[c(1,2,4,5)]
+
+plots <- list()
+for(i in 1:length(confounder_traits_sub)){
+
+  trait <- confounder_traits_sub[i]
+  trait_description <- confounder_trait_tibble_descript[which(confounder_trait_tibble_descript[["value"]]==trait),"outcome_description"][[1]]
+
+  if(trait=="738"){trait_description <- "Average total household \nincome before tax"}
+  if(trait=="189_irnt"){trait_description <- "Townsend deprivation index \nat recruitment"}
+  if(trait=="6160_1"){trait_description <- "Sports club or gym user"}
+
+  plot_data <- corr_impact_by_traits_causal %>% filter(exposure_ID==trait) %>% filter(exposure_ID!=outcome_ID)
+  if(dim(plot_data)[1]==0) next
+
+  plot_data <- replace_Neale_ID(plot_data, traits_corr2_filled, "outcome_ID", "Phenotype")
+  plot <-
+    ggplot2::ggplot(
+      data = plot_data,
+      ggplot2::aes(x = outcome_ID_couple_r,
+                   y = corr_due_to_confounding, Phenotype = Phenotype)
+    ) +
+    theme(plot.title = element_text(hjust = 0.5)) +
+    geom_point(alpha = 3 / 4) +
+
+    ggplot2::geom_errorbar(
+      ggplot2::aes(ymin = corr_due_to_confounding - corr_due_to_confounding_se, ymax = corr_due_to_confounding + corr_due_to_confounding_se),
+      colour = "grey",
+      width = 0
+    ) +
+    ggplot2::geom_errorbar(
+      ggplot2::aes(xmin = outcome_ID_couple_r - outcome_ID_couple_r_se, xmax = outcome_ID_couple_r + outcome_ID_couple_r_se),
+      colour = "grey",
+      width = 0
+    ) +
+
+    theme_minimal() +
+    ggplot2::labs(
+      x = paste0("Couple trait correlation"),
+      y = paste0("Couple correlation \ndue to confounding"),
+      title = paste0(trait_description)
+      #title = paste0("Impact of ", trait_description, "\n on couple correlation")
+    ) +
+    xlim(0, 1.0) +
+    theme(plot.title = element_text(size=11, hjust = 0.5)) + theme(axis.title = element_text(size = 10))  +
+    geom_abline(slope=1, intercept=0) +
+    coord_cartesian(ylim = c(0, NA))
+
+  plots[[i]] <- plot 
+}
+
+i <- 5
+
+figures_grid <- plot_grid(plotlist = plots, nrow=2, ncol = 2, byrow = F, labels="AUTO")
+figures_grid
+
+```
+
 # Comparison of paths from index to partner.
@@
 We sought to compare the different paths from $X_i$ to $Y_p$ using the estimates computed as $\omega$, $\rho$ and $\gamma$.
-
-```{r proxyMR_figure2, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600}
-proxyMR_prod_comparison_fig_yiyp_adj
-```
 
@@
 
-The results are summarized below:
+The results are summarized below (and provided in table "omega_regression_estimates" in the excel document):
 
@@
 
-omega_regression_estimates <- result %>% dplyr::select(outcome, everything())
+omega_regression_estimates <- result %>% dplyr::select(outcome, everything()) %>% filter(outcome !="gam_rho")
 
@@
 
-pairwise.t.test(data_boxplot$beta, data_boxplot$data, p.adjust = "none")
-
-#cmpr <- list(c("Omega", "Gamma_Rho"), c("Gamma","Gamma_Rho"), c("Gamma_Rho","Gamma_Rho_resid"), c("Rho", "Gamma_Rho"))
+cmpr <- list(c("Omega","Rho"), c("Gamma","Rho"), c("Rho", "Gamma_Rho_resid"))
 
@@
 data_boxplot <- data_boxplot %>% filter(data != "Gamma_Rho")
+pairwise.t.test(data_boxplot$beta, data_boxplot$data, p.adjust = "none", paired = T)
 
@@
   geom_boxplot(outlier.shape = NA) +
-  coord_cartesian(ylim=c(-0.5, 0.5)) +
+
+  stat_compare_means(data = data_boxplot, comparisons = cmpr, label.y = c(0.6, 0.4, 0.5), method = "t.test" , paired =T) + #,
+                     # label = "p.signif",
+                     # symnum.args = list(cutpoints = c(0, 0.05, 1),
+                     #                    symbols = c("*", "ns"))) +
+
+  coord_cartesian(ylim=c(-0.5, 0.75)) +
   scale_fill_viridis(discrete = TRUE, alpha=0.6, option="A") +
@@
 
-#+ #"Gamma_Rho" = "\u03B3 + \u03C1",
-
-  # stat_compare_means(comparisons = cmpr, tip.length=0.01,
-  #                    label = "p.signif",
-  #                    symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1),
-  #                                       symbols = c("****", "***", "**", "*", "ns"))) 
-  #
-
-boxplot
-
-figures_grid <- plot_grid(plotlist = list(proxyMR_prod_comparison_fig_yiyp_adj, boxplot), nrow=1, ncol = 2, byrow = F)
+```
+
+```{r xy_boxplot_combined, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE, warning = FALSE, fig.height = 10}
+
+figures_grid <- plot_grid(plotlist = list(proxyMR_prod_comparison_fig_yiyp_adj, boxplot), nrow=2, ncol = 1, byrow = T)
 figures_grid
@@
 Trait pairs were pruned such that:  if there is a pair A-B and C-D and if [`max(corr(A,C)*corr(B,D),corr(A,D)*corr(B,C))`] > `r corr_trait_threshold` then one pair is dropped.
-Pairs were prioritized based on the $\omega$ p-value.
+Pairs were prioritized based on the $\omega$ p-value (this table corresponds to "paths_summary" in the excel document).
 
@@
 write.xlsx_wrapper(nominally_sig_sex_differences, file=file_name, sheetName="nominally_sig_sex_diff", append=TRUE, row.names=FALSE)
-write.xlsx_wrapper(binned_pheno_corr_sig.DT, file=file_name, sheetName="pheno_binned_corrs", append=TRUE, row.names=FALSE)
-write.xlsx_wrapper(corr_impact_ratios_summary, file=file_name, sheetName="corr_impact_ratios_summary", append=TRUE, row.names=FALSE)
+write.xlsx_wrapper(pheno_corr_pearson_sig.DT, file=file_name, sheetName="pheno_binned_corrs", append=TRUE, row.names=FALSE)
+write.xlsx_wrapper(corr_impact_by_traits_causal_summary, file=file_name, sheetName="corr_impact_ratios_summary", append=TRUE, row.names=FALSE)
 write.xlsx_wrapper(omega_regression_estimates, file=file_name, sheetName="omega_regression_estimates", append=TRUE, row.names=FALSE)

### 2022-04-20 analysis/manuscript.Rmd
wflow_publish("analysis/manuscript.Rmd")
adding manuscript tables

@@
 
-interesting_phenotypes <- c("Basal metabolic rate", "Place of birth in UK - north co-ordinate", 
+interesting_phenotypes <- c("Place of birth in UK - north co-ordinate", 
                             "Systolic blood pressure, automated reading", "Standing height", 
@@
                                                        color = factor(diff_p_sig)),) +
+
+  ggplot2::geom_errorbar(
+    ggplot2::aes(ymin = MR_est - 1.96*IVW_se, ymax = MR_est + 1.96*IVW_se),
+    colour = "grey",
+    width = 0
+  ) +
+  ggplot2::geom_errorbar(
+    ggplot2::aes(xmin = Correlation - 1.96*couple_r_se, xmax = Correlation + 1.96*couple_r_se),
+    colour = "grey",
+    width = 0
+  ) +
+
   geom_point(alpha = 1.2) +
-  scale_color_manual(values = custom_col[c(1,2)], labels=c("Non-significant","BF-signifncant difference")) +
+  scale_color_manual(values = custom_col[c(1,2)], labels=c("Non-significant","Bonferroni-significant difference")) +
   #geom_text(show.legend = FALSE) +
   geom_abline(slope=1, intercept=0) +
-
-  ggplot2::geom_errorbar(
-    ggplot2::aes(ymin = MR_est - IVW_se, ymax = MR_est + IVW_se),
-    colour = "grey",
-    width = 0
-  ) +
-  ggplot2::geom_errorbar(
-    ggplot2::aes(xmin = Correlation - couple_r_se, xmax = Correlation + couple_r_se),
-    colour = "grey",
-    width = 0
-  ) +
 
@@
   ggplot2::labs(x = paste0("Couple correlation"),
-                y = paste("Standardized MR effect")) +
+                y = paste("MR-estimate")) +
   theme(legend.position="bottom") + theme(legend.title=element_blank()) +
   geom_text_repel(data = subset(corr_vs_MR_effects_dat, label_phenos==TRUE),
-                  nudge_y = -0.25 - subset(corr_vs_MR_effects_dat, label_phenos==TRUE)$MR_est, size = 3.25,
-                  direction = "x", show.legend = FALSE )
+                  nudge_y = -0.35 - subset(corr_vs_MR_effects_dat, label_phenos==TRUE)$MR_est, size = 3.25,
+                  direction = "x", show.legend = FALSE, box.padding = 0.5 )
 
@@
 
-number_of_potential_confounders <- dim(corr_potential_trait_confounders_pos)[1]/dim(corr_vs_MR_effects_sig_dat)[1]
+number_of_potential_confounders <- dim(corr_potential_trait_confounders_pos)[1]/dim(corr_potential_trait_confounders_pos)[1]
 
@@
 sig_confounders <- corr_potential_trait_confounders_pos %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
-  arrange(-corr_due_to_confounding_ratio) %>%
+  arrange(-corr_due_to_confounding) %>%
   mutate(total_potential_confounders = n()) %>%
@@
                    potential_confounders = paste0(exposure_description, collapse = "; "), 
-                   max_C = max(corr_due_to_confounding), 
-                   min_C = min(corr_due_to_confounding)) %>%
+                   max_C = max(corr_due_to_confounding)) %>%
   dplyr::rename(trait_ID = outcome_ID) %>%
@@
 sig_confounders_top5 <- corr_potential_trait_confounders_pos %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
-  arrange(-corr_due_to_confounding_ratio) %>%
+  arrange(-corr_due_to_confounding) %>%
   mutate(total_potential_confounders = n()) %>%
@@
                    top5_potential_confounders = paste0(exposure_description, collapse = "; "), 
-                   max_C = max(corr_due_to_confounding), 
-                   min_C = min(corr_due_to_confounding)) %>%
+                   max_C = max(corr_due_to_confounding)) %>%
   dplyr::rename(trait_ID = outcome_ID) %>%
@@
 
-```{r max_C_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}
+```{r max_C_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE, fig.height=10, dpi=600}
 
@@
   mutate(Z_prod_se = sqrt(Z_prod_var)) %>%
-  mutate(same_trait_corr_minus_MR_est_se = sqrt(variance_of_sum(outcome_ID_IVW_se, outcome_ID_couple_r_se)))
-
-max_C_plot <- ggplot2::ggplot(data = max_C_data, ggplot2::aes(x = corr_due_to_confounding,
-                                                       y = same_trait_corr_minus_MR_est)) +
+  mutate(same_trait_corr_minus_MR_est_se = sqrt(variance_of_sum(outcome_ID_IVW_se, outcome_ID_couple_r_se))) %>%
+  mutate(Phenotype = outcome_description) %>% 
+  mutate(Phenotype = case_when(Phenotype == "Basal metabolic rate" ~ "BMR",
+                               Phenotype == "Systolic blood pressure, automated reading" ~ "SBP",
+                               Phenotype == "Pulse rate, automated reading" ~ "Pulse rate",
+                               Phenotype == "Body fat percentage" ~ "BFP",
+                               Phenotype == "Standing height" ~ "Height",
+                               Phenotype == "Reticulocyte percentage" ~ "Reticulocyte %",
+
+                               #Phenotype == "Weight" & outcome_ID == "21002_irnt" ~ "Weight",
+                               Phenotype == "Forced vital capacity (FVC), Best measure" ~ "FVC",
+                               Phenotype == "Red blood cell (erythrocyte) count" ~ "RBC count",
+                               Phenotype == "Place of birth in UK - north co-ordinate" ~ "Place of birth, NC",
+                               TRUE ~ Phenotype)) %>%
+  mutate(label_phenos = case_when(outcome_description %in% interesting_phenotypes ~ TRUE, TRUE ~ FALSE))
+
+max_C_plot <- ggplot2::ggplot(data = max_C_data, 
+                              ggplot2::aes(x = corr_due_to_confounding, 
+                                           y = same_trait_corr_minus_MR_est, 
+                                           label = Phenotype)) +
+
+  ggplot2::geom_errorbar(
+    ggplot2::aes(ymin = same_trait_corr_minus_MR_est - 1.96*same_trait_corr_minus_MR_est_se, 
+                 ymax = same_trait_corr_minus_MR_est + 1.96*same_trait_corr_minus_MR_est_se),
+    colour = "grey",
+    width = 0
+  ) +
+  ggplot2::geom_errorbar(
+    ggplot2::aes(xmin = corr_due_to_confounding - 1.96*Z_prod_se, 
+                 xmax = corr_due_to_confounding + 1.96*Z_prod_se),
+    colour = "grey",
+    width = 0
+  ) +
+
   geom_point(alpha = 1.2) +
@@
 
-  ggplot2::geom_errorbar(
-    ggplot2::aes(ymin = same_trait_corr_minus_MR_est - same_trait_corr_minus_MR_est_se, ymax = same_trait_corr_minus_MR_est + same_trait_corr_minus_MR_est_se),
-    colour = "grey",
-    width = 0
-  ) +
-  ggplot2::geom_errorbar(
-    ggplot2::aes(xmin = corr_due_to_confounding - Z_prod_se, xmax = corr_due_to_confounding + Z_prod_se),
-    colour = "grey",
-    width = 0
-  ) +
-
   theme_minimal() +
@@
   theme(legend.title=element_blank()) +
-  #geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
-  theme_ipsum()
+  geom_text_repel(data = subset(max_C_data, label_phenos==TRUE),
+                  nudge_y = -0.25 - subset(max_C_data, label_phenos==TRUE)$same_trait_corr_minus_MR_est, size = 3.25,
+                  direction = "x", show.legend = FALSE, box.padding = 0.5)
 
 max_C_plot
+
+## combine `max_C_plot` with `corr_vs_MR_effects_plot`
+
+corr_vs_MR_comb <- plot_grid(plotlist = list(corr_vs_MR_effects_plot, max_C_plot), nrow=2, ncol = 1, byrow = T, labels="AUTO")
+
+corr_vs_MR_comb
 
@@
 nominally_sig_sex_differences <- AM_result_sig_nom.DT %>% pivot_wider(names_from = exposure_sex, values_from = c(IVW_beta, IVW_se, IVW_pval))
-# write.table(nominally_sig_sex_differences, "docs/manuscript_tables/nominally_sig_sex_differences.txt", row.names = F, sep=",", quote = F)
 
@@
 
-pheno_corr_pearson_sig <- binned_pheno_corr_summary %>% filter(bin_slope_pval_pearson < 0.05/num_tests_by_PCs) %>% arrange(group)
+pheno_corr_pearson_sig <- binned_pheno_corr_summary %>% filter(bin_slope_pval_pearson < 0.05/num_tests_by_PCs) %>% arrange(desc(group))
 
@@
 
-```{r, pheno_binned_figs, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, out.width = "200%", dpi=600}
+```{r, pheno_binned_figs, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, fig.height=7, dpi=600}
 
@@
 
+## originally ran the confounder analysis using correlations only, but switched to use causal effects.
+
 corr_impact_by_coords_ratio <- corr_impact_by_coords %>% 
@@
 
-corr_impact_by_traits_ratio <- replace_Neale_ID(corr_impact_by_traits_ratio, traits_corr2_filled, "confounder_trait", "confounder_trait_description") %>% ungroup() %>%
+corr_impact_by_traits_ratio <- replace_Neale_ID(corr_impact_by_traits_ratio, traits_corr2_filled, "confounder_trait", "confounder_trait_description") %>% 
+  ungroup() %>%
   dplyr::select(-confounder_trait) %>% dplyr::rename(confounder_trait = confounder_trait_description)
@@
 
-# write.table(corr_impact_ratios_summary, "docs/manuscript_tables/corr_impact_ratios_summary.txt", row.names = F, sep=",", quote = F)
 ```
@@
 
-confounder_traits_sub <- confounder_traits[c(1,2,4,5)]
+confounder_traits_sub <- confounder_traits[c(1,2,5,10,15)]
 
 corr_impact_by_traits_causal_summary <- corr_impact_by_traits_causal %>% group_by(exposure_ID) %>%
-   dplyr::summarise(mean_ratio = mean(corr_due_to_confounding_ratio), median_ratio = median(corr_due_to_confounding_ratio), sd_ratio = sd(corr_due_to_confounding_ratio), number_of_traits = n())
+   dplyr::summarise(mean_ratio = mean(corr_due_to_confounding_ratio), 
+                    median_ratio = median(corr_due_to_confounding_ratio), sd_ratio = sd(corr_due_to_confounding_ratio), 
+                    number_of_traits = n())
 
@@
 
-corr_impact_by_traits_causal_summary %>%
+corr_impact_by_PCs_summary <- corr_impact_by_PCs %>% 
+  dplyr::select(Neale_pheno_ID, trait_couple_corr, trait_couple_corr_se, corr_due_to_confounding_all, corr_due_to_confounding_all_se) %>%
+  unique() %>%
+  mutate(confounder_trait = "PCs") %>%
+  mutate(trait_ratio = corr_due_to_confounding_all/trait_couple_corr) %>%
+  group_by(confounder_trait) %>%
+  dplyr::summarise(mean_ratio = mean(trait_ratio), median_ratio = median(trait_ratio), sd_ratio = sd(trait_ratio), number_of_traits = n()) %>%
+  mutate(confounder_trait_ID = NA)
+
+corr_impact_by_coords_summary <- corr_impact_by_traits_causal %>% filter(exposure_ID=="129_irnt" | exposure_ID=="130_irnt") %>% 
+  filter(!(outcome_ID=="129_irnt" | outcome_ID=="130_irnt"))  %>%
+  group_by(outcome_ID) %>% mutate(corr_due_to_confounding_all = sum(corr_due_to_confounding)) %>%
+  mutate(corr_due_to_confounding_all_se = sqrt(sum(corr_due_to_confounding_se^2))) %>%
+  dplyr::select(outcome_ID, outcome_ID_couple_r, outcome_ID_couple_r_se, corr_due_to_confounding_all, corr_due_to_confounding_all_se) %>% unique() %>%
+  mutate(confounder_trait = "Birth place coordinates") %>%
+  mutate(trait_ratio = corr_due_to_confounding_all/outcome_ID_couple_r) %>%
+  group_by(confounder_trait) %>%
+  dplyr::summarise(mean_ratio = mean(trait_ratio), median_ratio = median(trait_ratio), sd_ratio = sd(trait_ratio), number_of_traits = n()) %>%
+  mutate(confounder_trait_ID = NA)
+
+corr_impact_summary <- rbind(corr_impact_by_traits_causal_summary, corr_impact_by_coords_summary)
+
+corr_impact_summary %>%
   kbl() %>%
@@
 ```
-
-*A note on the number of traits:*
-
-We tested the impact of `trait_interest` on couple correlation amongst all traits in our couple MR framework (number of traits = `r dim(household_MR_summary_AM)[1]`) with the exception of birth place coordinates (2 each [N and E]), we also removed the trait itself from the confounder analysis where applicable (in the case of traits `20016_irnt`, `738` and `845`).
-[Trait `189_irnt` was not included in our final couple MR analysis but could still be evaluated here because no MR-estimates were used.]
 
@@
     ) +
+
+    ggplot2::geom_errorbar(
+      ggplot2::aes(ymin = corr_due_to_confounding - 1.96*corr_due_to_confounding_se, ymax = corr_due_to_confounding + 1.96*corr_due_to_confounding_se),
+      colour = "grey",
+      width = 0
+    ) +
+    ggplot2::geom_errorbar(
+      ggplot2::aes(xmin = trait_j_couple_corr - 1.96*trait_j_couple_corr_se, xmax = trait_j_couple_corr + 1.96*trait_j_couple_corr_se),
+      colour = "grey",
+      width = 0
+    ) +
+
     theme(plot.title = element_text(hjust = 0.5)) +
     geom_point(alpha = 3 / 4) +
-
-    ggplot2::geom_errorbar(
-      ggplot2::aes(ymin = corr_due_to_confounding - corr_due_to_confounding_se, ymax = corr_due_to_confounding + corr_due_to_confounding_se),
-      colour = "grey",
-      width = 0
-    ) +
-    ggplot2::geom_errorbar(
-      ggplot2::aes(xmin = trait_j_couple_corr - trait_j_couple_corr_se, xmax = trait_j_couple_corr + trait_j_couple_corr_se),
-      colour = "grey",
-      width = 0
-    ) +
 
@@
 
-```{r PC_confounder_plot, echo = FALSE, eval = FALSE, message=FALSE}
+```{r PC_confounder_plot, echo = FALSE, eval = TRUE, message=FALSE}
 
@@
   theme(plot.title = element_text(hjust = 0.5)) +
-  geom_point(alpha = 3 / 4) +
 
   ggplot2::geom_errorbar(
-    ggplot2::aes(ymin = corr_due_to_confounding_all - corr_due_to_confounding_all_se, ymax = corr_due_to_confounding_all + corr_due_to_confounding_all_se),
+    ggplot2::aes(ymin = corr_due_to_confounding_all - 1.96*corr_due_to_confounding_all_se, ymax = corr_due_to_confounding_all + 1.96*corr_due_to_confounding_all_se),
     colour = "grey",
@@
   ggplot2::geom_errorbar(
-    ggplot2::aes(xmin = trait_couple_corr - trait_couple_corr_se, xmax = trait_couple_corr + trait_couple_corr_se),
+    ggplot2::aes(xmin = trait_couple_corr - 1.96*trait_couple_corr_se, xmax = trait_couple_corr + 1.96*trait_couple_corr_se),
     colour = "grey",
@@
   ) +
+
+  geom_point(alpha = 3 / 4) +
 
@@
   xlim(0, 1.0) +
-  theme(plot.title = element_text(size=11, hjust = 0.5)) + theme(axis.title = element_text(size = 10))
-
-i <- 5
-plots[[i]] <- plot_PCs
-
-```
-
-```{r coord_confounder_plot, echo = FALSE, eval = FALSE, message=FALSE, out.width = "200%", dpi=600}
-
-plot_data <- corr_impact_by_coords %>% dplyr::select(Neale_pheno_ID, trait_couple_corr, corr_due_to_confounding_all, corr_due_to_confounding_all_se, trait_couple_corr_se, trait_couple_corr_se) %>% unique()
-
-plot_coords <-
+  theme(plot.title = element_text(size=11, hjust = 0.5)) + theme(axis.title = element_text(size = 10))  +
+  geom_abline(slope=1, intercept=0) +
+  coord_cartesian(ylim = c(0, NA))
+
+```
+
+```{r coord_confounder_plot, echo = FALSE, eval = TRUE, message=FALSE, out.width = "200%", dpi=600}
+
+trait_description <- "Place of birth coordinates"
+
+plot_data <- corr_impact_by_traits_causal %>% filter(exposure_ID=="129_irnt" | exposure_ID=="130_irnt") %>% 
+  filter(!(outcome_ID=="129_irnt" | outcome_ID=="130_irnt"))  %>%
+  group_by(outcome_ID) %>% mutate(corr_due_to_confounding_all = sum(corr_due_to_confounding)) %>%
+  mutate(corr_due_to_confounding_all_se = sqrt(sum(corr_due_to_confounding_se^2))) %>%
+  dplyr::select(outcome_ID, outcome_ID_couple_r, outcome_ID_couple_r_se, corr_due_to_confounding_all, corr_due_to_confounding_all_se) %>% unique()
+
+plot_data <- replace_Neale_ID(plot_data, traits_corr2_filled, "outcome_ID", "Phenotype")
+coord_plot <-
   ggplot2::ggplot(
     data = plot_data,
-    ggplot2::aes(x = trait_couple_corr,
-                 y = corr_due_to_confounding_all)
+    ggplot2::aes(x = outcome_ID_couple_r,
+                 y = corr_due_to_confounding_all, Phenotype = Phenotype)
   ) +
   theme(plot.title = element_text(hjust = 0.5)) +
-  geom_point(alpha = 3 / 4) +
 
   ggplot2::geom_errorbar(
-    ggplot2::aes(ymin = corr_due_to_confounding_all - corr_due_to_confounding_all_se, ymax = corr_due_to_confounding_all + corr_due_to_confounding_all_se),
+    ggplot2::aes(ymin = corr_due_to_confounding_all - 1.96*corr_due_to_confounding_all_se, ymax = corr_due_to_confounding_all + 1.96*corr_due_to_confounding_all_se),
     colour = "grey",
@@
   ggplot2::geom_errorbar(
-    ggplot2::aes(xmin = trait_couple_corr - trait_couple_corr_se, xmax = trait_couple_corr + trait_couple_corr_se),
+    ggplot2::aes(xmin = outcome_ID_couple_r - 1.96*outcome_ID_couple_r_se, xmax = outcome_ID_couple_r + 1.96*outcome_ID_couple_r_se),
     colour = "grey",
@@
     y = paste0("Couple correlation \ndue to confounding"),
-    title = paste0("Place of birth \ncoordinates")
+    title = paste0(trait_description)
     #title = paste0("Impact of ", trait_description, "\n on couple correlation")
@@
   xlim(0, 1.0) +
-  theme(plot.title = element_text(size=11, hjust = 0.5)) + theme(axis.title = element_text(size = 10))
-
-i <- 6
-plots[[i]] <- plot_coords
-
-figures_grid <- plot_grid(plotlist = plots, nrow=3, ncol = 2, byrow = T, labels="AUTO", label_size = 12)
-figures_grid
+  theme(plot.title = element_text(size=11, hjust = 0.5)) + theme(axis.title = element_text(size = 10))  +
+  geom_abline(slope=1, intercept=0) +
+  coord_cartesian(ylim = c(0, NA))
 
@@
 
-confounder_traits_sub <- confounder_traits[c(1,2,4,5)]
+confounder_traits_sub <- confounder_traits[c(1,2,5,10,15)]
 
@@
     theme(plot.title = element_text(hjust = 0.5)) +
-    geom_point(alpha = 3 / 4) +
 
     ggplot2::geom_errorbar(
-      ggplot2::aes(ymin = corr_due_to_confounding - corr_due_to_confounding_se, ymax = corr_due_to_confounding + corr_due_to_confounding_se),
+      ggplot2::aes(ymin = corr_due_to_confounding - 1.96*corr_due_to_confounding_se, ymax = corr_due_to_confounding + 1.96*corr_due_to_confounding_se),
       colour = "grey",
@@
     ggplot2::geom_errorbar(
-      ggplot2::aes(xmin = outcome_ID_couple_r - outcome_ID_couple_r_se, xmax = outcome_ID_couple_r + outcome_ID_couple_r_se),
+      ggplot2::aes(xmin = outcome_ID_couple_r - 1.96*outcome_ID_couple_r_se, xmax = outcome_ID_couple_r + 1.96*outcome_ID_couple_r_se),
       colour = "grey",
@@
     ) +
+
+    geom_point(alpha = 3 / 4) +
 
@@
 
-i <- 5
-
-figures_grid <- plot_grid(plotlist = plots, nrow=2, ncol = 2, byrow = F, labels="AUTO")
+i <- 6
+plots[[i]] <- coord_plot 
+figures_grid <- plot_grid(plotlist = plots, nrow=3, ncol = 2, byrow = F, labels="AUTO")
 figures_grid
@@
 data <- proxyMR_comparison_summary_yiyp_adj_SNPmeta
+
+data <- data %>%
+  mutate(
+    rho_larger = case_when(
+      gam_vs_rho_BF_sig == TRUE & rho_beta > gam_beta ~ TRUE,
+      gam_vs_rho_BF_sig == TRUE &
+        rho_beta < gam_beta ~ FALSE,
+      gam_vs_rho_BF_sig == FALSE ~ NA
+    )
+  ) %>%
+  mutate(omega_larger = case_when((omega_beta - rho_beta - gam_beta) > max(rho_beta, gam_beta) ~ TRUE, 
+                                  TRUE ~ FALSE))
+
 data$rho_resid <- resid(lm(rho_beta ~ gam_beta, data = data))
@@
 ```
+
+Of note, we found `r  table(data$gam_vs_rho_BF_sig)[[2]]` relationships which were significantly different between $\rho$ and  $\gamma$, whereby `r table(data$rho_larger)[[2]]` (`r (table(data$rho_larger)[[2]]/table(data$gam_vs_rho_BF_sig)[[2]])*100`%) showed larger effects through $\rho$ and the other `r table(data$rho_larger)[[1]]` (`r (table(data$rho_larger)[[1]]/table(data$gam_vs_rho_BF_sig)[[2]])*100`%) showed larger effects through $\gamma$.
 
@@
 
-figures_grid <- plot_grid(plotlist = list(proxyMR_prod_comparison_fig_yiyp_adj, boxplot), nrow=2, ncol = 1, byrow = T)
+figures_grid <- plot_grid(plotlist = list(proxyMR_prod_comparison_fig_yiyp_adj, boxplot), nrow=2, ncol = 1, byrow = T, labels = c('', 'E'))
 figures_grid
@@
 
-file_name <- "docs/manuscript_tables/manuscript_tables.xlsx"
+main_tables_file <- "docs/manuscript_tables/manuscript_tables.xlsx"
+supplementary_tables_file <- "docs/manuscript_tables/supplementary_tables.xlsx"
 
@@
 
-write.xlsx_wrapper(corr_vs_MR_effects_sig_dat, file=file_name, sheetName="sig_corr_vs_MR", row.names=FALSE)
-write.xlsx_wrapper(sig_confounders, file=file_name, sheetName="sig_confounders", append=TRUE, row.names=FALSE)
-write.xlsx_wrapper(AM_result_sig.DT, file=file_name, sheetName="BF_sig_same_trait_MR", append=TRUE, row.names=FALSE)
-write.xlsx_wrapper(nominally_sig_sex_differences, file=file_name, sheetName="nominally_sig_sex_diff", append=TRUE, row.names=FALSE)
-write.xlsx_wrapper(pheno_corr_pearson_sig.DT, file=file_name, sheetName="pheno_binned_corrs", append=TRUE, row.names=FALSE)
-write.xlsx_wrapper(corr_impact_by_traits_causal_summary, file=file_name, sheetName="corr_impact_ratios_summary", append=TRUE, row.names=FALSE)
-write.xlsx_wrapper(omega_regression_estimates, file=file_name, sheetName="omega_regression_estimates", append=TRUE, row.names=FALSE)
-write.xlsx_wrapper(paths_summary, file=file_name, sheetName="paths_summary", append=TRUE, row.names=FALSE)
-
-```
+## Main tables 
+write.xlsx_wrapper(corr_impact_summary, file=main_tables_file, sheetName="Table 1 corr_impact_ratios_summary", row.names=FALSE)
+write.xlsx_wrapper(omega_regression_estimates, file=main_tables_file, sheetName="Table 2 omega_regression_estimates", append=TRUE, row.names=FALSE)
+
+## Supplementary tables
+write.xlsx_wrapper(AM_result_sig.DT, file=supplementary_tables_file, sheetName="Table 1 BF_sig_same_trait_MR", append=TRUE, row.names=FALSE)
+write.xlsx_wrapper(nominally_sig_sex_differences, file=supplementary_tables_file, sheetName="Table 2 nominally_sig_sex_diff", append=TRUE, row.names=FALSE)
+write.xlsx_wrapper(pheno_corr_pearson_sig.DT, file=supplementary_tables_file, sheetName="Table 3 pheno_binned_corrs", append=TRUE, row.names=FALSE)
+write.xlsx_wrapper(corr_vs_MR_effects_sig_dat, file=supplementary_tables_file, sheetName="Table 4 sig_corr_vs_MR", row.names=FALSE)
+write.xlsx_wrapper(sig_confounders, file=supplementary_tables_file, sheetName="Table 5 sig_confounders", append=TRUE, row.names=FALSE)
+write.xlsx_wrapper(paths_summary, file=supplementary_tables_file, sheetName="Table 6 paths_summary", append=TRUE, row.names=FALSE)
+
+```

### 2022-04-20 analysis/update_meeting_15_10_2021.Rmd
add some final manuscript files

@@
 
-Decided best to also include all $X$ IVs in the model as well. MRMR model is: $Y_p \sim Y_i + X_i$.
+Decided best to also include all $X$ IVs in the model as well. MVMR model is: $Y_p \sim Y_i + X_i$.
 
@@
 
-## merge the list of dataframes 
 corr_impact_by_PCs.DT <- replace_Neale_ID(corr_impact_by_PCs, traits_corr2_filled, "Neale_pheno_ID", "outcome_description")
@@
 
-```{r coordin_traits_confounding, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
+```{r coord_traits_confounding, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
 
-## merge the list of dataframes 
 corr_impact_by_coords.DT <- replace_Neale_ID(corr_impact_by_coords, traits_corr2_filled, "Neale_pheno_ID", "outcome_description")

### 2022-04-20 analysis/project_status.Rmd

---
title: "Project status"
author: "Jenny Sjaarda"
date: "2021-09-16"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

This notebook displays the status of the project workflow, that is, whether everything is computationally up to date. It displays the status as at the time the notebook is executed, so if the status has changed since this notebook was last executed that fact won't be reflected in this rendered notebook. This notebook is currently intended to be run manually when you feel the need, for example, when returning to the project.

# `{renv}` status

`{renv}` tracks the used `R` packages and their versions.
```{r renv}
# Get {targets} to tell {renv} which packages it uses
targets::tar_renv() 
renv::status()
```

# `{targets}` status
`{targets}` is used to track the computational dependencies between the
project components and orchestrate recalculation when any components are
changed.

Each of the meta pipelines has leaves that are reports rendered from
`Rmd` files that are managed jointly by `{targets}` and `{workflowr}`.
`{targets}` is aware of the upstream dependencies (data and functions)
and the `Rmd` file itself. If any of these is invalidated `tar_make()`
will rebuild the report using `wflow_build()`
.
`{targets}` is *not* aware of the rendered meta report files. The
rendered report files are managed exclusively by `{workflowr}`. This
means that you can use `{workflowr}` to build and publish the report
without making `{targets}` think that the report is out of date because
`{workflowr}` has re-rendered it.
```{r targets}
# outline
targets::tar_glimpse()
# excruciating detail
targets::tar_visnetwork()
```

# `{workflowr}` status
`{workflowr}` is used to render the *meta* documents
(`analysis/m_*.Rmd`) and build the project website.
`{workflowr}` is only concerned with the modification and `git` commit
dates of the `Rmd` and `HTML` files, whereas `{targets}` is concerned
with the computational dependencies. Thus, you should rely on the
`{targets}` status to ensure the meta reports are computationally up to
date, then rely on the `{workflowr}` status to manage the building and
publication of the entire website.

```{r workflowr}
workflowr::wflow_status()
```

### 2022-04-20 analysis/manuscript.Rmd
Build site.

@@
 ## Supplementary tables
-write.xlsx_wrapper(AM_result_sig.DT, file=supplementary_tables_file, sheetName="Table 1 BF_sig_same_trait_MR", append=TRUE, row.names=FALSE)
-write.xlsx_wrapper(nominally_sig_sex_differences, file=supplementary_tables_file, sheetName="Table 2 nominally_sig_sex_diff", append=TRUE, row.names=FALSE)
-write.xlsx_wrapper(pheno_corr_pearson_sig.DT, file=supplementary_tables_file, sheetName="Table 3 pheno_binned_corrs", append=TRUE, row.names=FALSE)
-write.xlsx_wrapper(corr_vs_MR_effects_sig_dat, file=supplementary_tables_file, sheetName="Table 4 sig_corr_vs_MR", row.names=FALSE)
-write.xlsx_wrapper(sig_confounders, file=supplementary_tables_file, sheetName="Table 5 sig_confounders", append=TRUE, row.names=FALSE)
-write.xlsx_wrapper(paths_summary, file=supplementary_tables_file, sheetName="Table 6 paths_summary", append=TRUE, row.names=FALSE)
-
-```
+write.xlsx_wrapper(AM_result_sig.DT, file=supplementary_tables_file, sheetName="Supp Table 1 BF_sig_same_trait_MR", row.names=FALSE)
+write.xlsx_wrapper(nominally_sig_sex_differences, file=supplementary_tables_file, sheetName="Supp Table 2 nominally_sig_sex_diff", append=TRUE, row.names=FALSE)
+write.xlsx_wrapper(pheno_corr_pearson_sig.DT, file=supplementary_tables_file, sheetName="Supp Table 3 pheno_binned_corrs", append=TRUE, row.names=FALSE)
+write.xlsx_wrapper(corr_vs_MR_effects_sig_dat, file=supplementary_tables_file, sheetName="Supp Table 4 sig_corr_vs_MR", append = TRUE, row.names=FALSE)
+write.xlsx_wrapper(sig_confounders, file=supplementary_tables_file, sheetName="Supp Table 5 sig_confounders", append=TRUE, row.names=FALSE)
+write.xlsx_wrapper(paths_summary, file=supplementary_tables_file, sheetName="Supp Table 6 paths_summary", append=TRUE, row.names=FALSE)
+
+```

### 2022-04-20 analysis/AM_MR_summary.Rmd
wflow_publish("analysis/AM_MR_summary.Rmd")

@@
 tar_load(household_MR_AM_FvsM_fig)
+tar_load(household_MR_binned_het_sex_spec)
 
@@
 library(knitr)
-knitr::include_graphics("assets/single_trait_MR.png", error = FALSE)
+knitr::include_graphics("assets/single_trait_model.png", error = FALSE)
 ```
@@
 
-AM_sig_traits <- household_MR_summary_AM %>% filter(IVW_meta_pval < 0.05/num_tests_by_PCs) %>% pull(exposure_ID) %>% unique()
+AM_sig_traits <- household_MR_summary_AM %>% filter(IVW_pval < 0.05/num_tests_by_PCs) %>% pull(exposure_ID) %>% unique()
 AM_result_sig <- household_MR_summary_AM %>% filter(exposure_ID %in% AM_sig_traits)
@@
 household_MR_binned_het_descript <- household_MR_binned_het_temp2
+
+AM_sex_spec_result_sig <-
+  household_MR_binned_het_sex_spec %>% filter(same_trait == TRUE) %>% filter(exposure_ID %in% AM_sig_traits) %>% filter(sex_het_pval < 0.05 /
+                                                                                                                          num_tests_by_PCs_AM_sig)
 
@@
 
-AM_result_sig.DT <- AM_result_sig %>% dplyr::select(exposure_description, IVW_meta_beta, IVW_meta_pval) %>% unique()
-DT::datatable(AM_result_sig.DT, width = "100%") %>% formatSignif(columns=c('IVW_meta_beta', 'IVW_meta_pval'), digits=3)
+AM_result_sig.DT <- AM_result_sig %>% dplyr::select(exposure_description, IVW_beta, IVW_pval) %>% unique()
+DT::datatable(AM_result_sig.DT, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval'), digits=3)
+
+AM_sig_sex_het_nom <- household_MR_binned_het_sex_spec %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% filter(sex_het_pval < 0.05)
+AM_sig_sex_het_nom <- replace_Neale_ID(AM_sig_sex_het_nom, traits_corr2_filled,"exposure_ID", "exposure_description") %>%
+  dplyr::rename(trait_description = exposure_description) %>%
+  dplyr::rename(trait_ID = exposure_ID) %>%
+  dplyr::select(trait_description, trait_ID, everything())
+
+AM_result_sig_nom.DT <- AM_sig_sex_het_nom %>% dplyr::select(trait_description, trait_ID, exposure_sex, IVW_beta, IVW_se, IVW_pval, sex_het_pval) %>% unique()
 
@@
 - A similar adjustment for multiple hypothesis testing was applied as above, resulting in `r num_tests_by_PCs_AM_sig` number of effective tests. 
-- The table below displays the results that significantly differ among sexes (p < `0.05/`r num_tests_by_PCs_AM_sig``).
+- After adjusting for multiple hypothesis testing (p < `0.05/`r num_tests_by_PCs_AM_sig``), `r dim(AM_sex_spec_result_sig)[1]` traits showed significant differences amongst sexes.
+- The table below shows the *nominally significant* results at `p < 0.05` (of which there are `r dim(AM_result_sig_nom.DT)[1]/2`).
 
-```{r AM_sex_het, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
+```{r AM_sex_het_nom, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
 
-AM_sig_sex_het <- AM_result_sig %>% filter(IVW_sex_het_pval < 0.05/num_tests_by_PCs_AM_sig)
+DT::datatable(AM_result_sig_nom.DT, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'sex_het_pval'), digits=3)
 
-AM_result_sig.DT <- AM_sig_sex_het %>% dplyr::select(exposure_description, IVW_beta, IVW_pval, IVW_meta_beta, IVW_meta_pval, IVW_sex_het_pval) %>% unique()
-DT::datatable(AM_result_sig.DT, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'IVW_meta_beta', 'IVW_meta_pval', 'IVW_sex_het_pval'), digits=3)
+nominally_sig_sex_differences <- AM_result_sig_nom.DT %>% pivot_wider(names_from = exposure_sex, values_from = c(IVW_beta, IVW_se, IVW_pval))
+
+t.test(nominally_sig_sex_differences$IVW_beta_male, nominally_sig_sex_differences$IVW_beta_female, paired = T)
 
@@
 
-Only `r dim(AM_result_sig.DT)[1]/2` was significant after adjustment. The table below shows the results significant at `p < 0.05`.
+In general, found that female estimates are on average larger than males (result of the paired t-test shown below).
 
-```{r AM_sex_het_nom, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
+```{r sex_het_m_vs_f, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
 
-AM_sig_sex_het_nom <- AM_result_sig %>% filter(IVW_sex_het_pval < 0.05)
-AM_result_sig_nom.DT <- AM_sig_sex_het_nom %>% dplyr::select(exposure_description, IVW_beta, IVW_pval, IVW_meta_beta, IVW_meta_pval, IVW_sex_het_pval) %>% unique()
-DT::datatable(AM_result_sig_nom.DT, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'IVW_meta_beta', 'IVW_meta_pval', 'IVW_sex_het_pval'), digits=3)
+nominally_sig_sex_differences <- AM_result_sig_nom.DT %>% pivot_wider(names_from = exposure_sex, values_from = c(IVW_beta, IVW_se, IVW_pval))
 
-```
+t.test(nominally_sig_sex_differences$IVW_beta_male, nominally_sig_sex_differences$IVW_beta_female, paired = T)
 
-In general, the male to female estimate is slightly larger than the female to male estimate, as shown in the figure below.
-
-```{r sex_het_fig, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
-household_MR_AM_FvsM_fig
 ```

### 2022-09-01 R/functions.R
remove filtering on correlation with trait

@@ -3626,7 +3626,7 @@ find_potential_trait_confounders_pos <- function(Neale_pheno_ID, Neale_pheno_ID_
     mutate(sig_confounder = ifelse(IVW_pval < 0.05/num_tests_by_PCs, TRUE, FALSE)) %>%
     mutate(exposure_ID_phes = gsub("_irnt", "", exposure_ID)) %>%
     left_join(corr_mat_traits_sub, by = c("exposure_ID_phes" = "exposure_ID")) %>%
-    filter(abs(between_trait_correlation) < corr_trait_threshold) %>% arrange(IVW_pval) %>%
+    arrange(IVW_pval) %>%
     filter(sig_confounder) %>%
     filter(exposure_ID_AM_IVW_beta > 0) %>%
     filter(exposure_ID_AM_IVW_pval < 0.05/n())
@@ -3693,7 +3693,7 @@ find_potential_trait_confounders_neg <- function(Neale_pheno_ID, Neale_pheno_ID_
     mutate(sig_confounder = ifelse(IVW_pval < 0.05/num_tests_by_PCs, TRUE, FALSE)) %>%
     mutate(exposure_ID_phes = gsub("_irnt", "", exposure_ID)) %>%
     left_join(corr_mat_traits_sub, by = c("exposure_ID_phes" = "exposure_ID")) %>%
-    filter(abs(between_trait_correlation) < corr_trait_threshold) %>% arrange(IVW_pval) %>%
+    arrange(IVW_pval) %>%
     filter(sig_confounder) %>%
     filter(exposure_ID_AM_IVW_beta < 0) %>%
     filter(exposure_ID_AM_IVW_pval < 0.05/n())

### 2022-09-01 analysis/manuscript.Rmd
change C plot to be sum of C not max-C

--- 2022-04-20 analysis/manuscript.Rmd
@@
 tar_load(corr_potential_trait_confounders_pos)
+tar_load(corr_potential_trait_confounders_pos_MVMR_summary)
 tar_load(traits_corr)
@@
 
-sig_confounders <- corr_potential_trait_confounders_pos %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
-  arrange(-corr_due_to_confounding) %>%
+sig_confounders <- corr_potential_trait_confounders_pos_MVMR_summary %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
+  arrange(-corr_due_to_confounding_MVMR) %>%
   mutate(total_potential_confounders = n()) %>%
@@
                    potential_confounders = paste0(exposure_description, collapse = "; "), 
-                   max_C = max(corr_due_to_confounding)) %>%
+                   max_C = max(corr_due_to_confounding_MVMR)) %>%
   dplyr::rename(trait_ID = outcome_ID) %>%
@@
 
-sig_confounders_top5 <- corr_potential_trait_confounders_pos %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
-  arrange(-corr_due_to_confounding) %>%
+sig_confounders_top5 <- corr_potential_trait_confounders_pos_MVMR_summary %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
+  arrange(-corr_due_to_confounding_MVMR) %>%
   mutate(total_potential_confounders = n()) %>%
@@
                    top5_potential_confounders = paste0(exposure_description, collapse = "; "), 
-                   max_C = max(corr_due_to_confounding)) %>%
+                   max_C = max(corr_due_to_confounding_MVMR)) %>%
   dplyr::rename(trait_ID = outcome_ID) %>%
@@
 
+write.table(corr_potential_trait_confounders_pos_MVMR_summary, "output/tables/C_summ_MVMR_result.csv")
+# 
+# corr_potential_trait_confounders_pos_MVMR_summary <- corr_potential_trait_confounders_pos_MVMR_summary %>% 
+#   group_by(outcome_ID) %>%
+#   mutate(corr_due_to_confounding_MVMR_summ = sum(corr_due_to_confounding_MVMR)) %>%
+#   mutate(corr_due_to_confounding_summ = sum(corr_due_to_confounding))
+
 ```
@@
 
-max_C_data <- corr_potential_trait_confounders_pos %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
-  arrange(-corr_due_to_confounding_ratio) %>%
+max_C_data <- corr_potential_trait_confounders_pos %>% filter(sig_confounder) %>% 
+  #arrange(-corr_due_to_confounding_ratio) %>%
   mutate(total_potential_confounders = n()) %>%
-  slice_head(n = 1) %>%
+  #slice_head(n = 1) %>%
   left_join(couple_MR_vs_trait_corr %>% 
@@
   mutate(same_trait_corr_minus_MR_est = abs(outcome_ID_couple_r)- abs(outcome_ID_IVW_beta)) %>%
-
   dplyr::rename(Z_beta = IVW_beta) %>%
@@
   mutate(Z_prod_se = sqrt(Z_prod_var)) %>%
+  group_by(outcome_ID) %>% 
+  mutate(corr_due_to_confounding_summ = sum(corr_due_to_confounding)) %>%
+  mutate(corr_due_to_confounding_summ_SE = sqrt(sum(Z_prod_var))) %>%
   mutate(same_trait_corr_minus_MR_est_se = sqrt(variance_of_sum(outcome_ID_IVW_se, outcome_ID_couple_r_se))) %>%
@@
   ggplot2::geom_errorbar(
-    ggplot2::aes(ymin = same_trait_corr_minus_MR_est - 1.96*same_trait_corr_minus_MR_est_se, 
-                 ymax = same_trait_corr_minus_MR_est + 1.96*same_trait_corr_minus_MR_est_se),
+    ggplot2::aes(ymin = same_trait_corr_minus_MR_est - 1.96*corr_due_to_confounding_summ_SE, 
+                 ymax = same_trait_corr_minus_MR_est + 1.96*corr_due_to_confounding_summ_SE),
     colour = "grey",
@@
   ggplot2::geom_errorbar(
-    ggplot2::aes(xmin = corr_due_to_confounding - 1.96*Z_prod_se, 
-                 xmax = corr_due_to_confounding + 1.96*Z_prod_se),
+    ggplot2::aes(xmin = corr_due_to_confounding_summ - 1.96*Z_prod_se, 
+                 xmax = corr_due_to_confounding_summ + 1.96*Z_prod_se),
     colour = "grey",
@@
 
+```{r large_results_table}
+
+traits_description <- traits_final %>% dplyr::select(Neale_pheno_ID, description, variable_type, category, r2, N_pairs, num_IVs_pass_het) %>% as_tibble() %>% 
+  dplyr::rename(num_IVs = num_IVs_pass_het) %>%
+  dplyr::rename(n_pairs = N_pairs) %>%
+  dplyr::rename(pheno_couple_r2 = r2) %>% dplyr::rename(exposure_ID = Neale_pheno_ID)
+
+# changed diff_p name 
+MR_result_table <- corr_vs_MR_effects_dat %>% dplyr::select(exposure_ID, Correlation, couple_r_se, n_exposure, n_outcome, MR_est, IVW_se, IVW_pval, diff_p) %>%
+  dplyr::rename(pheno_couple_r = Correlation) %>%
+  dplyr::rename(pheno_couple_r_se = couple_r_se) %>%
+  dplyr::rename(IVW_beta = MR_est) %>%
+  dplyr::rename(couple_r_vs_IVW_diff_pval = diff_p)
+
+AM_sex_spec_result <-
+  household_MR_binned_het_sex_spec %>% filter(same_trait == TRUE) %>% dplyr::select(exposure_ID, exposure_sex, IVW_beta, IVW_se, IVW_pval, sex_het_pval) %>% unique()
+
+AM_binned_result <- 
+  household_MR_binned_het_SNPmeta %>% filter(same_trait) %>% dplyr::select(exposure_ID, grouping_var, bin_slope_beta_wt, bin_slope_se_wt, bin_slope_pval_wt) %>% 
+  mutate(grouping_var_rename = case_when(grouping_var=="age_even_bins" ~ "age",
+                                         grouping_var=="time_together_even_bins" ~ "time_together")) %>% dplyr::select(-grouping_var) %>%
+  dplyr::rename(bin_slope_beta_MR_wt = bin_slope_beta_wt) %>%
+  dplyr::rename(bin_slope_se_MR_wt = bin_slope_se_wt) %>%
+  dplyr::rename(bin_slope_pval_MR_wt = bin_slope_pval_wt)
+
+sex_differences_summary <- AM_sex_spec_result %>% pivot_wider(names_from = exposure_sex, values_from = c(IVW_beta, IVW_se, IVW_pval)) 
+binned_summary <- AM_binned_result %>% pivot_wider(names_from = grouping_var_rename, values_from = c(bin_slope_beta_MR_wt, bin_slope_se_MR_wt, bin_slope_pval_MR_wt))
+
+binned_corr_summary <- binned_pheno_corr_summary %>% 
+  mutate(grouping_var_rename = case_when(group=="age_even_bins" ~ "age",
+                                         group=="time_together_even_bins" ~ "time_together")) %>% unique() %>% dplyr::select(-group) %>%
+  pivot_wider(names_from = grouping_var_rename, values_from = c(bin_slope_beta_pearson, bin_slope_se_pearson, 
+                                                                bin_slope_pval_pearson,
+                                                                bin_slope_beta_spearman, bin_slope_se_spearman, bin_slope_pval_spearman)) %>%
+  dplyr::rename(exposure_ID = trait)
+
+df_list <- list(traits_description, MR_result_table, sex_differences_summary, binned_summary, binned_corr_summary)
+final_table <- df_list %>% reduce(full_join, by='exposure_ID') %>%
+  dplyr::rename(trait_ID = exposure_ID) %>%
+  dplyr::rename(trait_description = description) %>%
+  dplyr::select(trait_description, trait_ID, everything())
+
+```
+
 ```{r write_tables, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
@@
 ## Supplementary tables
-write.xlsx_wrapper(AM_result_sig.DT, file=supplementary_tables_file, sheetName="Table 1 BF_sig_same_trait_MR", append=TRUE, row.names=FALSE)
-write.xlsx_wrapper(nominally_sig_sex_differences, file=supplementary_tables_file, sheetName="Table 2 nominally_sig_sex_diff", append=TRUE, row.names=FALSE)
-write.xlsx_wrapper(pheno_corr_pearson_sig.DT, file=supplementary_tables_file, sheetName="Table 3 pheno_binned_corrs", append=TRUE, row.names=FALSE)
-write.xlsx_wrapper(corr_vs_MR_effects_sig_dat, file=supplementary_tables_file, sheetName="Table 4 sig_corr_vs_MR", row.names=FALSE)
-write.xlsx_wrapper(sig_confounders, file=supplementary_tables_file, sheetName="Table 5 sig_confounders", append=TRUE, row.names=FALSE)
-write.xlsx_wrapper(paths_summary, file=supplementary_tables_file, sheetName="Table 6 paths_summary", append=TRUE, row.names=FALSE)
-
-```
+write.xlsx_wrapper(final_table, file=supplementary_tables_file, sheetName="Supp Table 1 complete_summary", row.names=FALSE)
+write.xlsx_wrapper(AM_result_sig.DT, file=supplementary_tables_file, sheetName="Supp Table 2 BF_sig_same_trait_MR", append=TRUE, row.names=FALSE)
+write.xlsx_wrapper(nominally_sig_sex_differences, file=supplementary_tables_file, sheetName="Supp Table 3 nominally_sig_sex_diff", row.names=FALSE)
+write.xlsx_wrapper(pheno_corr_pearson_sig.DT, file=supplementary_tables_file, sheetName="Supp Table 4 pheno_binned_corrs", append=TRUE, row.names=FALSE)
+write.xlsx_wrapper(corr_vs_MR_effects_sig_dat, file=supplementary_tables_file, sheetName="Supp Table 5 sig_corr_vs_MR", append = TRUE, row.names=FALSE)
+write.xlsx_wrapper(sig_confounders, file=supplementary_tables_file, sheetName="Supp Table 6 sig_confounders", append=TRUE, row.names=FALSE)
+write.xlsx_wrapper(paths_summary, file=supplementary_tables_file, sheetName="Supp Table 7 paths_summary", append=TRUE, row.names=FALSE)
+
+```

### 2022-09-01 analysis/manuscript.Rmd
fix small typo

@@
 write.table(corr_potential_trait_confounders_pos_MVMR_summary, "output/tables/C_summ_MVMR_result.csv")
-# 
-# corr_potential_trait_confounders_pos_MVMR_summary <- corr_potential_trait_confounders_pos_MVMR_summary %>% 
-#   group_by(outcome_ID) %>%
-#   mutate(corr_due_to_confounding_MVMR_summ = sum(corr_due_to_confounding_MVMR)) %>%
-#   mutate(corr_due_to_confounding_summ = sum(corr_due_to_confounding))
+
+corr_potential_trait_confounders_pos_MVMR_summary <- corr_potential_trait_confounders_pos_MVMR_summary %>% 
+  group_by(outcome_ID) %>%
+  mutate(corr_due_to_confounding_MVMR_summ = sum(corr_due_to_confounding_MVMR)) %>%
+  mutate(corr_due_to_confounding_summ = sum(corr_due_to_confounding))
 
@@
   ggplot2::geom_errorbar(
-    ggplot2::aes(ymin = same_trait_corr_minus_MR_est - 1.96*corr_due_to_confounding_summ_SE, 
-                 ymax = same_trait_corr_minus_MR_est + 1.96*corr_due_to_confounding_summ_SE),
+    ggplot2::aes(ymin = same_trait_corr_minus_MR_est - 1.96*same_trait_corr_minus_MR_est_se, 
+                 ymax = same_trait_corr_minus_MR_est + 1.96*same_trait_corr_minus_MR_est_se),
     colour = "grey",
@@
   ggplot2::geom_errorbar(
-    ggplot2::aes(xmin = corr_due_to_confounding_summ - 1.96*Z_prod_se, 
-                 xmax = corr_due_to_confounding_summ + 1.96*Z_prod_se),
+    ggplot2::aes(xmin = corr_due_to_confounding_summ - 1.96*corr_due_to_confounding_summ_SE, 
+                 xmax = corr_due_to_confounding_summ + 1.96*corr_due_to_confounding_summ_SE),
     colour = "grey",

### 2022-09-02 analysis/manuscript.Rmd
fix small typo
remove duplicate rows on `max_C_data`

@@
                                TRUE ~ Phenotype)) %>%
-  mutate(label_phenos = case_when(outcome_description %in% interesting_phenotypes ~ TRUE, TRUE ~ FALSE))
+  mutate(label_phenos = case_when(outcome_description %in% interesting_phenotypes ~ TRUE, TRUE ~ FALSE)) %>% 
+  dplyr::select(Phenotype, label_phenos, same_trait_corr_minus_MR_est, corr_due_to_confounding_summ, 
+                same_trait_corr_minus_MR_est_se, corr_due_to_confounding_summ_SE) %>% unique()
 
 max_C_plot <- ggplot2::ggplot(data = max_C_data, 
-                              ggplot2::aes(x = corr_due_to_confounding, 
+                              ggplot2::aes(x = corr_due_to_confounding_summ, 
                                            y = same_trait_corr_minus_MR_est, 

### 2022-09-02 analysis/manuscript.Rmd
change label to C sum

@@
   theme_minimal() +
-  ggplot2::labs(x = paste0("max C"),
+  ggplot2::labs(x = paste0("C sum"),
                 y = paste("Correlation - MR-estimate")) +

### 2022-09-07 analysis/manuscript.Rmd
change label for "C-sum"

@@
   theme_minimal() +
-  ggplot2::labs(x = paste0("C sum"),
+  ggplot2::labs(x = paste0("C-sum"),
                 y = paste("Correlation - MR-estimate")) +

### 2022-09-07 analysis/manuscript.Rmd
update supplemental tables

@@
 tar_load(proxyMR_comparison_summary_yiyp_adj_SNPmeta_prune)
-
-data <- proxyMR_comparison_summary_yiyp_adj_SNPmeta_prune
+tar_load(proxyMR_comparison_summary_yiyp_adj_SNPmeta)
+
+data <- proxyMR_comparison_summary_yiyp_adj_SNPmeta
 data$rho_resid <- resid(lm(rho_beta ~ gam_beta, data = data))
@@
   dplyr::rename(trait_description = description) %>%
-  dplyr::select(trait_description, trait_ID, everything())
+  dplyr::select(trait_description, trait_ID, variable_type, category, n_pairs, 
+                pheno_couple_r2, pheno_couple_r, pheno_couple_r_se, everything())
 
@@
 write.xlsx_wrapper(AM_result_sig.DT, file=supplementary_tables_file, sheetName="Supp Table 2 BF_sig_same_trait_MR", append=TRUE, row.names=FALSE)
-write.xlsx_wrapper(nominally_sig_sex_differences, file=supplementary_tables_file, sheetName="Supp Table 3 nominally_sig_sex_diff", row.names=FALSE)
+write.xlsx_wrapper(nominally_sig_sex_differences, file=supplementary_tables_file, sheetName="Supp Table 3 nominally_sig_sex_diff", append=TRUE, row.names=FALSE)
 write.xlsx_wrapper(pheno_corr_pearson_sig.DT, file=supplementary_tables_file, sheetName="Supp Table 4 pheno_binned_corrs", append=TRUE, row.names=FALSE)

### 2022-09-07 analysis/manuscript.Rmd
change add C_summ column to supp tables

@@
                    potential_confounders = paste0(exposure_description, collapse = "; "), 
-                   max_C = max(corr_due_to_confounding_MVMR)) %>%
+                   C_summ = corr_due_to_confounding_summ[1]) %>%
   dplyr::rename(trait_ID = outcome_ID) %>%
@@
 write.xlsx_wrapper(final_table, file=supplementary_tables_file, sheetName="Supp Table 1 complete_summary", row.names=FALSE)
-write.xlsx_wrapper(AM_result_sig.DT, file=supplementary_tables_file, sheetName="Supp Table 2 BF_sig_same_trait_MR", append=TRUE, row.names=FALSE)
-write.xlsx_wrapper(nominally_sig_sex_differences, file=supplementary_tables_file, sheetName="Supp Table 3 nominally_sig_sex_diff", append=TRUE, row.names=FALSE)
-write.xlsx_wrapper(pheno_corr_pearson_sig.DT, file=supplementary_tables_file, sheetName="Supp Table 4 pheno_binned_corrs", append=TRUE, row.names=FALSE)
-write.xlsx_wrapper(corr_vs_MR_effects_sig_dat, file=supplementary_tables_file, sheetName="Supp Table 5 sig_corr_vs_MR", append = TRUE, row.names=FALSE)
+#write.xlsx_wrapper(AM_result_sig.DT, file=supplementary_tables_file, sheetName="Supp Table 2 BF_sig_same_trait_MR", append=TRUE, row.names=FALSE)
+#write.xlsx_wrapper(nominally_sig_sex_differences, file=supplementary_tables_file, sheetName="Supp Table 3 nominally_sig_sex_diff", append=TRUE, row.names=FALSE)
+#write.xlsx_wrapper(pheno_corr_pearson_sig.DT, file=supplementary_tables_file, sheetName="Supp Table 4 pheno_binned_corrs", append=TRUE, row.names=FALSE)
+#write.xlsx_wrapper(corr_vs_MR_effects_sig_dat, file=supplementary_tables_file, sheetName="Supp Table 5 sig_corr_vs_MR", append = TRUE, row.names=FALSE)
 write.xlsx_wrapper(sig_confounders, file=supplementary_tables_file, sheetName="Supp Table 6 sig_confounders", append=TRUE, row.names=FALSE)

### 2022-09-07 analysis/manuscript.Rmd
fix small typo

@@
   dplyr::summarize(description = outcome_description[1],
+
                    total_potential_confounders = total_potential_confounders[1],
                    potential_confounders = paste0(exposure_description, collapse = "; "), 
-                   C_summ = corr_due_to_confounding_summ[1]) %>%
+                   max_C = max(corr_due_to_confounding_MVMR)) %>%
   dplyr::rename(trait_ID = outcome_ID) %>%
@@
 
+sig_confounders <- corr_potential_trait_confounders_pos_MVMR_summary %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
+  arrange(-corr_due_to_confounding_MVMR) %>%
+  mutate(total_potential_confounders = n()) %>%
+  dplyr::summarize(description = outcome_description[1],
+
+                   total_potential_confounders = total_potential_confounders[1],
+                   potential_confounders = paste0(exposure_description, collapse = "; "), 
+                   max_C = max(corr_due_to_confounding_MVMR)) %>%
+  dplyr::rename(trait_ID = outcome_ID) %>%
+  dplyr::rename(trait_description = description) %>% 
+  dplyr::select(trait_description, trait_ID, everything())
+
 ```
@@
 
-```{r max_C_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE, fig.height=10, dpi=600}
-
-max_C_data <- corr_potential_trait_confounders_pos %>% filter(sig_confounder) %>% 
+```{r summ_C_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE, fig.height=10, dpi=600}
+
+summ_C_data <- corr_potential_trait_confounders_pos %>% filter(sig_confounder) %>% 
   #arrange(-corr_due_to_confounding_ratio) %>%
@@
 
-max_C_plot <- ggplot2::ggplot(data = max_C_data, 
+max_C_plot <- ggplot2::ggplot(data = summ_C_data, 
                               ggplot2::aes(x = corr_due_to_confounding_summ, 

### 2022-09-09 analysis/manuscript.Rmd
change name to summ C from max C

@@
 
-max_C_plot <- ggplot2::ggplot(data = summ_C_data, 
+summ_C_plot <- ggplot2::ggplot(data = summ_C_data, 
                               ggplot2::aes(x = corr_due_to_confounding_summ, 
@@
   theme(legend.title=element_blank()) +
-  geom_text_repel(data = subset(max_C_data, label_phenos==TRUE),
-                  nudge_y = -0.25 - subset(max_C_data, label_phenos==TRUE)$same_trait_corr_minus_MR_est, size = 3.25,
+  geom_text_repel(data = subset(summ_C_data, label_phenos==TRUE),
+                  nudge_y = -0.25 - subset(summ_C_data, label_phenos==TRUE)$same_trait_corr_minus_MR_est, size = 3.25,
                   direction = "x", show.legend = FALSE, box.padding = 0.5)
 
-max_C_plot
+summ_C_plot
 

### 2022-09-09 analysis/manuscript.Rmd
fix typo

@@
 
-## combine `max_C_plot` with `corr_vs_MR_effects_plot`
-
-corr_vs_MR_comb <- plot_grid(plotlist = list(corr_vs_MR_effects_plot, max_C_plot), nrow=2, ncol = 1, byrow = T, labels="AUTO")
+## combine `summ_C_plot` with `corr_vs_MR_effects_plot`
+
+corr_vs_MR_comb <- plot_grid(plotlist = list(corr_vs_MR_effects_plot, summ_C_plot), nrow=2, ncol = 1, byrow = T, labels="AUTO")
 

### 2022-09-09 analysis/manuscript.Rmd
add summ C to output

@@
 
-sig_confounders <- corr_potential_trait_confounders_pos_MVMR_summary %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
-  arrange(-corr_due_to_confounding_MVMR) %>%
+sig_confounders <- corr_potential_trait_confounders_pos %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
+  arrange(-corr_due_to_confounding) %>%
   mutate(total_potential_confounders = n()) %>%
@@
                    potential_confounders = paste0(exposure_description, collapse = "; "), 
-                   max_C = max(corr_due_to_confounding_MVMR)) %>%
+                   summ_C = sum(corr_due_to_confounding)) %>%
   dplyr::rename(trait_ID = outcome_ID) %>%
@@
 #write.xlsx_wrapper(corr_vs_MR_effects_sig_dat, file=supplementary_tables_file, sheetName="Supp Table 5 sig_corr_vs_MR", append = TRUE, row.names=FALSE)
-write.xlsx_wrapper(sig_confounders, file=supplementary_tables_file, sheetName="Supp Table 6 sig_confounders", append=TRUE, row.names=FALSE)
-write.xlsx_wrapper(paths_summary, file=supplementary_tables_file, sheetName="Supp Table 7 paths_summary", append=TRUE, row.names=FALSE)
-
-```
+write.xlsx_wrapper(sig_confounders, file=supplementary_tables_file, sheetName="Supp Table 3 sig_confounders", append=TRUE, row.names=FALSE)
+write.xlsx_wrapper(paths_summary, file=supplementary_tables_file, sheetName="Supp Table 4 paths_summary", append=TRUE, row.names=FALSE)
+
+```

### 2022-11-14 analysis/manuscript.Rmd
reformatting for misc thing for final publications

@@
 
-```{r setup, include = FALSE}
+```{r setup, include = FALSE, dev=c('png', 'pdf')}
 
@@
 library(xlsx)
+library(rstatix)
+
+knitr::opts_chunk$set(dev =c('png', 'pdf'))
+
+```
+
 ```
@@
   #geom_text(aes(x=1,y=2,label=""),show.legend=F) +
-  ggplot2::labs(x = paste0("Couple correlation"),
+  ggplot2::labs(x = paste0("Partner correlation"),
                 y = paste("MR-estimate")) +
@@
 
+sig_confounder_plus_SE <- left_join(sig_confounders, summ_C_data %>% dplyr::select(outcome_ID, corr_due_to_confounding_summ_SE), by = c("trait_ID"="outcome_ID")) %>%
+dplyr::rename(summ_C_se = summ_C)
+
 summ_C_plot <- ggplot2::ggplot(data = summ_C_data, 
@@
 
+stat.test <- data_boxplot %>% 
+  t_test(beta ~ data, comparisons = cmpr, paired = T) %>%
+  mutate(myformatted.p = paste0(pretty_round(p))) %>% 
+  mutate(y.position = c(0.6, 0.4, 0.5))
+
 boxplot <- data_boxplot %>%
@@
 
-  stat_compare_means(data = data_boxplot, comparisons = cmpr, label.y = c(0.6, 0.4, 0.5), method = "t.test" , paired =T) + #,
-                     # label = "p.signif",
-                     # symnum.args = list(cutpoints = c(0, 0.05, 1),
-                     #                    symbols = c("*", "ns"))) +
+  stat_pvalue_manual(stat.test, label = "myformatted.p") +
+
+  # stat_compare_means(data = data_boxplot, comparisons = cmpr, label.y = c(0.6, 0.4, 0.5), method = "t.test" , paired =T,
+  #                   label = "p.format") +
+  #                   # symnum.args = list(cutpoints = c(0, 0.05, 1),
+  #                   #                    symbols = c("*", "ns"))) +
 
@@
 #write.xlsx_wrapper(corr_vs_MR_effects_sig_dat, file=supplementary_tables_file, sheetName="Supp Table 5 sig_corr_vs_MR", append = TRUE, row.names=FALSE)
-write.xlsx_wrapper(sig_confounders, file=supplementary_tables_file, sheetName="Supp Table 3 sig_confounders", append=TRUE, row.names=FALSE)
-write.xlsx_wrapper(paths_summary, file=supplementary_tables_file, sheetName="Supp Table 4 paths_summary", append=TRUE, row.names=FALSE)
-
-```
+write.xlsx_wrapper(sig_confounder_plus_SE, file=supplementary_tables_file, sheetName="Supp Table 2 sig_confounders", append=TRUE, row.names=FALSE)
+write.xlsx_wrapper(paths_summary, file=supplementary_tables_file, sheetName="Supp Table 3 paths_summary", append=TRUE, row.names=FALSE)
+
+```

### 2022-11-15 analysis/manuscript.Rmd
save figures as pdf

@@
 
-```{r setup, include = FALSE, dev=c('png', 'pdf')}
+```{r setup, include = FALSE}
 
@@
 
-knitr::opts_chunk$set(dev =c('png', 'pdf'))
+#knitr::opts_chunk$set(dev =c('png', 'pdf'))
 
@@
 pearson_vs_spearman_plot
+ggsave("docs/figure/manuscript.Rmd/pearson_vs_spearman_plot.pdf", pearson_vs_spearman_plot, device = cairo_pdf)
+
 ```
@@
 
-  theme_minimal() +
+  #theme_minimal() +
 
@@
   geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
-  theme_ipsum()
+  theme_ipsum_rc()
 
 pheno_corr_plot
+
+ggsave("docs/figure/manuscript.Rmd/pheno_corr_plot.pdf", pheno_corr_plot, device = cairo_pdf)
 
@@
 corr_vs_MR_effects_plot
+
+ggsave("docs/figure/manuscript.Rmd/corr_vs_MR_effects_plot.pdf", corr_vs_MR_effects_plot, device = cairo_pdf)
 
@@
 
+ggsave("docs/figure/manuscript.Rmd/summ_C_plot.pdf", summ_C_plot, device = cairo_pdf)
+
 ## combine `summ_C_plot` with `corr_vs_MR_effects_plot`
@@
 corr_vs_MR_comb
+
+ggsave("docs/figure/manuscript.Rmd/corr_vs_MR_comb.pdf", corr_vs_MR_comb, device = cairo_pdf)
 
@@
 
+ggsave("docs/figure/manuscript.Rmd/pheno_binned_figures.pdf", pheno_binned_figures, device = cairo_pdf)
+
 ```
@@
 figures_grid
+
 ```
@@
 figures_grid
+
+ggsave("docs/figure/manuscript.Rmd/causal_confounder_plot.pdf", figures_grid, device = cairo_pdf)
 
@@
   scale_fill_viridis(discrete = TRUE, alpha=0.6, option="A") +
-  theme_ipsum() +
+  theme_ipsum_rc() +
   theme(
@@
 figures_grid
+
+ggsave("docs/figure/manuscript.Rmd/boxplot_combined.pdf", figures_grid, device = cairo_pdf)
 
